### 2017-03-20 draft-new/draft.tex
2017/03/20
2017/03/20

% !TeX spellcheck = en_US	
\documentclass[10pt,journal,compsoc]{IEEEtran}

\usepackage{color}

% *** CITATION PACKAGES ***
\ifCLASSOPTIONcompsoc
\usepackage[nocompress]{cite}
\else
\usepackage{cite}
\fi

\begin{document}

\title{Building and Checking Suffix Array in External Memory}

\author{Yi~Wu,
	Ge~Nong,
	Wai~Hong~Chan,
	and~Bin~Lao
	\IEEEcompsocitemizethanks{
		\IEEEcompsocthanksitem Y. Wu, G. Nong (corresponding author) and B. Lao are with the Department of Computer Science, Sun Yat-sen University, Guangzhou 510275, China. E-mails: wu.yi.christian@gmail.com, issng@mail.sysu.edu.cn, Laobin@mail3.sysu.edu.cn.

		\IEEEcompsocthanksitem Wai Hong Chan (corresponding author) is with the Department of Mathematics and Information Technology, The Education University of Hong Kong, Hong Kong. E-mail: waihchan@ied.edu.hk.
}}% <-this % stops a space

\IEEEtitleabstractindextext{%
\begin{abstract}

Suffix array~(SA) can be built within linear time and space by means of the induced sorting~(IS) method. The recently proposed IS suffix sorter fSAIS has a performance competitive to that of the state-of-the-art parallel construction algorithms in external memory, which indicates a great potential for improving the other disk-based IS alternatives by engineering them carefully. In view of this, we redesign the reduction phase of DSA-IS by using a new substring sorting and naming method and implement the variant, called DSA-IS+, by using several space-saving techinques previously applied to fSAIS.

A constructed SA should be checked to ensure its correctness. We present in this paper a method that enables any IS suffix sorter to perform building and checking simultaneously, where the consumption for checking is negligible in comparison with that for building in our experiments.

\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Suffix array, construction and verification, external memory.
\end{IEEEkeywords}}

% make the title area
\maketitle

\IEEEdisplaynontitleabstractindextext

\IEEEpeerreviewmaketitle

\section{Introduction}\label{sec:introduction}

The suffix array~(SA)~\cite{Manber1993} is an essential data structure for string processing and information retrieval. Among the existing internal-memory algorithms for SA construction, SA-IS~\cite{Nong11} achieves the optimal time and space complexities using the induced sorting principle, where the key operation is to retrieve the preceding characters of sorted suffixes in order for inducing the lexical order of unsorted ones. To handle massive datasets , several external memory algorithms have been proposed for building massive suffix arrays in recent years, e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014}, pSAscan~\cite{Karkkainen2015}, eSAIS~\cite{Bingmann2013}, EM-SA-DS~\cite{Nong2014} and DSA-IS~\cite{Nong2015}. Among them, the latter three algorithms are based on the induced sorting~(IS) method described in SA-IS~\cite{Nong2011}.

 This algorithm employs the IS method to induce the lexical order of all the suffixes from a selected subset.

 This algorithm uses the induced sorting method to

In the past decades, great effort has been taken to study efficient algorithms for SA construction. th

extensive works have been put on designing time and space efficient suffix sorting algorithms

data structure that has been widely used in many string processing applications, e.g., biological sequence alignment, time series analysis and text clustering. Given an input string, traversing its suffix tree can be emulated by using the corresponding enhanced suffix array~\cite{Abouelhodaa2004}, which mainly consists of the suffix and the longest common prefix arrays. It has been realized that the application scope of an index mainly depends on the construction speed and the space requirement. This leads to intensive works on designing time and space efficient suffix sorting algorithms over the past decade, assuming different computation models such as internal memory, external memory, parallel and distributed models. Particularly,

The basic idea behind the induced sorting method is to induce the lexicographical order of all the substrings/suffixes from a sorted subset of substrings/suffixes. Following the idea, an IS-based suffix sorting algorithm is typically comprised of a reduction phase for sorting and naming substrings to reduce a string $x[0,n)$ to a short string $x1[0,n1)$ with $n1 \le \frac{1}{2}n$ and an induction phase for sorting suffixes to induce ${\sf SA}(x)$ from ${\sf SA}(x1)$. During the two phases, the key operation is to retrieve the preceding character of a sorted substring/suffix. This can be done very quickly when $x$ is fully accommodated in the internal memory, but will become slow when $x$ resides in the external memory, as each operation takes a random disk access. For a high I/O efficiency, eSAIS, EM-SA-DS and DSA-IS use different auxiliary data structures to retrieve the preceding characters in a disk-friendly way. Particularly, both eSAIS and EM-SA-DS split a long substring into pieces and represent each piece by a fixed-size tuple, while DSA-IS does not. With an elaborate arrangement of the I/O operations, the programs for these three algorithms are competitive with those for others in terms of both time and space efficiencies.

Among the existing internal-memory suffix sorters, SA-IS~\cite{Nong11} achieves the optimal time and space complexities using the induced sorting method, the key operation of which is to retrieve the preceding characters of sorted suffixes in order for inducing the lexical order of unsorted ones. In the past five years, several works adapted SA-IS to design suffix sorters specific for EM models~\cite{Nong14, Nong15, Karkkainen2014, Liu2015}. These variants use different approaches to retrieve the preceding characters by sequential I/O operations. However, they all suffer from a space bottleneck for taking at least twice disk space as SA-IS on real-world datasets. Recently, it was presented in~\cite{xxx} a new engineering version of SA-IS that achieves nearly optimal space efficiency, indicating a great potential for improving the other disk-based IS alternatives by engineering them carefully.

Spae

\section{Conclusion}
xxx

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,bibfile}

\end{document}

### 2017-03-20 tc-rejected/paper.tex
2017/03/20
2017/03/20

\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage[nocompress]{cite}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{url}
\usepackage{multirow}
\usepackage[ruled]{algorithm2e}
\usepackage{enumitem}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{observation}[theorem]{Observation}

\begin{document}

\title{Efficient Suffix Array Construction and Checking in External Memory}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{Yi~Wu,
        Ge~Nong,
        Wai~Hong~Chan,
        and~Bin~Lao
\IEEEcompsocitemizethanks{
\IEEEcompsocthanksitem Y. Wu, G. Nong (corresponding author) and B. Lao are with the Department of Computer Science, Sun Yat-sen University, Guangzhou 510275, China. E-mails: wu.yi.christian@gmail.com, issng@mail.sysu.edu.cn, Laobin@mail3.sysu.edu.cn.

\IEEEcompsocthanksitem Wai Hong Chan (corresponding author) is with the Department of Mathematics and Information Technology, The Education University of Hong Kong, Hong Kong. E-mail: waihchan@ied.edu.hk.
}}% <-this % stops a space

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Advanced Demo of IEEEtran.cls for IEEE Computer Society Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2015 IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society journal
% papers don't need this extra clearance.)

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}

A suffix array (SA) must be checked to ensure its correctness. The existing SA checker verifies an SA by performing two passes of integer sorts that are time and space consuming especially when running in the external memory. Currently, building and checking an SA are done in sequence, where the time and the space bottlenecks are caused by building and checking the SA, respectively.
In this paper, a fast and lightweight checking method is proposed to enable any induced sorting (IS) suffix sorting algorithm to verify a suffix array while it is being built. For performance evaluation, we first redesign the reduction phase of the external memory IS algorithm DSA-IS by employing the new substring sorting and naming methods to produce an optimized alternative DSA-IS+, and then apply the new checking method to DSA-IS+ for building and checking an SA simultaneously. Our experimental results indicate that, in comparison with that for building, the checking time and space complexities of eSAIS are substantial, but that of DSA-IS+ are negligible. As a result, DSA-IS+ can build and check an SA in a higher total time and space efficiencies.

\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
suffix array, construction and verification, external memory.
\end{IEEEkeywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when compsoc mode
% is not selected <OR> if conference mode is selected - because compsoc
% conference papers position the abstract like regular (non-compsoc)
% papers do!
\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}

\subsection{Background}
\IEEEPARstart{T}{he} suffix array~(SA)~\cite{Manber1993} is a data structure that has been widely used in many string processing applications, e.g., biological sequence alignment, time series analysis and text clustering. Given an input string, traversing its suffix tree can be emulated by using the corresponding enhanced suffix array~\cite{Abouelhodaa2004}, which mainly consists of the suffix and the longest common prefix arrays. It has been realized that the application scope of an index mainly depends on the construction speed and the space requirement. This leads to intensive works on designing time and space efficient suffix sorting algorithms over the past decade, assuming different computation models such as internal memory, external memory, parallel and distributed models. Particularly, to keep pace with the fast development of the data sampling techniques, several external memory algorithms have been proposed for building massive suffix arrays in recent years, e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014}, pSAscan~\cite{Karkkainen2015}, eSAIS~\cite{Bingmann2013}, EM-SA-DS~\cite{Nong2014} and DSA-IS~\cite{Nong2015}. Among them, the latter three algorithms are based on the induced sorting~(IS) method described in SA-IS~\cite{Nong2011}.

The basic idea behind the induced sorting method is to induce the lexicographical order of all the substrings/suffixes from a sorted subset of substrings/suffixes. Following the idea, an IS-based suffix sorting algorithm is typically comprised of a reduction phase for sorting and naming substrings to reduce a string $x[0,n)$ to a short string $x1[0,n1)$ with $n1 \le \frac{1}{2}n$ and an induction phase for sorting suffixes to induce ${\sf SA}(x)$ from ${\sf SA}(x1)$. During the two phases, the key operation is to retrieve the preceding character of a sorted substring/suffix. This can be done very quickly when $x$ is fully accommodated in the internal memory, but will become slow when $x$ resides in the external memory, as each operation takes a random disk access. For a high I/O efficiency, eSAIS, EM-SA-DS and DSA-IS use different auxiliary data structures to retrieve the preceding characters in a disk-friendly way. Particularly, both eSAIS and EM-SA-DS split a long substring into pieces and represent each piece by a fixed-size tuple, while DSA-IS does not. With an elaborate arrangement of the I/O operations, the programs for these three algorithms are competitive with those for others in terms of both time and space efficiencies.

\subsection{Problems to Solve}
Although the above three IS-based algorithms achieve great performance in comparison with prior arts, there still exist several drawbacks to be dealt with.

Firstly, the space requirement remains at a high level for all the three algorithms. In details, the peak disk usages for eSAIS, EM-SA-DS and DSA-IS are around $24n$, $31n$ and $20n$, respectively, the minimum of which is about three times as that for the lightweight suffix sorter SAscan (which I/O volume is super-linear). It is identified that DSA-IS reaches the peak disk usage when induced sorting substrings during the reduction phase. This reveals a need for designing new substring sorting and naming methods to improve the space efficiency in order to scale the problem size that can be tackled.

Secondly, a suffix array must be checked to ensure its correctness for later use. Currently, an SA built by each of these algorithms is verified using the SA checker described in~\cite{Karkkainen2003}, which performs two passes of external memory integer sorts that charge large disk space and long cpu time. As a result, the peak disk usage for both eSAIS and DSA-IS rises up to $26n$ while the time consumption also grows by 20 percent as before, leading to a substantial performance degradation especially when processing large-scale datasets. Therefore, it is desired to reduce the check overhead in case it becomes a performance bottleneck.

\subsection{Our Contribution}

Our contribution in this paper mainly includes the following aspects.

\begin{itemize}

	\item We redesign the reduction phase of DSA-IS by employing new methods for sorting and naming substrings. In our experiments, the program for the enhanced algorithm, called DSA-IS+, runs as fast as that for eSAIS while its peak disk usage is about two-thirds as that of eSAIS.

	\item We design a fast and lightweight SA checking method that can be seamlessly integrated into any IS-based suffix sorting algorithm. For performance evaluation, we combine this method with DSA-IS+ to produce a unified solution that can be employed to construct and check SA at the same time. It was observed that, by exploiting the use of our checking method, the performance overhead for verification is considerable small in comparison with that for construction. In details, the peak disk usage for the unified solution remains at the same level of that for DSA-IS+, while the growth in both time and I/O volume is negligible.

\end{itemize}

The rest of this paper is organized as follows. Section~\ref{sec:preliminary} gives the preliminaries in our presentation. Section~\ref{sec:dsais_plus} describes the substring sorting and naming methods for DSA-IS+, Section~\ref{sec:sachecker} the checking method and Section~\ref{sec:experiments} the experimental results. Finally, we come to the conclusion in Section~\ref{sec:conclusion}.

\section{Preliminaries}\label{sec:preliminary}

\subsection{Basic Notations}

Given an input string $x[0,n)=x[0]x[1]...x[n - 1]$ drawn from a totally ordered alphabet $\delta$, we assume $x[n - 1]$ to be the smallest character in $\delta$ that appears in $x$ only once. For short, we use ${\sf sub}(x,i)$ and ${\sf suf}(x,i)$ to denote a substring running from $x[i]$ to the leftmost LMS character on its right side and a suffix running from $x[i]$ to $x[n - 1]$, respectively. To follow lists the basic notations used in our presentation.

\begin{itemize}[itemindent = 0ex]
	\item L-type, S-type and LMS character/substring/suffix. We say $x[i]$ is S-type if (1) $i = n - 1$ or (2) $x[i] < x[i + 1]$ or (3) $x[i] = x[i+1]$ and $x[i+1]$ is S-type; otherwise, $x[i]$ is L-type. Furthermore, if $x[i]$ and $x[i - 1]$ are respectively S-type and L-type, then $x[i]$ is also an LMS character. Moreover, if $x[i]$ is L-type, S-type or LMS, then ${\sf sub}(x,i)$ and ${\sf suf}(x,i)$ are L-type, S-type or LMS, respectively.

	\item preceding and succeeding character/substring/suffix. We say $x[i - 1]$, ${\sf sub}(x, i - 1)$ and ${\sf suf}(x, i - 1)$ are the preceding character, substring and suffix for $x[i]$, ${\sf sub}(x, i)$ and ${\sf suf}(x, i)$, respectively. Similarly, we say $x[i + 1]$, ${\sf sub}(x, i + 1)$ and ${\sf suf}(x, i + 1)$ are the succeeding character, substring and suffix for $x[i]$, ${\sf sub}(x, i)$ and ${\sf suf}(x, i)$, respectively.

	\item ${\sf SA}$ and ${\sf STRA}$. We respectively use ${\sf SA}$ and ${\sf STRA}$ to denote the suffix and substring arrays. Specifically, both ${\sf SA}(x)$ and ${\sf STRA}(x)$ are permutations of $[0, n)$ such that ${\sf suf}(x, {\sf SA}(x)[i]) < {\sf suf}(x, {\sf SA}(x)[i + 1])$ and ${\sf sub}(x, {\sf SA}(x)[i]) \le {\sf sub}(x, {\sf SA}(x)[i + 1])$ in lexicographical order for all $i \in [0, n - 1)$.

	\item ${\sf BKT}$. Suffixes in ${\sf SA}(x)$ are naturally grouped into buckets according to their head characters. Given ${\sf BKT}({\sf SA}(x), ch)$, the bucket occupies a consecutive range in ${\sf SA}(x)$ that gathers all the suffixes with a head character $ch$. Furthermore, a bucket, say ${\sf BKT}({\sf SA}(x), ch)$, can be subdivided into ${\sf BKT_L}({\sf SA}(x), ch)$ and ${\sf BKT_S}({\sf SA}(x), ch)$ that only contain L-type and S-type suffixes in ${\sf BKT}({\sf SA}(x), ch)$, respectively. Similarly, we define ${\sf BKT}({\sf STRA}(x), ch)$, ${\sf BKT_L}({\sf STRA}(x), ch)$ and ${\sf BKT_S}({\sf STRA}(x), ch)$ for ${\sf STRA}(x)$.

	\item ${\sf ISA}$. Inverse suffix array, satisfying ${\sf ISA}(x)[{\sf SA}(x)[i]] = i$ for all $i \in [0,n)$.

	\item ${\sf PA}$. Position array, where ${\sf PA}(x)[i]$ records the position index of the $i$-th LMS substring from the left side of $x$.

\end{itemize}

\subsection{An Overview of DSA-IS}

As with other IS-based suffix sorting algorithms, DSA-IS mainly consists of a reduction phase followed by an induction phase. The induced sorting method is adopted in both two phases to sort substrings and suffixes, respectively.

\subsubsection{Reduction Phase}

Assume $x[0,n)$ and ${\sf STRA}(x)$ can be wholly accommodated into RAM, SA-IS performs the following three steps in the reduction phase to determine the lexicographical order of substrings.

\begin{itemize}[itemindent = 5ex]
	\item [step 1:] Scan $x$ rightward with $i$ decreasing from $n - 1$ to $0$. For each scanned LMS character $x[i]$, insert $i$ into the rightmost empty position of ${\sf BKT_S}({\sf STRA}(x), ch)$.

	\item [step 2:] Scan ${\sf STRA}(x)$ leftward with $i$ increasing from $0$ to $n - 1$. For each scanned $p = {\sf STRA}(x)[i]$, if $ch = x[p - 1]$ is the starting character of an L-type substring, then insert $p - 1$ into the leftmost empty position of ${\sf BKT_L}({\sf STRA}(x), ch)$.

	\item [step 3:] Scan ${\sf STRA}(x)$ rightward with $i$ decreasing from $n- 1$ to $0$. For each scanned $p = {\sf STRA}(x)[i]$, if $ch = x[p - 1]$ is the starting character of an S-type substring, then insert $p - 1$ into the rightmost empty position of ${\sf BKT_S}({\sf STRA}(x), ch)$.
\end{itemize}

The key operation of steps 2-3 is to retrieve the preceding character $x[p - 1]$ for currently scanned substring starting at $x[{\sf STRA}(x)[i]]$, where each operation takes a random access to the internal memory. A naive solution for extending the above 3-step procedure to external memory models is to replace each memory access with a disk access. However, this is not practical due to the large overhead for $\mathcal{O}(n)$ disk accesses. To reduce the I/O complexity, DSA-IS introduces the alternative ${\sf DSTRA}(x)$ for ${\sf STRA}(x)$ and employs I/O buffers to cache ${\sf DSTRA}(x)$ and other data structures for amortizing the disk access overhead, where the involved data structures are defined as below.

\begin{itemize}[itemindent = 0ex]

	\item ${\sf DSTRA}$, ${\sf DSTRA_L}$ and ${\sf DSTRA_S}$. DSAITEM arrays. For an input $x$, each item of ${\sf DSTRA}(x)$ associates with a substring ${\sf sub}(x, p)$ and mainly consists of the following three components:
	\begin{itemize}[itemindent = 0ex]
		\item $p$: position index for ${\sf sub}(x, p)$.
		\item $c$: head character for ${\sf sub}(x, p)$, that is $x[p]$.
		\item $t$: type of the preceding character $x[p - 1]$, set as 0 or 1 for L-type or S-type, respectively.
	\end{itemize}

	${\sf DSTRA_L}(x)$ and ${\sf DSTRA_S}(x)$ only contain the items of L-type and S-type substrings in ${\sf DSTRA}(x)$, respectively.

	\item ${\sf LMSDATA}$. A string array. Each item stores the characters of an LMS substring.

	\item ${\sf LMSNAME}$. An integer array. Each item stores the name of an LMS substring, where the name represents the substring's lexicographical order among all.

\end{itemize}

Notice that the above data structures organize the items in the external memory according to the lexicographical order of their corresponding substrings.

% Algorithm
\begin{algorithm}
	\SetAlgoNoLine
	\KwIn{$x$}
	\KwOut{$x1$}

	step 1: Partition $x$ into blocks $b_1, b_2, ..., b_k$.~\label{alg:dsais_reduction_phase:1} \\
	step 2: Induced sort the substrings of $b_i$ for all $i \in [1, k]$ to compute ${\sf DSTRA_L}(b_i)$, ${\sf DSTRA_S}(b_i)$ and ${\sf LMSDATA}(b_i)$. ~\label{alg:dsais_reduction_phase:2}\\
	step 3: Induced sort the substrings of $x$ from ${\sf DSTRA_L}(b_i)$, ${\sf DSTRA_S}(b_i)$ and ${\sf LMSDATA}(b_i)$ to compute ${\sf DSTRA}(x)$ and ${\sf LMSDATA}(x)$. Meanwhile, scan ${\sf DSTRA}(x)$ to compute ${\sf DSTRA_{LMS}}(x)$.~\label{alg:dsais_reduction_phase:3}\\
	step 4: Scan ${\sf DSTRA_{LMS}}(x)$ and ${\sf LMSDATA}(x)$ to name the sorted LMS substrings for producing ${\sf LMSNAME}(x)$. ~\label{alg:dsais_reduction_phase:4}\\
	step 5: Compute the reduced string $x1$ from ${\sf LMSNAME}(x)$ and ${\sf DSTRA_{LMS}}(x)$. ~\label{alg:dsais_reduction_phase:5}\\

	\caption{The Reduction Phase for DSA-IS}
	\label{alg:dsais_reduction_phase}
\end{algorithm}

The algorithmic framework of the reduction phase for DSA-IS is shown in Algorithm~\ref{alg:dsais_reduction_phase}, where each step is sketched below.

Step 1 scans $x$ leftward to sequentially retrieve the LMS substrings and inserts them one by one into the blocks with a capacity $m = \mathcal{O}(\sf M)$, where ${\sf M}$ is the size of the internal memory. In this way, each block is either a single-block composed of one LMS substring or a multi-block composed of at least two successive LMS substrings. Particularly, each multi-block contains no more than $m$ characters such that it can be processed in the internal memory during the two phases, while a single-block may not be so.

Step 2 adopts different strategies to tackle $b_i$ with respect to whether the block is a single-block or a multi-block. Specifically, if $b_i$ is a multi-block, then it computes ${\sf STRA}(b_i)$ by conducting the substring sorting algorithm of SA-IS and scans ${\sf STRA}(b_i)$ and $b_i$ to obtain ${\sf DSTRA_L}(b_i)$, ${\sf DSTRA_S}(b_i)$ and ${\sf LMSDATA}(b_i)$; otherwise, the substrings in the block are already sorted and it directly scans $b_i$ to obtain the three arrays.

Step 3 also computes ${\sf DSTRA}(x)$ following the induced sorting method. In details, when inducing substrings in ${\sf DSTRA}(x)$, it retrieves the head character of the L-type/S-type preceding substring for the currently scanned substring from the external memory by using sequential I/O operations. This is feasible because the corresponding DSAITEM of the preceding substring is located at the leftmost unvisited position in ${\sf DSTRA_L}(b_i)/{\sf DSTRA_S}(b_i)$, where the subscript $i$ of the block that contains the target item can be determined in $\mathcal{O}(1)$ time using $\mathcal{O}(\frac{n}{m})$ space~\cite{Nong2015}.

Step 4 names the sorted LMS substrings according to their lexicographical order. For each pair of the neighboring substrings in ${\sf DSTRA_{LMS}}(x)$, it conducts a string comparison to check if they are equal, where their characters can be sequentially retrieved from ${\sf LMSDATA}(x)$. If yes, then the two substrings have a same name; otherwise, the name for the latter is greater than the former by one.

Step 5 sorts the names in ${\sf LMSNAME}(x)$ by ${\sf DSTRA_{LMS}}(x)[i].p$ for all $i \in [0, n1)$ to produce the reduced string $x1$.

\subsubsection{Induction Phase}

% Algorithm
\begin{algorithm}
	\SetAlgoNoLine
	\KwIn{$x$, $\sf{PA}(x)$, ${\sf DSA}(x1)$}
	\KwOut{${\sf DSA}(x)$}

	Step 1: Compute ${\sf DSA_{LMS}}(x)$ from ${\sf DSA}(x1)$ and decompose ${\sf DSA_{LMS}}(x)$ into ${\sf DSA_{LMS}}(b_i)$ for all $i \in [1, k]$.~\label{alg:dsais_induction_phase:1}\\
	Step 2: Induced sort the suffixes of $b_i$ from ${\sf DSA_{LMS}}(b_i)$ to compute ${\sf DSA_L}(b_i)$ and ${\sf DSA_S}(b_i)$ for all $i \in [1, k]$.~\label{alg:dsais_induction_phase:2}\\
	Step 3: Induced sort the suffixes of $x$ from ${\sf DSA_L}(b_i)$ and ${\sf DSA_S}(b_i)$ to compute ${\sf DSA}(x)$.~\label{alg:dsais_induction_phase:3} \\

	\caption{The Induction Phase for DSA-IS}
	\label{alg:dsais_induction_phase}
\end{algorithm}

As shown in Algorithm~\ref{alg:dsais_induction_phase}, the induction phase of DSA-IS consists of three steps, where the data structures for the induction phase are similar to their counterparts for the reduction phase:

\begin{itemize}[itemindent = 0ex]
	\item ${\sf DSA_{LMS}}$. A DSAITEM array. For an input $x$, each item of ${\sf DSA_{LMS}}(x)$ is associated with an LMS suffix of $x$.

	\item ${\sf DSA}$, ${\sf DSA_L}$ and ${\sf DSA_S}$. DSAITEM arrays. For an input $x$, each item of ${\sf DSA}(x)$ is associated with a suffix of $x$. ${\sf DSA_L}(x)$ and ${\sf DSA_S}(x)$ only contain the items of L-type and S-type suffixes in ${\sf DSA}(x)$, respectively.

\end{itemize}

The first step in Algorithm~\ref{alg:dsais_induction_phase} is to induce the order of LMS substrings of $x$. For the purpose, it sorts the items of ${\sf DSA}(x1)$ by ${\sf DSA}(x1)[i].p$ for all $i\in [0, n1)$ to compute ${\sf ISA}(x1)$ and then sorts the items of ${\sf PA}(x)$ by ${\sf ISA}(x1)[j]$ for all $j\in [0,n1)$ to compute ${\sf DSA_{LMS}}(x)$. Afterward, it decomposes ${\sf DSA_{LMS}}(x)$ into ${\sf DSA_{LMS}}(b_i)$ for all $i \in [1, k]$ and reuses steps 2-3 of the reduction phase to compute ${\sf DSA}(x)$ by replacing ${\sf DSTRA_L}(b_i)$ and ${\sf DSTRA_S}(b_i)$ with ${\sf DSA_L}(b_i)$ and ${\sf DSA_S}(b_i)$, respectively.

\section{Details of DSA-IS+} \label{sec:dsais_plus}

It was observed from our experiments that, the program for DSA-IS outperforms that for eSAIS with respect to the space efficiency, but it runs slower than the latter due to the large I/O volume for sorting and naming substrings. In this section, we present two methods that can  substantially reduce the construction time and I/O volume of the reduction phase without a sacrifice in its high space efficiency. We demonstrate in Section~\ref{sec:experiments} that, by exploiting the use of these methods, the optimized version of DSA-IS, namely DSA-IS+, runs as fast as eSAIS and its peak disk usage is two-thirds as that of the latter.

\subsection{Method A} \label{subsec:dsais_plus_method_a}
Recall that in Algorithm~\ref{alg:dsais_reduction_phase}, DSA-IS first sorts LMS substrings in steps 2-3 and then names the substrings in their sorted order in step 4. These two procedures can be merged by making use of the following fact: $\forall i, j, \in [0, n)$, ${\sf sub}(x, i)$ and ${\sf sub}(x, j)$ are equal if and only if $x[i] = x[j]$ and ${\sf sub}(x, i + 1) = {\sf sub}(x, j + 1)$. Specifically, for any two neighboring substrings in ${\sf DSTR}(x)$, say ${\sf DSTR}(x)[p]$ and ${\sf DSTR}(x)[p + 1]$, there must have ${\sf DSTR}(x)[p] \le {\sf DSTR}(x)[p + 1]$. Hence, the only information required for naming the two substrings is to check if they are equal or not. Following the fact describe above, this can be trivially done by adding two fields $r1$ and $r2$ to each item of ${\sf DSTR}(x)$, where $r1$ and $r2$ respectively record the names of the corresponding substring and the succeeding substring. More specifically, assume all the items of LMS characters are already inserted into the corresponding buckets of ${\sf DSTR}(x)$, it takes the following two steps to sort and name substrings at the same time:

\begin{itemize}[itemindent = 5ex]
	\item[step 1:] Initially set $r = 0$. Scan ${\sf DSTR}(x)$ rightward. For each scanned item $e1$ and previously scanned item $e2$, if $e1.r2 = e2.r2$ and $e1.ch = e2.ch$ then set $e1.r1 = e2.r1$; otherwise, set $e1.r1 = r$ and increase $r$ by one. If $e1.p > 0$ and $e1.t = 0$, then determine the block $b_i$ that contains ${\sf sub}(x, e1.p - 1)$, retrieve the leftmost unvisited item $e3$ from ${\sf DSTR_L}(b_i)$, set $e3.r2 = e1.r1$ and insert $e3$ into the leftmost empty position of ${\sf BKT_L}({\sf DSTR}(x), e3.c)$.

	\item[step 2:] Initially set $r = n - 1$. Scan ${\sf DSTR}(x)$ leftward. For each scanned item $e1$ and previously scanned item $e2$, if $e1.r2 = e2.r2$ and $e1.ch = e2.ch$ then set $e1.r1 = e2.r1$; otherwise, set $e1.r1 = r$ and decrease $r$ by one. If $e1.p > 0$ and $e1.t = 1$, then determine the block $b_i$ that contains ${\sf sub}(x, e1.p - 1)$, retrieve the leftmost unvisited item $e3$ from ${\sf DSTR_S}(b_i)$, set $e3.r2 = e1.r1$ and insert $e3$ into the rightmost empty position of ${\sf BKT_S}({\sf DSTR}(x), e3.c)$.

\end{itemize}

At the end of step 2, all the LMS substrings are already sorted and their names can be obtained by copying $r1$ from the items of ${\sf DSTR}(x)$.

\subsection{Method B} \label{subsec:dsais_plus_method_b}
A solution to reducing the I/O volume is to employ a multi-way merge algorithm for combining sorted LMS substrings of all the blocks, where the task of the merger is to sort and name LMS substrings by literally comparing their characters. A string comparison can be done very quickly if the involved two substrings are wholly loaded into RAM. However, there may exist some LMS substrings that cannot be accommodated in the internal memory. For these long substrings, the method described in Section~\ref{subsec:dsais_plus_method_a} is employed. Following the ideas, we revise steps 2-4 of Algorithm~\ref{alg:dsais_reduction_phase} as below, where $i \in [1, k]$.

\begin{itemize}[itemindent = 5ex]
	\item [step 2':] Induced sort the substrings of $b_i$ to compute ${\sf LSA_L}(b_i)$, ${\sf LSA_S}(b_i)$, ${\sf DSTRA_L'}(b_i)$ and ${\sf DSTRA_S'}(b_i)$.

	\item [step 3':] Induced sort the substrings embraced in the long LMS substrings of $x$ from ${\sf LSA_L}(b_i)$, ${\sf DSTRA_L'}(b_i)$ and ${\sf DSTRA_S'}(b_i)$ to compute ${\sf LSA_L}(x)$.

	\item [step 4':] Merge ${\sf LSA_L}(x)$ and ${\sf LSA_S}(b_i)$ by using a min heap to compute ${\sf LMSNAME}(x)$.

\end{itemize}

The definitions of the newly introduced notations are given below.

\begin{itemize}[itemindent = 0ex]
	\item long/short LMS substring and threshold value $D$. All the LMS substrings are classified into two categories: long and short. Specifically, an LMS substring is short if it contains no more than $D$ characters; otherwise, it is long.

	\item ${\sf DSTR'}$, ${\sf DSTRA_L'}$ and ${\sf DSTRA_S'}$. DSAITEM arrays. For an input $x$, ${\sf DSTR'}(x)$ only contains the items in ${\sf DSTR}(x)$ that associate with substrings embraced in long LMS substrings. We say ${\sf sub}(x, p)$ embraces ${\sf sub}(x, q)$ if the two substrings end with the same LMS character and $p \le q$. ${\sf DSTRA_L'}(x)$ and ${\sf DSTRA_S'}(x)$ respectively contain the items in ${\sf DSTRA'}(x)$ that associate with L-type and S-type substrings.

	\item ${\sf LSA_L}$ and ${\sf LSA_S}$. LSAITEM arrays. For an input $x$, each item of ${\sf LSA_L}(x)/{\sf LSA_S}(x)$ associates with a long/short LMS substring ${\sf sub}(x, p)$ and mainly consists of four components:

	\begin{itemize}
		\item $p$: position index for ${\sf sub}(x, p)$.

		\item $s$: leftmost $D$ characters of ${\sf sub}(x, p)$, if it is long; otherwise, all the characters.

		\item $t$: type of the last character in $s$, set as 0 or 1 for L-type or S-type, respectively.

		\item $r$: name of ${\sf sub}(x, p)$.

	\end{itemize}

\end{itemize}

We describe each step detailedly in what follows.

\vspace{2ex} \noindent {\bf Step 2'}

\noindent This step consists of two substeps:

\begin{itemize}[itemindent = 1ex]
	\item [(a)] Compute ${\sf DSTRA_L'}(b_i)$ and ${\sf DSTRA_S'}(b_i)$ by reusing the algorithm for computing ${\sf DSTRA_L}(b_i)$ and ${\sf DSTRA_S}(b_i)$ in step 2 of Algorithm~\ref{alg:dsais_reduction_phase}.

	\item [(b)] Scan ${\sf DSTRA_S'}(b_i)$ rightward. For each scanned item $e1$ with $e1.t = 0$, create an LSAITEM $e2$ for the current LMS substring, compute $e2.t$ and $e2.s$ by visiting $b_i$, and insert $e2$ into ${\sf LSA_L}(b_i)$ or ${\sf LSA_S}(b_i)$ according to its length.
\end{itemize}

\vspace{2ex} \noindent {\bf Step 3'}

\noindent This step consists of three substeps:

\begin{itemize}[itemindent = 1ex]
	\item [(a)] Scan $x$ leftward with $i$ decreasing from $n - 1$ to $0$. For each scanned character $x[i]$, if $x[i]$ is an ending character of a long LMS substring, insert ${\sf sub}(x, i)$ into the rightmost empty position in ${\sf BKT_S}({\sf DSTRA'}(x), x[i])$.

	\item [(b)] Reuse step 1 of method A to sort and name the L-type substrings embraced in long LMS substrings of $x$ by replacing ${\sf DSTR}(x)$ with ${\sf DSTR'}(x)$.

	\item [(c)] Reuse step 2 of method A to sort and name the S-type substrings embraced in long LMS substrings of $x$ by replacing ${\sf DSTR}(x)$ with ${\sf DSTR'}(x)$. For each scanned item $e1$ that associates with an LMS substring, determine the block $b_i$ that contains ${\sf sub}(x, e1.p)$, retrieve the leftmost unvisited item $e4$ from ${\sf LSA_L}(b_i)$, set $e4.r = e1.r1$ and insert $e4$ to the front of ${\sf LSA_L}(x)$.

\end{itemize}

\vspace{2ex} \noindent {\bf Step 4'}

\noindent Initialize the heap by inserting the leftmost item of each sorted sequence of ${\sf LSA_L}(x)$ and $\{{\sf LSA_S}(b_1),...,{\sf LSA_S}(b_k)\}$. Sequentially pop the items in the heap. For currently popped item $e1$, insert into the heap the next item $e2$ of the same sequence and compare $e1$ with the previously popped item $e3$ to determine the names of their corresponding LMS substrings. In this step, each string comparison for heap-sorting items and naming substrings observes the following rules, where $e1$ and $e2$ are items to be compared.

\begin{itemize}[itemindent = 5ex]

	\item [rule 1:] If both items belong to $\{{\sf LSA_S}(b_1),...,{\sf LSA_S}(b_k)\}$, then literally compare $e1.s$ with $e2.s$ to determine their order.

	\item [rule 2:] If both items belong to ${\sf LSA_L}(x)$, then directly compare $e1.r$ with $e2.r$ to determine their order.

	\item [rule 3:] If one belongs to ${\sf LSA_L}(x)$ and the other belongs to $\{{\sf LSA_S}(b_1),...,{\sf LSA_S}(b_k)\}$, then literally compare $e1.s$ with $e2.s$. If equal, then continue to compare $e1.t$ with $e2.t$.

\end{itemize}

Clearly, the time overhead of each string comparison is upper-bounded by the threshold value $D$.

\section{Lightweight SA checker} \label{sec:sachecker}

It is essential for a suffix sorting algorithm to check the correctness of the output suffix arrays. Several existing SA construction algorithms verify their outputs based on Lemma~\ref{lemma:1}~\cite{Karkkainen2003}.

\begin{lemma} \label{lemma:1}
	An array ${\sf SA}[0, n)$ is the suffix array of a string $x[0, n)$ if and only if the following conditions are satisfied for all $i, j \in [0, n)$ and $i\ne j$:\\
	(1) ${\sf SA}[i], {\sf SA}[j] \in [0, n)$ and ${\sf SA}[i] \ne {\sf SA}[j]$. \\
	(2)	${\sf ISA}[i] < {\sf ISA}[j] \Leftrightarrow (x[i], {\sf ISA}[i + 1]) < (x[j], {\sf ISA}[j + 1])$.
\end{lemma}

This SA checker performs two passes of integer sorts, where the first pass checks if ${\sf SA}[0, n)$ is a permutation of $[0, n)$ for condition (1) and the second pass checks if the order of suffixes in each SA bucket is correct for condition (2). Although this method runs fast in the internal memory, the required disk space and I/O volume are considerably huge when it sorts items using the external memory, resulting in a non-negligible performance overhead. Take eSAIS for example, the peak disk usages for constructing and checking SA are around 24n and 27n, respectively, while the overall I/O volume also increases by 50n for conducting the two integer sorts.

\subsection{Details}\label{subsec:sachecker:details}

We describe a new SA checking method that can be seamlessly integrated into any IS-based induced sorting algorithm, which has almost no influence on the overall performance. The main idea is to verify the two conditions listed in Lemma~\ref{lemma:1} during the induction phase. In what follows, we first describe two observations that are always true in the process of induced sorting suffixes, where ${\sf SA_{LMS1}}(x)$ is the position index array for the sorted LMS suffixes produced by sorting ${\sf PA}(x)$ with ${\sf ISA}(x1)$ at the beginning of the induction phase, and ${\sf SA_{LMS1}}(x)[u]$ and ${\sf SA_{LMS1}}(x)[v]$ are the ending position indexes of the LMS substrings that embrace $x[{\sf SA}(x)[i]]$ and $x[{\sf SA}(x)[j]]$, respectively.

\begin{observation} \label{observation:1} $\forall i, j \in [0, n)$, ${\sf SA}(x)[i] = {\sf SA}(x)[j] \; and \; i \ne j \Leftrightarrow {\sf SA_{LMS1}}(x)[u] = {\sf SA_{LMS1}}(x)[v] \; and \; u \ne v$.
\end{observation}

\begin{IEEEproof}
	Omit.
\end{IEEEproof}

\begin{observation} \label{observation:2} $\forall i, j \in [0, n)$, $i < j \; and \; {\sf suf}(x, {\sf SA}(x)[i]) > {\sf suf}(x, {\sf SA}(x)[j]) \Rightarrow u < v \; and \; {\sf suf}(x, {\sf SA_{LMS1}}(x)[u]) > {\sf suf}(x, {\sf SA_{LMS1}}(x)[v])$.
\end{observation}

\begin{IEEEproof} \label{proof:1}
	An IS-based SA construction algorithm induces the order of suffixes by utilizing the fact that ${\sf suf}(x, p_1) < {\sf suf}(x, q_1) \Leftrightarrow (x[p_1], {\sf suf}(x, p_1 + 1)) < (x[q_1], {\sf suf}(x, q_1 + 1)$. The algorithm guarantees that ${\sf ISA}(x)[p_1] < {\sf ISA}(x)[q_1] \Leftrightarrow (x[p_1], {\sf ISA}(x)[p_1 + 1]) < (x[q_1], {\sf ISA}(x)[q_1 + 1])$. This indicates that ${\sf suf}(x,p_1) < {\sf suf}(x, q_1)$ and ${\sf ISA}(x)[p_1] > {\sf ISA}(x)[q_1]$ only if ${\sf suf}(x, p_2) < {\sf suf}(x, q_2)$ and ${\sf ISA_{LMS1}}(x)[p_2] > {\sf ISA_{LMS1}}(x)[q_2]$, where $p_2$ and $q_2$ are the position indexes of the leftmost LMS characters on the right side of $x[p_1]$ and $x[q_1]$, respectively.
\end{IEEEproof}

The above observations are used to prove the correctness of the following statement.

\begin{theorem} \label{theorem:1}
	For any IS-based induced sorting algorithm, the output array ${\sf SA}(x)$ is the suffix array of the input string $x[0, n)$ if and only if the following conditions are satisfied:

	(1) $\forall i, j \in [0, n)$ and $i \ne j$, ${\sf SA_{LMS1}}(x)[i] \ne {\sf SA_{LMS1}}(x)[j]$.

	(2) ${\sf SA_{LMS1}(x)} = {\sf SA_{LMS2}(x)}$, where ${\sf SA_{LMS2}}$ is the position index array for the sorted LMS suffixes retrieved from ${\sf SA}(x)$ at the end of the induction phase.
\end{theorem}

\begin{IEEEproof} \label{proof:2}
	We only prove the sufficiency as the necessity is clear.

	Suppose ${\sf SA}(x)$ is not a permutation of $[0, n)$. Let $i \ne j$ and ${\sf SA}(x)[i] = {\sf SA}(x)[j]$, then there must exist $u, v$ satisfying $u \ne v$ and ${\sf SA_{LMS1}}(x)[u] = {\sf SA_{LMS1}}(x)[v]$ according to Observation~\ref{observation:1}. This violates the first condition in Theorem~\ref{theorem:1}.

	Suppose ${\sf suf}(x, i) > {\sf suf}(x, j)$ and ${\sf ISA}(x)[i] < {\sf ISA}(x)[j]$. Then, there must exist $u, v$ satisfying ${\sf suf}(x, u) > {\sf suf}(x, v)$ and ${\sf ISA_{LMS1}}(x)[u] < {\sf ISA_{LMS1}}(x)[v]$ according to Observation~\ref{observation:2}. Let $(u_0, v_0)$ be the pair that maximize $u + v$, then we have ${\sf ISA_{LMS2}}(x)[u_0] > {\sf ISA_{LMS2}}(x)[v_0]$. This violates the second condition in Theorem~\ref{theorem:1}.
\end{IEEEproof}

To verify the conditions in Theorem~\ref{theorem:1}, we can design a checker based on the following idea. The first condition can be checked after sorting ${\sf PA}(x)$ by ${\sf ISA}(x1)$ at the beginning of the induction phase. For the second condition, a naive method that takes $\mathcal{O}(n)$ time and space is to copy ${\sf SA_{LMS1}}(x)$ and ${\sf SA_{LMS2}}(x)$ and then compare their items in sequence at the end of the induction phase. An alternative for lowering the overhead is to compute the fingerprints for ${\sf SA_{LMS1}}(x)$ and ${\sf SA_{LMS2}}(x)$ by using fingerprint functions and then compare the fingerprints instead of the items in these arrays to determine their equality. Obviously, ${\sf SA_{LMS1}}(x)$ and ${\sf SA_{LMS2}}(x)$ have a common fingerprint if they are identical, but the inverse is not always true. Fortunately, the probability of a false match can be reduced to a completely negligible level by properly setting the parameters for the calculation formula. This leads us to Theorem~\ref{theorem:2}.

\begin{theorem} \label{theorem:2}
	For any IS-based induced sorting algorithm, the output array ${\sf SA}(x)$ is the suffix array of the input string $x[0, n)$ with a high probability if the following conditions are satisfied:

	(1) $\forall i, j$, ${\sf SA_{LMS1}}(x)[i] \ne {\sf SA_{LMS1}}(x)[j]$.

	(2) ${\sf FP(SA_{LMS1}(x))} = {\sf FP(SA_{LMS2}(x))}$.
\end{theorem}

\subsection{Implementation} \label{subsec:sachecker:implementation}
We choose the approach presented in~\cite{Karp1987} to compute the fingerprints in need. Specifically, for a given array $x[0, n)$, the fingerprint ${\sf FP}(x)[0, n)$ can be iteratively computed by the formula ${\sf FP}(x)[i,j] = \sum_{p = i}^{j} \alpha^{j-p} \cdot x[p] \, mod \, \beta$, where $\beta$ is a large prime and $\alpha$ is an integer randomly chosen from domain $[1,\beta)$. In this way, we can slightly adapt Algorithm~\ref{alg:dsais_induction_phase} to compute ${\sf FP}(\sf SA_{LMS1}(x))$ by scanning ${\sf DSA_{LMS}}(x)$ in step 1 and ${\sf FP}(\sf SA_{LMS2}(x))$ by scanning ${\sf DSA}(x)$ during the time when inducing S-type suffixes in step 3, respectively. Obviously, the time, space and I/O complexities of the implementation are linearly related to $n$.

\section{Experiments} \label{sec:experiments}

For performance evaluation, we engineer DSA-IS and its optimized DSA-IS+, by using the STXXL's containers~(sorter, priority queue, vector and stream) to conduct efficient I/O operations in the external memory. As reported in~\cite{Karkkainen2014}, SAscan outperforms eSAIS in practice when the internal memory capacity $M$ is adequate. However, the time and space complexities of the program for SAscan are $\widetilde{\mathcal{O}}(n^2/ M)$ and $\widetilde{\Omega}(n^2/ M)$, respectively, thus it will suffer from a performance degradation as the size of the input string gets larger.\footnote{The authors in~\cite{Karkkainen2015} says, "SAscan needs just 7.5n bytes of disk space but because of its $\widetilde{\mathcal{O}}(n^2/ M)$ time complexity, it is competitive with eSAIS only when the input is less than about five times the size of RAM.".} On the other hand, the programs for DSA-IS, DSA-IS+ and eSAIS\footnote{https://panthema.net/2012/1119-eSAIS-Inducing-Suffix-and-LCP-Arrays-in-External-Memory/} are not only of linear I/O volume and space complexities in theory, but also commonly implemented by the STXXL library. \textbf{Therefore, we use eSAIS rather than SAscan as a baseline for evaluating the performance of DSA-IS and DSA-IS+.} The experimental platform is a desktop computer equipped with an Intel Xeon E3-1220 V2 CPU, 4GiB RAM and 500GiB HD. The programs are complied by gcc/g++ 4.8.4 with -O3 options under Ubuntu 14.04 64-bit operating system.

In our experiments, three performance metrics are investigated for the programs running on the corpora listed in Table~\ref{tbl:corpora}, where each metric is measured as a mean of two runs of the programs.

\begin{itemize}
	\item construction time~(CT): the running time, in units of microseconds per character.
	\item peak disk usage~(PDU): the maximum disk space requirement, in units of bytes per character.
	\item I/O volume~(IOV): as the term suggests, in units of bytes per character.
\end{itemize}

%Table
\renewcommand\arraystretch{1.3}
\begin{table*}[!t]
\caption{Corpus, $n$ in Gi, 1 byte per character} 
\label{tbl:corpora}
\centering
\begin{tabular}{|l|c|c|p{10cm}|}
	\hline
	Corpora & \multicolumn{1}{c|}{$n$} & \multicolumn{1}{c|}{$\|\Sigma\|$} & Description \\\hline
	guten & 22.5 & 256 & Gutenberg, at \url{http://algo2.iti.kit.edu/bingmann/esais-corpus}.\\\hline 				
	enwiki & 74.7 & 256 & Enwiki, at \url{https://dumps.wikimedia.org/enwiki}, dated as 16/05/01. \\\hline	
	proteins & 1.1 & 27 & Swissprot database, at \url{http://pizzachili.dcc.uchile.cl/texts/protein}, dated as 06/12/15. \\\hline
	uniprot & 2.5 & 96 & UniProt Knowledgebase release 4.0, at \url{ftp://ftp.expasy.org/databases/.../complete}, dated as 16/05/11. \\\hline
	genome & 2.9 & 6 & Human genome data, used in Dementiev et al.~\cite{Dementiev2008a}, at \url{http://algo2.iti.kit.edu/dementiev/esuffix/instsances.shtml.} \\\hline
\end{tabular}
\end{table*}

\subsection{Performance Evaluation on Construction Algorithms}
Figure~\ref{fig:performance_analysis1} illustrates the performance of the programs for DSA-IS and eSAIS in terms of the investigated three metrics, where guten\_14G and enwiki\_14G are prefixes of the corresponding corpus. As depicted, the peak disk usage for DSA-IS is around $16n$ on the two datasets, which is two-thirds of that for eSAIS. However, eSAIS outperforms DSA-IS with respect to the construction time, where the speed gap between them is mainly due to the different I/O volumes taken by the two algorithms. In order for a deep insight, we show in Table~\ref{tbl:volume_cmp} the statistics for I/O volumes of the algorithms in their reduction and induction phases. It can be observed that, although the I/O volumes spent in the induction phase are similar, eSAIS is far more I/O-efficient than DSA-IS when sorting substrings during the reduction phase. More specifically, the ratio of reduction and induction volumes for DSA-IS is about 0.80 while the counterpart for eSAIS is 0.23.

\begin{figure}[htbp!]
	\centering
	\subfigure{
		\label{subfig:pdu_cmp1}
		\includegraphics[width = 0.9\columnwidth]{pdu_cmp1}
		}
	\hfil
	\subfigure{
		\label{subfig:iov_cmp1}
		\includegraphics[width = 0.9\columnwidth]{iov_cmp1}
		}
	\hfil
	\subfigure{
		\label{subfig:ct_cmp1}
		\includegraphics[width = 0.9\columnwidth]{ct_cmp1}
		}
	\caption{Experimental results for DSA-IS and eSAIS on guten\_14G and enwiki\_14G in terms of peak disk usage, I/O volume and construction time.}
	\label{fig:performance_analysis1}
\end{figure}

% Table
\begin{table*}%
\caption{Comparison of Reduction and Induction I/O Volumes Amongst DSA-IS, DSA-IS+ and eSAIS on enwiki}
\label{tbl:volume_cmp}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|}
	\hline
	\multicolumn{1}{|c}{} & \multicolumn{4}{|c|}{eSAIS} & \multicolumn{4}{c|}{DSA-IS} & \multicolumn{4}{c|}{DSA-IS+ ($D_1 = 8$, $D_2 = 10$)}\\\hline
	\hline
	Size & Red. & Ind. & Total & Ratio & Red. & Ind. & Total & Ratio & Red. & Ind. & Total & Ratio\\\hline
	1G & 36.6 & 132.8 & 169.4 & 0.27 & 81.3 & 105.9 & 187.2 & 0.76 & 39.1 & 104.3 & 143.4 & 0.37\\\hline
	2G & 36.0 & 141.9 & 177.9 & 0.25 & 83.5 & 108.2 & 191.7 & 0.77 & 37.8 & 106.4 & 144.2 & 0.35\\\hline
	4G & 35.6 & 152.1 & 187.7 & 0.23 & 94.3 & 118.9 & 213.2 & 0.79 & 42.4 & 117.0 & 159.4 & 0.36\\\hline
	8G & 35.2 & 165.7 & 200.9 & 0.21 & 107.8 & 132.3 & 240.1 & 0.81 & 43.7 & 129.9 & 173.6 & 0.33\\\hline
	14G & 35.0 & 172.1 & 207.1 & 0.20 & 121.9 & 146.6 & 268.5 & 0.83 & 43.8 & 142.3 & 186.1 & 0.30\\\hline
\end{tabular}
\end{table*}%

The above phenomenon can be explained as follows. During the reduction phase, DSA-IS partitions the input string $x$ into multiple blocks and induced sorts the substrings of each block in the internal memory. Then it reuses the induced sorting method to sort all the substrings of $x$ in the external memory by merging the block-wise results residing on disks. In contrast with DSA-IS, eSAIS splits $x$ into fixed-size partitions and sorts the substrings of each partition in the internal memory by calling std::sort. Then, it generates the global result by merging the partial ones with an internal memory heap. This leads to a higher I/O efficiency against DSA-IS.

We attempt to reduce the I/O overhead for DSA-IS by using the substring sorting and naming methods described in Section~\ref{sec:dsais_plus}. Our program for the enhanced algorithm DSA-IS+ introduces two parameters $D_1$ and $D_2$ to respectively specify the threshold value $D$ for classifying the substrings in $x$ and the reduced strings $\{x1, x2, ...\}$ according to the instructions in Section~\ref{subsec:dsais_plus_method_b}. As observed from Table~\ref{tbl:volume_cmp}, the total I/O volumes for DSA-IS+ and eSAIS are similar and the ratio of the reduction and induction I/O volumes for either of them is no more than half of that for DSA-IS. This substantial improvement leads to a great speedup against DSA-IS. Figure~\ref{fig:performance_analysis2} shows the performance trends of DSA-IS+ and eSAIS on guten and enwiki as the corpora size increases from $1G$ to $14G$, which indicates that the speed of DSA-IS+ is similar to that of eSAIS and its peak disk usage remains at a low level as the same as that of DSA-IS.

\begin{figure*}[htbp]
\centering
\subfigure[guten]{
	\begin{minipage}[b]{0.45\textwidth}
		\includegraphics[width=1\textwidth]{pdu_cmp2_guten} \\
		\includegraphics[width=1\textwidth]{iov_cmp2_guten} \\
		\includegraphics[width=1\textwidth]{ct_cmp2_guten}
	\end{minipage}
}
\subfigure[enwiki]{
	\begin{minipage}[b]{0.45\textwidth}
		\includegraphics[width=1\textwidth]{pdu_cmp2_enwiki} \\
		\includegraphics[width=1\textwidth]{iov_cmp2_enwiki} \\
		\includegraphics[width=1\textwidth]{ct_cmp2_enwiki}
	\end{minipage}
}
\caption{Experimental results for DSA-IS, DSA-IS+ and eSAIS on guten and enwiki in terms of peak disk usage, I/O volume and construction time, where $D_1 = 8$, $D_2 = 10$ and the input size varies in \{1, 2, 4, 8, 14\} GiB. }
\label{fig:performance_analysis2}
\end{figure*}

%As with other external memory suffix sorting algorithms, DSA-IS+ is I/O-intensive and its construction speed is presumed to be linearly proportional to the I/O volume. However, our experiments indicate that the ratio of the reduction and induction time is much higher than that of the reduction and induction volume. This phenomenon is possibly due to the performance of the STXXL library. For implementation simplicity, we organize the item of DSAITEM arrays by using the sorter and priority queue containers provided by STXXL in our programs. This

A series of experiments are conducted to investigate the effect of $D_1$ and $D_2$ on the performance of DSA-IS+. As shown in Figure~\ref{fig:performance_analysis3}, the time consumption of DSA-IS+ gets smaller as $D_2$ becomes larger, while the peak disk usage and I/O volume are almost unchanged. Furthermore, the data of Table~\ref{tbl:effect_of_D} indicates that a small variation on the values of $D_1$ and $D_2$ can incur a significant fluctuation on the speed of DSA-IS+. For instance, when changing $(D_1, D_2)$ from $(8, 10)$ to $(16, 16)$, the construction time for the uniprot dataset decreases from 2.64 to 2.24 microseconds per character. This is because the proportions of long LMS substrings of $x$ and $\{x1, x2,...\}$ decrease when $D_1$ and $D_2$ increase. Recall that, after partitioning $x$ into blocks, DSA-IS+ sorts the long LMS substrings of each block in RAM and merges them by an external memory sorter, which is time-consuming when the number of long LMS substrings is large. In our program, the ratio of long and short LMS substrings can be controlled by adjusting the values of $D_1$ and $D_2$. However, when $D_1$ and $D_2$ get larger, the min heap also takes more time to merge the sorted short and long LMS substrings, as each string comparison takes $\mathcal{O}(D_1)$ and $\mathcal{O}(D_2)$ time in the first and recursion levels, respectively. Fortunately, Table~\ref{tbl:long_short_distribution} shows that the majority of LMS substrings in real datasets are considerably short. In practice, we can set $D_1$ and $D_2$ according to the statistics of the length distribution of LMS substrings, where the statistics can be collected when partitioning $x$ at the beginning of the reduction phase.

\begin{table*}[htbp]
\caption{Effects of $D_1$ and $D_2$ for DSA-IS+}
\label{tbl:effect_of_D}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c||c|c|c|c|c|}
	\hline
	\multicolumn{1}{|c}{} & \multicolumn{3}{|c}{eSAIS} & \multicolumn{10}{|c|}{DSA-IS+}\\\hline
	\hline
	Corpora & PDU & IOV & CT & $D_1$ & $D_2$ & PDU & IOV & CT  & $D_1$ & $D_2$ & PDU & IOV & CT\\\hline
	uniprot & 22.71 & 162.01 & 2.50 & 8 & 10 & 14.27 & 146.10 & 2.64 & 16 & 16 & 14.27 & 144.90 & 2.24 \\\hline
	proteins & 24.09 & 172.29 & 2.33 & 8 & 10 & 16.12 & 147.92 & 2.28 & 20 & 12 & 16.12 & 147.85 & 2.14 \\\hline
	genome & 22.64 & 157.41 & 2.15 & 8 & 10 & 15.34 & 142.05 & 2.75 & 20 & 12 & 15.34 & 140.89 & 2.25 \\\hline
\end{tabular}
\end{table*}%

\begin{figure}[htbp]
	\centering
	\subfigure{
		\centering
		\includegraphics[width=0.9\columnwidth]{pdu_cmp3}
		\label{subfig:pdu_cmp3}
	}
	\hfil
	\subfigure{
		\centering
		\includegraphics[width=0.9\columnwidth]{iov_cmp3}
		\label{subfig:iov_cmp3}
	}
	\hfil
	\subfigure{
		\centering
		\includegraphics[width=0.9\columnwidth]{ct_cmp3}
		\label{subfig:ct_cmp3}
	}
	\caption{Experimental results for DSA-IS+ and eSAIS on guten in terms of peak disk usage, I/O volume and construction time, where $D_1 = 8$, $D_2$ ranges in \{6, 8, 10, 12\} and the input size varies in \{1, 2, 4, 8, 14\} GiB. }
	\label{fig:performance_analysis3}
\end{figure}

\begin{table*}[htbp]
\caption{Statistics on the Length Distribution of Long and Short LMS Substrings for DSA-IS+}
\label{tbl:long_short_distribution}	
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
	\hline
	\multicolumn{11}{|c|}{level 0} \\\hline
	Corpora & $D_1$ & long & short & total & ratio & $D_1$ & long & short & total & ratio\\\hline
	uniprot & 8 & 30243605 & 785997413 & 816241018 & 0.963 & 16 & 3850755 & 812390263 & 816241018 & 0.999\\\hline
	proteins & 8 & 1687673 & 377404329 & 379092002 & 0.999 & 20 & 10930 & 379081072 & 379092002 & 0.999\\\hline
	genome & 8 & 19446841 & 773670266 & 793117107 & 0.975 & 20 & 509998 & 792607109 & 793117107 & 0.999\\\hline
	\multicolumn{11}{|c|}{level 1}\\\hline
	Corpora & $D_2$ & long & short & total & ratio & $D_2$ & long & short & total & ratio\\\hline
	uniprot & 10 & 11849 & 263748775 & 263760624 & 0.995 & 16 & 380 & 263760244 & 263760624 & 0.999\\\hline
	proteins & 10 & 18126 & 123772731 & 123790857 & 0.996 & 20 & 2937 & 123787920 & 123790857 & 0.999\\\hline
	genome & 10 & 311607 & 248243216 & 248554823 & 0.999 & 20 & 153845 & 248400978 & 248554823 & 0.999\\\hline
\end{tabular}
\end{table*}%

\subsection{Performance Evaluation on Checking Methods}

\begin{table*}[htbp]
	\caption{Overall Performance of DSA-IS+ and eSAIS with CM1 and CM2 on enwiki\_8G}
	\label{tbl:check_overhead}
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|}
		\hline
		& \multicolumn{3}{|c|}{DSA-IS+ ($D_1 = 8$, $D_2 = 10$)} & \multicolumn{3}{c|}{eSAIS}\\\hline
		Checking Method & PDU & IOV & CT & PDU & IOV & CT \\\hline
		CM1 & 17.93  & 180.01 & 3.10 & - & - & - \\\hline
		CM2 & 26.00 & 231.68 & 4.00 & 27.00 & 246.93 & 3.74 \\\hline
		no check & 17.93  & 173.67	& 3.17 & 23.85 & 200.93 & 3.10 \\\hline
	\end{tabular}
\end{table*}%

We make a performance comparison of the proposed checking method and the work presented in~\cite{Karkkainen2003}, which are denoted by CM1 and CM2, respectively. For performance evaluation, we integrate CM1 into the program of DSA-IS+ following Section~\ref{subsec:sachecker:implementation} and reuse the implementation for CM2 in the program of eSAIS to verify the outputs of DSA-IS+ and eSAIS. Table~\ref{tbl:check_overhead} gives a glimpse of the performance overhead for the two checking methods. It can be seen that, when checking the suffix array of enwiki\_8G by using CM2, both time and I/O volume for verification are about one-fifth of that for construction. On the other hand, CM1 has almost no negative impact on the overall performance. As demonstrated in lines 3 and 5, the peak disk usage for CM1 is no more than that for DSA-IS+ and the increase in the I/O volume can be ignored. It worth mentioning that the data of the running time for the combination of DSA-IS+ and CM1 is a bit shorter than that for the plain DSA-IS+. This is a normal phenomenon due to the fluctuating I/O performance, which in turn indicates that the time overhead consumed by CM1 is also negligible.

\section{Conclusion} \label{sec:conclusion}

For better performance, we redesign the reduction phase of DSA-IS by employing two methods for sorting and naming substrings. The program of the enhanced algorithm DSA-IS+ is engineered by the STXXL library to achieve a high I/O efficiency. Our experiments indicate that DSA-IS+ requires only around $16/24=0.67$ peak disk usage as that for eSAIS and runs as fast as the latter on various real datasets.

We describe an SA checker of linear time and space complexities that can be used to verify a suffix array during the time when it is being built. In combination with DSA-IS+, our checking method almost has no influence on the overall performance and is more efficient than the existing work in terms of time, space and I/O efficiencies.

% Bibliography
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,bibfile}

\end{document}

### 2017-03-30 draft-new/draft.tex
2017/-3/30

@@
 
-Suffix array~(SA) can be built within linear time and space by means of the induced sorting~(IS) method.
-The recently proposed IS suffix sorter fSAIS has a performance competitive to that of the state-of-the-art parallel construction algorithms in external memory, which indicates a great potential for improving the other disk-based IS alternatives by engineering them carefully.
-In view of this, we redesign the reduction phase of DSA-IS by using a new substring sorting and naming method and implement the variant, called DSA-IS+, by using several space-saving techinques previously applied to fSAIS.
-
-A constructed SA should be checked to ensure its correctness.
-We present in this paper a method that enables any IS suffix sorter to perform building and checking simultaneously, where the consumption for checking is negligible in comparison with that for building in our experiments.
+Currently, there exist several external-memory suffix sorting algorithms that build a suffix array~(SA) using the induced sorting~(IS) method, where the biggest difference between them is the way of retrieving the preceding characters of sorted suffixes when inducing the order of unsorted ones.
+It was reported recently that a careful engineering of these algorithms can build an SA using nearly optimal disk space, with the time and I/O complexities less than the state-of-the-art SA construction solutions.
+In this paper, we further employ the IS method to design an SA checking algorithm that verifies a suffix array when it is being built by any IS suffix sorting algorithms, where the time, space and I/O consumptions for checking is negligible in comparison with that for building in our experiments.
+This indicates that the IS method can serve as a basis for developing an efficient solution to the situations where checking is a must for a constructed SA.
 
@@
 \begin{IEEEkeywords}
-Suffix array, construction and verification, external memory.
+Suffix array, in-place verification, external memory.
 \end{IEEEkeywords}}
@@
 \section{Introduction}\label{sec:introduction}
+
+For a text drawn from a constant or integer alphabet, its suffix array can be built within linear time and space by the internal-memory algorithm SA-IS~\cite{Nong11}.
+
+This algorithm sort suffixes using the induced sorting method, which sorts two suffixes by comparing their heading characters and the lexical order of the ones starting at the next positions.
+
+ the induced sorting principle. The  main idea behind the IS method is to
+
+ the main idea of which is to induce the lexicographical order of all the suffixes from that of a sorted subset.
+Recently, the IS method has been successfully applied to designing three suffix sorting algorithms~eSAIS~\cite{Bingmann12}, SAIS-PQ~\cite{Liu15} and DSA-IS~\cite{Nong15} for external memory models.
+These disk-based variants commonly use a priority queue to simulate SA-IS, but they differ from each other in the way of retrieving the preceding character of a sorted suffix when inserting it
+
+ when inserting the unsorted ones into the SA.
+
+several works successfully applied the IS method to designing efficient SA construction algorithms on external memory models, where the biggest difference between them is the way of retrieving the preceding characters of sorted suffixes for inducing the order of unsorted ones.
+
+The internal-memory IS algorithms for sorting suffixes can build an SA in linear time and space~\cite{Nong11}.
+Recently, several works successfully applied the IS method to construct suffix arrays for massive datasets using external memory.
+
+For an given string, its suffix array can be built in linear time and space using the induced sorting method.
+
+Suffix array can be
+
+The algorithm SA-IS is currently the fastest SA construction algorithm
+
+SA-IS is currently the fastest
+
+The suffix array~(SA)~\cite{Manber1993} can be built in linear time and space using the induced sorting method~\cite{Nong11}. Currently, there are three works
+
+The suffix array~(SA)~\cite{Manber1993} can be built in linear time and space
+
+These algorithms have been implemented for demonstration and experiment purposes,
 

### 2017-04-06 draft-new/draft.tex
2017/04/06

@@
 
-\title{Building and Checking Suffix Array in External Memory}
+\title{Building and In-Place Checking Suffix Array in External Memory}
 
@@
 
-Currently, there exist several external-memory suffix sorting algorithms that build a suffix array~(SA) using the induced sorting~(IS) method, where the biggest difference between them is the way of retrieving the preceding characters of sorted suffixes when inducing the order of unsorted ones.
-It was reported recently that a careful engineering of these algorithms can build an SA using nearly optimal disk space, with the time and I/O complexities less than the state-of-the-art SA construction solutions.
-In this paper, we further employ the IS method to design an SA checking algorithm that verifies a suffix array when it is being built by any IS suffix sorting algorithms, where the time, space and I/O consumptions for checking is negligible in comparison with that for building in our experiments.
-This indicates that the IS method can serve as a basis for developing an efficient solution to the situations where checking is a must for a constructed SA.
+The induced sorting~(IS) method has been used to design disk-based algorithms for suffix array~(SA) construction.
+A recent engineering of these algorithms achieved nearly optimal disk space, with both the I/O and the time complexities better than the current fastest external-memory suffix sorters.
+In this paper, we propose a checking algorithm that enables any IS suffix sorting algorithms to check an SA when it is being constructed.
+Our experimental results indicate that the time, space and I/O volume for verification by our checking program is negiligible in comparison with that for construction by our
+
+Recently, these algorithms have been carefully engineered to achieve nearly optimal disk space, with both the I/O and the time complexities better than the current fastest external-memory suffix sorters.
+In this paper, we propose a checking algorithm that enables any IS suffix sorting algorithm to build and check an SA simultaneously.
+Our experimental results indicate that, in comparison with that for building, the time, space and I/O consumptions for checking is negligible.
+This convinence us that the proposed checking algorithm can be combined with the existing IS suffix sorting algorithms to develop efficient solutions for the situations where checking is a must for a constructed SA.
 
@@
 
-For a text drawn from a constant or integer alphabet, its suffix array can be built within linear time and space by the internal-memory algorithm SA-IS~\cite{Nong11}.
+On RAM models, the suffix array for any text can be built within linear time and space using the SA-IS algorithm~\cite{Nong11}.
+Recently, this algorithm has been successfully adapted to design three external-memory suffix sorting algorithms~eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}.　To avoid random I/O accesses, these algorithms adopt different approaches to retrieve the head characters of unsorted suffixes for inducing their lexical order.
+However, they all suffer from a disk space bottleneck and
+
+Although these algorithms can achieve
+
+These algorithms use different approaches to retrieve the head characters of suffixes to be sorted when  when inducing their lexical order from the sorted ones. Although the experimental
 

### 2017-04-06 draft-new/draft.tex
2017/04/07

--- 2017-03-20 draft-new/draft.tex
@@
 
-Suffix array~(SA) can be built within linear time and space by means of the induced sorting~(IS) method.
-The recently proposed IS suffix sorter fSAIS has a performance competitive to that of the state-of-the-art parallel construction algorithms in external memory, which indicates a great potential for improving the other disk-based IS alternatives by engineering them carefully.
-In view of this, we redesign the reduction phase of DSA-IS by using a new substring sorting and naming method and implement the variant, called DSA-IS+, by using several space-saving techinques previously applied to fSAIS.
-
-A constructed SA should be checked to ensure its correctness.
-We present in this paper a method that enables any IS suffix sorter to perform building and checking simultaneously, where the consumption for checking is negligible in comparison with that for building in our experiments.
+The induced sorting~(IS) method has been used to design disk-based algorithms for suffix array~(SA) construction.
+A recent engineering of these algorithms can achieve nearly optimal disk space and has a higher scalability over the state-of-the-art suffix sorters in terms of time and I/O complexities.
+In this paper, we propose a checking method that enables any IS suffix sorting algorithms to verify an SA when it is being built.
+For performance analysis, we integrate the checking method into the IS suffix sorting algorithm DSA-IS and implement the adapted algorithm to investigate the checking overhead.
+The experimental results demonstrate that the time, space and I/O consmputions for verification is negligible in comparison with that for construction, indicating that the combination of the proposed checking and the IS building methods can constitute an efficient solution for the situations where building and checking must be done simultaneously.
 
@@
 
-The suffix array~(SA)~\cite{Manber1993} is an essential data structure for string processing and information retrieval.
-Among the existing internal-memory algorithms for SA construction, SA-IS~\cite{Nong11} achieves the optimal time and space complexities using the induced sorting principle, where the key operation is to retrieve the preceding characters of sorted suffixes in order for inducing the lexical order of unsorted ones.
-To handle massive datasets , several external memory algorithms have been proposed for building massive suffix arrays in recent years, e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014}, pSAscan~\cite{Karkkainen2015}, eSAIS~\cite{Bingmann2013}, EM-SA-DS~\cite{Nong2014} and DSA-IS~\cite{Nong2015}.
-Among them, the latter three algorithms are based on the induced sorting~(IS) method described in SA-IS~\cite{Nong2011}.
+For any text drawn from a constant or integer alphabet, its suffix array can be built on RAM in linear time and space by the SA-IS algorithm~\cite{Nong11}, which adopts the IS method to induce the lexical order of unsorted suffixes from those already sorted.
+In the past five years, the IS method has been reused to design three disk-based suffix sorting algorithms~eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15} that employ different approaches to retrieve the heading characters of unsorted suffixes by efficient I/O operations.
+These algorithms are competitive with the existing alternatives in terms of time and I/O complexities, but their naive implementations consume around 20 times the input size, leading to a performance bottleneck.
+A	 recent work in~\cite{xxx} presented a careful engineering of these algorithms, which takes less than $8n$ disk space for an input consisting of $n \le 2^{40}$ characters and runs faster than the state-of-the-art suffix sorter pSAscan~\cite{Karkkainen2015} when $n$ is considerably larger than the available RAM space.
+This reveals that the IS method can serve as a basis for potentially fastest suffix sorting solutions on EM models.
 
- This algorithm employs the IS method to induce the lexical order of all the suffixes from a selected subset.
+An SA should be verified before use to detect potential computation errors caused by hardware malfunctions or implementation bugs.
+Currently, the existing SA checker performs two runs of integer sorts and runs slowly in external memory.
+We propose in this paper a lightweight checking method that enables any disk-based IS suffix sorting algorithms to verify an SA when it is being built.
+From our experiments, the time, space and I/O volume for verification is negligible in comparison with that for construction.
 
- This algorithm uses the induced sorting method to
+The rest of this paper is organized as following.
+We first introduce the preliminaries for this paper in Section~\ref{sec:preliminaries} and describe the idea of the proposed checking method in Section~\ref{sec:idea}.
+Then, we demonstrate our experimental results in Section~\ref{sec:experiments} and conclude the remarks in Section~\ref{sec:conclusion}.
 
-In the past decades, great effort has been taken to study efficient algorithms for SA construction. th
+\section{Preliminaries}~\label{sec:preliminaries}
 
-extensive works have been put on designing time and space efficient suffix sorting algorithms
+\section{Checking Idea}~\label{sec:idea}
 
-data structure that has been widely used in many string processing applications, e.g., biological sequence alignment, time series analysis and text clustering.
-Given an input string, traversing its suffix tree can be emulated by using the corresponding enhanced suffix array~\cite{Abouelhodaa2004}, which mainly consists of the suffix and the longest common prefix arrays.
-It has been realized that the application scope of an index mainly depends on the construction speed and the space requirement.
-This leads to intensive works on designing time and space efficient suffix sorting algorithms over the past decade, assuming different computation models such as internal memory, external memory, parallel and distributed models.
-Particularly,
+\section{Experiments}~\label{sec:experiments}
 
-The basic idea behind the induced sorting method is to induce the lexicographical order of all the substrings/suffixes from a sorted subset of substrings/suffixes.
-Following the idea, an IS-based suffix sorting algorithm is typically comprised of a reduction phase for sorting and naming substrings to reduce a string $x[0,n)$ to a short string $x1[0,n1)$ with $n1 \le \frac{1}{2}n$ and an induction phase for sorting suffixes to induce ${\sf SA}(x)$ from ${\sf SA}(x1)$.
-During the two phases, the key operation is to retrieve the preceding character of a sorted substring/suffix.
-This can be done very quickly when $x$ is fully accommodated in the internal memory, but will become slow when $x$ resides in the external memory, as each operation takes a random disk access.
-For a high I/O efficiency, eSAIS, EM-SA-DS and DSA-IS use different auxiliary data structures to retrieve the preceding characters in a disk-friendly way.
-Particularly, both eSAIS and EM-SA-DS split a long substring into pieces and represent each piece by a fixed-size tuple, while DSA-IS does not.
-With an elaborate arrangement of the I/O operations, the programs for these three algorithms are competitive with those for others in terms of both time and space efficiencies.
-
-Among the existing internal-memory suffix sorters, SA-IS~\cite{Nong11} achieves the optimal time and space complexities using the induced sorting method, the key operation of which is to retrieve the preceding characters of sorted suffixes in order for inducing the lexical order of unsorted ones.
-In the past five years, several works adapted SA-IS to design suffix sorters specific for EM models~\cite{Nong14, Nong15, Karkkainen2014, Liu2015}.
-These variants use different approaches to retrieve the preceding characters by sequential I/O operations.
-However, they all suffer from a space bottleneck for taking at least twice disk space as SA-IS on real-world datasets.
-Recently, it was presented in~\cite{xxx} a new engineering version of SA-IS that achieves nearly optimal space efficiency, indicating a great potential for improving the other disk-based IS alternatives by engineering them carefully.
-
-Spae
-
-\section{Conclusion}
-xxx
+\section{Conclusion}~\label{sec:conlusion}
 

### 2017-04-21 new-draft/draft.tex
2017/04/21

--- 2017-03-20 tc-rejected/paper.tex
@@
+% !TeX spellcheck = en_US
+
 \documentclass[10pt,journal,compsoc]{IEEEtran}
-\usepackage[nocompress]{cite}
 \usepackage{graphicx}
-\usepackage{subfigure}
+\usepackage[ruled, linesnumbered]{algorithm2e}
 \usepackage{url}
+\usepackage{epstopdf}
+\usepackage{indentfirst}
+\usepackage[tight,footnotesize]{subfigure}
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{multirow}
-\usepackage[ruled]{algorithm2e}
-\usepackage{enumitem}
+\usepackage{color}
+\usepackage{enumerate}
 
@@
 \newtheorem{observation}[theorem]{Observation}
+\newtheorem{corollary}[theorem]{Corollary}
+
+% *** CITATION PACKAGES ***
+\ifCLASSOPTIONcompsoc
+\usepackage[nocompress]{cite}
+\else
+\usepackage{cite}
+\fi
 
@@
 
-\title{Efficient Suffix Array Construction and Checking in External Memory}
-%
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
-%
-%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
-% lists the Computer Society journals use for "first footnote" author
-% affiliations. Use \IEEEcompsocthanksitem which works much like \item
-% for each affiliation group. When not in compsoc mode,
-% \IEEEcompsocitemizethanks becomes like \thanks and
-% \IEEEcompsocthanksitem becomes a line break with idention. This
-% facilitates dual compilation, although admittedly the differences in the
-% desired content of \author between the different types of papers makes a
-% one-size-fits-all approach a daunting prospect. For instance, compsoc 
-% journal papers have the author affiliations above the "Manuscript
-% received ..."  text while in non-compsoc journals this is reversed. Sigh.
+\title{Building and Checking Suffix Array Using Induced Sorting Method}
 
@@
 
-% note the % following the last \IEEEmembership and also \thanks - 
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-% 
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those "LaTeX things". For
-% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
-% "AB" then you have to do: "\textbf{A}\textbf{B}"
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
-
-% The paper headers
-%\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
-%{Shell \MakeLowercase{\textit{et al.}}: Bare Advanced Demo of IEEEtran.cls for IEEE Computer Society Journals}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-% 
-% *** Note that you probably will NOT want to include the author's ***
-% *** name in the headers of peer review papers.                   ***
-% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
-% you desire.
-
-% The publisher's ID mark at the bottom of the page is less important with
-% Computer Society journal papers as those publications place the marks
-% outside of the main text columns and, therefore, unlike regular IEEE
-% journals, the available text space is not reduced by their presence.
-% If you want to put a publisher's ID mark on the page you can do it like
-% this:
-%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
-% or like this to get the Computer Society new two part style.
-%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2015 IEEE}%
-%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
-% Remember, if you use this you must call \IEEEpubidadjcol in the second
-% column for its text to clear the IEEEpubid mark (Computer Society journal
-% papers don't need this extra clearance.)
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% for Computer Society papers, we must declare the abstract and index terms
-% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
-% command as these need to go into the title area created by \maketitle.
-% As a general rule, do not put math, special symbols or citations
-% in the abstract or keywords.
 \IEEEtitleabstractindextext{% 
@@
 
-A suffix array (SA) must be checked to ensure its correctness.
-The existing SA checker verifies an SA by performing two passes of integer sorts that are time and space consuming especially when running in the external memory.
-Currently, building and checking an SA are done in sequence, where the time and the space bottlenecks are caused by building and checking the SA, respectively.
-In this paper, a fast and lightweight checking method is proposed to enable any induced sorting (IS) suffix sorting algorithm to verify a suffix array while it is being built.
-For performance evaluation, we first redesign the reduction phase of the external memory IS algorithm DSA-IS by employing the new substring sorting and naming methods to produce an optimized alternative DSA-IS+, and then apply the new checking method to DSA-IS+ for building and checking an SA simultaneously.
-Our experimental results indicate that, in comparison with that for building, the checking time and space complexities of eSAIS are substantial, but that of DSA-IS+ are negligible.
-As a result, DSA-IS+ can build and check an SA in a higher total time and space efficiencies.
+In the past decade, the induced sorting~(IS) method has been successively used to design disk-based suffix array~(SA) construction algorithms that have better time and I/O complexities than the existing alternatives.
+However, their current programs commonly suffer from a performance bottleneck due to large disk space in need.
+Recently, it was reported a carefully engineered IS suffix sorter that takes less than $8n$ disk space for constructing an SA encoded by 40-bit integers.
+This indicates a great potential for improving space efficiency of the prior arts, such as DSA-IS and SAIS-PQ.
+In this paper, we present a checking method that enables any IS suffix sorting algorithms to check an SA when it is being built.
+For performance analysis, we integrate the proposed method into DSA-IS and implement the enhanced algorithm to evaluate the checking overhead, where the program design of DSA-IS has been optimized for a better speed and smaller peak disk use.
+From our experiments, the time, space and I/O consumptions for verification is negligible in comparison with that for construction.
+In view of the high performance of this checker and the existing IS suffix sorters, we are convinced that our checking method, in accompanied with the state-of-the-art IS suffix sorting algorithm, can constitute an efficient solution to the situations where checking is a must for a constructed SA.
 
@@
 \begin{IEEEkeywords}
-suffix array, construction and verification, external memory.
+Suffix array, construction and verification, external memory.
 \end{IEEEkeywords}}
@@
 
-% To allow for easy dual compilation without having to reenter the
-% abstract/keywords data, the \IEEEtitleabstractindextext text will
-% not be used in maketitle, but will appear (i.e., to be "transported")
-% here as \IEEEdisplaynontitleabstractindextext when compsoc mode
-% is not selected <OR> if conference mode is selected - because compsoc
-% conference papers position the abstract like regular (non-compsoc)
-% papers do!
 \IEEEdisplaynontitleabstractindextext
-% \IEEEdisplaynontitleabstractindextext has no effect when using
-% compsoc under a non-conference mode.
 
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
 
-\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
+\section{Introduction}\label{sec:introduction}
 
-\subsection{Background}
-\IEEEPARstart{T}{he} suffix array~(SA)~\cite{Manber1993} is a data structure that has been widely used in many string processing applications, e.g., biological sequence alignment, time series analysis and text clustering.
-Given an input string, traversing its suffix tree can be emulated by using the corresponding enhanced suffix array~\cite{Abouelhodaa2004}, which mainly consists of the suffix and the longest common prefix arrays.
-It has been realized that the application scope of an index mainly depends on the construction speed and the space requirement.
-This leads to intensive works on designing time and space efficient suffix sorting algorithms over the past decade, assuming different computation models such as internal memory, external memory, parallel and distributed models.
-Particularly, to keep pace with the fast development of the data sampling techniques, several external memory algorithms have been proposed for building massive suffix arrays in recent years, e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014}, pSAscan~\cite{Karkkainen2015}, eSAIS~\cite{Bingmann2013}, EM-SA-DS~\cite{Nong2014} and DSA-IS~\cite{Nong2015}.
-Among them, the latter three algorithms are based on the induced sorting~(IS) method described in SA-IS~\cite{Nong2011}.
+Given a string drawn from a constant or integer alphabet, its suffix array can be built by the internal-memory algorithm SA-IS~\cite{Nong11} in linear time and space.
+So far, the IS method has been also applied to designing three external-memory suffix sorting algorithms~eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}, which have better time and I/O complexities than other alternatives, e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}.
+However, the peak disk uses of their current programs are at least 15 times the sizes of input strings, resulting in a performance bottleneck.
+Recently, it was reported in~\cite{Karkkainen2017} a new IS suffix sorting algorithm fSAIS with a careful engineering that consumes nearly optimal disk space and runs faster than the parallel suffix sorter pSAscan on massive datasets.
+The algorithmic design of fSAIS is highly dependent on that of the prior arts.
+Particularly, it employs a sorting method similar to that of DSA-IS for inducing suffixes/substrings and the naming method of SAIS-PQ for producing reduced strings.
+The substantial improvement on space efficiency is mainly achieved by means of a monotone priority queue  for sorting substrings/suffixes and a garbage collector for quickly recycling disk space occupied by data no longer to use.
+As a result, its implementation takes less than $8n$ disk space for any input string of a size $n \le 2^{40}$, indicating a great potential for optimizing the programs for DSA-IS and SAIS-PQ.
+In this paper, we make our first attempt to redesign DSA-IS using new substring sorting and naming methods.
+The experimental results indicate that our program for the enhanced algorithm consumes about two-thirds as much disk space as that for eSAIS and runs faster than the latter on real world datasets.
+In addition, we also give a discussion on further improving our program to achieve a better space efficiency using the techniques for implementing fSAIS.
 
-The basic idea behind the induced sorting method is to induce the lexicographical order of all the substrings/suffixes from a sorted subset of substrings/suffixes.
-Following the idea, an IS-based suffix sorting algorithm is typically comprised of a reduction phase for sorting and naming substrings to reduce a string $x[0,n)$ to a short string $x1[0,n1)$ with $n1 \le \frac{1}{2}n$ and an induction phase for sorting suffixes to induce ${\sf SA}(x)$ from ${\sf SA}(x1)$.
-During the two phases, the key operation is to retrieve the preceding character of a sorted substring/suffix.
-This can be done very quickly when $x$ is fully accommodated in the internal memory, but will become slow when $x$ resides in the external memory, as each operation takes a random disk access.
-For a high I/O efficiency, eSAIS, EM-SA-DS and DSA-IS use different auxiliary data structures to retrieve the preceding characters in a disk-friendly way.
-Particularly, both eSAIS and EM-SA-DS split a long substring into pieces and represent each piece by a fixed-size tuple, while DSA-IS does not.
-With an elaborate arrangement of the I/O operations, the programs for these three algorithms are competitive with those for others in terms of both time and space efficiencies.
+To detect potential computation errors caused by implementation bugs and/or hardware malfunctions, a constructed SA should be verified to ensure its correctness before use.
+This checking function has been included into various widespread software packages like SA-IS, DC3 and eSAIS.
+Particularly, the latter two verify an SA using the checking method presented in~\cite{Dementiev2008a}, which mainly consists of two passes of external-memory sorts involving $\mathcal{O}(n)$ fixed-size tuples with an integer key.
+In this paper, we propose a new checking method that enables any IS suffix sorting algorithm to check an SA when it is being built.
+For performance analysis, we integrate the proposed method into the optimized DSA-IS and implement the combined algorithm, called DSA-IS+, to evaluate the checking overhead.
+From our experiments, the time, space and I/O consumptions for verification is negligible in comparison with that for construction.
+In view of the high performance of this checker and the existing IS suffix sorters, we believe that our checking method, together with the state-of-the-art IS suffix sorting algorithm, can constitute an efficient solution to the situations where checking is a must for a constructed SA.
 
-\subsection{Problems to Solve}
-Although the above three IS-based algorithms achieve great performance in comparison with prior arts, there still exist several drawbacks to be dealt with.
+The rest of the paper is organized as follows.
+We first describe in Sections~\ref{sec:build_sa} and~\ref{sec:check_sa} the adapted algorithmic design of DSA-IS and the proposed checking method, respectively.
+Then, we demonstrate the experimental results in Section~\ref{sec:experiments} and draw the concluding remarks in Section~\ref{sec:conclusion}.
 
-Firstly, the space requirement remains at a high level for all the three algorithms.
-In details, the peak disk usages for eSAIS, EM-SA-DS and DSA-IS are around $24n$, $31n$ and $20n$, respectively, the minimum of which is about three times as that for the lightweight suffix sorter SAscan (which I/O volume is super-linear).
-It is identified that DSA-IS reaches the peak disk usage when induced sorting substrings during the reduction phase.
-This reveals a need for designing new substring sorting and naming methods to improve the space efficiency in order to scale the problem size that can be tackled.
+\section{SA Builder}~\label{sec:build_sa}
 
-Secondly, a suffix array must be checked to ensure its correctness for later use.
-Currently, an SA built by each of these algorithms is verified using the SA checker described in~\cite{Karkkainen2003}, which performs two passes of external memory integer sorts that charge large disk space and long cpu time.
-As a result, the peak disk usage for both eSAIS and DSA-IS rises up to $26n$ while the time consumption also grows by 20 percent as before, leading to a substantial performance degradation especially when processing large-scale datasets.
-Therefore, it is desired to reduce the check overhead in case it becomes a performance bottleneck.
+\subsection{Preliminaries}
 
-\subsection{Our Contribution}
+Given an input string $x[0, n)$ with a unique smallest character $x[n - 1]$, we denote by ${\sf suf}(i)$ a suffix starting with $x[i]$ and ${\sf sub}(i,j)$ a substring running from $x[i]$ to $x[j]$.
+The following notations are used in our presentation for denotation convenience.
 
-Our contribution in this paper mainly includes the following aspects.
+Character/substring/suffix classification.
+Characters of $x$ are classified into two categories: L-type and S-type.
+To be specific, $x[i]$ is S-type if (1) $i = n - 1$ or (2) $x[i] = x[i + 1]$ and $x[i + 1]$ is S-type; otherwise, $x[i]$ is L-type.
+Particularly, if $x[i]$ and $x[i - 1]$ are respectively S-type and L-type, then $x[i]$ is also S*-type.
+In addition, a substring/suffix is of the same type as its heading character.
+We use an integer array $t$ to indicate the type of characters, where $t[i] = 1$ or $0$ if $x[i] $ S-type or L-type, respectively.
 
-\begin{itemize}
+Predecessor and Successor.
+For any two neighboring characters $x[i]$ and $x[i + 1]$, we say $x[i]$ is the predecessor of $x[i + 1]$ while the latter is the successor of the former.
+Similarly, there also exists a predecessor-successor relationship between ${\sf suf}(i)/{\sf sub}(i, j)$ and ${\sf suf}(i + 1)/{\sf sub}(i + 1, j)$.
 
-	\item We redesign the reduction phase of DSA-IS by employing new methods for sorting and naming substrings.
-In our experiments, the program for the enhanced algorithm, called DSA-IS+, runs as fast as that for eSAIS while its peak disk usage is about two-thirds as that of eSAIS.
-
-	\item We design a fast and lightweight SA checking method that can be seamlessly integrated into any IS-based suffix sorting algorithm.
-For performance evaluation, we combine this method with DSA-IS+ to produce a unified solution that can be employed to construct and check SA at the same time.
-It was observed that, by exploiting the use of our checking method, the performance overhead for verification is considerable small in comparison with that for construction.
-In details, the peak disk usage for the unified solution remains at the same level of that for DSA-IS+, while the growth in both time and I/O volume is negligible.
-
-\end{itemize}
-
-The rest of this paper is organized as follows.
-Section~\ref{sec:preliminary} gives the preliminaries in our presentation.
-Section~\ref{sec:dsais_plus} describes the substring sorting and naming methods for DSA-IS+, Section~\ref{sec:sachecker} the checking method and Section~\ref{sec:experiments} the experimental results.
-Finally, we come to the conclusion in Section~\ref{sec:conclusion}.
-
-\section{Preliminaries}\label{sec:preliminary}
-
-\subsection{Basic Notations}
-
-Given an input string $x[0,n)=x[0]x[1]...x[n - 1]$ drawn from a totally ordered alphabet $\delta$, we assume $x[n - 1]$ to be the smallest character in $\delta$ that appears in $x$ only once.
-For short, we use ${\sf sub}(x,i)$ and ${\sf suf}(x,i)$ to denote a substring running from $x[i]$ to the leftmost LMS character on its right side and a suffix running from $x[i]$ to $x[n - 1]$, respectively.
-To follow lists the basic notations used in our presentation.
-
-\begin{itemize}[itemindent = 0ex]
-	\item L-type, S-type and LMS character/substring/suffix.
-We say $x[i]$ is S-type if (1) $i = n - 1$ or (2) $x[i] < x[i + 1]$ or (3) $x[i] = x[i+1]$ and $x[i+1]$ is S-type; otherwise, $x[i]$ is L-type.
-Furthermore, if $x[i]$ and $x[i - 1]$ are respectively S-type and L-type, then $x[i]$ is also an LMS character.
-Moreover, if $x[i]$ is L-type, S-type or LMS, then ${\sf sub}(x,i)$ and ${\sf suf}(x,i)$ are L-type, S-type or LMS, respectively.
-
-	\item preceding and succeeding character/substring/suffix.
-We say $x[i - 1]$, ${\sf sub}(x, i - 1)$ and ${\sf suf}(x, i - 1)$ are the preceding character, substring and suffix for $x[i]$, ${\sf sub}(x, i)$ and ${\sf suf}(x, i)$, respectively.
-Similarly, we say $x[i + 1]$, ${\sf sub}(x, i + 1)$ and ${\sf suf}(x, i + 1)$ are the succeeding character, substring and suffix for $x[i]$, ${\sf sub}(x, i)$ and ${\sf suf}(x, i)$, respectively.
-
-	\item ${\sf SA}$ and ${\sf STRA}$.
-We respectively use ${\sf SA}$ and ${\sf STRA}$ to denote the suffix and substring arrays.
-Specifically, both ${\sf SA}(x)$ and ${\sf STRA}(x)$ are permutations of $[0, n)$ such that ${\sf suf}(x, {\sf SA}(x)[i]) < {\sf suf}(x, {\sf SA}(x)[i + 1])$ and ${\sf sub}(x, {\sf SA}(x)[i]) \le {\sf sub}(x, {\sf SA}(x)[i + 1])$ in lexicographical order for all $i \in [0, n - 1)$.
-
-	\item ${\sf BKT}$.
-Suffixes in ${\sf SA}(x)$ are naturally grouped into buckets according to their head characters.
-Given ${\sf BKT}({\sf SA}(x), ch)$, the bucket occupies a consecutive range in ${\sf SA}(x)$ that gathers all the suffixes with a head character $ch$.
-Furthermore, a bucket, say ${\sf BKT}({\sf SA}(x), ch)$, can be subdivided into ${\sf BKT_L}({\sf SA}(x), ch)$ and ${\sf BKT_S}({\sf SA}(x), ch)$ that only contain L-type and S-type suffixes in ${\sf BKT}({\sf SA}(x), ch)$, respectively.
-Similarly, we define ${\sf BKT}({\sf STRA}(x), ch)$, ${\sf BKT_L}({\sf STRA}(x), ch)$ and ${\sf BKT_S}({\sf STRA}(x), ch)$ for ${\sf STRA}(x)$.
-
-	\item ${\sf ISA}$. Inverse suffix array, satisfying ${\sf ISA}(x)[{\sf SA}(x)[i]] = i$ for all $i \in [0,n)$.
-
-	\item ${\sf PA}$. Position array, where ${\sf PA}(x)[i]$ records the position index of the $i$-th LMS substring from the left side of $x$.
-
-\end{itemize}
-
-\subsection{An Overview of DSA-IS}
-
-As with other IS-based suffix sorting algorithms, DSA-IS mainly consists of a reduction phase followed by an induction phase.
-The induced sorting method is adopted in both two phases to sort substrings and suffixes, respectively.
-
-\subsubsection{Reduction Phase}
-
-Assume $x[0,n)$ and ${\sf STRA}(x)$ can be wholly accommodated into RAM, SA-IS performs the following three steps in the reduction phase to determine the lexicographical order of substrings.
-
-\begin{itemize}[itemindent = 5ex]
-	\item [step 1:] Scan $x$ rightward with $i$ decreasing from $n - 1$ to $0$. For each scanned LMS character $x[i]$, insert $i$ into the rightmost empty position of ${\sf BKT_S}({\sf STRA}(x), ch)$.
-
-	\item [step 2:] Scan ${\sf STRA}(x)$ leftward with $i$ increasing from $0$ to $n - 1$.
-For each scanned $p = {\sf STRA}(x)[i]$, if $ch = x[p - 1]$ is the starting character of an L-type substring, then insert $p - 1$ into the leftmost empty position of ${\sf BKT_L}({\sf STRA}(x), ch)$.
-
-	\item [step 3:] Scan ${\sf STRA}(x)$ rightward with $i$ decreasing from $n- 1$ to $0$.
-For each scanned $p = {\sf STRA}(x)[i]$, if $ch = x[p - 1]$ is the starting character of an S-type substring, then insert $p - 1$ into the rightmost empty position of ${\sf BKT_S}({\sf STRA}(x), ch)$.
-\end{itemize}
-
-The key operation of steps 2-3 is to retrieve the preceding character $x[p - 1]$ for currently scanned substring starting at $x[{\sf STRA}(x)[i]]$, where each operation takes a random access to the internal memory.
-A naive solution for extending the above 3-step procedure to external memory models is to replace each memory access with a disk access.
-However, this is not practical due to the large overhead for $\mathcal{O}(n)$ disk accesses.
-To reduce the I/O complexity, DSA-IS introduces the alternative ${\sf DSTRA}(x)$ for ${\sf STRA}(x)$ and employs I/O buffers to cache ${\sf DSTRA}(x)$ and other data structures for amortizing the disk access overhead, where the involved data structures are defined as below.
-
-\begin{itemize}[itemindent = 0ex]
-
-	\item ${\sf DSTRA}$, ${\sf DSTRA_L}$ and ${\sf DSTRA_S}$.
-DSAITEM arrays.
-For an input $x$, each item of ${\sf DSTRA}(x)$ associates with a substring ${\sf sub}(x, p)$ and mainly consists of the following three components:
-	\begin{itemize}[itemindent = 0ex]
-		\item $p$: position index for ${\sf sub}(x, p)$.
-		\item $c$: head character for ${\sf sub}(x, p)$, that is $x[p]$.
-		\item $t$: type of the preceding character $x[p - 1]$, set as 0 or 1 for L-type or S-type, respectively.
-	\end{itemize}
-
-	${\sf DSTRA_L}(x)$ and ${\sf DSTRA_S}(x)$ only contain the items of L-type and S-type substrings in ${\sf DSTRA}(x)$, respectively.
-
-	\item ${\sf LMSDATA}$. A string array. Each item stores the characters of an LMS substring.
-
-	\item ${\sf LMSNAME}$. An integer array. Each item stores the name of an LMS substring, where the name represents the substring's lexicographical order among all.
-
-\end{itemize}
-
-Notice that the above data structures organize the items in the external memory according to the lexicographical order of their corresponding substrings.
+\subsection{Overview of DSA-IS} \label{subsec:improvement}
 
 % Algorithm
-\begin{algorithm}
+\begin{algorithm*}
 	\SetAlgoNoLine
 	\KwIn{$x$}
-	\KwOut{$x1$}
+	\KwOut{$x_1$}
 
-	step 1: Partition $x$ into blocks $b_1, b_2, ..., b_k$.~\label{alg:dsais_reduction_phase:1} \\
-	step 2: Induced sort the substrings of $b_i$ for all $i \in [1, k]$ to compute ${\sf DSTRA_L}(b_i)$, ${\sf DSTRA_S}(b_i)$ and ${\sf LMSDATA}(b_i)$. ~\label{alg:dsais_reduction_phase:2}\\
-	step 3: Induced sort the substrings of $x$ from ${\sf DSTRA_L}(b_i)$, ${\sf DSTRA_S}(b_i)$ and ${\sf LMSDATA}(b_i)$ to compute ${\sf DSTRA}(x)$ and ${\sf LMSDATA}(x)$.
-Meanwhile, scan ${\sf DSTRA}(x)$ to compute ${\sf DSTRA_{LMS}}(x)$.~\label{alg:dsais_reduction_phase:3}\\
-	step 4: Scan ${\sf DSTRA_{LMS}}(x)$ and ${\sf LMSDATA}(x)$ to name the sorted LMS substrings for producing ${\sf LMSNAME}(x)$. ~\label{alg:dsais_reduction_phase:4}\\
-	step 5: Compute the reduced string $x1$ from ${\sf LMSNAME}(x)$ and ${\sf DSTRA_{LMS}}(x)$. ~\label{alg:dsais_reduction_phase:5}\\
+	$r := 0$ \label{alg1:1}
 
-	\caption{The Reduction Phase for DSA-IS}
-	\label{alg:dsais_reduction_phase}
-\end{algorithm}
+	\For{$i := n - 1; i \ge 0; i := i - 1$}{
 
-The algorithmic framework of the reduction phase for DSA-IS is shown in Algorithm~\ref{alg:dsais_reduction_phase}, where each step is sketched below.
+		\If{$t[i] = 1$ and $t[i - 1] = 0$}{
 
-Step 1 scans $x$ leftward to sequentially retrieve the LMS substrings and inserts them one by one into the blocks with a capacity $m = \mathcal{O}(\sf M)$, where ${\sf M}$ is the size of the internal memory.
-In this way, each block is either a single-block composed of one LMS substring or a multi-block composed of at least two successive LMS substrings.
-Particularly, each multi-block contains no more than $m$ characters such that it can be processed in the internal memory during the two phases, while a single-block may not be so.
+			$PQ_1.{\sf push}(\langle x[i], t[i], r, i\rangle)$ // sort S*-type characters 
+		}
+	}\label{alg1:2}
 
-Step 2 adopts different strategies to tackle $b_i$ with respect to whether the block is a single-block or a multi-block.
-Specifically, if $b_i$ is a multi-block, then it computes ${\sf STRA}(b_i)$ by conducting the substring sorting algorithm of SA-IS and scans ${\sf STRA}(b_i)$ and $b_i$ to obtain ${\sf DSTRA_L}(b_i)$, ${\sf DSTRA_S}(b_i)$ and ${\sf LMSDATA}(b_i)$; otherwise, the substrings in the block are already sorted and it directly scans $b_i$ to obtain the three arrays.
+	\While{$PQ_1.{\sf notEmpty}()$}{ \label{alg1:3}
 
-Step 3 also computes ${\sf DSTRA}(x)$ following the induced sorting method.
-In details, when inducing substrings in ${\sf DSTRA}(x)$, it retrieves the head character of the L-type/S-type preceding substring for the currently scanned substring from the external memory by using sequential I/O operations.
-This is feasible because the corresponding DSAITEM of the preceding substring is located at the leftmost unvisited position in ${\sf DSTRA_L}(b_i)/{\sf DSTRA_S}(b_i)$, where the subscript $i$ of the block that contains the target item can be determined in $\mathcal{O}(1)$ time using $\mathcal{O}(\frac{n}{m})$ space~\cite{Nong2015}.
+		$top = PQ_1.{\sf top}(), PQ_1.{\sf pop()}, r:= r + 1$
 
-Step 4 names the sorted LMS substrings according to their lexicographical order.
-For each pair of the neighboring substrings in ${\sf DSTRA_{LMS}}(x)$, it conducts a string comparison to check if they are equal, where their characters can be sequentially retrieved from ${\sf LMSDATA}(x)$.
-If yes, then the two substrings have a same name; otherwise, the name for the latter is greater than the former by one.
+		$pos = top.4th, ch = top.1st, pre\_ch = {\sf getPrecCh}(pos)$  // retrieve the heading character of the predecessor
 
-Step 5 sorts the names in ${\sf LMSNAME}(x)$ by ${\sf DSTRA_{LMS}}(x)[i].p$ for all $i \in [0, n1)$ to produce the reduced string $x1$.
+		\If{$top.2nd = 0$ and $pre\_ch \ge ch$ or $top.2nd = 1$ and $pre\_ch > ch$}{
 
-\subsubsection{Induction Phase}
+			$PQ_1.{\sf push}(\langle pre\_ch, 1, r, pos  - 1 \rangle)$ // induce the L-type predecessor
+		}
+		\Else{
+			$PQ_2.{\sf push}(\langle ch, 1, r, pos \rangle)$ // keep the sorted L*-type substrings
+		}
+	} \label{alg1:4}
 
-% Algorithm
-\begin{algorithm}
-	\SetAlgoNoLine
-	\KwIn{$x$, $\sf{PA}(x)$, ${\sf DSA}(x1)$}
-	\KwOut{${\sf DSA}(x)$}
+	$r := n$  \label{alg1:5}
+	\While {$PQ_2.{\sf notEmpty}()$}{
 
-	Step 1: Compute ${\sf DSA_{LMS}}(x)$ from ${\sf DSA}(x1)$ and decompose ${\sf DSA_{LMS}}(x)$ into ${\sf DSA_{LMS}}(b_i)$ for all $i \in [1, k]$.~\label{alg:dsais_induction_phase:1}\\
-	Step 2: Induced sort the suffixes of $b_i$ from ${\sf DSA_{LMS}}(b_i)$ to compute ${\sf DSA_L}(b_i)$ and ${\sf DSA_S}(b_i)$ for all $i \in [1, k]$.~\label{alg:dsais_induction_phase:2}\\
-	Step 3: Induced sort the suffixes of $x$ from ${\sf DSA_L}(b_i)$ and ${\sf DSA_S}(b_i)$ to compute ${\sf DSA}(x)$.~\label{alg:dsais_induction_phase:3} \\
+		$top = PQ_2.{\sf top}(), PQ_2.{\sf pop()}, r := r - 1$
 
-	\caption{The Induction Phase for DSA-IS}
-	\label{alg:dsais_induction_phase}
-\end{algorithm}
+		$pos = top.4th, ch = top.1st, pre\_ch = {\sf getPrecCh}(pos)$  // retrieve the heading character of the predecessor
 
-As shown in Algorithm~\ref{alg:dsais_induction_phase}, the induction phase of DSA-IS consists of three steps, where the data structures for the induction phase are similar to their counterparts for the reduction phase:
+		\If{$top.2nd = 1$ and $pre\_ch \le ch$ or $top.2nd = 0$ and $pre\_ch < ch$}{
 
-\begin{itemize}[itemindent = 0ex]
-	\item ${\sf DSA_{LMS}}$. A DSAITEM array. For an input $x$, each item of ${\sf DSA_{LMS}}(x)$ is associated with an LMS suffix of $x$.
+			$PQ_2.{\sf push}(\langle pre\_ch, 1, r, pos  - 1\rangle)$ // induce the S-type predecessor
+		}
+		\Else{
 
-	\item ${\sf DSA}$, ${\sf DSA_L}$ and ${\sf DSA_S}$.
-DSAITEM arrays.
-For an input $x$, each item of ${\sf DSA}(x)$ is associated with a suffix of $x$.
-${\sf DSA_L}(x)$ and ${\sf DSA_S}(x)$ only contain the items of L-type and S-type suffixes in ${\sf DSA}(x)$, respectively.
+			$Vec.{\sf push\_back}(pos)$ // keep the sorted S*-type substrings
+		}
+	} \label{alg1:6}
 
-\end{itemize}
+	$last\_str := 0, last\_pos = n - 1, name = 0$ \label{alg1:7}
 
-The first step in Algorithm~\ref{alg:dsais_induction_phase} is to induce the order of LMS substrings of $x$.
-For the purpose, it sorts the items of ${\sf DSA}(x1)$ by ${\sf DSA}(x1)[i].p$ for all $i\in [0, n1)$ to compute ${\sf ISA}(x1)$ and then sorts the items of ${\sf PA}(x)$ by ${\sf ISA}(x1)[j]$ for all $j\in [0,n1)$ to compute ${\sf DSA_{LMS}}(x)$.
-Afterward, it decomposes ${\sf DSA_{LMS}}(x)$ into ${\sf DSA_{LMS}}(b_i)$ for all $i \in [1, k]$ and reuses steps 2-3 of the reduction phase to compute ${\sf DSA}(x)$ by replacing ${\sf DSTRA_L}(b_i)$ and ${\sf DSTRA_S}(b_i)$ with ${\sf DSA_L}(b_i)$ and ${\sf DSA_S}(b_i)$, respectively.
+	\While {$Vec.{\sf notEmpty}()$}{
 
-\section{Details of DSA-IS+} \label{sec:dsais_plus}
+		$pos = Vec.{\sf pop\_back}(), str := {\sf getStr}(pos)$ // get S*-type substring
 
-It was observed from our experiments that, the program for DSA-IS outperforms that for eSAIS with respect to the space efficiency, but it runs slower than the latter due to the large I/O volume for sorting and naming substrings.
-In this section, we present two methods that can  substantially reduce the construction time and I/O volume of the reduction phase without a sacrifice in its high space efficiency.
-We demonstrate in Section~\ref{sec:experiments} that, by exploiting the use of these methods, the optimized version of DSA-IS, namely DSA-IS+, runs as fast as eSAIS and its peak disk usage is two-thirds as that of the latter.
+		\If{$last\_str \ne str$}{
 
-\subsection{Method A} \label{subsec:dsais_plus_method_a}
-Recall that in Algorithm~\ref{alg:dsais_reduction_phase}, DSA-IS first sorts LMS substrings in steps 2-3 and then names the substrings in their sorted order in step 4.
-These two procedures can be merged by making use of the following fact: $\forall i, j, \in [0, n)$, ${\sf sub}(x, i)$ and ${\sf sub}(x, j)$ are equal if and only if $x[i] = x[j]$ and ${\sf sub}(x, i + 1) = {\sf sub}(x, j + 1)$.
-Specifically, for any two neighboring substrings in ${\sf DSTR}(x)$, say ${\sf DSTR}(x)[p]$ and ${\sf DSTR}(x)[p + 1]$, there must have ${\sf DSTR}(x)[p] \le {\sf DSTR}(x)[p + 1]$.
-Hence, the only information required for naming the two substrings is to check if they are equal or not.
-Following the fact describe above, this can be trivially done by adding two fields $r1$ and $r2$ to each item of ${\sf DSTR}(x)$, where $r1$ and $r2$ respectively record the names of the corresponding substring and the succeeding substring.
-More specifically, assume all the items of LMS characters are already inserted into the corresponding buckets of ${\sf DSTR}(x)$, it takes the following two steps to sort and name substrings at the same time:
+			$name := name + 1$ // literally compare two lexicographically neighboring substrings 
+		}
 
-\begin{itemize}[itemindent = 5ex]
-	\item[step 1:] Initially set $r = 0$.
-Scan ${\sf DSTR}(x)$ rightward.
-For each scanned item $e1$ and previously scanned item $e2$, if $e1.r2 = e2.r2$ and $e1.ch = e2.ch$ then set $e1.r1 = e2.r1$; otherwise, set $e1.r1 = r$ and increase $r$ by one.
-If $e1.p > 0$ and $e1.t = 0$, then determine the block $b_i$ that contains ${\sf sub}(x, e1.p - 1)$, retrieve the leftmost unvisited item $e3$ from ${\sf DSTR_L}(b_i)$, set $e3.r2 = e1.r1$ and insert $e3$ into the leftmost empty position of ${\sf BKT_L}({\sf DSTR}(x), e3.c)$.
+		$Sorter.{\sf push}(\langle pos, name\rangle)$
 
-	\item[step 2:] Initially set $r = n - 1$.
-Scan ${\sf DSTR}(x)$ leftward.
-For each scanned item $e1$ and previously scanned item $e2$, if $e1.r2 = e2.r2$ and $e1.ch = e2.ch$ then set $e1.r1 = e2.r1$; otherwise, set $e1.r1 = r$ and decrease $r$ by one.
-If $e1.p > 0$ and $e1.t = 1$, then determine the block $b_i$ that contains ${\sf sub}(x, e1.p - 1)$, retrieve the leftmost unvisited item $e3$ from ${\sf DSTR_S}(b_i)$, set $e3.r2 = e1.r1$ and insert $e3$ into the rightmost empty position of ${\sf BKT_S}({\sf DSTR}(x), e3.c)$.
+		$last\_pos = pos, last\_str = str$
+	} \label{alg1:8}
 
-\end{itemize}
+	$Sorter.sort()$  \label{alg1:9}
 
-At the end of step 2, all the LMS substrings are already sorted and their names can be obtained by copying $r1$ from the items of ${\sf DSTR}(x)$.
+	\For{$i := 0; Sorter.{\sf notEmpty}(); i:= i + 1$}{
 
-\subsection{Method B} \label{subsec:dsais_plus_method_b}
-A solution to reducing the I/O volume is to employ a multi-way merge algorithm for combining sorted LMS substrings of all the blocks, where the task of the merger is to sort and name LMS substrings by literally comparing their characters.
-A string comparison can be done very quickly if the involved two substrings are wholly loaded into RAM.
-However, there may exist some LMS substrings that cannot be accommodated in the internal memory.
-For these long substrings, the method described in Section~\ref{subsec:dsais_plus_method_a} is employed.
-Following the ideas, we revise steps 2-4 of Algorithm~\ref{alg:dsais_reduction_phase} as below, where $i \in [1, k]$.
+		$x_1[i] = Sorter.2nd, Sorter.{\sf next}()$ // produce the reduced string
 
-\begin{itemize}[itemindent = 5ex]
-	\item [step 2':] Induced sort the substrings of $b_i$ to compute ${\sf LSA_L}(b_i)$, ${\sf LSA_S}(b_i)$, ${\sf DSTRA_L'}(b_i)$ and ${\sf DSTRA_S'}(b_i)$.
+	} \label{alg1:10}
 
-	\item [step 3':] Induced sort the substrings embraced in the long LMS substrings of $x$ from ${\sf LSA_L}(b_i)$, ${\sf DSTRA_L'}(b_i)$ and ${\sf DSTRA_S'}(b_i)$ to compute ${\sf LSA_L}(x)$.
+	\caption{The Reduction Phase for an IS Suffix Sorting Algorithm.}
 
-	\item [step 4':] Merge ${\sf LSA_L}(x)$ and ${\sf LSA_S}(b_i)$ by using a min heap to compute ${\sf LMSNAME}(x)$.
+	\label{alg:1}
+\end{algorithm*}
 
-\end{itemize}
+An IS suffix sorting algorithm mainly consists of a reduction phase for sorting S*-type substrings and an induction phase for sorting suffixes.
+Both two phases execute the same process to accomplish the sorting tasks, while the reduction phase additionally performs a naming process to produce a reduced string $x_1$.
+For a better understanding, we illustrate in Algorithm~\ref{alg:1} the details of the reduction phase and summarize the framework below, where $PQ_1$ and $PQ_2$ are two priority queues with a sorting key consisting of the first three components.
 
-The definitions of the newly introduced notations are given below.
+\begin{enumerate}[S1]
 
-\begin{itemize}[itemindent = 0ex]
-	\item long/short LMS substring and threshold value $D$.
-All the LMS substrings are classified into two categories: long and short.
-Specifically, an LMS substring is short if it contains no more than $D$ characters; otherwise, it is long.
+\item Scan $x$ to find S*-type characters and insert them into $PQ_1$. (lines~\ref{alg1:1}-\ref{alg1:2}) \label{reduction_phase_framework:1}
 
-	\item ${\sf DSTR'}$, ${\sf DSTRA_L'}$ and ${\sf DSTRA_S'}$.
-DSAITEM arrays.
-For an input $x$, ${\sf DSTR'}(x)$ only contains the items in ${\sf DSTR}(x)$ that associate with substrings embraced in long LMS substrings.
-We say ${\sf sub}(x, p)$ embraces ${\sf sub}(x, q)$ if the two substrings end with the same LMS character and $p \le q$.
-${\sf DSTRA_L'}(x)$ and ${\sf DSTRA_S'}(x)$ respectively contain the items in ${\sf DSTRA'}(x)$ that associate with L-type and S-type substrings.
+\item Sequentially access the sorted substrings popped from $PQ_1$ in ascending order to induce their L-type predecessors, where each induced L-type substring is pushed into $PQ_1$ and each scanned L*-type substring is pushed into $PQ_2$ for later use.
+(lines~\ref{alg1:3}-\ref{alg1:4}) \label{reduction_phase_framework:2}
 
-	\item ${\sf LSA_L}$ and ${\sf LSA_S}$.
-LSAITEM arrays.
-For an input $x$, each item of ${\sf LSA_L}(x)/{\sf LSA_S}(x)$ associates with a long/short LMS substring ${\sf sub}(x, p)$ and mainly consists of four components:
+\item Sequentially access the sorted substrings popped from $PQ_2$ in descending order to induce their S-type predecessors, where each induced S-type substring is pushed into $PQ_2$ and each scanned S*-type substring is appended to a vector $Vec$.
+(lines~\ref{alg1:5}-\ref{alg1:6}) \label{reduction_phase_framework:3}
 
-	\begin{itemize}
-		\item $p$: position index for ${\sf sub}(x, p)$.
+\item Compare the sorted S*-type substrings in $Vec$ to assign them names, where the name for each substring indicates it relative order among all.
+(lines~\ref{alg1:7}-\ref{alg1:8}) \label{reduction_phase_framework:4}
 
-		\item $s$: leftmost $D$ characters of ${\sf sub}(x, p)$, if it is long; otherwise, all the characters.
+\item Sort the names in position order to produce the reduced string $x_1$. (lines~\ref{alg1:9}-\ref{alg1:10}) \label{reduction_phase_framework:5}
 
-		\item $t$: type of the last character in $s$, set as 0 or 1 for L-type or S-type, respectively.
+\end{enumerate}
 
-		\item $r$: name of ${\sf sub}(x, p)$.
+It should be noticed that S\ref{reduction_phase_framework:2} and S\ref{reduction_phase_framework:3} induce the order of unsorted substrings from that of their heading characters and successors.
+When finished these two steps, all the S*-type substrings are sorted in their lexical order and stored in $Vec$.
+S\ref{reduction_phase_framework:4} performs a naming process to literally compare each pair of neighbors in the vector and assign them the same or different names with respect to whether they are equal or not.
+After that, S\ref{reduction_phase_framework:5} replaces each S*-type substring with its name to produce the reduced string $x_1$.
 
-	\end{itemize}
+Algorithm~\ref{alg:1} can be implemented in external memory by sequential I/O operations except when calling ${\sf getPrecCh}$ or ${\sf getStr}$.
+These two functions invoke a random disk access per execution and thus may lead to a performance bottleneck.
+The existing disk-based IS suffix sorting algorithms adopt different approaches to amortize the I/O overhead.
+For DSA-IS, it employs an internal-memory preprocessor to prefetch the required data and sort them in their access order before use.
+In general, it first divides the input string into $\mathcal{O}(n/M)$ blocks consisting of one or multiple successive S*-type substrings.
+Then, it loads each block into RAM and sorts the S*-type substrings in the block following the same idea of S\ref{reduction_phase_framework:2}-S\ref{reduction_phase_framework:5}, where it records the returns of ${\sf getPrecCh}$ and ${\sf getStr}$ during the sort.
+Afterward, it executes Algorithm~\ref{alg:1} to sort all the S*-type substrings and retrieves the preceding characters and the target substrings from the block-wise results by sequential I/O operations.
 
-\end{itemize}
+\subsection{Improvements on Design of DSA-IS}
 
-We describe each step detailedly in what follows.
+Recall that Algorithm~\ref{alg:1} performs the sorting and naming processes in sequence.
+Our first improvement is to reuse the substring naming method originally proposed for SAIS-PQ to combine the execution of these two processes during the reduction phase.
+In Algorithm~\ref{alg:1}, two priority queues are employed to sort substrings according to their heading characters and the rank of their successors, where the rank value indicates the access order of the corresponding substring.
+Notice that the lexical order of any two substrings/suffixes can be determined by that of their heading characters and successors as well, thus the priority queues can sort substrings by their names instead of their ranks.
+The name of currently scanned L-type/S-type substring can be computed by comparing its heading character and the name of its successor with that of the last scanned L-type/S-type substring.
+If both are equal, then the two substrings have the same name; otherwise, the former is assigned a name bigger/smaller than the latter.
 
-\vspace{2ex} \noindent {\bf Step 2'}
+From our experiments, sorting substrings by the priority queues takes long CPU time and large I/O volume.
+To reduce the consumption, we present a new method that combines the sorted S*-type substrings of all the blocks by a merger, which maintains a buffer for each block to preload the sorted S*-type substrings into RAM and an internal-memory heap to compare the substrings belonging to different blocks by literal comparisons.
+For the substrings too long to be wholly loaded into RAM, this method first uses the priority queue to sort them in external memory and then employs the merger to compare them with the short substrings based on the following property: suppose $sub_a$ and $sub_b$ are respectively long and short substrings, then $sub_a < sub_b$ if (1) $sub_a[0, D) = sub_b$ or (2) $sub_a[0, k) = sub_b[0, k)$ and $sub_af_a[k] < sub_b[k]$, where $D = \|sub_b\|$ and $k < D$.
+This indicates that we only need to fetch the first $D$ characters of a long substring into RAM when comparing it with a short substring containing $D$ characters.
 
-\noindent This step consists of two substeps:
+\subsection{Discussion} \label{subsec:further_discussion}
 
-\begin{itemize}[itemindent = 1ex]
-	\item [(a)] Compute ${\sf DSTRA_L'}(b_i)$ and ${\sf DSTRA_S'}(b_i)$ by reusing the algorithm for computing ${\sf DSTRA_L}(b_i)$ and ${\sf DSTRA_S}(b_i)$ in step 2 of Algorithm~\ref{alg:dsais_reduction_phase}.
+As demonstrated in Section~\ref{sec:experiments}, our implementation for the enhanced DSA-IS has an advantage over that for eSAIS in terms of time, space and I/O volume, but its peak disk use remains at a high level.
+This is mainly because it fails to gradually clean up the temporary data when they are being used.
+On the contrary, fSAIS stores the temporary data using multiple files and deletes each file immediately when all that residing on it is longer to use.
+Therefore, it is reasonable to believe that our program can achieve a substantial improvement following the same technique.
+Currently, we are attempting to further improve the algorithmic and implementation designs of DSA-IS.
 
-	\item [(b)] Scan ${\sf DSTRA_S'}(b_i)$ rightward.
-For each scanned item $e1$ with $e1.t = 0$, create an LSAITEM $e2$ for the current LMS substring, compute $e2.t$ and $e2.s$ by visiting $b_i$, and insert $e2$ into ${\sf LSA_L}(b_i)$ or ${\sf LSA_S}(b_i)$ according to its length.
-\end{itemize}
+\section{SA Checker}~\label{sec:check_sa}
 
-\vspace{2ex} \noindent {\bf Step 3'}
+\subsection{Details}
 
-\noindent This step consists of three substeps:
+A constructed SA should be verified to ensure its correctness before use.
+This checking function can be found in some widespread software packages like SA-IS, DC3 and eSAIS, where the latter two perform verification based on Theorem~\ref{theorem:1}.
 
-\begin{itemize}[itemindent = 1ex]
-	\item [(a)] Scan $x$ leftward with $i$ decreasing from $n - 1$ to $0$.
-For each scanned character $x[i]$, if $x[i]$ is an ending character of a long LMS substring, insert ${\sf sub}(x, i)$ into the rightmost empty position in ${\sf BKT_S}({\sf DSTRA'}(x), x[i])$.
+\begin{theorem} \label{theorem:1}
+	$sa[0, n)$ is the SA for $x[0, n)$ if and only if the following conditions are satisfied:\\
+	(1) $sa$ is a permutation of $[0, n)$. \\	
+	(2) ${\sf r_i < r_j} \Leftrightarrow (x[i], r_{i + 1}) < (x[i], r_{j + 1})$ for $ i, j \in [0, n)$ and $i\ne j$, where $r_i$ and $r_j$ indicates the relative order of ${\sf suf}(i)$ and ${\sf suf}(j)$ among all the suffixes, respectively.
+\\
+\end{theorem}
 
-	\item [(b)] Reuse step 1 of method A to sort and name the L-type substrings embraced in long LMS substrings of $x$ by replacing ${\sf DSTR}(x)$ with ${\sf DSTR'}(x)$.
+\begin{IEEEproof}
+	The proof is available in~\cite{Dementiev2008a}.
+\end{IEEEproof}
 
-	\item [(c)] Reuse step 2 of method A to sort and name the S-type substrings embraced in long LMS substrings of $x$ by replacing ${\sf DSTR}(x)$ with ${\sf DSTR'}(x)$.
-For each scanned item $e1$ that associates with an LMS substring, determine the block $b_i$ that contains ${\sf sub}(x, e1.p)$, retrieve the leftmost unvisited item $e4$ from ${\sf LSA_L}(b_i)$, set $e4.r = e1.r1$ and insert $e4$ to the front of ${\sf LSA_L}(x)$.
+This method executes two passes of integer sorts and each sort involves $n$ fixed-size tuples, thus it takes large disk space and I/O volume when implemented in the external memory, resulting in a non-negligible checking overhead.
+Take eSAIS for example, the peak disk use for sequentially building and checking SA rises up to $27n$, while the I/O volume also increases by $50n$.
 
-\end{itemize}
-
-\vspace{2ex} \noindent {\bf Step 4'}
-
-\noindent Initialize the heap by inserting the leftmost item of each sorted sequence of ${\sf LSA_L}(x)$ and $\{{\sf LSA_S}(b_1),...,{\sf LSA_S}(b_k)\}$.
-Sequentially pop the items in the heap.
-For currently popped item $e1$, insert into the heap the next item $e2$ of the same sequence and compare $e1$ with the previously popped item $e3$ to determine the names of their corresponding LMS substrings.
-In this step, each string comparison for heap-sorting items and naming substrings observes the following rules, where $e1$ and $e2$ are items to be compared.
-
-\begin{itemize}[itemindent = 5ex]
-
-	\item [rule 1:] If both items belong to $\{{\sf LSA_S}(b_1),...,{\sf LSA_S}(b_k)\}$, then literally compare $e1.s$ with $e2.s$ to determine their order.
-
-	\item [rule 2:] If both items belong to ${\sf LSA_L}(x)$, then directly compare $e1.r$ with $e2.r$ to determine their order.
-
-	\item [rule 3:] If one belongs to ${\sf LSA_L}(x)$ and the other belongs to $\{{\sf LSA_S}(b_1),...,{\sf LSA_S}(b_k)\}$, then literally compare $e1.s$ with $e2.s$.
-If equal, then continue to compare $e1.t$ with $e2.t$.
-
-\end{itemize}
-
-Clearly, the time overhead of each string comparison is upper-bounded by the threshold value $D$.
-
-\section{Lightweight SA checker} \label{sec:sachecker}
-
-It is essential for a suffix sorting algorithm to check the correctness of the output suffix arrays.
-Several existing SA construction algorithms verify their outputs based on Lemma~\ref{lemma:1}~\cite{Karkkainen2003}.
+We propose a fast and lightweight SA checking method that can be seamlessly integrated into any IS suffix sorting algorithms for building and checking an SA simultaneously.
+Before presentation, we introduce two lemmas that will be used to prove the correctness of our checking method.
+Both two Lemmas demonstrate the fact that the lexical order of two suffixes can be determined by their heading characters and successors.
+Particularly, Lemma~\ref{lemma:2} is applied to any IS suffix sorting algorithm for inducing unsorted suffixes from sorted ones during the induction phase.
 
 \begin{lemma} \label{lemma:1}
-	An array ${\sf SA}[0, n)$ is the suffix array of a string $x[0, n)$ if and only if the following conditions are satisfied for all $i, j \in [0, n)$ and $i\ne j$:\\
-	(1) ${\sf SA}[i], {\sf SA}[j] \in [0, n)$ and ${\sf SA}[i] \ne {\sf SA}[j]$. \\
-	(2)	${\sf ISA}[i] < {\sf ISA}[j] \Leftrightarrow (x[i], {\sf ISA}[i + 1]) < (x[j], {\sf ISA}[j + 1])$.
+	For $i, j \in [0, n)$ and $i \ne j$, ${\sf suf}(i) < {\sf suf}(j) \Leftrightarrow (x[i], {\sf suf}(i + 1)) < (x[j], {\sf suf}(j + 1))$.
 \end{lemma}
 
-This SA checker performs two passes of integer sorts, where the first pass checks if ${\sf SA}[0, n)$ is a permutation of $[0, n)$ for condition (1) and the second pass checks if the order of suffixes in each SA bucket is correct for condition (2).
-Although this method runs fast in the internal memory, the required disk space and I/O volume are considerably huge when it sorts items using the external memory, resulting in a non-negligible performance overhead.
-Take eSAIS for example, the peak disk usages for constructing and checking SA are around 24n and 27n, respectively, while the overall I/O volume also increases by 50n for conducting the two integer sorts.
+\begin{lemma} \label{lemma:2}
+	For $i, j \in [0, n)$ and $i \ne j$, $r_i < r_j \Leftrightarrow (x[i], r_{i + 1}) < (x[j], r_{j + 1})$, where $r_i$ and $r_j$ indicates the relative order of ${\sf suf}(i)$ and ${\sf suf}(j)$ among all the suffixes, respectively.
+\end{lemma}
 
-\subsection{Details}\label{subsec:sachecker:details}
+In what follows, we use the above lemmas to prove Theorem~\ref{theorem:2}, where $sa^*_1$ and $sa^*_2$ are two integer arrays that record the starting positions of the sorted S*-type suffixes.
+To be specific, the former is computed from the SA for $x_1$ and the latter is retrieved from the final SA for $x$.
 
-We describe a new SA checking method that can be seamlessly integrated into any IS-based induced sorting algorithm, which has almost no influence on the overall performance.
-The main idea is to verify the two conditions listed in Lemma~\ref{lemma:1} during the induction phase.
-In what follows, we first describe two observations that are always true in the process of induced sorting suffixes, where ${\sf SA_{LMS1}}(x)$ is the position index array for the sorted LMS suffixes produced by sorting ${\sf PA}(x)$ with ${\sf ISA}(x1)$ at the beginning of the induction phase, and ${\sf SA_{LMS1}}(x)[u]$ and ${\sf SA_{LMS1}}(x)[v]$ are the ending position indexes of the LMS substrings that embrace $x[{\sf SA}(x)[i]]$ and $x[{\sf SA}(x)[j]]$, respectively.
-
-\begin{observation} \label{observation:1} $\forall i, j \in [0, n)$, ${\sf SA}(x)[i] = {\sf SA}(x)[j] \; and \; i \ne j \Leftrightarrow {\sf SA_{LMS1}}(x)[u] = {\sf SA_{LMS1}}(x)[v] \; and \; u \ne v$.
-\end{observation}
+\begin{theorem} \label{theorem:2}
+	For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for the input string $x[0,n)$ if and only if the following conditions are satisfied: \\
+	(1) $sa^*_1[i] \ne sa^*_1[j]$ for $i, j \in [0, n)$ and $i \ne j$. \\
+	(2) $sa^*_1 = sa^*_2$. \\
+\end{theorem}
 
 \begin{IEEEproof}
-	Omit.
+	We only prove the sufficiency as the necessity is clear.
+
+	Suppose $sa$ is not a permutation of $[0, n)$.
+Without loss of generality, let $i \ne j$ and $sa[i] = sa[j]$, then $u \ne v$ and $sa[u] = sa[v]$ according Lemma~\ref{lemma:2}, where ${\sf suf}(sa[u])$ and ${\sf suf}(sa[v])$ are the successors of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[j])$, respectively.
+By recursively replacing $(i, j)$ with $(u, v)$, we can find two elements $sa[i_0]$ and $sa[j_0]$ storing the starting position of the same S*-type suffix.
+This indicates that $sa^*_1$ contains duplicate elements, which violates condition (1) in Theorem~\ref{theorem:2}.
+
+	Suppose $i < j$ and ${\sf suf}(sa[i]) > {\sf suf}(sa[j])$.
+If $x[sa[i]] < x[sa[j]]$, then ${\sf suf}(sa[i]) < {\sf suf}(sa[j])$ according to Lemma~\ref{lemma:1}; otherwise, if $x[sa[i]] > x[sa[j]]$, then $i > j$ according to Lemma~\ref{lemma:2}.
+Both lead to a contradiction.
+Given that $x[sa[i]] = x[sa[j]]$, then $u < v$ and ${\sf suf}(u) > {\sf suf}(v)$, where ${\sf suf}(sa[u])$ and ${\sf suf}(sa[v])$ are the successors of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[j])$, respectively.
+By recursively replacing $(i, j)$ with $(u, v)$, we can find two elements $sa[i_0]$ and $sa[j_0]$ such that at least one stores the starting position of an S*-type suffix.
+We continue the proof following one of the three cases below:
+	\begin{enumerate}[(1)]
+	\item $x[sa[i_0]]$ is S*-type and $x[sa[j_0]]$ is L-type.
+In this case, $x[sa[i_0] + 1] > x[sa[j_0] + 1]$.
+Therefore, $u_0 > v_0$ according to Lemma~\ref{lemma:2}.
+This indicates that $(x[sa[i_0]], u_0) > (x[sa[j_0]], v_0)$ and thus $i_0 > j_0$ according to Lemma~\ref{lemma:2}.
+This leads to a contradiction.
+	\item $x[sa[i_0]]$ is L-type and $x[sa[j_0]]$ is S*-type.
+In this case, $x[sa[i_0] + 1] < x[sa[j_0] + 1]$.
+Therefore, ${\sf suf}(sa[u_0]) < {\sf suf}(sa[v_0])$ according to Lemma~\ref{lemma:1}.
+This indicates that $(x[sa[i_0]], {\sf suf}(sa[u_0])) < (x[sa[j_0]], {\sf suf}(sa[v_0])$ and thus ${\sf suf}(sa[i_0]) < {\sf suf}(sa[j_0])$ according to Lemma~\ref{lemma:1}.
+This leads to a contradiction.
+	\item $x[sa[i_0]]$ and $x[sa[j_0]]$ are both S*-type.
+Let $(i_0, j_0)$ be the pair that maximize $sa[i_0] + sa[j_0]$ such that $i_0 < j_0$ and ${\sf suf}(sa[i_0]) > {\sf suf}(sa[j_0])$, then $u_0 > v_0$ and ${\sf suf}(sa[u_0]) > {\sf suf}(sa[v_0])$.
+\footnote{Notice that $(u_0, v_0)$ must exist because the sentinel is a size-one S*-type substring smaller than others.} Because $x[i_0, u_0) = x[j_0, v_0)$, there must be $i_0 > j_0$ according to Lemma~\ref{lemma:2}.
+This contradiction indicates that $sa^*_1$ and $sa^*_2$ differ from each other, which violates condition (2) in Theorem~\ref{theorem:2}.
+	\end{enumerate}
+
 \end{IEEEproof}
 
-\begin{observation} \label{observation:2} $\forall i, j \in [0, n)$, $i < j \; and \; {\sf suf}(x, {\sf SA}(x)[i]) > {\sf suf}(x, {\sf SA}(x)[j]) \Rightarrow u < v \; and \; {\sf suf}(x, {\sf SA_{LMS1}}(x)[u]) > {\sf suf}(x, {\sf SA_{LMS1}}(x)[v])$.
-\end{observation}
+The first condition in Theorem~\ref{theorem:2} can be easily checked when computing $sa^*_1$ from $sa_1$.
+For the second condition, a naive method is to first make a copy of $sa^*_1$ and $sa^*_2$ when inducing $sa$ and then sequentially compare their elements after the induction phase.
+This method takes linear time, space and I/O overhead.
+A more efficient alternative is to compute and compare the fingerprints for these two arrays, denoted by ${\sf fp}(sa^*_1)$ and ${\sf fp}(sa^*_2)$, instead to determine their equality.
+Obviously, $sa^*_1$ and $sa^*_2$ share a common fingerprint if they are equal to each other, but the inverse is not always true.
+Fortunately, by using the fingerprinting function described in~\cite{Karp1987}, the probability of a false match can be reduced to a completely negligible level.
+The above discussion leads us to the following statement.
 
-\begin{IEEEproof} \label{proof:1}
-	An IS-based SA construction algorithm induces the order of suffixes by utilizing the fact that ${\sf suf}(x, p_1) < {\sf suf}(x, q_1) \Leftrightarrow (x[p_1], {\sf suf}(x, p_1 + 1)) < (x[q_1], {\sf suf}(x, q_1 + 1)$.
-The algorithm guarantees that ${\sf ISA}(x)[p_1] < {\sf ISA}(x)[q_1] \Leftrightarrow (x[p_1], {\sf ISA}(x)[p_1 + 1]) < (x[q_1], {\sf ISA}(x)[q_1 + 1])$.
-This indicates that ${\sf suf}(x,p_1) < {\sf suf}(x, q_1)$ and ${\sf ISA}(x)[p_1] > {\sf ISA}(x)[q_1]$ only if ${\sf suf}(x, p_2) < {\sf suf}(x, q_2)$ and ${\sf ISA_{LMS1}}(x)[p_2] > {\sf ISA_{LMS1}}(x)[q_2]$, where $p_2$ and $q_2$ are the position indexes of the leftmost LMS characters on the right side of $x[p_1]$ and $x[q_1]$, respectively.
-\end{IEEEproof}
+\begin{corollary} \label{theorem:3}
+	For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for the input string $x[0, n)$ with a high probability if the following conditions are satisfied:
+	(1) $sa^*_1[i] \ne sa^*_1[j]$ for $i, j \in [0, n)$ and $i \ne j$. \\
+	(2) ${\sf fp}(sa^*_1) = {\sf fp}(sa^*_2)$. \\
+\end{corollary}
 
-The above observations are used to prove the correctness of the following statement.
+\subsection{Example} \label{sec:check_sa:example}
 
-\begin{theorem} \label{theorem:1}
-	For any IS-based induced sorting algorithm, the output array ${\sf SA}(x)$ is the suffix array of the input string $x[0, n)$ if and only if the following conditions are satisfied:
-
-	(1) $\forall i, j \in [0, n)$ and $i \ne j$, ${\sf SA_{LMS1}}(x)[i] \ne {\sf SA_{LMS1}}(x)[j]$.
-
-	(2) ${\sf SA_{LMS1}(x)} = {\sf SA_{LMS2}(x)}$, where ${\sf SA_{LMS2}}$ is the position index array for the sorted LMS suffixes retrieved from ${\sf SA}(x)$ at the end of the induction phase.
-\end{theorem}
-
-\begin{IEEEproof} \label{proof:2}
-	We only prove the sufficiency as the necessity is clear.
-
-	Suppose ${\sf SA}(x)$ is not a permutation of $[0, n)$.
-Let $i \ne j$ and ${\sf SA}(x)[i] = {\sf SA}(x)[j]$, then there must exist $u, v$ satisfying $u \ne v$ and ${\sf SA_{LMS1}}(x)[u] = {\sf SA_{LMS1}}(x)[v]$ according to Observation~\ref{observation:1}.
-This violates the first condition in Theorem~\ref{theorem:1}.
-
-	Suppose ${\sf suf}(x, i) > {\sf suf}(x, j)$ and ${\sf ISA}(x)[i] < {\sf ISA}(x)[j]$.
-Then, there must exist $u, v$ satisfying ${\sf suf}(x, u) > {\sf suf}(x, v)$ and ${\sf ISA_{LMS1}}(x)[u] < {\sf ISA_{LMS1}}(x)[v]$ according to Observation~\ref{observation:2}.
-Let $(u_0, v_0)$ be the pair that maximize $u + v$, then we have ${\sf ISA_{LMS2}}(x)[u_0] > {\sf ISA_{LMS2}}(x)[v_0]$.
-This violates the second condition in Theorem~\ref{theorem:1}.
-\end{IEEEproof}
-
-To verify the conditions in Theorem~\ref{theorem:1}, we can design a checker based on the following idea.
-The first condition can be checked after sorting ${\sf PA}(x)$ by ${\sf ISA}(x1)$ at the beginning of the induction phase.
-For the second condition, a naive method that takes $\mathcal{O}(n)$ time and space is to copy ${\sf SA_{LMS1}}(x)$ and ${\sf SA_{LMS2}}(x)$ and then compare their items in sequence at the end of the induction phase.
-An alternative for lowering the overhead is to compute the fingerprints for ${\sf SA_{LMS1}}(x)$ and ${\sf SA_{LMS2}}(x)$ by using fingerprint functions and then compare the fingerprints instead of the items in these arrays to determine their equality.
-Obviously, ${\sf SA_{LMS1}}(x)$ and ${\sf SA_{LMS2}}(x)$ have a common fingerprint if they are identical, but the inverse is not always true.
-Fortunately, the probability of a false match can be reduced to a completely negligible level by properly setting the parameters for the calculation formula.
-This leads us to Theorem~\ref{theorem:2}.
-
-\begin{theorem} \label{theorem:2}
-	For any IS-based induced sorting algorithm, the output array ${\sf SA}(x)$ is the suffix array of the input string $x[0, n)$ with a high probability if the following conditions are satisfied:
-
-	(1) $\forall i, j$, ${\sf SA_{LMS1}}(x)[i] \ne {\sf SA_{LMS1}}(x)[j]$.
-
-	(2) ${\sf FP(SA_{LMS1}(x))} = {\sf FP(SA_{LMS2}(x))}$.
-\end{theorem}
-
-\subsection{Implementation} \label{subsec:sachecker:implementation}
-We choose the approach presented in~\cite{Karp1987} to compute the fingerprints in need.
-Specifically, for a given array $x[0, n)$, the fingerprint ${\sf FP}(x)[0, n)$ can be iteratively computed by the formula ${\sf FP}(x)[i,j] = \sum_{p = i}^{j} \alpha^{j-p} \cdot x[p] \, mod \, \beta$, where $\beta$ is a large prime and $\alpha$ is an integer randomly chosen from domain $[1,\beta)$.
-In this way, we can slightly adapt Algorithm~\ref{alg:dsais_induction_phase} to compute ${\sf FP}(\sf SA_{LMS1}(x))$ by scanning ${\sf DSA_{LMS}}(x)$ in step 1 and ${\sf FP}(\sf SA_{LMS2}(x))$ by scanning ${\sf DSA}(x)$ during the time when inducing S-type suffixes in step 3, respectively.
-Obviously, the time, space and I/O complexities of the implementation are linearly related to $n$.
+xxxx
 
@@
 
-% Bibliography
 \bibliographystyle{IEEEtran}

### 2017-04-24 new-draft/draft.tex
2017/04/24

@@
 This indicates a great potential for improving space efficiency of the prior arts, such as DSA-IS and SAIS-PQ.
-In this paper, we present a checking method that enables any IS suffix sorting algorithms to check an SA when it is being built.
-For performance analysis, we integrate the proposed method into DSA-IS and implement the enhanced algorithm to evaluate the checking overhead, where the program design of DSA-IS has been optimized for a better speed and smaller peak disk use.
+In this paper, we present two checking methods that enable any IS suffix sorting algorithms to check an SA when it is being built, where the second method can also be applied to ensure the correctness of an LCP array.
+For performance analysis, we integrate the proposed methods into DSA-IS and implement the enhanced algorithm to evaluate the checking overhead, where the program design of DSA-IS has been optimized for a better speed and smaller peak disk use.
 From our experiments, the time, space and I/O consumptions for verification is negligible in comparison with that for construction.
-In view of the high performance of this checker and the existing IS suffix sorters, we are convinced that our checking method, in accompanied with the state-of-the-art IS suffix sorting algorithm, can constitute an efficient solution to the situations where checking is a must for a constructed SA.
+In view of the high performance of these checkers and the existing IS suffix sorters, we are convinced that our checking methods, in accompanied with the state-of-the-art IS suffix sorting algorithm, constitute efficient solutions to the situations where checking is a must after construction.
 
@@
 \begin{IEEEkeywords}
-Suffix array, construction and verification, external memory.
+Suffix and LCP arrays, construction and verification, external memory.
 \end{IEEEkeywords}}
@@
 
-\section{SA Builder}~\label{sec:build_sa}
+\section{Builder}~\label{sec:build_sa}
 
@@
 
-\section{SA Checker}~\label{sec:check_sa}
+\section{Checkers}~\label{sec:check_sa}
 

### 2017-04-25 new-draft/draft.tex
2017/04/25

@@
 This indicates a great potential for improving space efficiency of the prior arts, such as DSA-IS and SAIS-PQ.
-In this paper, we present two checking methods that enable any IS suffix sorting algorithms to check an SA when it is being built, where the second method can also be applied to ensure the correctness of an LCP array.
-For performance analysis, we integrate the proposed methods into DSA-IS and implement the enhanced algorithm to evaluate the checking overhead, where the program design of DSA-IS has been optimized for a better speed and smaller peak disk use.
+We present two checking methods that enable any IS suffix sorting algorithms to check an SA when it is being built, where the second method can also be applied to ensuring the correctness of an LCP array.
+For performance analysis, we integrate the proposed methods into DSA-IS and implement the enhanced algorithm to evaluate the checking overhead, where the program design of DSA-IS has been optimized in advance for a better speed and smaller peak disk use.
 From our experiments, the time, space and I/O consumptions for verification is negligible in comparison with that for construction.
-In view of the high performance of these checkers and the existing IS suffix sorters, we are convinced that our checking methods, in accompanied with the state-of-the-art IS suffix sorting algorithm, constitute efficient solutions to the situations where checking is a must after construction.
+In view of the high performance of these checkers and the existing IS suffix sorters, we are convinced that our checking methods, in accompanied with the state-of-the-art IS suffix sorting algorithm, constitute efficient solutions for the situations where checking is a must after construction.
 
@@
 
-Given a string drawn from a constant or integer alphabet, its suffix array can be built by the internal-memory algorithm SA-IS~\cite{Nong11} in linear time and space.
-So far, the IS method has been also applied to designing three external-memory suffix sorting algorithms~eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}, which have better time and I/O complexities than other alternatives, e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}.
-However, the peak disk uses of their current programs are at least 15 times the sizes of input strings, resulting in a performance bottleneck.
-Recently, it was reported in~\cite{Karkkainen2017} a new IS suffix sorting algorithm fSAIS with a careful engineering that consumes nearly optimal disk space and runs faster than the parallel suffix sorter pSAscan on massive datasets.
-The algorithmic design of fSAIS is highly dependent on that of the prior arts.
-Particularly, it employs a sorting method similar to that of DSA-IS for inducing suffixes/substrings and the naming method of SAIS-PQ for producing reduced strings.
-The substantial improvement on space efficiency is mainly achieved by means of a monotone priority queue  for sorting substrings/suffixes and a garbage collector for quickly recycling disk space occupied by data no longer to use.
-As a result, its implementation takes less than $8n$ disk space for any input string of a size $n \le 2^{40}$, indicating a great potential for optimizing the programs for DSA-IS and SAIS-PQ.
-In this paper, we make our first attempt to redesign DSA-IS using new substring sorting and naming methods.
+Given a string drawn from any constant or integer alphabet	, its suffix array can be constructed in linear time and space by the internal-memory algorithm SA-IS~\cite{Nong11}.
+So far, the IS method has been also applied to designing three external-memory suffix sorting algorithms~eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}.
+Theoretically, these algorithms have better time and I/O complexities than the existing alternatives, e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}, but they all suffer from a space bottleneck due to the large disk space in use.
+Recently, it was reported in~\cite{Karkkainen2017} a new IS suffix sorting algorithm fSAIS with a careful engineering that achieves optimal peak disk use and fastest speed among the state-of-the-art suffix sorters on massive datasets.
+The algorithm design of fSAIS is highly dependent on that of the prior arts, where it uses a sorting method similar to that of DSA-IS for inducing suffixes/substrings and the naming method of SAIS-PQ for producing reduced strings.
+Its high space performance is mainly due to the use of a monotone priority queue for sorting substrings/suffixes and a garbage collector for quickly recycling disk space occupied by temporary data no longer to use.
+As a result, its implementation takes less than $8n$ disk space for any input string of a size $n \le 2^{40}$, which also indicates a great potential for optimizing the program designs of DSA-IS and SAIS-PQ.
+In this paper, we present our first attempt to redesign DSA-IS by means of new substring sorting and naming methods.
 The experimental results indicate that our program for the enhanced algorithm consumes about two-thirds as much disk space as that for eSAIS and runs faster than the latter on real world datasets.
-In addition, we also give a discussion on further improving our program to achieve a better space efficiency using the techniques for implementing fSAIS.
+The enhanced program can be further improved to achieve better time and space performance using the techniques for implementing fSAIS.
 
-To detect potential computation errors caused by implementation bugs and/or hardware malfunctions, a constructed SA should be verified to ensure its correctness before use.
+To detect potential computation errors caused by implementation bugs and/or hardware malfunctions, a suffix array should be verified before use.
 This checking function has been included into various widespread software packages like SA-IS, DC3 and eSAIS.
 Particularly, the latter two verify an SA using the checking method presented in~\cite{Dementiev2008a}, which mainly consists of two passes of external-memory sorts involving $\mathcal{O}(n)$ fixed-size tuples with an integer key.
-In this paper, we propose a new checking method that enables any IS suffix sorting algorithm to check an SA when it is being built.
-For performance analysis, we integrate the proposed method into the optimized DSA-IS and implement the combined algorithm, called DSA-IS+, to evaluate the checking overhead.
+In this paper, we propose two checking methods that enable any IS suffix sorting algorithm to verify an SA when it is being built, where the second method can also be applied to ensuring the correctness of an LCP array.
+For performance analysis, we integrate the proposed checking method into the optimized DSA-IS and implement the combined algorithms to evaluate the checking overhead.
 From our experiments, the time, space and I/O consumptions for verification is negligible in comparison with that for construction.
-In view of the high performance of this checker and the existing IS suffix sorters, we believe that our checking method, together with the state-of-the-art IS suffix sorting algorithm, can constitute an efficient solution to the situations where checking is a must for a constructed SA.
+In view of the high performance of these checkers and the existing IS suffix sorters, we believe that our checking methods, together with the state-of-the-art IS suffix sorting algorithm, can constitute efficient solutions for the situations where building and checking must be performed at the same time.
 
-The rest of the paper is organized as follows.
-We first describe in Sections~\ref{sec:build_sa} and~\ref{sec:check_sa} the adapted algorithmic design of DSA-IS and the proposed checking method, respectively.
-Then, we demonstrate the experimental results in Section~\ref{sec:experiments} and draw the concluding remarks in Section~\ref{sec:conclusion}.
+The rest of this paper is organized as follows.
+We first describe in Sections~\ref{sec:build_sa} and~\ref{sec:check_sa} the adapted algorithmic design of DSA-IS and the proposed checking methods, respectively.
+Then, we demonstrate the experimental results in Section~\ref{sec:experiments} and draw the conclusion in Section~\ref{sec:conclusion}.
 

### 2017-06-06 new-draft/draft.tex
2016/06/06

@@
 \newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{formula}[theorem]{Formula}
 
@@
 
-In the past decade, the induced sorting~(IS) method has been successively used to design disk-based suffix array~(SA) construction algorithms that have better time and I/O complexities than the existing alternatives.
-However, their current programs commonly suffer from a performance bottleneck due to large disk space in need.
-Recently, it was reported a carefully engineered IS suffix sorter that takes less than $8n$ disk space for constructing an SA encoded by 40-bit integers.
-This indicates a great potential for improving space efficiency of the prior arts, such as DSA-IS and SAIS-PQ.
-We present two checking methods that enable any IS suffix sorting algorithms to check an SA when it is being built, where the second method can also be applied to ensuring the correctness of an LCP array.
-For performance analysis, we integrate the proposed methods into DSA-IS and implement the enhanced algorithm to evaluate the checking overhead, where the program design of DSA-IS has been optimized in advance for a better speed and smaller peak disk use.
-From our experiments, the time, space and I/O consumptions for verification is negligible in comparison with that for construction.
-In view of the high performance of these checkers and the existing IS suffix sorters, we are convinced that our checking methods, in accompanied with the state-of-the-art IS suffix sorting algorithm, constitute efficient solutions for the situations where checking is a must after construction.
+%The induced sorting~(IS) method has been applied to designing several disk-based suffix array~(SA) construction algorithms.
+Although these algorithms have better time and I/O complexities than the other alternatives, their current programs commonly suffer from a bottleneck due to the large disk space in need.
+
+The induced sorting~(IS) method has been applied to designing efficient suffix array~(SA) construction algorithms on both internal and external memory models.
+In this paper, we propose two checking methods that enable any IS suffix sorters to check an SA when it is being built.
+The first method is capable of checking both suffix and LCP arrays for strings on constant alphabets, while the second method is specific for checking SA only but applicable to strings on both constant and integer alphabets.
+For performance analysis, we integrate the second checking method into the existing SA construction algorithm DSA-IS, where the design of DSA-IS is adapted by new substring sorting and naming methods for reducing its peak disk use to at most 14 times as input size.
+From our experimental results, the time, space and I/O consumptions for verification by our second method is negligible to that for construction by the adapted DSA-IS.
+This convinced that our checkers, together with the state-of-the-art IS suffix sorters, may constitute efficient solutions for building and checking SA at the same time.
 
@@
 
-Given a string drawn from any constant or integer alphabet	, its suffix array can be constructed in linear time and space by the internal-memory algorithm SA-IS~\cite{Nong11}.
-So far, the IS method has been also applied to designing three external-memory suffix sorting algorithms~eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}.
-Theoretically, these algorithms have better time and I/O complexities than the existing alternatives, e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}, but they all suffer from a space bottleneck due to the large disk space in use.
-Recently, it was reported in~\cite{Karkkainen2017} a new IS suffix sorting algorithm fSAIS with a careful engineering that achieves optimal peak disk use and fastest speed among the state-of-the-art suffix sorters on massive datasets.
-The algorithm design of fSAIS is highly dependent on that of the prior arts, where it uses a sorting method similar to that of DSA-IS for inducing suffixes/substrings and the naming method of SAIS-PQ for producing reduced strings.
-Its high space performance is mainly due to the use of a monotone priority queue for sorting substrings/suffixes and a garbage collector for quickly recycling disk space occupied by temporary data no longer to use.
-As a result, its implementation takes less than $8n$ disk space for any input string of a size $n \le 2^{40}$, which also indicates a great potential for optimizing the program designs of DSA-IS and SAIS-PQ.
-In this paper, we present our first attempt to redesign DSA-IS by means of new substring sorting and naming methods.
-The experimental results indicate that our program for the enhanced algorithm consumes about two-thirds as much disk space as that for eSAIS and runs faster than the latter on real world datasets.
-The enhanced program can be further improved to achieve better time and space performance using the techniques for implementing fSAIS.
+The suffix array of any constant or integer string can be built in linear time and RAM space by SA-IS~\cite{Nong11}.
+The induced-sorting method has been also applied to designing fast external-memory SA construction algorithms eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}.
+These three disk-based variants have better time complexities than the other alternatives~(e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}), but they all suffer from a bottleneck due to the large disk space in use.
+For example, when building SA encoded by 40-bit integers, the peak disk use for eSAIS, DSA-IS and SAIS-PQ are $24n$, $20n$ and $15n$, respectively.
+This situation changed due to the work presented in~\cite{Karkkainen2017}, which proposed a carefully engineered IS suffix sorter fSAIS that takes no more than $8n$ disk memory for any constant or integer string of size $n \le 2^{40}$~\cite{Karkkainen2017}.
+As reported, fSAIS reuses the substring/suffix sorting and naming methods of DSA-IS and SAIS-PQ, it achieves the remarkable space performance by means of a monotone priority queue for sorting fixed-size tuples and an I/O layer for allocating and deallocating disk space in a fine-grained way.
+The successful design of fSAIS shows a great potential for sharply decreasing the disk space requirements of the current programs for DSA-IS and SAIS-PQ.
+In this paper, we present our first attempt to improve DSA-IS by new substring sorting and naming methods.
+From the experiments, our program for the adapted algorithm DSA-IS+ takes $12n$ disk space in average and runs as fast as eSAIS on various real-world datasets.
+We will demonstrate that this program can be further improved using the optimization technique for implementing fSAIS.
 
-To detect potential computation errors caused by implementation bugs and/or hardware malfunctions, a suffix array should be verified before use.
-This checking function has been included into various widespread software packages like SA-IS, DC3 and eSAIS.
-Particularly, the latter two verify an SA using the checking method presented in~\cite{Dementiev2008a}, which mainly consists of two passes of external-memory sorts involving $\mathcal{O}(n)$ fixed-size tuples with an integer key.
-In this paper, we propose two checking methods that enable any IS suffix sorting algorithm to verify an SA when it is being built, where the second method can also be applied to ensuring the correctness of an LCP array.
-For performance analysis, we integrate the proposed checking method into the optimized DSA-IS and implement the combined algorithms to evaluate the checking overhead.
-From our experiments, the time, space and I/O consumptions for verification is negligible in comparison with that for construction.
-In view of the high performance of these checkers and the existing IS suffix sorters, we believe that our checking methods, together with the state-of-the-art IS suffix sorting algorithm, can constitute efficient solutions for the situations where building and checking must be performed at the same time.
+To avoid computation errors caused by implementation bugs and/or hardware malfunctions, an SA should be checked to ensure its correctness after construction.
+Currently, the software packages for DC3 and eSAIS provide users a checker based on the method proposed in~\cite{Dementiev2008a}, which performs two disk sorts for $\mathcal{O}(n)$ fixed-size tuples with an integer key.
+When running in external memory, this checker takes one-third as much time and I/O consumptions as the programs for DSA-IS and eSAIS, it also requires nearly the same disk space as the two SA builders.
+Against the background, we propose two checking methods that enable any IS suffix sorting algorithm to build and check SA at the same time.
+In comparison with the previous checking method, they are rather lightweight in terms of time, space and I/O volume.
+It should be noticed that these two methods are designed for different situations.
+In brief, the first method can check both suffix and LCP arrays, but it is specific for strings on constant alphabets.
+The second method can be applied to strings on constant and integer alphabets, but it can check SA only.
+As will be seen in our experiments, the consumption for checking by our second method is negligible compared with that for building by DSA-IS+.
 
 The rest of this paper is organized as follows.
-We first describe in Sections~\ref{sec:build_sa} and~\ref{sec:check_sa} the adapted algorithmic design of DSA-IS and the proposed checking methods, respectively.
-Then, we demonstrate the experimental results in Section~\ref{sec:experiments} and draw the conclusion in Section~\ref{sec:conclusion}.
+Section~\ref{sec:builder} describes the framework of IS suffix sorting algorithms and the idea of new substring sorting and naming methods, Section~\ref{sec:checkers} the designs of two checking methods, Sections~\ref{sec:experiments} the experimental results, and Section~\ref{sec:conclusion} the concluding remarks.
 
-\section{Builder}~\label{sec:build_sa}
+\section{Builder}~\label{sec:builder}
 
@@
 
-Given an input string $x[0, n)$ with a unique smallest character $x[n - 1]$, we denote by ${\sf suf}(i)$ a suffix starting with $x[i]$ and ${\sf sub}(i,j)$ a substring running from $x[i]$ to $x[j]$.
-The following notations are used in our presentation for denotation convenience.
+Consider a string $x[0,n)$ drawn from a full-ordered alphabet $\Sigma$.
+The ending character $x[n - 1]$ is supposed to be unique and lexicographically smaller than any others in $x$.
+For convenience, we denote by ${\sf suf}(i)$ the suffix running from $x[i]$ to the ending character and ${\sf sub}(i,j)$ the substring running from $x[i]$ to $x[j]$, respectively.
+The following notations are also used in our presentation.
 
-Character/substring/suffix classification.
-Characters of $x$ are classified into two categories: L-type and S-type.
-To be specific, $x[i]$ is S-type if (1) $i = n - 1$ or (2) $x[i] = x[i + 1]$ and $x[i + 1]$ is S-type; otherwise, $x[i]$ is L-type.
-Particularly, if $x[i]$ and $x[i - 1]$ are respectively S-type and L-type, then $x[i]$ is also S*-type.
-In addition, a substring/suffix is of the same type as its heading character.
-We use an integer array $t$ to indicate the type of characters, where $t[i] = 1$ or $0$ if $x[i] $ S-type or L-type, respectively.
+{\em Character/substring/suffix classification.} Characters in $x$ are classified into two categories: L-type and S-type.
+We say $x[i]$ is S-type if (1) $i = n - 1$ or (2) $x[i] = x[i + 1]$ and $x[i + 1]$ is S-type; otherwise $x[i]$ is L-type.
+Furthermore, if $x[i]$ and $x[i - 1]$ are respectively S-type/L-type and L-type/S-type, then $x[i]$ is also called S*-type/L*-type.
+An array $t$ records the type of all the characters in $x$, $t[i]$ is set to $1$ or $0$ if $x[i]$ is S-type or L-type, respectively.
+Each substring/suffix is of the same type as its heading character.
+In this paper, we only consider the substrings ending with an S*-type character.
 
-Predecessor and Successor.
-For any two neighboring characters $x[i]$ and $x[i + 1]$, we say $x[i]$ is the predecessor of $x[i + 1]$ while the latter is the successor of the former.
-Similarly, there also exists a predecessor-successor relationship between ${\sf suf}(i)/{\sf sub}(i, j)$ and ${\sf suf}(i + 1)/{\sf sub}(i + 1, j)$.
+{\em Predecessor and successor.} Given two characters $x[i]$ and $x[i + 1]$, $x[i]$ is the predecessor of $x[i + 1]$ and $x[i + 1]$ is the successor of $x[i]$.
+Accordingly, we define the predecessor-successor relationship between two suffixes or substrings starting with neighboring characters.
 
-\subsection{Overview of DSA-IS} \label{subsec:improvement}
+{\em Reduced string.} Split $x$ into multiple substrings such that each substring only contains two S*-type characters at the starting and ending positions.
+Assign each substring a name and replace them with their names to produce a reduced string $x_1$, where the name indicates the lexical order of the corresponding substring among all.
+
+{\em Suffix array.} The suffix array $sa$ indicates the lexicographical order of all the suffixes in $x$, where $sa[i]$ records the starting position of the $(i + 1)$-th smallest suffix.
+We also define the SA for $x_1$ by $sa_1$.
+
+\subsection{Introduction to IS Suffix Sorting Algorithms} \label{subsec:improvement}
 
 % Algorithm
-\begin{algorithm*}
+\begin{algorithm}
 	\SetAlgoNoLine
 	\KwIn{$x$}
-	\KwOut{$x_1$}
+	\KwOut{$sa$}
 
-	$r := 0$ \label{alg1:1}
+	/* Reduction Phase */ \\
+	Sort S*-type substrings by the IS method.
 
-	\For{$i := n - 1; i \ge 0; i := i - 1$}{
+	Name the sorted S*-type substrings to produce $x_1$.
 
-		\If{$t[i] = 1$ and $t[i - 1] = 0$}{
+	~\\
 
-			$PQ_1.{\sf push}(\langle x[i], t[i], r, i\rangle)$ // sort S*-type characters 
-		}
-	}\label{alg1:2}
+	/* Check Recursion Condition */ \\	
+	\If{exist two equal characters in $x_1$}{
 
-	\While{$PQ_1.{\sf notEmpty}()$}{ \label{alg1:3}
-
-		$top = PQ_1.{\sf top}(), PQ_1.{\sf pop()}, r:= r + 1$
-
-		$pos = top.4th, ch = top.1st, pre\_ch = {\sf getPrecCh}(pos)$  // retrieve the heading character of the predecessor
-
-		\If{$top.2nd = 0$ and $pre\_ch \ge ch$ or $top.2nd = 1$ and $pre\_ch > ch$}{
-
-			$PQ_1.{\sf push}(\langle pre\_ch, 1, r, pos  - 1 \rangle)$ // induce the L-type predecessor
-		}
-		\Else{
-			$PQ_2.{\sf push}(\langle ch, 1, r, pos \rangle)$ // keep the sorted L*-type substrings
-		}
-	} \label{alg1:4}
-
-	$r := n$  \label{alg1:5}
-	\While {$PQ_2.{\sf notEmpty}()$}{
-
-		$top = PQ_2.{\sf top}(), PQ_2.{\sf pop()}, r := r - 1$
-
-		$pos = top.4th, ch = top.1st, pre\_ch = {\sf getPrecCh}(pos)$  // retrieve the heading character of the predecessor
-
-		\If{$top.2nd = 1$ and $pre\_ch \le ch$ or $top.2nd = 0$ and $pre\_ch < ch$}{
-
-			$PQ_2.{\sf push}(\langle pre\_ch, 1, r, pos  - 1\rangle)$ // induce the S-type predecessor
-		}
-		\Else{
-
-			$Vec.{\sf push\_back}(pos)$ // keep the sorted S*-type substrings
-		}
-	} \label{alg1:6}
-
-	$last\_str := 0, last\_pos = n - 1, name = 0$ \label{alg1:7}
-
-	\While {$Vec.{\sf notEmpty}()$}{
-
-		$pos = Vec.{\sf pop\_back}(), str := {\sf getStr}(pos)$ // get S*-type substring
-
-		\If{$last\_str \ne str$}{
-
-			$name := name + 1$ // literally compare two lexicographically neighboring substrings 
+		Recursively call the reduction phase on $x_1$.
 	}
 
-		$Sorter.{\sf push}(\langle pos, name\rangle)$
+	~\\
 
-		$last\_pos = pos, last\_str = str$
-	} \label{alg1:8}
+	/* Induction Phase */ \\
+	Sort suffixes by the IS method.
 
-	$Sorter.sort()$  \label{alg1:9}
-
-	\For{$i := 0; Sorter.{\sf notEmpty}(); i:= i + 1$}{
-
-		$x_1[i] = Sorter.2nd, Sorter.{\sf next}()$ // produce the reduced string
-
-	} \label{alg1:10}
-
-	\caption{The Reduction Phase for an IS Suffix Sorting Algorithm.}
+	\caption{The general framework for an IS suffix sorting algorithm.}
 
 	\label{alg:1}
-\end{algorithm*}
+\end{algorithm}
 
-An IS suffix sorting algorithm mainly consists of a reduction phase for sorting S*-type substrings and an induction phase for sorting suffixes.
-Both two phases execute the same process to accomplish the sorting tasks, while the reduction phase additionally performs a naming process to produce a reduced string $x_1$.
-For a better understanding, we illustrate in Algorithm~\ref{alg:1} the details of the reduction phase and summarize the framework below, where $PQ_1$ and $PQ_2$ are two priority queues with a sorting key consisting of the first three components.
+As shown in Algorithm~\ref{alg:1}, an IS suffix sorting algorithm first performs a reduction phase to produce the reduced string $x_1$.
+If there exist two equal characters in $x_1$, then it recursively calls the reduction phase with $x_1$ as input; otherwise, all the S*-type suffixes in $x$ are already sorted and it performs an induction phase to produce $sa$ from these sorted suffixes.
+According to the IS principle, the order of two substrings/suffixes is determined by comparing their heading characters and successors in sequence during the execution of reduction/induction phase.
+Notice that this key operation takes multiple random accesses to $x$ and $sa$.
+It can be done very quickly if both $x$ and $sa$ are wholly loaded into RAM; otherwise, each access may take an individual I/O operation to external memory and thus becomes the bottleneck.
+
+To amortize the I/O overhead, eSAIS, DSA-IS and SAIS-PQ use different methods to obtain heading characters in a disk-friendly way.
+Particularly, DSA-IS first divides $x$ into multiple blocks and calls SA-IS to sort suffixes in each block at the beginning of induction phase, where the heading characters in need are written to external memory in their access order.
+Then, it organizes the sorted suffixes in a priority queue and visits them sequentially to induce the order of their predecessors following the IS principle.
+Because the required heading characters was already stored on disks, they can be read into RAM by sequential I/O operations when needed.
+This technique is also applied to sorting substrings during the reduction phase.
+Afterward, a naming process is conducted to check equality of lexicographically neighboring substrings for producing the reduced string $x_1$.
+To do this, DSA-IS maintains a copy for each substring in external memory and arranges them in their sorted order.
+As can be seen in Section~\ref{sec:experiments}, our program for DSA-IS requires less disk space than that for eSAIS, but it runs slower than the latter due to the large I/O volume for sorting and naming S*-type substrings.
+
+\subsection{Improvements on DSA-IS}
+
+We first describe a new method for sorting S*-type substrings during the reduction phase.
+Given an integer value $D$, all the S*-type substrings are classified into long and short categories with respect to whether or not they contain more than $D$ characters.
+Our sorting method executes the following three steps in sequence after splitting $x$ into blocks $\{b_1, b_2, ...\}$.
 
 \begin{enumerate}[S1]
+	\item Sort S*-type substrings in each block by calling the reduction phase of SA-IS.
+Copy the short substrings and store them on the disk in their sorted order.
+Copy the heading characters needed for sorting long substrings and store them on the disk in their access order.~\label{sort_and_name_method:1}
 
-\item Scan $x$ to find S*-type characters and insert them into $PQ_1$. (lines~\ref{alg1:1}-\ref{alg1:2}) \label{reduction_phase_framework:1}
+	\item Sort long S*-type substrings by the same way of DSA-IS. Copy the leftmost $D$ characters of the long substrings and store them on the disk in their sorted order.~\label{sort_and_name_method:2}
 
-\item Sequentially access the sorted substrings popped from $PQ_1$ in ascending order to induce their L-type predecessors, where each induced L-type substring is pushed into $PQ_1$ and each scanned L*-type substring is pushed into $PQ_2$ for later use.
-(lines~\ref{alg1:3}-\ref{alg1:4}) \label{reduction_phase_framework:2}
-
-\item Sequentially access the sorted substrings popped from $PQ_2$ in descending order to induce their S-type predecessors, where each induced S-type substring is pushed into $PQ_2$ and each scanned S*-type substring is appended to a vector $Vec$.
-(lines~\ref{alg1:5}-\ref{alg1:6}) \label{reduction_phase_framework:3}
-
-\item Compare the sorted S*-type substrings in $Vec$ to assign them names, where the name for each substring indicates it relative order among all.
-(lines~\ref{alg1:7}-\ref{alg1:8}) \label{reduction_phase_framework:4}
-
-\item Sort the names in position order to produce the reduced string $x_1$. (lines~\ref{alg1:9}-\ref{alg1:10}) \label{reduction_phase_framework:5}
-
+	\item Merge short and long S*-type substrings by a multi-way sorter.~\label{sort_and_name_method:3}
 \end{enumerate}
 
-It should be noticed that S\ref{reduction_phase_framework:2} and S\ref{reduction_phase_framework:3} induce the order of unsorted substrings from that of their heading characters and successors.
-When finished these two steps, all the S*-type substrings are sorted in their lexical order and stored in $Vec$.
-S\ref{reduction_phase_framework:4} performs a naming process to literally compare each pair of neighbors in the vector and assign them the same or different names with respect to whether they are equal or not.
-After that, S\ref{reduction_phase_framework:5} replaces each S*-type substring with its name to produce the reduced string $x_1$.
+Assume the number of blocks is $k$, S\ref{sort_and_name_method:1} copies the sorted short substrings in $b_i$ to an external memory vector $vs_i$ and S\ref{sort_and_name_method:2} copies the sorted long substrings to an external memory vector $vl$.
+Then, S\ref{sort_and_name_method:3} uses a sorter to merge these partial results, where the sorter maintains a minimal heap in RAM to determine the current smallest substring in $vs_1, vs_2, ..., vs_k$ and $vl$.
+For any two substrings in the heap, their lexicographical order can be determined in $\mathcal{O}(D)$ time by literally comparing their characters from left to right.
+This sorting method achieves a good performance when most of the S*-type substrings are short.
+This makes it feasible in practice because the average length of S*-type substrings in real-world datasets are typically small.
 
-Algorithm~\ref{alg:1} can be implemented in external memory by sequential I/O operations except when calling ${\sf getPrecCh}$ or ${\sf getStr}$.
-These two functions invoke a random disk access per execution and thus may lead to a performance bottleneck.
-The existing disk-based IS suffix sorting algorithms adopt different approaches to amortize the I/O overhead.
-For DSA-IS, it employs an internal-memory preprocessor to prefetch the required data and sort them in their access order before use.
-In general, it first divides the input string into $\mathcal{O}(n/M)$ blocks consisting of one or multiple successive S*-type substrings.
-Then, it loads each block into RAM and sorts the S*-type substrings in the block following the same idea of S\ref{reduction_phase_framework:2}-S\ref{reduction_phase_framework:5}, where it records the returns of ${\sf getPrecCh}$ and ${\sf getStr}$ during the sort.
-Afterward, it executes Algorithm~\ref{alg:1} to sort all the S*-type substrings and retrieves the preceding characters and the target substrings from the block-wise results by sequential I/O operations.
-
-\subsection{Improvements on Design of DSA-IS}
-
-Recall that Algorithm~\ref{alg:1} performs the sorting and naming processes in sequence.
-Our first improvement is to reuse the substring naming method originally proposed for SAIS-PQ to combine the execution of these two processes during the reduction phase.
-In Algorithm~\ref{alg:1}, two priority queues are employed to sort substrings according to their heading characters and the rank of their successors, where the rank value indicates the access order of the corresponding substring.
-Notice that the lexical order of any two substrings/suffixes can be determined by that of their heading characters and successors as well, thus the priority queues can sort substrings by their names instead of their ranks.
-The name of currently scanned L-type/S-type substring can be computed by comparing its heading character and the name of its successor with that of the last scanned L-type/S-type substring.
-If both are equal, then the two substrings have the same name; otherwise, the former is assigned a name bigger/smaller than the latter.
-
-From our experiments, sorting substrings by the priority queues takes long CPU time and large I/O volume.
-To reduce the consumption, we present a new method that combines the sorted S*-type substrings of all the blocks by a merger, which maintains a buffer for each block to preload the sorted S*-type substrings into RAM and an internal-memory heap to compare the substrings belonging to different blocks by literal comparisons.
-For the substrings too long to be wholly loaded into RAM, this method first uses the priority queue to sort them in external memory and then employs the merger to compare them with the short substrings based on the following property: suppose $sub_a$ and $sub_b$ are respectively long and short substrings, then $sub_a < sub_b$ if (1) $sub_a[0, D) = sub_b$ or (2) $sub_a[0, k) = sub_b[0, k)$ and $sub_af_a[k] < sub_b[k]$, where $D = \|sub_b\|$ and $k < D$.
-This indicates that we only need to fetch the first $D$ characters of a long substring into RAM when comparing it with a short substring containing $D$ characters.
+Next, we describe how to name S*-type substrings at the same time when they are being sorted.
+Given two S*-type substrings successively popped from the minimal heap, we make a literal comparison of them to check their equality if they are both short; otherwise, we use the naming method of SAIS-PQ to determine their equality in S\ref{sort_and_name_method:2}.
+Specifically,
+substrings are sorted by comparing their heading characters and the names of their sorted successors in S\ref{sort_and_name_method:2}, where two substrings are of the same name if and only if their heading characters and the names of their successors are both equal.
+By recording the names of long S*-type substrings in their sorted order, it takes constant time to check equality of two lexicographically neighboring long substrings in S\ref{sort_and_name_method:3}.
 
@@
 
-As demonstrated in Section~\ref{sec:experiments}, our implementation for the enhanced DSA-IS has an advantage over that for eSAIS in terms of time, space and I/O volume, but its peak disk use remains at a high level.
-This is mainly because it fails to gradually clean up the temporary data when they are being used.
-On the contrary, fSAIS stores the temporary data using multiple files and deletes each file immediately when all that residing on it is longer to use.
-Therefore, it is reasonable to believe that our program can achieve a substantial improvement following the same technique.
-Currently, we are attempting to further improve the algorithmic and implementation designs of DSA-IS.
+Section~\ref{sec:experiments} shows that, by using the new substring sorting and naming methods, our program for the enhanced  DSA-IS has an advantage over that for eSAIS in terms of space and I/O volume, but its average peak disk use is still higher than that for fSAIS by around $5n$.
+This is mainly due to the use of external-memory containers (vectors, sorters and priority queues) provided by the STXXL library~\cite{Dementiev2007}.
+More specifically, the STXXL's containers fail to free the disk space for saving temporary data immediately after they are no longer to use.
+This implementation issue can be solved by storing temporary data into multiple files and deleting each file when it is obsolete.
+In this way, our program can be further improved to achieve a space performance similar to that for fSAIS.
 
-\section{Checkers}~\label{sec:check_sa}
+\section{Checkers}~\label{sec:checkers}
 
-\subsection{Details}
-
-A constructed SA should be verified to ensure its correctness before use.
-This checking function can be found in some widespread software packages like SA-IS, DC3 and eSAIS, where the latter two perform verification based on Theorem~\ref{theorem:1}.
+\subsection{Prior Art} \label{sec:checkers:prior_art}
+The widespread software packages for DC3 and eSAIS provide users a checker to perform verification. Theorem~\ref{theorem:1} gives the main idea behind this checker.
 
@@
 
-This method executes two passes of integer sorts and each sort involves $n$ fixed-size tuples, thus it takes large disk space and I/O volume when implemented in the external memory, resulting in a non-negligible checking overhead.
-Take eSAIS for example, the peak disk use for sequentially building and checking SA rises up to $27n$, while the I/O volume also increases by $50n$.
+This checker performs two passes of integer sorts and each sort involves $n$ fixed-size tuples.
+When implemented in external memory, it requires large disk space and consumes high I/O volume, resulting in a non-negligible checking overhead.
+As will be seen in Section~\ref{sec:experiments}, the peak disk use for checking an SA encoded by 40-bit integers is around $21n$, while the I/O volume is no less than $50n$.
+In this section, we describe two lightweight SA checking methods based on the IS principle.
+Both of them can be seamlessly integrated into any IS suffix sorting algorithm to build and check SA simultaneously.
 
-We propose a fast and lightweight SA checking method that can be seamlessly integrated into any IS suffix sorting algorithms for building and checking an SA simultaneously.
-Before presentation, we introduce two lemmas that will be used to prove the correctness of our checking method.
-Both two Lemmas demonstrate the fact that the lexical order of two suffixes can be determined by their heading characters and successors.
-Particularly, Lemma~\ref{lemma:2} is applied to any IS suffix sorting algorithm for inducing unsorted suffixes from sorted ones during the induction phase.
+\subsection{Method A} \label{sec:checkers:method_a}
+
+\subsubsection{Preliminaries} \label{sec:checkers:method_a:preliminaries}
+
+{\em LCP array.} The LCP array for $sa$, denoted by $lcp$, satisfies that $lcp[0] = 0$ and $lcp[i] = \ell$ for $i \in [1, n)$, where $\ell$ is the length of the longest common prefix of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[i - 1])$.
+
+{\em Suffix and LCP buckets.} All the suffixes in $sa$ are naturally partitioned into multiple buckets.
+Each bucket contains suffixes starting with a same character and occupies a contiguous interval of $sa$.
+For example, ${\sf sa\_bkt}(c_0)$ gathers all the suffixes starting with $c_0$.
+This bucket can be further partitioned into two parts, where the left and right parts only contain the L-type and S-type suffixes, respectively.
+For short, we denote the left and right parts by ${\sf sa\_bkt_L}(c_0)$ and ${\sf sa\_bkt_S}(c_0)$, respectively.
+Accordingly, $lcp$ can be also partitioned into multiple buckets and ${\sf lcp\_bkt}(c_0)$ records the LCP-values of suffixes in ${\sf sa\_bkt}(c_0)$.
+Similarly, we define ${\sf lcp\_bkt_L}(c_0)$ and ${\sf lcp\_bkt_S}(c_0)$ for ${\sf sa\_bkt_L}(c_0)$ and ${\sf sa\_bkt_S}(c_0)$, respectively.
+
+{\em Suffix and LCP arrays for S*-type suffixes.} Let $n_1 = \|x_1\|$, we use $sa^*[0, n_1)$ and $lcp^*[0, n_1)$ to denote the suffix and LCP arrays for S*-type suffixes, respectively.
+The $i$-th element in $sa^*$ records the starting position of the $(i + 1)$-th smallest S*-type substring, while the $i$-th element in $lcp^*[i]$ records the LCP-value of ${\sf suf}(sa^*[i])$ and ${\sf suf}(sa^*[i - 1])$.
+
+\subsubsection{Idea} \label{sec:checkers:method_a:idea}
+
+During the induction phase, SA-IS first computes $sa^*$ by sorting the starting positions of all the S*-type suffixes with their ranks indicated by $sa_1$ and then induces $sa$ from $sa^*$ by S\ref{induction_phase:1}'-S\ref{induction_phase:3}'.
+
+\begin{enumerate}[S1']
+
+	\item Scan $sa^*$ rightward with $i$ decreasing from $n_1 - 1$ to $0$.
+For each scanned item $sa^*[i]$, insert it into the rightmost empty position of ${\sf sa\_bkt_S}(x[sa^*[i]])$.
+\label{induction_phase:1}
+
+	\item Scan $sa$ rightward with $i$ increasing from $0$ to $n - 1$.
+For each scanned non-empty item $sa[i]$, insert $sa[i] - 1$ into the leftmost empty position of ${\sf sa\_bkt_L}(x[sa[i] - 1])$ if $t[sa[i] - 1] = 0$.\label{induction_phase:2}
+
+	\item Scan $sa$ leftward with $i$ decreasing from $n - 1$ to $0$.
+For each scanned non-empty item $sa[i]$, insert $sa[i] - 1$ into the rightmost empty position of ${\sf sa\_bkt_S}(x[sa[i] - 1])$ if $t[sa[i] - 1] = 1$.\label{induction_phase:3}
+
+\end{enumerate}
+
+An example of the above inducing process is given in Fig.~\ref{fig:example1}.
+In line 6, we find the end of each bucket and insert the sorted S*-type suffixes into the corresponding buckets.
+Then, we find the head of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes.
+In lines 8-9, we check $\sf suf(14)$ (marked by the symbol $@$) and find its predecessor ${\sf suf(13)}$ is L-type, thus we put ${\sf suf(13)}$ into the current leftmost empty position in ${\sf sa\_bkt_L}(i)$.
+To step through $sa$ in this way, we get all the L-type suffixes sorted in line 17.
+After that, we find the end of each bucket and scan $sa$ leftward for inducing the order of S-type suffixes in lines 18-29.
+When scanning $sa[14]$, we see $x[3]$ is S-type and thus put ${\sf suf(3)}$ into the current rightmost empty position in ${\sf sa\_bkt_S}(i)$.
+Following the same, we get all the S-type suffixes sorted in line 29.
+
+\begin{figure}[t]
+	\centering
+
+	\includegraphics[width = 1\columnwidth]{example.pdf}
+
+	\caption{An example for inducing $sa$ from $sa^*$.}
+
+	\label{fig:example1}
+
+\end{figure}
+
+As stated below, S\ref{induction_phase:1}'-S\ref{induction_phase:3}' also constitute the sufficient and necessary conditions for a correct $sa$.
+
+\begin{lemma} \label{lemma:3}
+	For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied: \\ 
+	(1) $sa^*$ is correct. \\
+	(2) $sa[i]$ is equal to the value calculated by SS\ref{induction_phase:1}-SS\ref{induction_phase:3}. \\
+
+\end{lemma}
+
+Suppose $sa^*$ is correct\footnote{The correctness of $sa^*$ can be checked by a sparse SA checker, such as~\cite{wu2017}.
+The checking processes for $sa^*$ and $sa$ can be executed in parallel.}, this lemma suggests a method to check $sa$ when it is induced from $sa^*$ during the induction phase.
+Specifically, in S\ref{induction_phase:2}' and S\ref{induction_phase:3}', each item induced into $sa$ will be scanned later to induce the order of its predecessor.
+If the induced and scanned values for each item are equal, then $sa$ is correct according to Lemma~\ref{lemma:3}.
+The problem to be solved here is that when an item is induced into a bucket, its value in $sa$ will not be scanned at once.
+Our solution for this is to integrate the building and checking processes into a whole, by ensuring the sequence of items induced into a bucket is identical to that scanned later.
+For the purpose, we increasingly compute the fingerprints of both sequences and check their equality at the end of the induction phase.
+If the two fingerprints for each bucket are equal, then the second condition of Lemma~\ref{lemma:3} will be seen with a high probability.
+As a result, $sa$ can be built and probabilistically checked at the same time, for this we have Corollary~\ref{corollary:1}.
+
+\begin{corollary} \label{corollary:1}
+	$sa[0, n)$ is the SA for the input string $x[0, n)$ with a high probability if the following conditions are satisfied for all $c \in \Sigma$: \\
+	(1) $sa^*$ is correct.\\
+	(2) The fingerprints of ${\sf sa_{I1}}(c)$ and ${\sf sa_{S1}}(c)$ are equal. \\
+	(3) The fingerprints of ${\sf sa_{I2}}(c)$ and ${\sf sa_{S2}}(c)$ are equal. \\
+
+\end{corollary}
+
+Notice that ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$.
+
+\subsection{Method B}\label{sec:checkers:method_b}
+
+We introduce Lemma~\ref{lemma:1} before the presentation of Method B.
+This statement has been applied to deriving Theorem~\ref{theorem:1} for comparing two suffixes given the order of their successors already known.
 
@@
 
-\begin{lemma} \label{lemma:2}
-	For $i, j \in [0, n)$ and $i \ne j$, $r_i < r_j \Leftrightarrow (x[i], r_{i + 1}) < (x[j], r_{j + 1})$, where $r_i$ and $r_j$ indicates the relative order of ${\sf suf}(i)$ and ${\sf suf}(j)$ among all the suffixes, respectively.
-\end{lemma}
-
-In what follows, we use the above lemmas to prove Theorem~\ref{theorem:2}, where $sa^*_1$ and $sa^*_2$ are two integer arrays that record the starting positions of the sorted S*-type suffixes.
-To be specific, the former is computed from the SA for $x_1$ and the latter is retrieved from the final SA for $x$.
+Below we describe the idea of Method B and prove its correctness using Lemma~\ref{lemma:1} and Theorem~\ref{theorem:1}, where $\overline{sa^*}$ is retrieved from $sa$ and it records the starting positions of all the S*-type suffixes in their sorted order.
 
 \begin{theorem} \label{theorem:2}
-	For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for the input string $x[0,n)$ if and only if the following conditions are satisfied: \\
-	(1) $sa^*_1[i] \ne sa^*_1[j]$ for $i, j \in [0, n)$ and $i \ne j$. \\
-	(2) $sa^*_1 = sa^*_2$. \\
+	For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied for all $i \in [0, n_1)$: \\
+	(1) $sa^*[i]$ is the starting position of an S*-type suffix and it differs from any other items in $sa^*$. \\
+	(2) $sa^*[i] = \overline{sa^*}[i]$. \\
 \end{theorem}
@@
 	Suppose $sa$ is not a permutation of $[0, n)$.
-Without loss of generality, let $i \ne j$ and $sa[i] = sa[j]$, then $u \ne v$ and $sa[u] = sa[v]$ according Lemma~\ref{lemma:2}, where ${\sf suf}(sa[u])$ and ${\sf suf}(sa[v])$ are the successors of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[j])$, respectively.
-By recursively replacing $(i, j)$ with $(u, v)$, we can find two elements $sa[i_0]$ and $sa[j_0]$ storing the starting position of the same S*-type suffix.
-This indicates that $sa^*_1$ contains duplicate elements, which violates condition (1) in Theorem~\ref{theorem:2}.
+Without loss of generality, let $i \ne j$ and $sa[i] = sa[j]$, then $u \ne v$ and $sa[u] = sa[v]$ according Theorem~\ref{theorem:1}, where ${\sf suf}(sa[u])$ and ${\sf suf}(sa[v])$ are the successors of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[j])$, respectively.
+By recursively replacing $(i, j)$ with $(u, v)$, we can find two items $sa[i_0]$ and $sa[j_0]$ storing the starting position of the same S*-type suffix, indicating that $sa^*$ contains duplicate items.
+This violates condition (1) of Theorem~\ref{theorem:2}.
 
 	Suppose $i < j$ and ${\sf suf}(sa[i]) > {\sf suf}(sa[j])$.
-If $x[sa[i]] < x[sa[j]]$, then ${\sf suf}(sa[i]) < {\sf suf}(sa[j])$ according to Lemma~\ref{lemma:1}; otherwise, if $x[sa[i]] > x[sa[j]]$, then $i > j$ according to Lemma~\ref{lemma:2}.
-Both lead to a contradiction.
-Given that $x[sa[i]] = x[sa[j]]$, then $u < v$ and ${\sf suf}(u) > {\sf suf}(v)$, where ${\sf suf}(sa[u])$ and ${\sf suf}(sa[v])$ are the successors of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[j])$, respectively.
+If $x[sa[i]] < x[sa[j]]$, then ${\sf suf}(sa[i]) < {\sf suf}(sa[j])$ according to Lemma~\ref{lemma:1}; otherwise, if $x[sa[i]] > x[sa[j]]$, then $i > j$ according to Theorem~\ref{theorem:1}.
+Both two conditions lead to a contradiction.
+Therefore, we must have $x[sa[i]] = x[sa[j]]$.
+If so, then $u < v$ and ${\sf suf}(u) > {\sf suf}(v)$, where ${\sf suf}(sa[u])$ and ${\sf suf}(sa[v])$ are the successors of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[j])$, respectively.
 By recursively replacing $(i, j)$ with $(u, v)$, we can find two elements $sa[i_0]$ and $sa[j_0]$ such that at least one stores the starting position of an S*-type suffix.
@@
 In this case, $x[sa[i_0] + 1] > x[sa[j_0] + 1]$.
-Therefore, $u_0 > v_0$ according to Lemma~\ref{lemma:2}.
-This indicates that $(x[sa[i_0]], u_0) > (x[sa[j_0]], v_0)$ and thus $i_0 > j_0$ according to Lemma~\ref{lemma:2}.
+Therefore, $u_0 > v_0$ according to Theorem~\ref{theorem:1}.
+This indicates that $(x[sa[i_0]], u_0) > (x[sa[j_0]], v_0)$ and thus $i_0 > j_0$ according to Theorem~\ref{theorem:1}.
 This leads to a contradiction.
@@
 Let $(i_0, j_0)$ be the pair that maximize $sa[i_0] + sa[j_0]$ such that $i_0 < j_0$ and ${\sf suf}(sa[i_0]) > {\sf suf}(sa[j_0])$, then $u_0 > v_0$ and ${\sf suf}(sa[u_0]) > {\sf suf}(sa[v_0])$.
-\footnote{Notice that $(u_0, v_0)$ must exist because the sentinel is a size-one S*-type substring smaller than others.} Because $x[i_0, u_0) = x[j_0, v_0)$, there must be $i_0 > j_0$ according to Lemma~\ref{lemma:2}.
-This contradiction indicates that $sa^*_1$ and $sa^*_2$ differ from each other, which violates condition (2) in Theorem~\ref{theorem:2}.
+\footnote{Notice that $(u_0, v_0)$ must exist because the sentinel is a size-one S*-type substring smaller than others.} Because $x[i_0, u_0) = x[j_0, v_0)$, there must be $i_0 > j_0$ according to Theorem~\ref{theorem:1}, thus $sa^*$ and $\overline{sa^*}$ differ from each other.
+This violates condition (2) of Theorem~\ref{theorem:2}.
 	\end{enumerate}
@@
 
-The first condition in Theorem~\ref{theorem:2} can be easily checked when computing $sa^*_1$ from $sa_1$.
-For the second condition, a naive method is to first make a copy of $sa^*_1$ and $sa^*_2$ when inducing $sa$ and then sequentially compare their elements after the induction phase.
-This method takes linear time, space and I/O overhead.
-A more efficient alternative is to compute and compare the fingerprints for these two arrays, denoted by ${\sf fp}(sa^*_1)$ and ${\sf fp}(sa^*_2)$, instead to determine their equality.
-Obviously, $sa^*_1$ and $sa^*_2$ share a common fingerprint if they are equal to each other, but the inverse is not always true.
-Fortunately, by using the fingerprinting function described in~\cite{Karp1987}, the probability of a false match can be reduced to a completely negligible level.
-The above discussion leads us to the following statement.
+The first condition of Theorem~\ref{theorem:2} is naturally satisfied when computing $sa^*$ from $sa_1$.
+A naive method for checking the second condition is to keep a copy of $\overline{sa^*}$ when computing $sa$ from $sa^*$ and compare it with $sa^*$ afterward.
+This takes linear time, space and I/O overhead.
+We prefer to check equality of them by comparing their fingerprints, which can be calculated during the scan of these arrays in S\ref{induction_phase:1}' and S\ref{induction_phase:3}'.
 
-\begin{corollary} \label{theorem:3}
-	For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for the input string $x[0, n)$ with a high probability if the following conditions are satisfied:
-	(1) $sa^*_1[i] \ne sa^*_1[j]$ for $i, j \in [0, n)$ and $i \ne j$. \\
-	(2) ${\sf fp}(sa^*_1) = {\sf fp}(sa^*_2)$. \\
+\begin{corollary} \label{corollary:2}
+	$sa[0, n)$ is the SA for the input string $x[0, n)$ with a high probability if the following conditions are satisfied for all $i \in [0, n_1)$: \\
+	(1) $sa^*[i]$ is the starting position of an S*-type suffix and it differs from any other items in $sa^*$. \\
+	(2) The fingerprints of $sa^*$ and $\overline{sa^*}$ are equal. \\
 \end{corollary}
 
-\subsection{Example} \label{sec:check_sa:example}
+In Fig.~\ref{fig:example1}, both $sa^*$ and $\overline{sa^*}$ contain 4 items $14, 8, 5, 2$ arranged in the same order.
+By using the fingerprinting function introduced in the next subsection, the fingerprints can be calculated in linear time using constant space.
 
-xxxx
+\subsection{Discussion}
+
+\begin{figure}[htbp!]
+	\centering
+
+	\includegraphics[width = 0.9\columnwidth]{example2.pdf}
+
+	\caption{An example for calculating fingerprints by Karp-Rabin fingerprinting function.}
+
+	\label{fig:example2}
+
+\end{figure}
+
+Methods A and B check equality of two integer arrays by comparing their fingerprints.
+In our programs, we choose to use the Karp-Rabin fingerprinting function to compute the fingerprints in need.
+As depicted in Fig.~\ref{fig:example2}, the fingerprint of $A$ is calculated according to Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$.
+It should be noticed that two equal arrays must have an identical fingerprint, but the inverse is not always true.
+Fortunately, the probability of a false match can be reduced to a negligible level if $L$ is set to a large value.
+
+\begin{formula} \label{formula:1}${\sf fp}(A[0, -1]) = 0$. 
+\end{formula}
+
+\begin{formula} \label{formula:2}${\sf fp}(A[0, i]) = {\sf fp}(A[0, i - 1]) \cdot {\delta} + A[i]\mod L$ for $i \ge 0$.
+\end{formula}
+
+We also point out that Method A can check $lcp$ when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}.
+This is done by comparing the sequence of items induced into an $lcp$ bucket with that scanned from the same bucket.
+If the two sequences of each bucket are equal, then $lcp$ is correctly the LCP array for $sa$.
 
@@
 
-For performance evaluation, we engineer DSA-IS and its optimized DSA-IS+, by using the STXXL's containers~(sorter, priority queue, vector and stream) to conduct efficient I/O operations in the external memory.
-As reported in~\cite{Karkkainen2014}, SAscan outperforms eSAIS in practice when the internal memory capacity $M$ is adequate.
-However, the time and space complexities of the program for SAscan are $\widetilde{\mathcal{O}}(n^2/ M)$ and $\widetilde{\Omega}(n^2/ M)$, respectively, thus it will suffer from a performance degradation as the size of the input string gets larger.\footnote{The authors in~\cite{Karkkainen2015} says, "SAscan needs just 7.5n bytes of disk space but because of its $\widetilde{\mathcal{O}}(n^2/ M)$ time complexity, it is competitive with eSAIS only when the input is less than about five times the size of RAM.".} On the other hand, the programs for DSA-IS, DSA-IS+ and eSAIS\footnote{https://panthema.net/2012/1119-eSAIS-Inducing-Suffix-and-LCP-Arrays-in-External-Memory/} are not only of linear I/O volume and space complexities in theory, but also commonly implemented by the STXXL library.
-\textbf{Therefore, we use eSAIS rather than SAscan as a baseline for evaluating the performance of DSA-IS and DSA-IS+.} The experimental platform is a desktop computer equipped with an Intel Xeon E3-1220 V2 CPU, 4GiB RAM and 500GiB HD.
-The programs are complied by gcc/g++ 4.8.4 with -O3 options under Ubuntu 14.04 64-bit operating system.
-
-In our experiments, three performance metrics are investigated for the programs running on the corpora listed in Table~\ref{tbl:corpora}, where each metric is measured as a mean of two runs of the programs.
+For implementation simplicity, we engineer DSA-IS and DSA-IS+ by the STXXL's containers~(vector, sorter, priority queue and stream).
+The experimental platform is a desktop computer equipped with an Intel Xeon E3-1220 V2 CPU, 4GiB RAM and 500GiB HD.
+All the programs are complied by gcc/g++ 4.8.4 with -O3 options under Ubuntu 14.04 64-bit operating system.
+In our experiments, three performance metrics are investigated for the programs running on the corpora listed in Table~\ref{tbl:corpora}, where each metric is measured as a mean of two runs.
 
@@
 	\item construction time~(CT): the running time, in units of microseconds per character.
-	\item peak disk usage~(PDU): the maximum disk space requirement, in units of bytes per character.
+	\item peak disk use~(PDU): the maximum disk space requirement, in units of bytes per character.
 	\item I/O volume~(IOV): as the term suggests, in units of bytes per character.
@@
 \renewcommand\arraystretch{1.3}
-\begin{table*}[!t]
+\begin{table}[!h]
 	\caption{Corpus, $n$ in Gi, 1 byte per character} 
@@
 	\centering
-	\begin{tabular}{|l|c|c|p{10cm}|}
+	\begin{tabular}{|l|c|c|p{5cm}|}
 		\hline
@@
 		uniprot & 2.5 & 96 & UniProt Knowledgebase release 4.0, at \url{ftp://ftp.expasy.org/databases/.../complete}, dated as 16/05/11. \\\hline
-		genome & 2.9 & 6 & Human genome data, used in Dementiev et al.~\cite{Dementiev2008a}, at \url{http://algo2.iti.kit.edu/dementiev/esuffix/instsances.shtml.} \\\hline
 	\end{tabular}
-\end{table*}
+\end{table}
 
-\subsection{Performance Evaluation on Construction Algorithms}
-Figure~\ref{fig:performance_analysis1} illustrates the performance of the programs for DSA-IS and eSAIS in terms of the investigated three metrics, where guten\_14G and enwiki\_14G are prefixes of the corresponding corpus.
-As depicted, the peak disk usage for DSA-IS is around $16n$ on the two datasets, which is two-thirds of that for eSAIS.
-However, eSAIS outperforms DSA-IS with respect to the construction time, where the speed gap between them is mainly due to the different I/O volumes taken by the two algorithms.
-In order for a deep insight, we show in Table~\ref{tbl:volume_cmp} the statistics for I/O volumes of the algorithms in their reduction and induction phases.
-It can be observed that, although the I/O volumes spent in the induction phase are similar, eSAIS is far more I/O-efficient than DSA-IS when sorting substrings during the reduction phase.
-More specifically, the ratio of reduction and induction volumes for DSA-IS is about 0.80 while the counterpart for eSAIS is 0.23.
+\subsection{Building Performance}
 
-\begin{figure}[htbp!]
+\begin{figure*}[t]
 	\centering
-	\subfigure{
-		\label{subfig:pdu_cmp1}
-		\includegraphics[width = 0.9\columnwidth]{pdu_cmp1}
+	\subfigure[enwiki]{
+		\begin{minipage}[b]{0.45\textwidth}
+			\includegraphics[width=1\textwidth]{construction_pdu_enwiki} \\
+			\includegraphics[width=1\textwidth]{construction_iov_enwiki} \\
+			\includegraphics[width=1\textwidth]{construction_ct_enwiki}
+		\end{minipage}
 	}
-	\hfil
-	\subfigure{
-		\label{subfig:iov_cmp1}
-		\includegraphics[width = 0.9\columnwidth]{iov_cmp1}
+	\subfigure[guten]{
+		\begin{minipage}[b]{0.45\textwidth}
+			\includegraphics[width=1\textwidth]{construction_pdu_guten} \\
+			\includegraphics[width=1\textwidth]{construction_iov_guten} \\
+			\includegraphics[width=1\textwidth]{construction_ct_guten}
+		\end{minipage}
 	}
-	\hfil
-	\subfigure{
-		\label{subfig:ct_cmp1}
-		\includegraphics[width = 0.9\columnwidth]{ct_cmp1}
+	\caption{A comparison of DSA-IS, DSA-IS+ and eSAIS on guten and enwiki in terms of peak disk usage, I/O volume and construction time, where $D = 4$ and the input size varies in \{1, 2, 4, 8, 16\} GiB.
 }
-	\caption{Experimental results for DSA-IS and eSAIS on guten\_14G and enwiki\_14G in terms of peak disk usage, I/O volume and construction time.}
-	\label{fig:performance_analysis1}
-\end{figure}
+	\label{fig:construction_performance1}
+\end{figure*}
 
@@
 \begin{table*}%
-	\caption{Comparison of Reduction and Induction I/O Volumes Amongst DSA-IS, DSA-IS+ and eSAIS on enwiki}
+	\caption{A Comparison of Reduction and Induction I/O Volumes Amongst DSA-IS, DSA-IS+ and eSAIS on enwiki}
 	\label{tbl:volume_cmp}
@@
 		\hline
-		\multicolumn{1}{|c}{} & \multicolumn{4}{|c|}{eSAIS} & \multicolumn{4}{c|}{DSA-IS} & \multicolumn{4}{c|}{DSA-IS+ ($D_1 = 8$, $D_2 = 10$)}\\\hline
+		\multicolumn{1}{|c}{} & \multicolumn{4}{|c|}{eSAIS} & \multicolumn{4}{c|}{DSA-IS} & \multicolumn{4}{c|}{DSA-IS+ ($D = 4$)}\\\hline
 		\hline
 		Size & Red. & Ind. & Total & Ratio & Red. & Ind. & Total & Ratio & Red. & Ind. & Total & Ratio\\\hline
-		1G & 36.6 & 132.8 & 169.4 & 0.27 & 81.3 & 105.9 & 187.2 & 0.76 & 39.1 & 104.3 & 143.4 & 0.37\\\hline
-		2G & 36.0 & 141.9 & 177.9 & 0.25 & 83.5 & 108.2 & 191.7 & 0.77 & 37.8 & 106.4 & 144.2 & 0.35\\\hline
-		4G & 35.6 & 152.1 & 187.7 & 0.23 & 94.3 & 118.9 & 213.2 & 0.79 & 42.4 & 117.0 & 159.4 & 0.36\\\hline
-		8G & 35.2 & 165.7 & 200.9 & 0.21 & 107.8 & 132.3 & 240.1 & 0.81 & 43.7 & 129.9 & 173.6 & 0.33\\\hline
-		14G & 35.0 & 172.1 & 207.1 & 0.20 & 121.9 & 146.6 & 268.5 & 0.83 & 43.8 & 142.3 & 186.1 & 0.30\\\hline
+		1G & 36.6 & 132.8 & 169.4 & 0.27 & 81.3 & 109.6 & 190.9 & 0.74 & 45.4 & 91.7 & 137.1 & 0.33\\\hline
+		2G & 36.0 & 141.9 & 177.9 & 0.25 & 83.5 & 111.6 & 195.1 & 0.75 & 47.2 & 93.4 & 140.6 & 0.34\\\hline
+		4G & 35.6 & 152.1 & 187.7 & 0.23 & 94.3 & 144.1 & 238.4 & 0.65 & 54.1 & 111.5  & 165.6 & 0.33\\\hline
+		8G & 35.2 & 165.7 & 200.9 & 0.21 & 107.8 & 159.6 & 267.4 & 0.68 & 60.1 & 122.1 & 182.2 & 0.33\\\hline
+		16G & 35.0 & 172.1 & 207.1 & 0.20 & 121.9 & 166.1 & 288.0 & 0.73 & 62.7 & 128.7 & 191.4 & 0.33\\\hline
 	\end{tabular}
@@
 
-The above phenomenon can be explained as follows.
-During the reduction phase, DSA-IS partitions the input string $x$ into multiple blocks and induced sorts the substrings of each block in the internal memory.
-Then it reuses the induced sorting method to sort all the substrings of $x$ in the external memory by merging the block-wise results residing on disks.
-In contrast with DSA-IS, eSAIS splits $x$ into fixed-size partitions and sorts the substrings of each partition in the internal memory by calling std::sort.
-Then, it generates the global result by merging the partial ones with an internal memory heap.
-This leads to a higher I/O efficiency against DSA-IS.
+Because fSAIS is not available online, we use eSAIS as a baseline for analyzing the performance of DSA-IS and DSA-IS+.
+Fig.~\ref{fig:construction_performance1} shows a comparison between the programs for these three algorithms in terms of the investigated metrics.
+As depicted, the program for DSA-IS requires less disk space than that for eSAIS when running on "enwiki" and "guten".
+In details, the peak disk use of DSA-IS and eSAIS are around $18n$ and $24n$, respectively.
+However, eSAIS runs much faster than DSA-IS due to the different I/O volumes.
+In order for a deep insight, we collect in Table~\ref{tbl:volume_cmp} the statistics of their I/O volumes in the reduction and induction phases.
+As can be seen, although DSA-IS and eSAIS have similar performances when sorting suffixes in the induction phase, the latter consumes much less I/O volume than the former when sorting substrings in the reduction phase.
+More specifically, the mean ratio of induction I/O volume to reduction I/O volume are $0.23$ and $0.71$ for them, respectively.
+We can also see from the same figure that DSA-IS+ achieves a substantial improvement against DSA-IS, it runs as fast as eSAIS and takes half as much disk space as the latter.
+This is because the reduction I/O volume for DSA-IS+ is only half as much as that for DSA-IS (Table~\ref{tbl:volume_cmp}).
+Notice that the new substring sorting and naming methods adopted by DSA-IS+ take effect when most of the S*-type substrings are short.
+From our experiments, given $D = 8$, the ratio of long S*-type substrings in the investigated corpus nearly approaches one hundred percent.
+Therefore,the proposed substring sorting and naming methods are practical for real-world datasets.
 
-We attempt to reduce the I/O overhead for DSA-IS by using the substring sorting and naming methods described in Section~\ref{sec:dsais_plus}.
-Our program for the enhanced algorithm DSA-IS+ introduces two parameters $D_1$ and $D_2$ to respectively specify the threshold value $D$ for classifying the substrings in $x$ and the reduced strings $\{x1, x2, ...\}$ according to the instructions in Section~\ref{subsec:dsais_plus_method_b}.
-As observed from Table~\ref{tbl:volume_cmp}, the total I/O volumes for DSA-IS+ and eSAIS are similar and the ratio of the reduction and induction I/O volumes for either of them is no more than half of that for DSA-IS.
-This substantial improvement leads to a great speedup against DSA-IS.
-Figure~\ref{fig:performance_analysis2} shows the performance trends of DSA-IS+ and eSAIS on guten and enwiki as the corpora size increases from $1G$ to $14G$, which indicates that the speed of DSA-IS+ is similar to that of eSAIS and its peak disk usage remains at a low level as the same as that of DSA-IS.
+\subsection{Checking Performance}
 
-\begin{figure*}[htbp]
+For evaluation, we integrate Method B into DSA-IS+ to constitute "Solution A" and compare it with "Solution B" composed of eSAIS and the existing checking method in~\cite{Dementiev2008a}.
+Fig.~\ref{fig:verification_performance} gives a glimpse of the performance of two solutions on various corpora.
+It can be observed that, the time, space and I/O volume for verification by Method B is negligible in comparison with that for construction by DSA-IS+, while the overhead for checking SA in Solution B is relatively large.
+Table~\ref{tbl:breakdown_solutionb} shows the performance breakdown of Solution B, where the checking time is one-fifth as the running time of the plain eSAIS and the peak disk use for verification is also a bit larger than that for construction.
+As a result, the combination of DSA-IS+ and Method B can build and check an SA in better total time and space.
+
+\begin{figure}[t]
 	\centering
-	\subfigure[guten]{
-		\begin{minipage}[b]{0.45\textwidth}
-			\includegraphics[width=1\textwidth]{pdu_cmp2_guten} \\
-			\includegraphics[width=1\textwidth]{iov_cmp2_guten} \\
-			\includegraphics[width=1\textwidth]{ct_cmp2_guten}
-		\end{minipage}
+	\subfigure{
+		\label{subfig:verification_pdu}
+		\includegraphics[width = 0.9\columnwidth]{verification_pdu}
 	}
-	\subfigure[enwiki]{
-		\begin{minipage}[b]{0.45\textwidth}
-			\includegraphics[width=1\textwidth]{pdu_cmp2_enwiki} \\
-			\includegraphics[width=1\textwidth]{iov_cmp2_enwiki} \\
-			\includegraphics[width=1\textwidth]{ct_cmp2_enwiki}
-		\end{minipage}
+	\hfil
+	\subfigure{
+		\label{subfig:verification_iov}
+		\includegraphics[width = 0.9\columnwidth]{verification_iov}
 	}
-	\caption{Experimental results for DSA-IS, DSA-IS+ and eSAIS on guten and enwiki in terms of peak disk usage, I/O volume and construction time, where $D_1 = 8$, $D_2 = 10$ and the input size varies in \{1, 2, 4, 8, 14\} GiB.
+	\hfil
+	\subfigure{
+		\label{subfig:verification_ct}
+		\includegraphics[width = 0.9\columnwidth]{verification_ct}
 	}
-	\label{fig:performance_analysis2}
-\end{figure*}
+	\caption{A comparison of DSA-IS and eSAIS on various corpora in terms of peak disk usage, I/O volume and construction time.}
+	\label{fig:verification_performance}
+\end{figure}
 
-%As with other external memory suffix sorting algorithms, DSA-IS+ is I/O-intensive and its construction speed is presumed to be linearly proportional to the I/O volume.
-However, our experiments indicate that the ratio of the reduction and induction time is much higher than that of the reduction and induction volume.
-This phenomenon is possibly due to the performance of the STXXL library.
-For implementation simplicity, we organize the item of DSAITEM arrays by using the sorter and priority queue containers provided by STXXL in our programs.
-This
+% Table
+\begin{table}%
+	\caption{Performance Breakdown of Solution B on various Corpora}
+	\label{tbl:breakdown_solutionb}
+	\centering
+	\begin{tabular}{|c|c|c|c|c|c|c|}
+		\hline
+		\multirow{2}{*}{Corpus} & \multicolumn{3}{|c}{checking} & \multicolumn{3}{|c|}{building} \\\cline{2-7}
+		& PDU & IOV & CT & PDU & IOV & CT \\\hline
+		enwiki\_16G & 26.0 & 53.0 & 0.71 & 23.5 & 205.6 & 3.49 \\\hline
+		guten\_16G & 26.0 & 53.0 & 0.79 & 23.4 & 195.2 & 3.20 \\\hline
+		uniprot & 25.9 & 53.0 & 0.74 & 22.7 & 162.0 & 2.50 \\\hline
+		proteins & 25.9 & 53.0 & 0.58 & 24.1 & 172.3 & 2.33 \\\hline
+	\end{tabular}
+\end{table}%
+
+According to Corollary~\ref{corollary:2}, Method A must check both $sa^*$ and $sa$ to accomplish verification.
+Similar to Method B, the overhead for checking $sa$ is mainly caused by fingerprint calculations and thus can be neglected.
+On the other hand, we can apply the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ within sorting complexity, where the time and space in need is proportional to the number of S*-type characters in $x$.
+Because the ratio of S*-type characters to all in $x$ is commonly one-third in real-world datasets, the checking process for $sa^*$ will not become the bottleneck for Method A.
+We also point out that this checking process can be parallelized with that for $sa$ to achieve a higher speed.
+
+\section{Conclusion} \label{sec:conclusion}
+
+In this paper, we made an attempt to improve the space performance of DSA-IS by new substring sorting and naming methods.
+For implementation convenience, we currently employ the STXXL's containers to perform reading, writing and sorting on the disk.
+The experimental results shows that our program for the adapted algorithm DSA-IS+ runs as fast as eSAIS and requires only half disk space as that for the latter on various real-world datasets.
+This program can be further optimized to approach the optimal space performance by means of the external-memory vector, sorter and priority queues supporting fine-grained disk space allocation and deallocation.
+
+We also proposed two methods that enable any IS suffix sorting algorithm to build and check SA simultaneously.
+The second method is rather lightweight in terms of that its time and space complexities are negligible compared with that of the existing IS suffix sorting/checking algorithms.
+We will describe in another paper a disk-based suffix sorting algorithm taking only $1n$ work space.
+By augmenting with this new suffix array builder, the proposed checking methods may potentially constitute the best solution for the situations where checking is a must after building.
+
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,bibfile}
+
+\end{document}
 
@@
 
-\subsection{Performance Evaluation on Checking Methods}
+\subsection{Checking Performance}
 
@@
 This is a normal phenomenon due to the fluctuating I/O performance, which in turn indicates that the time overhead consumed by CM1 is also negligible.
-
-\section{Conclusion} \label{sec:conclusion}
-
-For better performance, we redesign the reduction phase of DSA-IS by employing two methods for sorting and naming substrings.
-The program of the enhanced algorithm DSA-IS+ is engineered by the STXXL library to achieve a high I/O efficiency.
-Our experiments indicate that DSA-IS+ requires only around $16/24=0.67$ peak disk usage as that for eSAIS and runs as fast as the latter on various real datasets.
-
-We describe an SA checker of linear time and space complexities that can be used to verify a suffix array during the time when it is being built.
-In combination with DSA-IS+, our checking method almost has no influence on the overall performance and is more efficient than the existing work in terms of time, space and I/O efficiencies.
-
-\bibliographystyle{IEEEtran}
-\bibliography{IEEEabrv,bibfile}
-
-\end{document}

### 2017-06-06 dsaism_plus_checker/src/build.cpp

@@ -1,0 +1,68 @@
+//#include "dsais.h"
+
+#include "common.h"
+#include "dsais.h"
+
+#include <iostream>
+#include <chrono>
+#include <fstream>
+
+
+int main(int argc, char** argv){
+
+	// mount stxxl disk	
+	stxxl::config *cfg = stxxl::config::get_instance();
+
+	stxxl::disk_config disk("stxxl.tmp", 1024 * 1024 * 1024, "syscall autogrow unlink");
+
+	disk.direct = stxxl::disk_config::DIRECT_ON;
+
+	cfg->add_disk(disk);
+	
+	// statistics collection
+	stxxl::stats *Stats = stxxl::stats::get_instance();
+
+	stxxl::stats_data stats_begin(*Stats);
+
+	stxxl::block_manager *bm = stxxl::block_manager::get_instance();
+
+	// check if input params are legal
+	if (argc != 3) {
+
+		std::cerr << "two param required: input_path and output_path.\n";
+
+		exit(-1);
+	}	
+
+	// retrieve file name for input string
+	std::string s_fname(argv[1]); 
+	
+	// retrieve file name for output SA
+	std::string sa_fname(argv[2]);
+	
+	// compute input string's size
+	std::fstream s_stream(s_fname, std::fstream::in);
+
+	s_stream.seekg(0, std::ios_base::end);
+
+	uint64 s_size = s_stream.tellg();
+
+	// report static information
+	std::cerr << "s fname: " << s_fname << "\nsa fname: " << sa_fname << "\ns size: " << s_size / 1024 / 1024 << " MB" << std::endl;
+
+	// compute the SA for the given string
+	DSAIS<uint8, uint40> dsa(s_fname, sa_fname);	
+
+	dsa.run();
+
+	// output report
+	std::cerr << (stxxl::stats_data(*Stats) - stats_begin);
+
+	uint64 pdu = bm->get_maximum_allocation(); 
+
+	std::cerr << "pdu\t total: " << pdu << "\t per char:" << (double)pdu / s_size << std::endl;
+
+	uint64 io_volume = Stats->get_written_volume() + Stats->get_read_volume(); 
+
+	std::cerr << "io\t total: " << io_volume << "\t per char: " << (double)io_volume / s_size << std::endl;
+}

### 2017-06-06 dsaism_plus_checker/src/bytestream.h

@@ -1,0 +1,61 @@
+#ifndef BYTESTREAM_H
+#define BYTESTREAM_H
+
+
+#include "namespace.h"
+#include "mycommon.h"
+
+WTL_BEG_NAMESPACE
+
+class ByteStream {
+public:
+	//alias
+	//type of allocation strategy
+
+	//type of block identifier
+	typedef stxxl::BID<BLOCK_SIZE> bid_type;
+
+	//type of block identifer vector
+	typedef std::vector<bid_type> bid_vector_type;
+
+	//type of iterator for block identifier vector
+	typedef typename bid_vector_type::iterator bid_iterator_type;
+
+	//type of uint8 block
+	typedef stxxl::typed_block<BLOCK_SIZE, uint8> block_type;
+
+	//type of uint8 buf_ostream
+	typedef stxxl::buf_ostream<block_type, bid_iterator_type> buf_ostream_type;
+
+	//type of uint8 buf_istream
+	typedef stxxl::buf_istream<block_type, bid_iterator_type> buf_istream_type;
+
+	//type of vector for uint8 buf_ostream
+	typedef std::vector<buf_ostream_type*> buf_ostream_vector_type;
+
+	//type of vector for uint8 buf_istream 
+	typedef std::vector<buf_istream_type*> buf_istream_vector_type;
+
+	//data members
+	//put data into the buf_ostream
+	template<typename dataT> static buf_ostream_type& putData(buf_ostream_type & _os, const dataT & _data) {
+		for (size_t i = 0; i < sizeof(dataT) / sizeof(typename buf_ostream_type::block_type::type); ++i) {
+			_os << (((typename buf_ostream_type::block_type::type*)&_data)[i]);
+		}
+		return _os;
+	}
+
+	//get data from the buf_istream
+	template<typename dataT> static buf_istream_type& getData(buf_istream_type & _is, const dataT & _data) {
+		for (size_t i = 0; i < sizeof(dataT) / sizeof(typename buf_istream_type::block_type::type); ++i) {
+			_is >> (((typename buf_istream_type::block_type::type*)&_data)[i]);
+		}
+		return _is;
+	}
+};
+
+
+WTL_END_NAMESPACE
+
+
+#endif

### 2017-06-06 dsaism_plus_checker/src/check.cpp

@@ -1,0 +1,37 @@
+#include "common.h"
+#include "checker.h"
+
+#include <iostream>
+#include <chrono>
+#include <fstream>
+
+int main(int argc, char** argv){
+
+	// mount stxxl disk	
+	stxxl::config *cfg = stxxl::config::get_instance();
+
+	stxxl::disk_config disk("stxxl.tmp", 1024 * 1024 * 1024, "syscall autogrow unlink");
+
+	disk.direct = stxxl::disk_config::DIRECT_ON;
+
+	cfg->add_disk(disk);
+	
+	// check if input params are legal
+	if (argc != 3) {
+
+		std::cerr << "two param required: input_path and output_path.\n";
+
+		exit(-1);
+	}	
+
+	// retrieve file name for input string
+	std::string s_fname(argv[1]); 
+	
+	// retrieve file name for output SA
+	std::string sa_fname(argv[2]);
+	
+	// compute the SA for the given string
+	Checker<uint8, uint40> checker(s_fname, sa_fname);	
+
+	std::cerr << (checker.run() ? "right" : "wrong")<< std::endl;
+}

### 2017-06-06 dsaism_plus_checker/src/checker.h

@@ -1,0 +1,180 @@
+#ifndef MY_CHECKER_H
+#define MY_CHECHER_H
+
+#include "common.h"
+#include "tuple.h"
+#include "tuple_sorter.h"
+#include "io.h"
+
+/// \brief perform checking after construction
+///
+template<typename alphabet_type, typename offset_type>
+class Checker{
+
+private:
+
+	typedef typename ExVector<alphabet_type>::vector alphabet_vector_type;
+
+	typedef typename ExVector<offset_type>::vector offset_vector_type; 
+
+private:
+
+	std::string m_s_fname;
+
+	std::string m_sa_fname;
+
+public:
+
+	/// \brief ctor
+	///
+	Checker(const std::string & _s_fname, const std::string & _sa_fname) :
+		m_s_fname(_s_fname), m_sa_fname(_sa_fname) {}
+
+
+	/// \brief check
+	///
+	bool run() {
+
+		// sort <sa[i], i> by sa[i]
+		typedef Pair<offset_type, offset_type> pair_type;
+
+		typedef TupleAscCmp1<pair_type> pair_comparator_type;
+
+		typedef typename ExSorter<pair_type, pair_comparator_type>::sorter sorter_type;
+
+		sorter_type *sorter = new sorter_type(pair_comparator_type(), MAX_MEM / 2);
+
+		stxxl::syscall_file *sa_file = new stxxl::syscall_file(m_sa_fname, stxxl::syscall_file::RDWR | stxxl::syscall_file::DIRECT);
+
+		offset_vector_type *sa = new offset_vector_type(sa_file);
+
+		typename offset_vector_type::const_iterator it = sa->begin();
+
+		for (offset_type i = 1; it != sa->end(); ++it, ++i) {
+
+			sorter->push(pair_type(*it, i)); // start ranking from 1 (0 is reserved for the sentinel)
+		}
+
+		delete sa; sa = nullptr;
+
+		delete sa_file; sa_file = nullptr;
+
+		sorter->sort();
+
+		// sort <isa[i], s[i], isa[i + 1]> by isa[i]
+		typedef Triple<offset_type, alphabet_type, offset_type> triple_type;
+
+		typedef TupleAscCmp1<triple_type> triple_comparator_type;
+
+		typedef typename ExSorter<triple_type, triple_comparator_type>::sorter sorter_type2;
+
[…]

### 2017-06-06 dsaism_plus_checker/src/common.h

@@ -1,0 +1,56 @@
+#ifndef __COMMON_H
+#define __COMMON_H
+
+#include "stxxl/bits/common/uint_types.h"
+#include "stxxl/vector"
+#include "stxxl/bits/containers/queue.h"
+#include "stxxl/bits/containers/deque.h"
+#include "stxxl/bits/containers/priority_queue.h"
+#include "stxxl/bits/containers/sorter.h"
+#include "stxxl/bits/io/syscall_file.h"
+
+#include <limits>
+#include <vector>
+#include <utility>
+
+// alias for integral types
+using uint8 = stxxl::uint8;
+
+using uint16 = stxxl::uint16;
+
+using uint32 = stxxl::uint32;
+
+using uint40 = stxxl::uint40;
+
+using uint64 = stxxl::uint64;
+
+// L-type or S-type
+constexpr uint8 L_TYPE = 0;
+
+constexpr uint8 S_TYPE = 1;
+
+// LONG_TYPE or SHORT_TYPE
+constexpr uint8 SHORT_TYPE = 0;
+
+constexpr uint8 LONG_TYPE = 1;
+
+// D_LOW & D_HIGH
+constexpr uint8 D_LOW = 4;
+
+constexpr uint8 D_HIGH = 4;
+
+// memory allocation
+constexpr uint64 K_512 = 512 * 1024;
+
+constexpr uint64 K_1024 = 1024 * 1024;
+
+constexpr uint64 MAX_MEM = 3 * 1024 * K_1024;
+
+constexpr uint64 MAX_MEM2 = MAX_MEM;
+
+constexpr uint64 MAX_ITEM = 16 * K_1024;
+
+
+
+
+#endif // common_h

### 2017-06-06 dsaism_plus_checker/src/dsais.h

@@ -1,0 +1,2647 @@
+#ifndef MY_DSAIS_H
+#define MY_DSAIS_H
+
+#include "common.h"
+#include "io.h"
+#include "sais.h"
+#include "tuple.h"
+#include "tuple_sorter.h"
+#include "substr_sorter.h"
+#include "test.h"
+
+#include <string>
+#include <fstream>
+#include <cassert>
+
+//#define TEST_DEBUG1 // output
+//#define TEST_DEBUG2 // assertion
+
+template<typename alphabet_type, typename offset_type, uint8 D>
+class DSAComputation;
+
+/// \brief portal to DSAComputation
+///
+template<typename alphabet_type, typename offset_type>
+class DSAIS{
+
+private:
+
+	typedef typename ExVector<alphabet_type>::vector alphabet_vector_type;
+
+	typedef typename ExVector<offset_type>::vector offset_vector_type;
+
+private:
+
+	const std::string m_s_fname; ///< input string file name
+	
+	const std::string m_sa_fname; ///< output SA file name
+
+public:
+
+	/// \brief ctor 
+	///
+	DSAIS(const std::string & _s_fname, const std::string & _sa_fname) : m_s_fname(_s_fname), m_sa_fname(_sa_fname) {}
+
+	/// \brief create DSAComputation
+	///
+	void run() {
+
+		// append a sentinel to the input string
+		stxxl::syscall_file *s_file = new stxxl::syscall_file(m_s_fname, stxxl::syscall_file::RDWR | stxxl::syscall_file::DIRECT);
+
+		alphabet_vector_type *s_origin = new alphabet_vector_type(s_file);
+
+		typename alphabet_vector_type::bufreader_type s_origin_reader(*s_origin);
+
+		alphabet_vector_type *s_target = new alphabet_vector_type(); s_target->resize(s_origin->size() + 1); // append a sentinel
+
+		typename alphabet_vector_type::bufwriter_type s_target_writer(*s_target);
+
+		uint64 s_origin_len = s_origin->size();
+
+		for (; !s_origin_reader.empty(); ++s_origin_reader) {
+
+			s_target_writer << *s_origin_reader;
+		}
+	
+		s_target_writer << alphabet_type(0);
+
+		s_target_writer.finish();
+
+		// clear
+		delete s_origin; s_origin = nullptr;
+
[…]

### 2017-06-06 dsaism_plus_checker/src/format.cpp

@@ -1,0 +1,39 @@
+#include "common.h"
+
+#include <iostream>
+#include <chrono>
+#include <fstream>
+
+#include "formatter.h"
+
+
+int main(int argc, char** argv){
+
+	// mount stxxl disk	
+	stxxl::config *cfg = stxxl::config::get_instance();
+
+	stxxl::disk_config disk("stxxl.tmp", 1024 * 1024 * 1024, "syscall autogrow unlink");
+
+	disk.direct = stxxl::disk_config::DIRECT_ON;
+
+	cfg->add_disk(disk);
+	
+	// check if input params are legal
+	if (argc != 3) {
+
+		std::cerr << "two param required: input_path and output_path.\n";
+
+		exit(-1);
+	}	
+
+	// retrieve file name for input string
+	std::string s_fname(argv[1]); 
+	
+	// retrieve file name for output SA
+	std::string s_target_fname(argv[2]);
+	
+	// compute the SA for the given string
+	Formatter<uint8> formatter(s_fname, s_target_fname);	
+
+	formatter.run();
+}

### 2017-06-06 dsaism_plus_checker/src/formatter.h

@@ -1,0 +1,70 @@
+#ifndef MY_FORMAT_H
+#define MY_FORMAT_H
+
+#include "common.h"
+#include "io.h"
+
+//#define TEST_DEBUG
+
+/// \brief format the input string
+///
+/// delete alpha_max, plus each remaining by 1
+template<typename alphabet_type>
+class Formatter{
+
+private:
+
+	typedef typename ExVector<alphabet_type>::vector alphabet_vector_type;
+
+private:
+
+	std::string m_s_fname;
+
+	std::string m_target_s_fname;
+
+public:
+	
+	/// \brief ctor
+	///
+	Formatter(const std::string & _s_fname, const std::string & _target_s_fname) : 
+			m_s_fname(_s_fname), m_target_s_fname(_target_s_fname) {}
+
+	/// \brief 
+	///
+	void run() {
+
+		stxxl::syscall_file *s_file = new stxxl::syscall_file(m_s_fname, stxxl::syscall_file::RDWR | stxxl::syscall_file::DIRECT);
+
+		alphabet_vector_type *s = new alphabet_vector_type(s_file);
+
+		typename alphabet_vector_type::const_iterator it = s->begin();
+
+		stxxl::syscall_file *s_target_file = new 
+			stxxl::syscall_file(m_target_s_fname, stxxl::syscall_file::CREAT | stxxl::syscall_file::RDWR | stxxl::syscall_file::DIRECT);
+
+		alphabet_vector_type *s_target = new alphabet_vector_type(s_target_file);
+
+		for (uint64 i = 0; it != s->end(); ++it, ++i) {
+
+			if (*it != std::numeric_limits<alphabet_type>::max()) {
+
+				s_target->push_back(*it + 1); // increment by 1
+			}
+			else {
+
+				// get rid of it
+			}
+		}
+
+		delete s; s = nullptr;
+
+		delete s_file; s_file = nullptr;
+
+		delete s_target; s_target = nullptr;
+
+		delete s_target_file; s_target_file = nullptr;
+	}
+};
+
+
+#endif

### 2017-06-06 dsaism_plus_checker/src/generate_error.sh

@@ -1,0 +1,8 @@
+for i in `seq 100000`
+do
+	dd if=enwiki_100M of=enwiki_10240 bs=10240 count=1 skip=$i
+
+	./process.sh enwiki_10240
+
+	echo $i
+done

### 2017-06-06 dsaism_plus_checker/src/io.h

@@ -1,0 +1,46 @@
+#ifndef MY_IO_H
+#define MY_IO_H
+
+#include "common.h"
+
+/// \brief stxxl vector generator
+///
+template<typename element_type>
+struct ExVector{
+
+	typedef typename stxxl::VECTOR_GENERATOR<element_type, 8 / sizeof(element_type) + 1, 2, K_512 * sizeof(element_type)>::result vector;
+};
+
+
+/// \brief stxxl sorter generator
+///
+template<typename tuple_type, typename comparator_type>
+struct ExSorter{
+
+	typedef typename stxxl::sorter<tuple_type, comparator_type> sorter;
+};
+
+
+/// \brief stxxl priority queue
+///
+template<typename tuple_type, typename comparator_type, uint32 mem_size, uint64 max_items, uint64 tune = 6>
+struct ExHeap{
+
+	typedef typename stxxl::PRIORITY_QUEUE_GENERATOR<tuple_type, comparator_type, mem_size, max_items, tune>::result heap;
+};
+
+/// \brief stxxl queue
+template<typename tuple_type>
+struct ExQueue{
+
+	typedef typename stxxl::queue<tuple_type> queue;
+};
+
+/// \brief stxxl deque
+template<typename tuple_type>
+struct ExDeque{
+
+	typedef typename stxxl::deque<tuple_type> deque;
+};
+
+#endif

### 2017-06-06 dsaism_plus_checker/src/losertree.h

@@ -1,0 +1,172 @@
+#include "common.h"
+
+template <typename Comparator>
+class LoserTree3Way
+{
+private:
+
+    /// marker inside loser tree to save equality comparisons
+    const uint32     EqualMark = 0x80000000;
+
+    /// marker inside loser tree for finished players
+    const uint32       DoneMark = 0x7FFFFFFF;
+
+    /// the comparator object of this tree
+    const Comparator&	m_cmp;
+
+    /// number of players
+    unsigned int	m_size;
+
+    /// the tree of size n-1
+    std::vector<uint32_t> m_tree;
+
+    /// saves the equality mark of top after extraction
+    bool		m_top_equal;
+
+    /// all players are done
+    bool                m_done;
+
+public:
+
+    /// Initialize loser tree and play initial games.
+    LoserTree3Way(const Comparator& cmp)
+        : m_cmp(cmp)
+    {
+    }
+
+    /// Play initial round of comparison games.
+    bool play_initial(unsigned int size)
+    {
+        m_size = size;
+        m_top_equal = false;
+        m_done = false;
+
+        // initialize loser tree of appropriate size (round up to power of two)
+        unsigned int treesize = (1 << (int)(log2(size - 1) + 2)) - 1;
+        if (treesize == 0) treesize = 1;
+        m_tree.resize(treesize, DoneMark);
+
+        // fill in lowest level: all ascending player numbers
+        int levelput = m_tree.size() / 2;
+        for (unsigned int i = 0; i < m_size; ++i)
+        {
+            m_tree[levelput + i] = m_cmp.exists(i) ? i : DoneMark;
+        }
+
+        unsigned int levelsize = levelput + 1;
+
+        // construct higher levels iteratively from bottom up
+        while ( levelsize > 1 )
+        {
+            levelsize = levelsize / 2;
+            int levelget = levelput;
+            levelput /= 2;
+
+            for (unsigned int i = 0; i < levelsize; ++i)
+            {
+                if ( m_tree[levelget + 2*i + 1] == DoneMark )
+                    m_tree[levelput + i] = m_tree[levelget + 2*i] & ~EqualMark;
+                else if ( m_tree[levelget + 2*i] == DoneMark )
+                    m_tree[levelput + i] = m_tree[levelget + 2*i + 1] & ~EqualMark;
+                else
+                {
+                    int cmp = m_cmp( m_tree[levelget + 2*i] & ~EqualMark, m_tree[levelget + 2*i + 1] & ~EqualMark );
+                    if (cmp < 0)
[…]

### 2017-06-06 dsaism_plus_checker/src/process.sh

@@ -1,0 +1,8 @@
+rm $1.format
+rm $1.format.sa5
+
+./format $1 $1.format
+
+./build $1.format $1.format.sa5
+
+./check $1.format $1.format.sa5

### 2017-06-06 dsaism_plus_checker/src/sacheckerlightweight.h

@@ -1,0 +1,36 @@
+#ifndef SACHECKERLIGHTWEIGHT_H
+#define SACHECKERLIGHTWEIGHT_H
+
+
+template<typename charT>
+class SaCheckerLightWeight{
+private:
+	const int32_t P = 2047483525;
+	const int32_t R = 1732938289;
+	
+	int32_t fp1;
+	int32_t fp2;
+
+public:
+
+	SaCheckerLightWeight(): fp1(0), fp2(0){}
+
+	void computeFP1(charT t) {
+		fp1 = ((int64_t)fp1 * R + t + 1) % P;
+	}
+
+	void computeFP2(charT t) {
+		fp2 = ((int64_t)fp2 * R + t + 1) % P;
+	}
+
+	bool validate() {
+		return (fp1 == fp2);
+	}	
+	
+	void print() {
+		std::cerr << "fp1: " << fp1 << " fp2: " << fp2 << std::endl;	
+	}
+};
+
+
+#endif

### 2017-06-06 dsaism_plus_checker/src/sais.h

@@ -1,0 +1,254 @@
+#ifndef MY_SAIS_H
+#define MY_SAIS_H
+
+#include "common.h"
+#include "utility.h"
+#include "io.h"
+
+
+//#define TEST_DEBUG2
+
+class SAComputation;
+
+/// \brief portal to SA computation on RAM
+///
+template<typename offset_type>
+class SAIS{
+
+private:
+
+	typedef typename ExVector<offset_type>::vector offset_vector_type;
+
+public:
+
+	SAIS(offset_vector_type * _s, offset_vector_type *&_sa_reverse); 
+
+};
+
+/// \brief compute SA for the input string residing on EM
+///
+/// \note alphabet_type = offset_type > uint32
+template<typename offset_type>
+SAIS<offset_type>::SAIS(offset_vector_type *_s, offset_vector_type*& _sa_reverse) {
+
+	// load _s into ram
+	// note that characters in _s are named in a condensed way
+	uint32 s_size = _s->size();
+
+	uint32 *s = new uint32[s_size];
+
+	typename offset_vector_type::bufreader_type *s_reader = new typename offset_vector_type::bufreader_type(*_s);
+
+	uint32 max_alpha = 0;
+
+	for (uint32 i = 0; i < s_size; ++i, ++(*s_reader)) {
+
+		s[i] = static_cast<uint32>(*(*s_reader)); // correct
+
+#ifdef TEST_DEBUG2
+		std::cerr << "_s[i]: " << *(*s_reader) << " s[i]: " << s[i] << std::endl;
+#endif
+		max_alpha = std::max(max_alpha, s[i]);
+	}	
+
+	delete s_reader; s_reader = nullptr;
+
+	// compute sa
+	uint32 *sa = new uint32[s_size];
+
+	SAComputation(s, s_size, max_alpha, sa);
+
+	// output sa reversely
+	_sa_reverse = new offset_vector_type();
+
+#ifdef TEST_DEBUG2
+
+	std::cerr << "sa: \n";
+#endif
+
+	for (uint32 i = s_size - 1; ; --i) {
+
+		_sa_reverse->push_back(static_cast<offset_type>(sa[i]));
+
+#ifdef TEST_DEBUG2
+
+		std::cerr << sa[i] << " ";
[…]

### 2017-06-06 dsaism_plus_checker/src/substr.h

@@ -1,0 +1,112 @@
+#ifndef MY_SUBSTR_H
+#define MY_SUBSTR_H
+
+//#define TEST_DEBUG
+
+#include "io.h"
+
+#if _MSC_VER
+#pragma pack(push, 1)
+#endif
+template<typename alphabet_type, uint D>
+struct Substr{
+
+private: 
+
+	typedef typename ExVector<alphabet_type>::vector alphabet_vector_type;
+
+	typedef typename ExVector<uint8>::vector uint8_vector_type;
+
+public:
+
+	std::vector<alphabet_type> data; // store at most D characters
+
+	bool is_short; // short or long
+	
+public:
+
+	/// \brief retrieve a non-size-one short substr
+	///
+	/// \return false is no more; otherwise, true
+	/// \note short substrs in a seq are separated by 0
+	bool deserializeShort(alphabet_vector_type * _substr_ch_seq, typename alphabet_vector_type::const_iterator & _substr_ch_it, 
+				typename uint8_vector_type::const_iterator & _substr_len_it) {
+
+		if (_substr_ch_it == _substr_ch_seq->end()) {
+
+			return false;
+		}
+
+		uint8 len = static_cast<uint8>(*_substr_len_it); // len < UINT8_MAX 
+
+		++_substr_len_it;
+
+		data.resize(len);
+
+		for (uint8 i = 0; i < len; ++i) { data[i] = *_substr_ch_it; ++_substr_ch_it; }
+
+		is_short = true;
+
+		return true;
+	}
+
+	/// \brief retrieve the first D characters from a long substr
+	///
+	void deserializeLong(typename alphabet_vector_type::const_iterator & _substr_ch_it) {
+
+		data.resize(D);
+
+		for (uint8 i = 0; i < D; ++i) { data[i] = *_substr_ch_it; ++_substr_ch_it; }	
+
+		is_short = false;
+	} 
+	
+	
+	/// \brief compare two substrs (at most one is long)
+	/// 
+	int cmp(const Substr &_b) const {
+
+		auto iter_a = data.begin(), iter_b = _b.data.begin();
+
+		for (; iter_a != data.end() && iter_b != _b.data.end(); ++iter_a, ++iter_b) {
+
+			if (*iter_a < *iter_b) return -1;
+
+			if (*iter_a > *iter_b) return +1;
+		}
+
+		if (iter_a != data.end()) return -1; // b is a prefix of a, a < b
+		
[…]

### 2017-06-06 dsaism_plus_checker/src/substr_sorter.h

@@ -1,0 +1,387 @@
+#ifndef SUBSTR_SORTER_H
+#define SUBSTR_SORTER_H
+
+
+#include "substr.h"
+#include "losertree.h"
+
+#define TEST_DEBUG_SUBSTR_SORTER
+
+template<typename alphabet_type, typename offset_type, uint8 D>
+class SubstrSorter{
+
+private:
+
+	typedef typename ExVector<alphabet_type>::vector alphabet_vector_type;
+
+	typedef typename ExVector<offset_type>::vector offset_vector_type;
+
+	typedef typename ExVector<uint8>::vector uint8_vector_type;
+
+	typedef typename ExVector<uint32>::vector uint32_vector_type;
+
+public:
+
+	struct StrCompare{
+
+	public:
+
+		std::vector<alphabet_vector_type*> & m_short_ch_seqs;
+
+		std::vector<typename alphabet_vector_type::const_iterator> m_short_ch_its; ///< short lms seqs
+
+		std::vector<uint8_vector_type*> & m_short_len_seqs;
+
+		std::vector<typename uint8_vector_type::const_iterator> m_short_len_its;
+
+		std::vector<alphabet_vector_type*> & m_long_ch_seqs;
+
+		std::vector<typename alphabet_vector_type::const_iterator> m_long_ch_its; ///< long lms seqs
+
+		uint8 m_seq_num; ///< number of seqs
+
+		std::vector<Substr<alphabet_type, D>> m_head; ///< current smallest for each seq
+
+		std::vector<bool> m_exists; ///< existence of each seq
+
+	public:
+		/// \brief ctor
+		///
+		StrCompare(std::vector<alphabet_vector_type*> &_short_ch_seqs, std::vector<uint8_vector_type*> & _short_len_seqs, std::vector<alphabet_vector_type*> &_long_ch_seqs) : m_short_ch_seqs(_short_ch_seqs), m_short_len_seqs(_short_len_seqs), m_long_ch_seqs(_long_ch_seqs) {
+
+			for (uint8 i = 0; i < m_short_ch_seqs.size(); ++i) {
+
+				m_short_ch_its.push_back(m_short_ch_seqs[i]->begin());
+
+				m_short_len_its.push_back(m_short_len_seqs[i]->begin());
+
+				m_long_ch_its.push_back(m_long_ch_seqs[i]->begin());
+			}
+		
[…]

### 2017-06-06 dsaism_plus_checker/src/test.h

@@ -1,0 +1,76 @@
+#ifndef MY_TEST_H
+#define MY_TEST_H
+
+#include "common.h"
+
+template<typename alphabet_vector_type>
+void test_output_vector(alphabet_vector_type *_vector) {
+
+	auto it = _vector->begin();
+
+	for (; it != _vector->end(); ++it) {
+
+		if (sizeof(*it) == sizeof(uint8)) {
+			
+			std::cerr << (uint32)*it << " ";
+		}
+		else {
+
+			std::cerr << *it << " ";
+		}
+	}
+
+	std::cerr << std::endl;
+}
+
+template<typename BlockInfo>
+void test_output_blocks_info(std::vector<BlockInfo> & _blocks_info) {
+
+	for (uint8 i = 0; i < _blocks_info.size(); ++i) {
+
+		const BlockInfo & bi = _blocks_info[i];
+
+		std::cerr << "id: " << (uint32)bi.m_id;
+	
+		std::cerr << " bpos: " << bi.m_beg_pos;
+	
+		std::cerr << " epos: " << bi.m_end_pos;
+
+		std::cerr << " lms_num: " << bi.m_lms_num;
+
+		std::cerr << " size: " << bi.m_size;
+
+		std::cerr << " cap: " << bi.m_capacity << std::endl;
+
+	}
+}
+
+template<typename T>
+void test_output_array_ram(T * _arr, uint32 _size) {
+
+	for (uint32 i = 0; i < _size; ++i) {
+
+		if (sizeof (T) == sizeof(uint8)) {
+
+			std::cerr << (uint32)_arr[i] <<" ";
+		}
+		else {
+			
+			std::cerr << _arr[i] << " ";
+		}
+	}
+
+	std::cerr << std::endl;
+}
+
+template<typename T>
+void test_output_element(T & _elem) {
+
+	std::cerr << _elem << " ";
+}
+
+
+
+
+
+#endif // MY_TEST_H

### 2017-06-06 dsaism_plus_checker/src/tuple.h

@@ -1,0 +1,141 @@
+#ifndef MY_TUPLE_H
+#define MY_TUPLE_H
+
+/// \brief structure pair
+///
+template<typename T1, typename T2>
+struct Pair{
+
+	T1 first;
+
+	T2 second;
+
+	/// \brief constructor, default
+	Pair(): first(0), second(0) {}
+
+	/// \brief constructor, copy
+	Pair(const Pair& _item) : first(_item.first), second(_item.second) {}
+
+	/// \brief constructor, component
+	Pair(const T1& _first, const T2& _second) : first(_first), second(_second) {}
+
+	/// \brief min value
+	static Pair& min_value() {
+
+		static Pair min_val = Pair();
+
+		return min_val;
+	}
+
+	/// \brief max vlaue
+	static Pair& max_value() {
+
+		static Pair max_val = Pair(std::numeric_limits<T1>::max(), std::numeric_limits<T2>::max());
+
+		return max_val;
+	}
+
+	/// \brief output
+	friend std::ostream& operator << (std::ostream _os, const Pair& _item) {
+
+		return _os << "first: " << _item.first << " second: " << _item.second << std::endl;
+
+	}
+}__attribute__((packed));
+
+
+/// \brief structure triple
+///
+template<typename T1, typename T2, typename T3>
+struct Triple{
+
+	T1 first;
+
+	T2 second;
+
+	T3 third;
+
+	/// \brief constructor, default
+	Triple(): first(0), second(0), third(0) {}
+
+	/// \brief constructor, copy
+	Triple(const Triple& _item) : first(_item.first), second(_item.second), third(_item.third) {}
+
+	/// \brief constructor, component
+	Triple(const T1& _first, const T2& _second, const T3& _third) : first(_first), second(_second), third(_third) {}
+
+	/// \brief min value
+	static Triple& min_value() {
+
+		static Triple min_val = Triple();
+
+		return min_val;
+	}
+
+	/// \brief max value
+	static Triple& max_value() {
+
+		static Triple max_val = Triple(std::numeric_limits<T1>::max(), std::numeric_limits<T2>::max(), std::numeric_limits<T3>::max());
+
+		return max_val;
+	}
+
+	/// \brief output
+	friend std::ostream& operator << (std::ostream _os, const Triple& _item) {
+
+		return _os << "first: " << _item.first << " second: " << _item.second << " third: " << _item.third << std::endl;
+	}
+
+}__attribute__((packed));
+
+
[…]

### 2017-06-06 dsaism_plus_checker/src/tuple_sorter.h

@@ -1,0 +1,155 @@
+#ifndef MY_TUPLE_SORTER_H
+#define MY_TUPLE_SORTER_H
+
+#include "common.h"
+	
+#include "stxxl/sorter"
+
+
+/// \brief sort tuples by 1st component in ascending order
+///
+template<typename tuple_type>
+struct TupleAscCmp1{
+
+	bool operator()(const tuple_type &_a, const tuple_type &_b) const {
+
+		return _a.first < _b.first;
+	}
+
+	tuple_type min_value() const {
+
+		return tuple_type::min_value();
+	}
+
+	tuple_type max_value() const {
+
+		return tuple_type::max_value();
+	}
+};
+
+/// \brief sort tuples by 1st component in descending order
+///
+template<typename tuple_type>
+struct TupleDscCmp1{
+
+	bool operator()(const tuple_type &_a, const tuple_type &_b) const {
+
+		return _a.first > _b.first;
+	}
+
+	tuple_type min_value() const {
+
+		return tuple_type::max_value();
+	}
+
+	tuple_type max_value() const {
+
+		return tuple_type::min_value();
+	}
+};
+
+/// \brief sort tuples by 1st component in ascending order
+///
+template<typename tuple_type>
+struct TupleAscCmp2{
+
+	bool operator()(const tuple_type &_a, const tuple_type &_b) const {
+
+		if (_a.first == _b.first) return _a.second < _b.second;
+
+		return _a.first < _b.first;
+	}
+
+	tuple_type min_value() const {
+
+		return tuple_type::min_value();
+	}
+
+	tuple_type max_value() const {
+
+		return tuple_type::max_value();
+	}
+};
+
+/// \brief sort tuples by 1st component in descending order
+///
+template<typename tuple_type>
+struct TupleDscCmp2{
+
+	bool operator()(const tuple_type &_a, const tuple_type &_b) const {
+
+		if (_a.first == _b.first) return _a.second > _b.second;
+
+		return _a.first > _b.first;
+	}
+
+	tuple_type min_value() const {
+
+		return tuple_type::max_value();
+	}
+
+	tuple_type max_value() const {
+
+		return tuple_type::min_value();
+	}
+};
+/// \brief sort tuples by <1st, 2nd, 3rd> components in ascending order
+///
+template<typename tuple_type>
+struct TupleAscCmp3{
+
+
+	bool operator()(const tuple_type &_a, const tuple_type &_b) const {
[…]

### 2017-06-06 dsaism_plus_checker/src/utility.h

@@ -1,0 +1,60 @@
+#ifndef MY_UTILITY_H
+#define MY_UTILITY_H
+
+#include "common.h"
+
+
+static const uint8 MASK[] = { 0x80, 0x40, 0x20, 0x10, 0x08, 0x04, 0x02, 0x01 };
+
+/// \brief data wrapper for a bit array in RAM
+///
+class BitWrapper{
+
+private:
+
+	char * data;
+
+public:
+
+	BitWrapper(char * _data = nullptr) : data(_data) {}
+
+	bool get(size_t _idx) {
+
+		return  (data[_idx / 8] & MASK[_idx % 8]) ? 1 : 0;
+	}
+
+	void set(size_t _idx, bool _val) {
+
+		data[_idx / 8] = _val ? (MASK[_idx % 8] | data[_idx / 8]) : ((~MASK[_idx % 8]) & data[_idx / 8]);
+	}
+};
+
+
+
+static const uint8 MASK2[] = { 0xC0, 0x30, 0x0C, 0x03 };
+
+static const uint8 SHIFT2[] = { 6,4,2,0 };
+
+/// \brief data warpper for a 2-bit array in RAM
+///
+class Bit2Wrapper {
+private:
+
+	char * data;
+
+public:
+
+	Bit2Wrapper(char* _data = nullptr) : data(_data) {}
+
+	uint8 get(size_t _idx) {
+
+		return (data[_idx / 4] >> SHIFT2[_idx % 4]) & MASK2[3];
+	}
+
+	void set(size_t _idx, uint8 _val) {
+
+		data[_idx / 4] = (_val << SHIFT2[_idx % 4]) | (data[_idx / 4] & ~MASK2[_idx % 4]);
+	}
+};
+
+#endif //MY_UTILITY_H

### 2017-06-27 new-draft/draft.tex
20170628

@@
 
-%The induced sorting~(IS) method has been applied to designing several disk-based suffix array~(SA) construction algorithms.
-Although these algorithms have better time and I/O complexities than the other alternatives, their current programs commonly suffer from a bottleneck due to the large disk space in need.
-
-The induced sorting~(IS) method has been applied to designing efficient suffix array~(SA) construction algorithms on both internal and external memory models.
-In this paper, we propose two checking methods that enable any IS suffix sorters to check an SA when it is being built.
-The first method is capable of checking both suffix and LCP arrays for strings on constant alphabets, while the second method is specific for checking SA only but applicable to strings on both constant and integer alphabets.
-For performance analysis, we integrate the second checking method into the existing SA construction algorithm DSA-IS, where the design of DSA-IS is adapted by new substring sorting and naming methods for reducing its peak disk use to at most 14 times as input size.
-From our experimental results, the time, space and I/O consumptions for verification by our second method is negligible to that for construction by the adapted DSA-IS.
-This convinced that our checkers, together with the state-of-the-art IS suffix sorters, may constitute efficient solutions for building and checking SA at the same time.
+Suffix array~(SA) can be built by the induced sorting~(IS) method on both internal and external memory models.
+We propose two methods that enable any IS suffix sorter to build and check an SA simultaneously.
+The first method can check both suffix and longest common prefix~(LCP) arrays for strings drawn from constant alphabets, while the second method is applicable to checking SA for strings of any alphabets.
+By combining our methods with the Karp-Rabin fingerprinting function, we design two probablistic algorithms that perform verification with a neligible error probability.
+These algorithms are rather lightweight in terms of time and space.
+Particularly, the algorithm designed by the second method only takes linear time and constant space when running in external memory.
+For a further study, we integrate the checking algorithm designed by the second method into the existing suffix sorter DSA-IS to evaluate the checking overhead, where the time, space and I/O consumptions for verification are considerably smaller than that for construction in the experiments.
+This convinces us that the proposed checking methods, together with the state-of-the-art IS suffix sorters, could constitute efficient solutions for both construction and verification.
 
@@
 
-The suffix array of any constant or integer string can be built in linear time and RAM space by SA-IS~\cite{Nong11}.
-The induced-sorting method has been also applied to designing fast external-memory SA construction algorithms eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}.
-These three disk-based variants have better time complexities than the other alternatives~(e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}), but they all suffer from a bottleneck due to the large disk space in use.
-For example, when building SA encoded by 40-bit integers, the peak disk use for eSAIS, DSA-IS and SAIS-PQ are $24n$, $20n$ and $15n$, respectively.
-This situation changed due to the work presented in~\cite{Karkkainen2017}, which proposed a carefully engineered IS suffix sorter fSAIS that takes no more than $8n$ disk memory for any constant or integer string of size $n \le 2^{40}$~\cite{Karkkainen2017}.
-As reported, fSAIS reuses the substring/suffix sorting and naming methods of DSA-IS and SAIS-PQ, it achieves the remarkable space performance by means of a monotone priority queue for sorting fixed-size tuples and an I/O layer for allocating and deallocating disk space in a fine-grained way.
-The successful design of fSAIS shows a great potential for sharply decreasing the disk space requirements of the current programs for DSA-IS and SAIS-PQ.
-In this paper, we present our first attempt to improve DSA-IS by new substring sorting and naming methods.
-From the experiments, our program for the adapted algorithm DSA-IS+ takes $12n$ disk space in average and runs as fast as eSAIS on various real-world datasets.
-We will demonstrate that this program can be further improved using the optimization technique for implementing fSAIS.
+A suffix array can be built in linear time and space by the internal-memory algorithm SA-IS~\cite{Nong11}.
+According to the IS principle, the relative order of two unsorted suffixes is determined by sequentially comparing the heading characters and the sorted succeeding suffixes starting with their second characters.
+Recently, the IS method has been applied to designing three disk-based suffix sorting algorithms eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}, which have a better time complexity than the other alternatives~(e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}) but suffer from a bottleneck due to the large disk space for obtaining the heading characters of unsorted suffixes and the ranks of their successors in a disk-friendly way.
+More specifically, the average peak disk use to construct an SA encoded by 40-bit integers for pSAscan is $7.5n$, while that for eSAIS, DSA-IS and SAIS-PQ are $24n$, $20n$ and $15n$, respectively.
+The poor space performance is mainly because that the current programs for these algorithms fail to free the disk space for temporary data even when the data is no longer to use.
+A dramatic improvement can be achieved by splitting a file into multiple pieces and deleting each piece immediately when the temporary data in it is not needed any more.
+This technique has been used to implement a new IS suffix sorting algorithm fSAIS~\cite{Karkkainen2017}.
+As reported, the engineering of fSAIS consumes no more than $8n$ disk space for constructing an SA encoded by 40-bit integers, indicating a great potential for optimizing our programs for DSA-IS and SAIS-PQ.
 
-To avoid computation errors caused by implementation bugs and/or hardware malfunctions, an SA should be checked to ensure its correctness after construction.
-Currently, the software packages for DC3 and eSAIS provide users a checker based on the method proposed in~\cite{Dementiev2008a}, which performs two disk sorts for $\mathcal{O}(n)$ fixed-size tuples with an integer key.
-When running in external memory, this checker takes one-third as much time and I/O consumptions as the programs for DSA-IS and eSAIS, it also requires nearly the same disk space as the two SA builders.
-Against the background, we propose two checking methods that enable any IS suffix sorting algorithm to build and check SA at the same time.
-In comparison with the previous checking method, they are rather lightweight in terms of time, space and I/O volume.
-It should be noticed that these two methods are designed for different situations.
-In brief, the first method can check both suffix and LCP arrays, but it is specific for strings on constant alphabets.
-The second method can be applied to strings on constant and integer alphabets, but it can check SA only.
-As will be seen in our experiments, the consumption for checking by our second method is negligible compared with that for building by DSA-IS+.
+A constructed SA should be checked to avoid computation errors caused by implementation bugs and/or hardware malfunctions.
+Currently, the software packages for DC3 and eSAIS provide users a checker designed by the method proposed in~\cite{Dementiev2008a}.
+The overhead for this checker is rather high because it takes two external-memory sorts for arranging $\mathcal{O}(n)$ fixed-size tuples according to their integer keys.
+Against the background, we propose in this paper two methods that enable any IS suffix sorter to check an SA when it is being built.
+The first method can check both suffix and LCP arrays for strings drawn from constant alphabets, while the second method can check SA for strings drawn from any alphabets.
+We augment these methods with the Karp-Rabin fingerprinting function~\cite{Karp1987} to design two probabilistic algorithms, in terms of that their checking results are wrong with a negligible probability.
+These algorithms are rather lightweight in terms of time and space.
+Particularly, the algorithm designed by the second method only takes linear time and constant space to run using external memory.
+For a further study, we integrate this algorithm into DSA-IS to evaluate its checking overhead, where the design of DSA-IS has been adapted by new substring sorting and naming methods for a better performance.
+As demonstrated in our experiments, the time, space and I/O consumptions for verification by this algorithm are negligible to that for construction by the adapted DSA-IS.
+We are convinced that the proposed SA checkers, together with the state-of-the-art IS suffix sorters, could constitute efficient solutions for both construction and verification.
 
 The rest of this paper is organized as follows.
-Section~\ref{sec:builder} describes the framework of IS suffix sorting algorithms and the idea of new substring sorting and naming methods, Section~\ref{sec:checkers} the designs of two checking methods, Sections~\ref{sec:experiments} the experimental results, and Section~\ref{sec:conclusion} the concluding remarks.
+Section~\ref{sec:preliminaries} introduces some basic notations and symbols.
+Section~\ref{sec:builder} gives an overview of IS suffix sorting algorithms and the details of new substring sorting and naming methods.
+Section~\ref{sec:checkers} describes the ideas of checking methods and algorithms.
+Sections~\ref{sec:experiments} and~\ref{sec:conclusion} show the experimental results and the concluding remarks, respectively.
+
+\section{Preliminaries}~\label{sec:preliminaries}
+
+Given a string $x[0,n)$ drawn from a full-ordered alphabet $\Sigma$, we assume the ending character $x[n - 1]$ to be unique and lexicographically smaller than any other characters in $x$.
+For convenience, we denote by ${\sf suf}(i)$ and ${\sf sub}(i, j)$ the suffix running from $x[i]$ to $x[n-  1]$ and the substring running from $x[i]$ to $x[j]$, respectively.
+The following notations are also used in our presentation.
+
+Characters in $x$ are classified into two categories: L-type and S-type.
+We say $x[i]$ is S-type if (1) $i = n - 1$ or (2) $x[i] = x[i + 1]$ and $x[i + 1]$ is S-type; otherwise $x[i]$ is L-type.
+Further, if $x[i]$ and $x[i - 1]$ are respectively S-type/L-type and L-type/S-type, then $x[i]$ is also called S*-type/L*-type.
+We use an array $t$ to record the type of all the characters in $x$, where $t[i] = 1$ or $0$ if $x[i]$ is S-type or L-type, respectively.
+A substring or suffix is of the same type as its heading character.
+In this paper, we only consider substrings ending with an S*-type character.
+
+Given two characters $x[i]$ and $x[i + 1]$, we say $x[i]$ is the predecessor of $x[i + 1]$ and $x[i + 1]$ is the successor of $x[i]$.
+Accordingly, we define the predecessor-successor relationship between ${\sf suf}(i)/{\sf sub}(i, j)$ and ${\sf suf}(i + 1)/{\sf sub}(i + 1, j)$.
+
+We partition $x$ into S*-type substrings and replace each substring with its name to produce the reduced string $x_1$, where the name indicates the rank of the corresponding S*-type substring among all.
+
+The suffix array $sa$ indicates the lexical order of all the suffixes in $x$, where $sa[i]$ records the starting position of the $(i + 1)$-th smallest suffix.
+We also define the suffix array $sa_1$ for $x_1$ following the same way.
+
+The LCP array $lcp$ indicates the LCP-values of all the neighboring suffixes in $sa$, where $lcp[0] = 0$ and $lcp[i] = \ell$ for $i \in [1, n)$, where $\ell$ is the length of longest common prefix of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[i - 1])$.
+
+All the suffixes in $sa$ are naturally partitioned into multiple buckets.
+Each bucket occupies a contiguous interval in $sa$ and contains all the suffixes starting with a certain character.
+Without loss of generality, we denote by ${\sf sa\_bkt}(c_0)$ the bucket for suffixes starting with $c_0$.
+Notice that ${\sf sa\_bkt}(c_0)$ can be further divided into two parts, where the left part ${\sf sa\_bkt_L}(c_0)$ and the right part ${\sf sa\_bkt_S}(c_0)$ contain the L-type and S-type suffixes, respectively.
+We also partition $lcp$ into buckets such that ${\sf lcp\_bkt}(c_0)$, ${\sf lcp\_bkt_L}(c_0)$ and ${\sf lcp\_bkt_S}(c_0)$ record the LCP-values of suffixes in ${\sf sa\_bkt}(c_0)$, ${\sf sa\_bkt_L}(c_0)$ and ${\sf sa\_bkt_S}(c_0)$, respectively.
+
+Let $n_1 = \|x_1\|$, we use $sa^*[0, n_1)$ and $lcp^*[0, n_1)$ to denote the suffix and LCP arrays for S*-type suffixes in $x$, respectively.
+Specifically, $sa^*[i]$ records the starting position of the $(i + 1)$-th smallest S*-type substring, while $lcp^*[i]$ records the LCP-value of ${\sf suf}(sa^*[i])$ and ${\sf suf}(sa^*[i - 1])$.
 
 \section{Builder}~\label{sec:builder}
-
-\subsection{Preliminaries}
-
-Consider a string $x[0,n)$ drawn from a full-ordered alphabet $\Sigma$.
-The ending character $x[n - 1]$ is supposed to be unique and lexicographically smaller than any others in $x$.
-For convenience, we denote by ${\sf suf}(i)$ the suffix running from $x[i]$ to the ending character and ${\sf sub}(i,j)$ the substring running from $x[i]$ to $x[j]$, respectively.
-The following notations are also used in our presentation.
-
-{\em Character/substring/suffix classification.} Characters in $x$ are classified into two categories: L-type and S-type.
-We say $x[i]$ is S-type if (1) $i = n - 1$ or (2) $x[i] = x[i + 1]$ and $x[i + 1]$ is S-type; otherwise $x[i]$ is L-type.
-Furthermore, if $x[i]$ and $x[i - 1]$ are respectively S-type/L-type and L-type/S-type, then $x[i]$ is also called S*-type/L*-type.
-An array $t$ records the type of all the characters in $x$, $t[i]$ is set to $1$ or $0$ if $x[i]$ is S-type or L-type, respectively.
-Each substring/suffix is of the same type as its heading character.
-In this paper, we only consider the substrings ending with an S*-type character.
-
-{\em Predecessor and successor.} Given two characters $x[i]$ and $x[i + 1]$, $x[i]$ is the predecessor of $x[i + 1]$ and $x[i + 1]$ is the successor of $x[i]$.
-Accordingly, we define the predecessor-successor relationship between two suffixes or substrings starting with neighboring characters.
-
-{\em Reduced string.} Split $x$ into multiple substrings such that each substring only contains two S*-type characters at the starting and ending positions.
-Assign each substring a name and replace them with their names to produce a reduced string $x_1$, where the name indicates the lexical order of the corresponding substring among all.
-
-{\em Suffix array.} The suffix array $sa$ indicates the lexicographical order of all the suffixes in $x$, where $sa[i]$ records the starting position of the $(i + 1)$-th smallest suffix.
-We also define the SA for $x_1$ by $sa_1$.
 
@@
 
-	\caption{The general framework for an IS suffix sorting algorithm.}
+	\caption{The overview of IS suffix sorting algorithms	.}
 
@@
 As shown in Algorithm~\ref{alg:1}, an IS suffix sorting algorithm first performs a reduction phase to produce the reduced string $x_1$.
-If there exist two equal characters in $x_1$, then it recursively calls the reduction phase with $x_1$ as input; otherwise, all the S*-type suffixes in $x$ are already sorted and it performs an induction phase to produce $sa$ from these sorted suffixes.
-According to the IS principle, the order of two substrings/suffixes is determined by comparing their heading characters and successors in sequence during the execution of reduction/induction phase.
-Notice that this key operation takes multiple random accesses to $x$ and $sa$.
-It can be done very quickly if both $x$ and $sa$ are wholly loaded into RAM; otherwise, each access may take an individual I/O operation to external memory and thus becomes the bottleneck.
+If there exist two equal characters in $x_1$, then it recursively calls the reduction phase with $x_1$ as input; otherwise, all the S*-type suffixes in $x$ are already sorted and it performs an induction phase to produce the final $sa$.
+During the execution of two phases, the order of substrings/suffixes is determined by comparing the heading characters and the ranks of their successors according to the IS principle.
+This involves frequent random accesses to $x$ and $sa$ for retrieving the heading characters of unsorted substrings/suffixes and the ranks of their sorted successors.
+It can be done very quickly if both $x$ and $sa$ are wholly loaded into RAM; otherwise, each disk access takes an individual I/O operation, leading to a severe performance degradation.
+This problem is solved in DSA-IS by conducting two steps:
 
-To amortize the I/O overhead, eSAIS, DSA-IS and SAIS-PQ use different methods to obtain heading characters in a disk-friendly way.
-Particularly, DSA-IS first divides $x$ into multiple blocks and calls SA-IS to sort suffixes in each block at the beginning of induction phase, where the heading characters in need are written to external memory in their access order.
-Then, it organizes the sorted suffixes in a priority queue and visits them sequentially to induce the order of their predecessors following the IS principle.
-Because the required heading characters was already stored on disks, they can be read into RAM by sequential I/O operations when needed.
-This technique is also applied to sorting substrings during the reduction phase.
-Afterward, a naming process is conducted to check equality of lexicographically neighboring substrings for producing the reduced string $x_1$.
-To do this, DSA-IS maintains a copy for each substring in external memory and arranges them in their sorted order.
-As can be seen in Section~\ref{sec:experiments}, our program for DSA-IS requires less disk space than that for eSAIS, but it runs slower than the latter due to the large I/O volume for sorting and naming S*-type substrings.
+\begin{enumerate}[S1]
+	\item Split $x$ into blocks and sort substrings/suffixes of each block by calling SA-IS, the heading characters in use are copied to external memory in their access order.
+\label{dsais_sorting_method:1}
+	\item Sort the substrings/suffixes of $x$ by their heading characters and the ranks of their successors.
+Organize the sorted substrings/suffixes in an external-memory heap and induce their predecessors into the heap according to the IS principle, where the heading characters of the induced substrings/suffixes are retrieved from external memory by sequential I/O operations.\label{dsais_sorting_method:2}
+\end{enumerate}
+
+As shown in Section~\ref{sec:experiments}, our program for DSA-IS requires less disk space than that for eSAIS, but the former runs slower than the latter caused by the large I/O volume for sorting and naming S*-type substrings during the reduction phase.
+For this, we present new substring sorting and naming methods in the next subsection.
 
@@
 
-We first describe a new method for sorting S*-type substrings during the reduction phase.
-Given an integer value $D$, all the S*-type substrings are classified into long and short categories with respect to whether or not they contain more than $D$ characters.
-Our sorting method executes the following three steps in sequence after splitting $x$ into blocks $\{b_1, b_2, ...\}$.
+All the S*-type substrings are classified into long and short categories with respect to whether or not containing more than $D$ characters.
+The new substring sorting method mainly consists of the three steps below:
 
-\begin{enumerate}[S1]
-	\item Sort S*-type substrings in each block by calling the reduction phase of SA-IS.
-Copy the short substrings and store them on the disk in their sorted order.
-Copy the heading characters needed for sorting long substrings and store them on the disk in their access order.~\label{sort_and_name_method:1}
+\begin{enumerate}[S1']
+	\item Sort the long of each block by S\ref{dsais_sorting_method:1}. During the process, copy the short to external memory in their sorted order.~\label{dsaism_sorting_method:1}
 
-	\item Sort long S*-type substrings by the same way of DSA-IS. Copy the leftmost $D$ characters of the long substrings and store them on the disk in their sorted order.~\label{sort_and_name_method:2}
+	\item Sort the long of $x$ by S\ref{dsais_sorting_method:2}. During the process, copy the leftmost $D$ characters of the long to external memory in their sorted order.~\label{dsaism_sorting_method:2}
 
-	\item Merge short and long S*-type substrings by a multi-way sorter.~\label{sort_and_name_method:3}
+	\item Merge the short and long by a multi-way sorter.~\label{dsaism_sorting_method:3}
 \end{enumerate}
 
-Assume the number of blocks is $k$, S\ref{sort_and_name_method:1} copies the sorted short substrings in $b_i$ to an external memory vector $vs_i$ and S\ref{sort_and_name_method:2} copies the sorted long substrings to an external memory vector $vl$.
-Then, S\ref{sort_and_name_method:3} uses a sorter to merge these partial results, where the sorter maintains a minimal heap in RAM to determine the current smallest substring in $vs_1, vs_2, ..., vs_k$ and $vl$.
-For any two substrings in the heap, their lexicographical order can be determined in $\mathcal{O}(D)$ time by literally comparing their characters from left to right.
-This sorting method achieves a good performance when most of the S*-type substrings are short.
-This makes it feasible in practice because the average length of S*-type substrings in real-world datasets are typically small.
+After the first two steps, the short S*-type substrings of each block and the long S*-type substrings of $x$ are sorted and separately organized as a sequence in external memory.
+Assume $x$ is split into $k$ blocks, the multi-way sorter in S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap to cache the current smallest of each sequence and continually retrieve the top item from the heap to determine the lexical order of substrings from different sequences.
+The heap contains at most $k + 1$ substrings at any point of time and it compares any two substrings in $\mathcal{O}(D)$ time by a literal string comparison\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}.
+Theoretically, the sorting method has a good performance if the majority of S*-type substrings in $x$ are short with a small $D$.
+This is commonly satisfied in practice, because the average length of S*-type substrings is typically small in real-world datasets.
 
-Next, we describe how to name S*-type substrings at the same time when they are being sorted.
-Given two S*-type substrings successively popped from the minimal heap, we make a literal comparison of them to check their equality if they are both short; otherwise, we use the naming method of SAIS-PQ to determine their equality in S\ref{sort_and_name_method:2}.
-Specifically,
-substrings are sorted by comparing their heading characters and the names of their sorted successors in S\ref{sort_and_name_method:2}, where two substrings are of the same name if and only if their heading characters and the names of their successors are both equal.
-By recording the names of long S*-type substrings in their sorted order, it takes constant time to check equality of two lexicographically neighboring long substrings in S\ref{sort_and_name_method:3}.
+In what follows, we describe a method for naming the S*-type substrings during the above sorting process.
+The key point here is to check equality of two substrings successively popped from the heap in S\ref{dsaism_sorting_method:3}' as fast as possible.
+If either of the two substrings is short, our solution is to literally compare them in $\mathcal{O}(D)$ time.
+Otherwise, if both of them are long, then we assign names to all the long S*-type substrings using the technique proposed for SAIS-PQ in S\ref{dsaism_sorting_method:2}' and compare two successively popped long S*-type substrings by their names.
+This technique is based on the idea that two substrings are of the same name if their heading characters and the names of their successors are both equal\footnote{We refer interested readers to~\cite{Liu15} for more information.}.
 
-\subsection{Discussion} \label{subsec:further_discussion}
-
-Section~\ref{sec:experiments} shows that, by using the new substring sorting and naming methods, our program for the enhanced  DSA-IS has an advantage over that for eSAIS in terms of space and I/O volume, but its average peak disk use is still higher than that for fSAIS by around $5n$.
-This is mainly due to the use of external-memory containers (vectors, sorters and priority queues) provided by the STXXL library~\cite{Dementiev2007}.
-More specifically, the STXXL's containers fail to free the disk space for saving temporary data immediately after they are no longer to use.
-This implementation issue can be solved by storing temporary data into multiple files and deleting each file when it is obsolete.
-In this way, our program can be further improved to achieve a space performance similar to that for fSAIS.
+The experimental results in Section~\ref{sec:experiments} indicate that, by using the new substring sorting and naming methods, the adapted DSA-IS, called DSA-IS+, only takes half as much disk space as eSAIS.
+Also, the peak disk use of the program for DSA-IS+ can be further reduced to a level comparable to that of the program for fSAIS.
 
@@
 \subsection{Prior Art} \label{sec:checkers:prior_art}
-The widespread software packages for DC3 and eSAIS provide users a checker to perform verification. Theorem~\ref{theorem:1} gives the main idea behind this checker.
+
+The software packages for DC3 and eSAIS provide users a checker to ensure the correctness of their outputs. The following statements describe the idea behind this checker.
+
+\begin{lemma} \label{lemma:1}
+	For $i, j \in [0, n)$ and $i \ne j$, ${\sf suf}(i) < {\sf suf}(j) \Leftrightarrow (x[i], {\sf suf}(i + 1)) < (x[j], {\sf suf}(j + 1))$.
+\end{lemma}
 
@@
 	(1) $sa$ is a permutation of $[0, n)$. \\	
-	(2) ${\sf r_i < r_j} \Leftrightarrow (x[i], r_{i + 1}) < (x[i], r_{j + 1})$ for $ i, j \in [0, n)$ and $i\ne j$, where $r_i$ and $r_j$ indicates the relative order of ${\sf suf}(i)$ and ${\sf suf}(j)$ among all the suffixes, respectively.
+	(2) ${\sf r_i < r_j} \Leftrightarrow (x[i], r_{i + 1}) < (x[i], r_{j + 1})$ for $ i, j \in [0, n)$ and $i\ne j$, where $r_i$ and $r_j$ represent the ranks of ${\sf suf}(i)$ and ${\sf suf}(j)$ among all the suffixes, respectively.
 \\
@@
 \begin{IEEEproof}
-	The proof is available in~\cite{Dementiev2008a}.
+	Theorem~\ref{theorem:1} is derived from Lemma~\ref{lemma:1}. The proof is available in~\cite{Dementiev2008a}.
 \end{IEEEproof}
@@
 This checker performs two passes of integer sorts and each sort involves $n$ fixed-size tuples.
-When implemented in external memory, it requires large disk space and consumes high I/O volume, resulting in a non-negligible checking overhead.
-As will be seen in Section~\ref{sec:experiments}, the peak disk use for checking an SA encoded by 40-bit integers is around $21n$, while the I/O volume is no less than $50n$.
-In this section, we describe two lightweight SA checking methods based on the IS principle.
-Both of them can be seamlessly integrated into any IS suffix sorting algorithm to build and check SA simultaneously.
+When implemented in external memory, it consumes large disk space and I/O volume, resulting in a high checking overhead.
+For example, we can observe from Section~\ref{sec:experiments} that, the peak disk use and the I/O volume for checking an SA encoded by 40-bit integers are around $26n$ and $53n$, respectively.
+In this section,  we describe two SA checking methods based on the IS principle, both of them can be seamlessly integrated into any IS suffix sorting algorithm to build and check an SA at the same time.
 
@@
 
-\subsubsection{Preliminaries} \label{sec:checkers:method_a:preliminaries}
+Given that each S*-type substring is unique in $x$, SA-IS first computes $sa^*$ by sorting the starting positions of all the S*-type suffixes with their ranks indicated by $sa_1$ and then induces $sa$ from $sa^*$ by S\ref{induction_phase:1}''-S\ref{induction_phase:3}'' during the inducting phase.
 
-{\em LCP array.} The LCP array for $sa$, denoted by $lcp$, satisfies that $lcp[0] = 0$ and $lcp[i] = \ell$ for $i \in [1, n)$, where $\ell$ is the length of the longest common prefix of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[i - 1])$.
+\begin{enumerate}[S1'']
 
-{\em Suffix and LCP buckets.} All the suffixes in $sa$ are naturally partitioned into multiple buckets.
-Each bucket contains suffixes starting with a same character and occupies a contiguous interval of $sa$.
-For example, ${\sf sa\_bkt}(c_0)$ gathers all the suffixes starting with $c_0$.
-This bucket can be further partitioned into two parts, where the left and right parts only contain the L-type and S-type suffixes, respectively.
-For short, we denote the left and right parts by ${\sf sa\_bkt_L}(c_0)$ and ${\sf sa\_bkt_S}(c_0)$, respectively.
-Accordingly, $lcp$ can be also partitioned into multiple buckets and ${\sf lcp\_bkt}(c_0)$ records the LCP-values of suffixes in ${\sf sa\_bkt}(c_0)$.
-Similarly, we define ${\sf lcp\_bkt_L}(c_0)$ and ${\sf lcp\_bkt_S}(c_0)$ for ${\sf sa\_bkt_L}(c_0)$ and ${\sf sa\_bkt_S}(c_0)$, respectively.
-
-{\em Suffix and LCP arrays for S*-type suffixes.} Let $n_1 = \|x_1\|$, we use $sa^*[0, n_1)$ and $lcp^*[0, n_1)$ to denote the suffix and LCP arrays for S*-type suffixes, respectively.
-The $i$-th element in $sa^*$ records the starting position of the $(i + 1)$-th smallest S*-type substring, while the $i$-th element in $lcp^*[i]$ records the LCP-value of ${\sf suf}(sa^*[i])$ and ${\sf suf}(sa^*[i - 1])$.
-
-\subsubsection{Idea} \label{sec:checkers:method_a:idea}
-
-During the induction phase, SA-IS first computes $sa^*$ by sorting the starting positions of all the S*-type suffixes with their ranks indicated by $sa_1$ and then induces $sa$ from $sa^*$ by S\ref{induction_phase:1}'-S\ref{induction_phase:3}'.
-
-\begin{enumerate}[S1']
-
-	\item Scan $sa^*$ rightward with $i$ decreasing from $n_1 - 1$ to $0$.
+	\item Clear $sa$.
+Scan $sa^*$ leftward with $i$ decreasing from $n_1 - 1$ to $0$.
 For each scanned item $sa^*[i]$, insert it into the rightmost empty position of ${\sf sa\_bkt_S}(x[sa^*[i]])$.
@@
 
-	\item Scan $sa$ leftward with $i$ decreasing from $n - 1$ to $0$.
+	\item Clear ${\sf sa\_bkt_S}(c)$ for $c \in \Sigma$.
+Scan $sa$ leftward with $i$ decreasing from $n - 1$ to $0$.
 For each scanned non-empty item $sa[i]$, insert $sa[i] - 1$ into the rightmost empty position of ${\sf sa\_bkt_S}(x[sa[i] - 1])$ if $t[sa[i] - 1] = 1$.\label{induction_phase:3}
@@
 
-An example of the above inducing process is given in Fig.~\ref{fig:example1}.
-In line 6, we find the end of each bucket and insert the sorted S*-type suffixes into the corresponding buckets.
-Then, we find the head of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes.
+We illustrate a running example of the above process in Fig.~\ref{fig:example1} and show more details for a better understanding.
+In line 6, we find the ending position of each bucket and insert the sorted S*-type suffixes into the right part of the corresponding buckets.
+Afterward, we first find the heading position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes as following.
 In lines 8-9, we check $\sf suf(14)$ (marked by the symbol $@$) and find its predecessor ${\sf suf(13)}$ is L-type, thus we put ${\sf suf(13)}$ into the current leftmost empty position in ${\sf sa\_bkt_L}(i)$.
 To step through $sa$ in this way, we get all the L-type suffixes sorted in line 17.
-After that, we find the end of each bucket and scan $sa$ leftward for inducing the order of S-type suffixes in lines 18-29.
+After that, we find the ending position of each bucket and scan $sa$ leftward for inducing the order of S-type suffixes in lines 18-28.
 When scanning $sa[14]$, we see $x[3]$ is S-type and thus put ${\sf suf(3)}$ into the current rightmost empty position in ${\sf sa\_bkt_S}(i)$.
-Following the same, we get all the S-type suffixes sorted in line 29.
+Following the same, we get all the S-type suffixes sorted in line 28.
 
@@
 
-As stated below, S\ref{induction_phase:1}'-S\ref{induction_phase:3}' also constitute the sufficient and necessary conditions for a correct $sa$.
+As stated below, S\ref{induction_phase:1}''-S\ref{induction_phase:3}'' also constitute the sufficient and necessary conditions for a correct $sa$.
 
@@
 	(1) $sa^*$ is correct. \\
-	(2) $sa[i]$ is equal to the value calculated by SS\ref{induction_phase:1}-SS\ref{induction_phase:3}. \\
+	(2) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:3}''. \\
 
@@
 Suppose $sa^*$ is correct\footnote{The correctness of $sa^*$ can be checked by a sparse SA checker, such as~\cite{wu2017}.
-The checking processes for $sa^*$ and $sa$ can be executed in parallel.}, this lemma suggests a method to check $sa$ when it is induced from $sa^*$ during the induction phase.
-Specifically, in S\ref{induction_phase:2}' and S\ref{induction_phase:3}', each item induced into $sa$ will be scanned later to induce the order of its predecessor.
+It should be noticed that the checking processes for $sa^*$ and $sa$ can be executed in parallel.}, this lemma suggests a method to check $sa$ when it is induced from $sa^*$ during the induction phase.
+Specifically, in S\ref{induction_phase:2}'' and S\ref{induction_phase:3}'', each item induced into $sa$ will be scanned later to induce the order of its predecessor.
 If the induced and scanned values for each item are equal, then $sa$ is correct according to Lemma~\ref{lemma:3}.
 The problem to be solved here is that when an item is induced into a bucket, its value in $sa$ will not be scanned at once.
-Our solution for this is to integrate the building and checking processes into a whole, by ensuring the sequence of items induced into a bucket is identical to that scanned later.
-For the purpose, we increasingly compute the fingerprints of both sequences and check their equality at the end of the induction phase.
+One solution for this is to record the induced and scanned values separately during the building process and check their equality afterward.
+Our solution is to integrate the building and checking processes into a whole, by ensuring the sequence of items induced into a bucket is identical to that scanned later.
+For the purpose, we can use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end of induction phase.
 If the two fingerprints for each bucket are equal, then the second condition of Lemma~\ref{lemma:3} will be seen with a high probability.
-As a result, $sa$ can be built and probabilistically checked at the same time, for this we have Corollary~\ref{corollary:1}.
+As a result, $sa$ can be built and probabilistically checked at the same time.
+Following the previous discussion, we design a probabilistic algorithm based on  Corollary~\ref{corollary:1}.
 
@@
 
+We point out that Method A can be also applied to checking $lcp$ when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}.
+In general, this is done by comparing the sequences of induced and scanned items for all the buckets in $lcp$.
+
 \subsection{Method B}\label{sec:checkers:method_b}
 
-We introduce Lemma~\ref{lemma:1} before the presentation of Method B.
-This statement has been applied to deriving Theorem~\ref{theorem:1} for comparing two suffixes given the order of their successors already known.
+{\color{red} 1. $\overline{sa^*}$ is a subset of the final $sa$, what is $sa^*[i]$? These two symbols have not been defined clearly and are confusing.
 
-\begin{lemma} \label{lemma:1}
-	For $i, j \in [0, n)$ and $i \ne j$, ${\sf suf}(i) < {\sf suf}(j) \Leftrightarrow (x[i], {\sf suf}(i + 1)) < (x[j], {\sf suf}(j + 1))$.
-\end{lemma}
+2.
+Notice: The whole algorithm for induced sorting suffixes consists of many processes, the inducing process is only one of them.
+The checking takes place only at the top level by reusing the inducing process there.
+For the checking algorithm to work, it must be correctly implemented, this implies that the inducing process at the top level is also implicitly assumed to be correctly implemented.
+This needs to be specified/assumed for clarity.
+}
 
-Below we describe the idea of Method B and prove its correctness using Lemma~\ref{lemma:1} and Theorem~\ref{theorem:1}, where $\overline{sa^*}$ is retrieved from $sa$ and it records the starting positions of all the S*-type suffixes in their sorted order.
+{ \color{blue} The definition of $sa^*$ was introduced in Section~\ref{sec:preliminaries}. An example was given on the left for better understanding.}
 
+Our second Method is based on Theorem~\ref{theorem:2}, where $\overline{sa^*}$ is a subset of the final $sa$ for recording the starting positions of all the S*-type suffixes in $x$.
+{\color{blue}  We prove this theorem using Lemma~\ref{lemma:1} and Theorem~\ref{theorem:1}.
 \begin{theorem} \label{theorem:2}
 	For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied for all $i \in [0, n_1)$: \\
-	(1) $sa^*[i]$ is the starting position of an S*-type suffix and it differs from any other items in $sa^*$. \\
+	(1) $sa^*[i]$ is a permutation of all the S*-type suffixes and  $sa^*[0]=n-1$. \\
 	(2) $sa^*[i] = \overline{sa^*}[i]$. \\
@@
 
-	Suppose $sa$ is not a permutation of $[0, n)$.
-Without loss of generality, let $i \ne j$ and $sa[i] = sa[j]$, then $u \ne v$ and $sa[u] = sa[v]$ according Theorem~\ref{theorem:1}, where ${\sf suf}(sa[u])$ and ${\sf suf}(sa[v])$ are the successors of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[j])$, respectively.
-By recursively replacing $(i, j)$ with $(u, v)$, we can find two items $sa[i_0]$ and $sa[j_0]$ storing the starting position of the same S*-type suffix, indicating that $sa^*$ contains duplicate items.
-This violates condition (1) of Theorem~\ref{theorem:2}.
+	Case 1: Suppose all the S*-type suffixes are correctly sorted, $sa$ is correct by Lemma \ref{lemma:3}.
 
-	Suppose $i < j$ and ${\sf suf}(sa[i]) > {\sf suf}(sa[j])$.
-If $x[sa[i]] < x[sa[j]]$, then ${\sf suf}(sa[i]) < {\sf suf}(sa[j])$ according to Lemma~\ref{lemma:1}; otherwise, if $x[sa[i]] > x[sa[j]]$, then $i > j$ according to Theorem~\ref{theorem:1}.
-Both two conditions lead to a contradiction.
-Therefore, we must have $x[sa[i]] = x[sa[j]]$.
-If so, then $u < v$ and ${\sf suf}(u) > {\sf suf}(v)$, where ${\sf suf}(sa[u])$ and ${\sf suf}(sa[v])$ are the successors of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[j])$, respectively.
-By recursively replacing $(i, j)$ with $(u, v)$, we can find two elements $sa[i_0]$ and $sa[j_0]$ such that at least one stores the starting position of an S*-type suffix.
-We continue the proof following one of the three cases below:
-	\begin{enumerate}[(1)]
-	\item $x[sa[i_0]]$ is S*-type and $x[sa[j_0]]$ is L-type.
-In this case, $x[sa[i_0] + 1] > x[sa[j_0] + 1]$.
-Therefore, $u_0 > v_0$ according to Theorem~\ref{theorem:1}.
-This indicates that $(x[sa[i_0]], u_0) > (x[sa[j_0]], v_0)$ and thus $i_0 > j_0$ according to Theorem~\ref{theorem:1}.
-This leads to a contradiction.
-	\item $x[sa[i_0]]$ is L-type and $x[sa[j_0]]$ is S*-type.
-In this case, $x[sa[i_0] + 1] < x[sa[j_0] + 1]$.
-Therefore, ${\sf suf}(sa[u_0]) < {\sf suf}(sa[v_0])$ according to Lemma~\ref{lemma:1}.
-This indicates that $(x[sa[i_0]], {\sf suf}(sa[u_0])) < (x[sa[j_0]], {\sf suf}(sa[v_0])$ and thus ${\sf suf}(sa[i_0]) < {\sf suf}(sa[j_0])$ according to Lemma~\ref{lemma:1}.
-This leads to a contradiction.
-	\item $x[sa[i_0]]$ and $x[sa[j_0]]$ are both S*-type.
-Let $(i_0, j_0)$ be the pair that maximize $sa[i_0] + sa[j_0]$ such that $i_0 < j_0$ and ${\sf suf}(sa[i_0]) > {\sf suf}(sa[j_0])$, then $u_0 > v_0$ and ${\sf suf}(sa[u_0]) > {\sf suf}(sa[v_0])$.
-\footnote{Notice that $(u_0, v_0)$ must exist because the sentinel is a size-one S*-type substring smaller than others.} Because $x[i_0, u_0) = x[j_0, v_0)$, there must be $i_0 > j_0$ according to Theorem~\ref{theorem:1}, thus $sa^*$ and $\overline{sa^*}$ differ from each other.
-This violates condition (2) of Theorem~\ref{theorem:2}.
-	\end{enumerate}
-
+	Case 2: {\color{red} Notice that the inducing process at the top level is correctly implemented by Assumption xxx.} 
+	Suppose any two S*-type suffixes are not correctly sorted, i.e.
+${\sf suf}(sa^*[i_0])>{\sf suf}(sa^*[j_0])$ and $i_0<j_0$.
+By condition (2), we have ${\sf suf}(\overline{sa^*}[i_0])>{\sf suf}(\overline{sa^*}[j_0])$.
+Because the order of ${\sf suf}(\overline{sa^*}[i_0])$ and ${\sf suf}(\overline{sa^*}[j_0])$ is induced from ${\sf suf}(sa^*[i_1])$ and ${\sf suf}(sa^*[j_1])$, where ${\sf sub}(\overline{sa^*}[i_0], sa^*[i_1])$ and ${\sf sub}(\overline{sa^*}[j_0], sa^*[j_1])$ are two LMS-substrings, there must be ${\sf suf}(sa^*[i_1])>{\sf suf}(sa^*[j_1])$ and $i_1<j_1$.
+Repeating this reasoning process, because condition (1), we must see ${\sf suf}(sa^*[i_k])>{\sf suf}(sa^*[j_k])$ and ${\sf suf}(sa^*[i_{k+1}])<{\sf suf}(sa^*[j_{k+1}])$, where $i_k<j_k$ and $i_{k+1}<j_{k+1}$.
+However, given ${\sf suf}(sa^*[i_{k+1}])<{\sf suf}(sa^*[j_{k+1}])$, the inducing process will produce ${\sf suf}(\overline{sa^*}[i_k])<{\sf suf}(\overline{sa^*}[j_k])$ , which implies ${\sf suf}(sa^*[i_k])<{\sf suf}(sa^*[j_k])$ because condition (2), leading to a contradiction.
 \end{IEEEproof}
@@
 The first condition of Theorem~\ref{theorem:2} is naturally satisfied when computing $sa^*$ from $sa_1$.
-A naive method for checking the second condition is to keep a copy of $\overline{sa^*}$ when computing $sa$ from $sa^*$ and compare it with $sa^*$ afterward.
+A naive solution for checking the second condition is to keep a copy of $\overline{sa^*}$ when computing $sa$ from $sa^*$ and compare it with $sa^*$ afterward.
 This takes linear time, space and I/O overhead.
-We prefer to check equality of them by comparing their fingerprints, which can be calculated during the scan of these arrays in S\ref{induction_phase:1}' and S\ref{induction_phase:3}'.
+We prefer to check equality of them by comparing their fingerprints, which can be calculated during the scan of these arrays in S\ref{induction_phase:1}'' and S\ref{induction_phase:3}''.
 
@@
 	$sa[0, n)$ is the SA for the input string $x[0, n)$ with a high probability if the following conditions are satisfied for all $i \in [0, n_1)$: \\
-	(1) $sa^*[i]$ is the starting position of an S*-type suffix and it differs from any other items in $sa^*$. \\
+	(1) $sa^*[i]$ is a permutation of all the S*-type suffixes and  $sa^*[0]=n-1$. \\
 	(2) The fingerprints of $sa^*$ and $\overline{sa^*}$ are equal. \\
@@
 
-In Fig.~\ref{fig:example1}, both $sa^*$ and $\overline{sa^*}$ contain 4 items $14, 8, 5, 2$ arranged in the same order.
+As can be seen in Fig.~\ref{fig:example1}, both $sa^*$ and $\overline{sa^*}$ contain items $\{14, 8, 5, 2\}$ arranged in the same order.
 By using the fingerprinting function introduced in the next subsection, the fingerprints can be calculated in linear time using constant space.
+This implies another probabilistic algorithm designed by Method B.
 
-\subsection{Discussion}
+\subsection{Fingerprinting Function}
 
@@
 
-Methods A and B check equality of two integer arrays by comparing their fingerprints.
-In our programs, we choose to use the Karp-Rabin fingerprinting function to compute the fingerprints in need.
+Methods A and B check equality of two integer arrays by computing and comparing their fingerprints.
+In our programs, we select the Karp-Rabin fingerprinting function to compute the fingerprints in need.
 As depicted in Fig.~\ref{fig:example2}, the fingerprint of $A$ is calculated according to Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$.
 It should be noticed that two equal arrays must have an identical fingerprint, but the inverse is not always true.
-Fortunately, the probability of a false match can be reduced to a negligible level if $L$ is set to a large value.
+Fortunately, the error probability can be reduced to a negligible level by setting $L$ to a large value.
 
@@
 \end{formula}
-
-We also point out that Method A can check $lcp$ when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}.
-This is done by comparing the sequence of items induced into an $lcp$ bucket with that scanned from the same bucket.
-If the two sequences of each bucket are equal, then $lcp$ is correctly the LCP array for $sa$.
 
@@
 Notice that the new substring sorting and naming methods adopted by DSA-IS+ take effect when most of the S*-type substrings are short.
-From our experiments, given $D = 8$, the ratio of long S*-type substrings in the investigated corpus nearly approaches one hundred percent.
-Therefore,the proposed substring sorting and naming methods are practical for real-world datasets.
+From our experiments, given $D = 8$, the ratio of long S*-type substrings in the investigated corpus nearly approaches one hundred percent, indicating that these methods are practical for real-world datasets.
 
@@
 	}
-	\caption{A comparison of DSA-IS and eSAIS on various corpora in terms of peak disk usage, I/O volume and construction time.}
+	\caption{A comparison of Solutions A and B on various corpora in terms of peak disk usage, I/O volume and construction time, where $D = 4$.}
 	\label{fig:verification_performance}
@@
 Because the ratio of S*-type characters to all in $x$ is commonly one-third in real-world datasets, the checking process for $sa^*$ will not become the bottleneck for Method A.
-We also point out that this checking process can be parallelized with that for $sa$ to achieve a higher speed.
+To achieve a higher speed, we can also parallelize the checking processes for $sa$ and $sa^*$ using more computing resources.
+
+\subsection{Discussion}
+
+Rather than designing an I/O layer for efficient I/O operations, we currently use the containers provided by the STXXL library to perform reading, writing and sorting in external memory, these containers do not free the disk space for storing temporary data even if it is not needed any more, leading to a space consumption higher than our expectation.
+This is an implementation issue that can be solved by storing the data into multiple files and deleting each file when it is obsolete.
+In this way, our program can be further improved to achieve a space performance comparable to fSAIS.
 
@@
 
-In this paper, we made an attempt to improve the space performance of DSA-IS by new substring sorting and naming methods.
-For implementation convenience, we currently employ the STXXL's containers to perform reading, writing and sorting on the disk.
-The experimental results shows that our program for the adapted algorithm DSA-IS+ runs as fast as eSAIS and requires only half disk space as that for the latter on various real-world datasets.
-This program can be further optimized to approach the optimal space performance by means of the external-memory vector, sorter and priority queues supporting fine-grained disk space allocation and deallocation.
-
-We also proposed two methods that enable any IS suffix sorting algorithm to build and check SA simultaneously.
-The second method is rather lightweight in terms of that its time and space complexities are negligible compared with that of the existing IS suffix sorting/checking algorithms.
-We will describe in another paper a disk-based suffix sorting algorithm taking only $1n$ work space.
-By augmenting with this new suffix array builder, the proposed checking methods may potentially constitute the best solution for the situations where checking is a must after building.
+In this paper, we proposed two methods that enable any IS suffix sorting algorithm to build and check SA simultaneously.
+In particular, the probabilistic algorithm designed by the second method is rather lightweight in terms of that it takes negligible time and space compared with the existing IS suffix sorting and checking algorithms.
+We also made an attempt to improve the space performance of DSA-IS using new substring sorting and naming methods.
+The experimental results shows that our program for the adapted algorithm DSA-IS+ runs as fast as that for eSAIS and requires only half as much disk space as the latter on various real-world datasets.
+We also showed that the space performance of this program can be further optimized using some optimization techniques.
+In our next paper, we will propose another disk-based suffix sorting algorithm that takes only $1n$ work space to run.
 
@@
 \end{document}
-
-A series of experiments are conducted to investigate the effect of $D_1$ and $D_2$ on the performance of DSA-IS+.
-As shown in Figure~\ref{fig:performance_analysis3}, the time consumption of DSA-IS+ gets smaller as $D_2$ becomes larger, while the peak disk usage and I/O volume are almost unchanged.
-Furthermore, the data of Table~\ref{tbl:effect_of_D} indicates that a small variation on the values of $D_1$ and $D_2$ can incur a significant fluctuation on the speed of DSA-IS+.
-For instance, when changing $(D_1, D_2)$ from $(8, 10)$ to $(16, 16)$, the construction time for the uniprot dataset decreases from 2.64 to 2.24 microseconds per character.
-This is because the proportions of long LMS substrings of $x$ and $\{x1, x2,...\}$ decrease when $D_1$ and $D_2$ increase.
-Recall that, after partitioning $x$ into blocks, DSA-IS+ sorts the long LMS substrings of each block in RAM and merges them by an external memory sorter, which is time-consuming when the number of long LMS substrings is large.
-In our program, the ratio of long and short LMS substrings can be controlled by adjusting the values of $D_1$ and $D_2$.
-However, when $D_1$ and $D_2$ get larger, the min heap also takes more time to merge the sorted short and long LMS substrings, as each string comparison takes $\mathcal{O}(D_1)$ and $\mathcal{O}(D_2)$ time in the first and recursion levels, respectively.
-Fortunately, Table~\ref{tbl:long_short_distribution} shows that the majority of LMS substrings in real datasets are considerably short.
-In practice, we can set $D_1$ and $D_2$ according to the statistics of the length distribution of LMS substrings, where the statistics can be collected when partitioning $x$ at the beginning of the reduction phase.
-
-\begin{table*}[htbp]
-	\caption{Effects of $D_1$ and $D_2$ for DSA-IS+}
-	\label{tbl:effect_of_D}
-	\centering
-	\begin{tabular}{|c|c|c|c|c|c|c|c|c||c|c|c|c|c|}
-		\hline
-		\multicolumn{1}{|c}{} & \multicolumn{3}{|c}{eSAIS} & \multicolumn{10}{|c|}{DSA-IS+}\\\hline
-		\hline
-		Corpora & PDU & IOV & CT & $D_1$ & $D_2$ & PDU & IOV & CT  & $D_1$ & $D_2$ & PDU & IOV & CT\\\hline
-		uniprot & 22.71 & 162.01 & 2.50 & 8 & 10 & 14.27 & 146.10 & 2.64 & 16 & 16 & 14.27 & 144.90 & 2.24 \\\hline
-		proteins & 24.09 & 172.29 & 2.33 & 8 & 10 & 16.12 & 147.92 & 2.28 & 20 & 12 & 16.12 & 147.85 & 2.14 \\\hline
-		genome & 22.64 & 157.41 & 2.15 & 8 & 10 & 15.34 & 142.05 & 2.75 & 20 & 12 & 15.34 & 140.89 & 2.25 \\\hline
-	\end{tabular}
-\end{table*}%
-
-\begin{figure}[htbp]
-	\centering
-	\subfigure{
-		\centering
-		\includegraphics[width=0.9\columnwidth]{pdu_cmp3}
-		\label{subfig:pdu_cmp3}
-	}
-	\hfil
-	\subfigure{
-		\centering
-		\includegraphics[width=0.9\columnwidth]{iov_cmp3}
-		\label{subfig:iov_cmp3}
-	}
-	\hfil
-	\subfigure{
-		\centering
-		\includegraphics[width=0.9\columnwidth]{ct_cmp3}
-		\label{subfig:ct_cmp3}
-	}
-	\caption{Experimental results for DSA-IS+ and eSAIS on guten in terms of peak disk usage, I/O volume and construction time, where $D_1 = 8$, $D_2$ ranges in \{6, 8, 10, 12\} and the input size varies in \{1, 2, 4, 8, 14\} GiB.
-}
-	\label{fig:performance_analysis3}
-\end{figure}
-
-\begin{table*}[htbp]
-	\caption{Statistics on the Length Distribution of Long and Short LMS Substrings for DSA-IS+}
-	\label{tbl:long_short_distribution}	
-	\centering
-	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
-		\hline
-		\multicolumn{11}{|c|}{level 0} \\\hline
-		Corpora & $D_1$ & long & short & total & ratio & $D_1$ & long & short & total & ratio\\\hline
-		uniprot & 8 & 30243605 & 785997413 & 816241018 & 0.963 & 16 & 3850755 & 812390263 & 816241018 & 0.999\\\hline
-		proteins & 8 & 1687673 & 377404329 & 379092002 & 0.999 & 20 & 10930 & 379081072 & 379092002 & 0.999\\\hline
-		genome & 8 & 19446841 & 773670266 & 793117107 & 0.975 & 20 & 509998 & 792607109 & 793117107 & 0.999\\\hline
-		\multicolumn{11}{|c|}{level 1}\\\hline
-		Corpora & $D_2$ & long & short & total & ratio & $D_2$ & long & short & total & ratio\\\hline
-		uniprot & 10 & 11849 & 263748775 & 263760624 & 0.995 & 16 & 380 & 263760244 & 263760624 & 0.999\\\hline
-		proteins & 10 & 18126 & 123772731 & 123790857 & 0.996 & 20 & 2937 & 123787920 & 123790857 & 0.999\\\hline
-		genome & 10 & 311607 & 248243216 & 248554823 & 0.999 & 20 & 153845 & 248400978 & 248554823 & 0.999\\\hline
-	\end{tabular}
-\end{table*}%
-
-\subsection{Checking Performance}
-
-\begin{table*}[htbp]
-	\caption{Overall Performance of DSA-IS+ and eSAIS with CM1 and CM2 on enwiki\_8G}
-	\label{tbl:check_overhead}
-	\centering
-	\begin{tabular}{|c|c|c|c|c|c|c|}
-		\hline
-		& \multicolumn{3}{|c|}{DSA-IS+ ($D_1 = 8$, $D_2 = 10$)} & \multicolumn{3}{c|}{eSAIS}\\\hline
-		Checking Method & PDU & IOV & CT & PDU & IOV & CT \\\hline
-		CM1 & 17.93  & 180.01 & 3.10 & - & - & - \\\hline
-		CM2 & 26.00 & 231.68 & 4.00 & 27.00 & 246.93 & 3.74 \\\hline
-		no check & 17.93  & 173.67	& 3.17 & 23.85 & 200.93 & 3.10 \\\hline
-	\end{tabular}
-\end{table*}%
-
-We make a performance comparison of the proposed checking method and the work presented in~\cite{Karkkainen2003}, which are denoted by CM1 and CM2, respectively.
-For performance evaluation, we integrate CM1 into the program of DSA-IS+ following Section~\ref{subsec:sachecker:implementation} and reuse the implementation for CM2 in the program of eSAIS to verify the outputs of DSA-IS+ and eSAIS.
-Table~\ref{tbl:check_overhead} gives a glimpse of the performance overhead for the two checking methods.
-It can be seen that, when checking the suffix array of enwiki\_8G by using CM2, both time and I/O volume for verification are about one-fifth of that for construction.
-On the other hand, CM1 has almost no negative impact on the overall performance.
-As demonstrated in lines 3 and 5, the peak disk usage for CM1 is no more than that for DSA-IS+ and the increase in the I/O volume can be ignored.
-It worth mentioning that the data of the running time for the combination of DSA-IS+ and CM1 is a bit shorter than that for the plain DSA-IS+.
-This is a normal phenomenon due to the fluctuating I/O performance, which in turn indicates that the time overhead consumed by CM1 is also negligible.

### 2017-08-05 new-draft/draft.tex
2017/08/05

@@
 We propose two methods that enable any IS suffix sorter to build and check an SA simultaneously.
-The first method can check both suffix and longest common prefix~(LCP) arrays for strings drawn from constant alphabets, while the second method is applicable to checking SA for strings of any alphabets.
+The first method can check both suffix and longest common prefix~(LCP) arrays, while the second method is applicable to checking SA only.
 By combining our methods with the Karp-Rabin fingerprinting function, we design two probablistic algorithms that perform verification with a neligible error probability.
@@
 Particularly, the algorithm designed by the second method only takes linear time and constant space when running in external memory.
-For a further study, we integrate the checking algorithm designed by the second method into the existing suffix sorter DSA-IS to evaluate the checking overhead, where the time, space and I/O consumptions for verification are considerably smaller than that for construction in the experiments.
+For a further study, we integrate the checking algorithm designed by the second method into the existing suffix sorting algorithm DSA-IS to evaluate the checking overhead, where the time, space and I/O consumptions for verification are considerably smaller than that for construction in the experiments.
 This convinces us that the proposed checking methods, together with the state-of-the-art IS suffix sorters, could constitute efficient solutions for both construction and verification.
@@
 
-The software packages for DC3 and eSAIS provide users a checker to ensure the correctness of their outputs. The following statements describe the idea behind this checker.
+\color{red}{The software packages for DC3 and eSAIS provide users a checker to ensure the correctness of their outputs.
+Below we describe the main idea behind this checker and refer interested readers to ~\cite{Dementiev2008a} for more details.}
 
 \begin{lemma} \label{lemma:1}
-	For $i, j \in [0, n)$ and $i \ne j$, ${\sf suf}(i) < {\sf suf}(j) \Leftrightarrow (x[i], {\sf suf}(i + 1)) < (x[j], {\sf suf}(j + 1))$.
-\end{lemma}
-
-\begin{theorem} \label{theorem:1}
 	$sa[0, n)$ is the SA for $x[0, n)$ if and only if the following conditions are satisfied:\\
@@
 \\
-\end{theorem}
+\end{lemma}
 
-\begin{IEEEproof}
-	Theorem~\ref{theorem:1} is derived from Lemma~\ref{lemma:1}. The proof is available in~\cite{Dementiev2008a}.
+\begin{IEEEproof}The first condition indicates that $sa$ is a permutation of all the suffixes in $x$.
+The second condition indicates that the order of suffixes in $sa$ corresponds to that of their heading characters and successors.
+Because a comparison between any two suffixes can be done by comparing their heading characters and successors, the above conditions are sufficient and necessary for SA verification.
+
 \end{IEEEproof}
 
-This checker performs two passes of integer sorts and each sort involves $n$ fixed-size tuples.
-When implemented in external memory, it consumes large disk space and I/O volume, resulting in a high checking overhead.
-For example, we can observe from Section~\ref{sec:experiments} that, the peak disk use and the I/O volume for checking an SA encoded by 40-bit integers are around $26n$ and $53n$, respectively.
-In this section,  we describe two SA checking methods based on the IS principle, both of them can be seamlessly integrated into any IS suffix sorting algorithm to build and check an SA at the same time.
+The disk-based implementation of this checker conducts two passes of integer sorts and each sort arranges the order of $n$ fixed-size tuples using external memory.
+As can be observed from Section~\ref{sec:experiments}, the peak disk use and the I/O volume for an SA encoded by 40-bit integers are around $26n$ and $53n$, respectively.
 
-\subsection{Method A} \label{sec:checkers:method_a}
+\subsection{Proposals} \label{sec:checkers:proposals}
 
-Given that each S*-type substring is unique in $x$, SA-IS first computes $sa^*$ by sorting the starting positions of all the S*-type suffixes with their ranks indicated by $sa_1$ and then induces $sa$ from $sa^*$ by S\ref{induction_phase:1}''-S\ref{induction_phase:3}'' during the inducting phase.
+Recall that, an IS suffix sorting algorithm recursively perform the reduction phase to compute the reduced string $s_1$ until $s_1$ contains no duplicate characters.
+Afterward, it produces $sa_1$ from $s_1$ and recursively perform the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase executes mainly consists of the following steps:
 
 \begin{enumerate}[S1'']
+	\item sort the starting positions of all the S*-type suffixes with their ranks indicated by $sa_1$ to produce $sa^*$ \label{induction_phase:1}.
 
@@
 For each scanned item $sa^*[i]$, insert it into the rightmost empty position of ${\sf sa\_bkt_S}(x[sa^*[i]])$.
-\label{induction_phase:1}
+\label{induction_phase:2}
 
 	\item Scan $sa$ rightward with $i$ increasing from $0$ to $n - 1$.
-For each scanned non-empty item $sa[i]$, insert $sa[i] - 1$ into the leftmost empty position of ${\sf sa\_bkt_L}(x[sa[i] - 1])$ if $t[sa[i] - 1] = 0$.\label{induction_phase:2}
+For each scanned non-empty item $sa[i]$, insert $sa[i] - 1$ into the leftmost empty position of ${\sf sa\_bkt_L}(x[sa[i] - 1])$ if $t[sa[i] - 1] = 0$.\label{induction_phase:3}
 
@@
 Scan $sa$ leftward with $i$ decreasing from $n - 1$ to $0$.
-For each scanned non-empty item $sa[i]$, insert $sa[i] - 1$ into the rightmost empty position of ${\sf sa\_bkt_S}(x[sa[i] - 1])$ if $t[sa[i] - 1] = 1$.\label{induction_phase:3}
+For each scanned non-empty item $sa[i]$, insert $sa[i] - 1$ into the rightmost empty position of ${\sf sa\_bkt_S}(x[sa[i] - 1])$ if $t[sa[i] - 1] = 1$.\label{induction_phase:4}
 
@@
 
-We illustrate a running example of the above process in Fig.~\ref{fig:example1} and show more details for a better understanding.
-In line 6, we find the ending position of each bucket and insert the sorted S*-type suffixes into the right part of the corresponding buckets.
-Afterward, we first find the heading position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes as following.
-In lines 8-9, we check $\sf suf(14)$ (marked by the symbol $@$) and find its predecessor ${\sf suf(13)}$ is L-type, thus we put ${\sf suf(13)}$ into the current leftmost empty position in ${\sf sa\_bkt_L}(i)$.
+A running example of S\ref{induction_phase:2}-Sp\ref{induction_phase:4} is shown in Fig.~\ref{fig:example1}.
+Given $sa^*$ is already known, line 6 inserts each S*-type suffix into the right part of the corresponding bucket and arranges them according to their sorted order indicated by $sa^*$.
+For example, $\sf suf(8)$, $\sf suf(5)$, $\sf suf(2)$ are placed into the right part of $\sf sa\_bkt(i)$ and arranged in their sorted order indicated by $sa^*$.
+Then, we find the leftmost position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes.
+To do this, we first check $sa[0] = 14$ (marked by the symbol @) and induce the predecessor of $\sf suf(14)$ in lines 8-9.
+Because $x[13] = i$ is L-type, we put $\sf suf(13)$ into the current leftmost empty position in $\sf sa\_bkt_L(i)$.
 To step through $sa$ in this way, we get all the L-type suffixes sorted in line 17.
-After that, we find the ending position of each bucket and scan $sa$ leftward for inducing the order of S-type suffixes in lines 18-28.
-When scanning $sa[14]$, we see $x[3]$ is S-type and thus put ${\sf suf(3)}$ into the current rightmost empty position in ${\sf sa\_bkt_S}(i)$.
-Following the same, we get all the S-type suffixes sorted in line 28.
+Afterward, we find the rightmost position of each bucket and scan $sa$ leftward for inducing the order of S-type suffixes.
+When scanning $sa[14] = 4$ in lines 19-20, we see $x[3] = i$ is S-type and thus put $\sf suf(3)$ into the current rightmost empty position in $\sf sa\_bkt_S(i)$.
+Following the same idea, we get all the S-type suffixes sorted in line 28.
+Based on the discussion, we show in Lemma~\ref{lemma:2} a set of sufficient and necessary conditions for SA verification.
 
@@
 
-As stated below, S\ref{induction_phase:1}''-S\ref{induction_phase:3}'' also constitute the sufficient and necessary conditions for a correct $sa$.
+\begin{lemma} \label{lemma:2}
+    $sa[0, n)$ is the SA for $x[0, n)$ if and only if the following conditions are satisfied: \\
+    (1) $sa^*$ is correctly computed. \\
+    (2) S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented. \\
+\end{lemma}
+
+\begin{IEEEproof} Given two suffixes placed at $sa[i]$ and $sa[j]$ and their successors placed at $sa[p]$ and $sa[q]$, we prove the statement as follows.
+
+Because S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented, $i < j \iff (x[sa[i]], p) < (x[sa[j]], q)$. This satisfies the second condition of Lemma~\ref{lemma:1}.
+
+Further, suppose $sa[i] = sa[j]$ and $i \ne j$, then $sa[p] = sa[q]$ and $p \ne q$.
+Repeatedly replacing $(i, j)$ with $(p, q)$ until $\sf suf(sa[i])$ is S*-type, then $sa^*$ contain duplicate elements.
+Because $sa^*$ is correctly computed, each element in $sa^*$ must be unique.
+This leads to a contradiction.
+\end{IEEEproof}
+
+The above lemma indicates that, if the first condition of Lemma~\ref{lemma:2} is true, we can check the correctness of $sa^*$ instead of $sa$ at the top recursion level to perform SA verification.
+It should be noticed that the code snippet for the induction phase at the top recursion level only consists of tens of C++ code lines.
+This is considerable smaller than the whole program for a disk-based suffix sorting algorithm.
+Following the idea, we assume S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented and propose two methods for checking computation errors caused by implementation bugs and I/O errors caused by hardware malfunctions.
+The algorithms designed by these methods can be seamlessly integrated into any IS suffix sorting algorithms for building and checking an SA at the same time.
+
+\subsubsection{Method A} \label{sec:proposals:method_a}
+
+Our first method is directly extended from Lemma~\ref{lemma:2}.
 
 \begin{lemma} \label{lemma:3}
-	For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied: \\ 
+	Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied: \\
 	(1) $sa^*$ is correct. \\
-	(2) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:3}''. \\
+	(2) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}''. \\
 
@@
 
-Suppose $sa^*$ is correct\footnote{The correctness of $sa^*$ can be checked by a sparse SA checker, such as~\cite{wu2017}.
-It should be noticed that the checking processes for $sa^*$ and $sa$ can be executed in parallel.}, this lemma suggests a method to check $sa$ when it is induced from $sa^*$ during the induction phase.
-Specifically, in S\ref{induction_phase:2}'' and S\ref{induction_phase:3}'', each item induced into $sa$ will be scanned later to induce the order of its predecessor.
-If the induced and scanned values for each item are equal, then $sa$ is correct according to Lemma~\ref{lemma:3}.
-The problem to be solved here is that when an item is induced into a bucket, its value in $sa$ will not be scanned at once.
-One solution for this is to record the induced and scanned values separately during the building process and check their equality afterward.
-Our solution is to integrate the building and checking processes into a whole, by ensuring the sequence of items induced into a bucket is identical to that scanned later.
-For the purpose, we can use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end of induction phase.
+The first condition of Lemma~\ref{lemma:3} can be checked by a sparse SA checker like~\cite{wu2017}.
+Notice that each suffix induced into $sa$ will be latter scanned for inducing the order of its predecessor during S\ref{induction_phase:2}'' and S\ref{induction_phase:4}''.
+The second condition intends for detecting I/O errors and can be checked by determining whether the induced and scanned values for each suffix are equal.
+The problem to be solved here is that when a suffix is induced into a bucket, its corresponding value in $sa$ may not be scanned at once.
+Our solution is to integrate the checking process into the building process, by ensuring the sequence of values placed at a bucket is identical to that scanned later.
+For the purpose, we use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end of the induction phase.
 If the two fingerprints for each bucket are equal, then the second condition of Lemma~\ref{lemma:3} will be seen with a high probability.
 As a result, $sa$ can be built and probabilistically checked at the same time.
-Following the previous discussion, we design a probabilistic algorithm based on  Corollary~\ref{corollary:1}.
+It should be noticed that this method can be also applied to checking the longest common prefix array when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}.
+As a result, we design Algorithm~\ref{alg:2} to probabilistically check $sa$ and $lcp$ during the induction phase at the top recursion level.
 
-\begin{corollary} \label{corollary:1}
-	$sa[0, n)$ is the SA for the input string $x[0, n)$ with a high probability if the following conditions are satisfied for all $c \in \Sigma$: \\
-	(1) $sa^*$ is correct.\\
-	(2) The fingerprints of ${\sf sa_{I1}}(c)$ and ${\sf sa_{S1}}(c)$ are equal. \\
-	(3) The fingerprints of ${\sf sa_{I2}}(c)$ and ${\sf sa_{S2}}(c)$ are equal. \\
+\SetKwProg{Fn}{Function}{}{}
 
-\end{corollary}
+\begin{algorithm*}
 
-Notice that ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$.
+	\caption{The Algorithm Based on Lemma~\ref{lemma:3}.}
 
-We point out that Method A can be also applied to checking $lcp$ when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}.
-In general, this is done by comparing the sequences of induced and scanned items for all the buckets in $lcp$.
+	\label{alg:2}
 
-\subsection{Method B}\label{sec:checkers:method_b}
+	%\SetAlgoNoLine
+	\Fn{{\sf CheckByMethodA}}{
 
-{\color{red} 1. $\overline{sa^*}$ is a subset of the final $sa$, what is $sa^*[i]$? These two symbols have not been defined clearly and are confusing.
+        Check the correctness of $sa^*$ and $lcp^*$ by the method proposed in~\cite{wu2017}. \\
 
-2.
-Notice: The whole algorithm for induced sorting suffixes consists of many processes, the inducing process is only one of them.
-The checking takes place only at the top level by reusing the inducing process there.
-For the checking algorithm to work, it must be correctly implemented, this implies that the inducing process at the top level is also implicitly assumed to be correctly implemented.
-This needs to be specified/assumed for clarity.
+        Compute the fingerprints of ${\sf sa_{I1}}(c)$, ${\sf sa_{S1}}(c)$, ${\sf lcp_{I1}}(c)$ and ${\sf lcp_{S1}}(c)$ when inducing the order and the LCP-values of L-type suffixes. \\
+
+        Check if ${\sf sa_{I1}}(c) = {\sf sa_{S1}}(c)$ and ${\sf lcp_{I1}}(c) = {\sf lcp_{S1}}(c)$. \\
+
+        Compute the fingerprints of ${\sf sa_{I2}}(c)$, ${\sf sa_{S2}}(c)$, ${\sf lcp_{I2}}(c)$ and ${\sf lcp_{S2}}(c)$ when inducing the order and the LCP-values of S-type suffixes. \\
+
+        Check if ${\sf sa_{I2}}(c) = {\sf sa_{S2}}(c)$, ${\sf lcp_{I2}}(c) = {\sf lcp_{S2}}(c)$. \\
+	}
+	\end{algorithm*}
+
+\subsubsection{Method B}\label{sec:proposals:method_b}
+
+Method B uses a different way to check the correctness of $sa^*$ in Lemma~\ref{lemma:4}.
+Before our presentation, we first introduce a notation called $\overline{sa^*}$.
+The same as $sa^*$, $\overline{sa^*}$ also represents the suffix array for all the S*-type suffixes.
+Both of them are computed during the induction phase at the top recursion level of an IS suffix sorting algorithm.
+To be specific, $sa^*$ is computed from $sa_1$ and $\overline{sa^*}$ is retrieved from the constructed $sa$.
+
+{\color{red}
+\begin{lemma} \label{lemma:4}
+    Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied for all $i \in [0, n_1)$:
+
+	(1) $sa^*[i]$ is a permutation of all the S*-type suffixes. \\
+	(2) $sa^*[i] = \overline{sa^*}[i]$. \\
+	(3) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}''. \\
+
+\end{lemma}
+
+\begin{IEEEproof}
+
+    We only prove the sufficiency as the necessity is clear.
+
+    Case 1: Suppose all the S*-type suffixes are correctly sorted in $sa^*$, then $sa$ is right according to Lemma~\ref{lemma:3}.
+
+    Case 2: Suppose any two S*-type suffixes are not corrected sorted, i.e., $\sf suf(sa^*[i_0]) > \sf suf(sa^*[j_0])$ and $i_0 < j_0$.
+By condition (2), we have $\sf suf(\overline{sa^*}[i_0] > \sf suf(\overline{sa^*}[j_0])$.
+Given that the order of $\sf suf(\overline{sa^*}[i_0]$ and $\sf suf(\overline{sa^*}[j_0])$ are induced from ${\sf suf}(sa^*[i_1])$ and ${\sf suf}(sa^*[j_1])$, then ${\sf sub}(\overline{sa^*}[i_0], sa^*[i_1])$ and ${\sf sub}(\overline{sa^*}[j_0], sa^*[j_1])$ are two LMS-substrings and there must be ${\sf suf}(sa^*[i_1])>{\sf suf}(sa^*[j_1])$ and $i_1 < j_1$.
+Repeating this reasoning process, because condition (1), we must see ${\sf suf}(sa^*[i_k])>{\sf suf}(sa^*[j_k])$ and ${\sf suf}(sa^*[i_{k+1}])<{\sf suf}(sa^*[j_{k+1}])$, where $i_k < j_k$ and $i_{k+1} < j_{k+1}$.
+However, given ${\sf suf}(sa^*[i_{k+1}])<{\sf suf}(sa^*[j_{k+1}])$, the inducing process will produce ${\sf suf}(\overline{sa^*}[i_k])<{\sf suf}(\overline{sa^*}[j_k])$ , which implies ${\sf suf}(sa^*[i_k])<{\sf suf}(sa^*[j_k])$ because condition (2), leading to a contradiction.
+
+\end{IEEEproof}
 }
 
-{ \color{blue} The definition of $sa^*$ was introduced in Section~\ref{sec:preliminaries}. An example was given on the left for better understanding.}
+The first condition of Lemma~\ref{lemma:4} is naturally satisfied when computing $sa^*$ from $sa_1$.
+We reuse the fingerprinting function to probabilistically check the third condition.
+This method is also applied to checking the second condition by computing and comparing the fingerprints of $sa^*$ and $\overline{sa^*}$.
+Specifically, the fingerprint of $sa^*$ is calculated when it is being scanned in S$\ref{induction_phase:1}''$.
+Because $\overline{sa^*}$ is a subset of $sa$, the fingerprint of $\overline{sa^*}$ is calculated when the S-type suffixes are being induced in S$\ref{induction_phase:4}''$.
 
-Our second Method is based on Theorem~\ref{theorem:2}, where $\overline{sa^*}$ is a subset of the final $sa$ for recording the starting positions of all the S*-type suffixes in $x$.
-{\color{blue}  We prove this theorem using Lemma~\ref{lemma:1} and Theorem~\ref{theorem:1}.
-\begin{theorem} \label{theorem:2}
-	For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied for all $i \in [0, n_1)$: \\
-	(1) $sa^*[i]$ is a permutation of all the S*-type suffixes and  $sa^*[0]=n-1$. \\
-	(2) $sa^*[i] = \overline{sa^*}[i]$. \\
-\end{theorem}
+\SetKwProg{Fn}{Function}{}{}
 
-\begin{IEEEproof}
-	We only prove the sufficiency as the necessity is clear.
+\begin{algorithm*}
 
-	Case 1: Suppose all the S*-type suffixes are correctly sorted, $sa$ is correct by Lemma \ref{lemma:3}.
+	\caption{The Algorithm Based on Lemma~\ref{lemma:4}.}
 
-	Case 2: {\color{red} Notice that the inducing process at the top level is correctly implemented by Assumption xxx.} 
-	Suppose any two S*-type suffixes are not correctly sorted, i.e.
-${\sf suf}(sa^*[i_0])>{\sf suf}(sa^*[j_0])$ and $i_0<j_0$.
-By condition (2), we have ${\sf suf}(\overline{sa^*}[i_0])>{\sf suf}(\overline{sa^*}[j_0])$.
-Because the order of ${\sf suf}(\overline{sa^*}[i_0])$ and ${\sf suf}(\overline{sa^*}[j_0])$ is induced from ${\sf suf}(sa^*[i_1])$ and ${\sf suf}(sa^*[j_1])$, where ${\sf sub}(\overline{sa^*}[i_0], sa^*[i_1])$ and ${\sf sub}(\overline{sa^*}[j_0], sa^*[j_1])$ are two LMS-substrings, there must be ${\sf suf}(sa^*[i_1])>{\sf suf}(sa^*[j_1])$ and $i_1<j_1$.
-Repeating this reasoning process, because condition (1), we must see ${\sf suf}(sa^*[i_k])>{\sf suf}(sa^*[j_k])$ and ${\sf suf}(sa^*[i_{k+1}])<{\sf suf}(sa^*[j_{k+1}])$, where $i_k<j_k$ and $i_{k+1}<j_{k+1}$.
-However, given ${\sf suf}(sa^*[i_{k+1}])<{\sf suf}(sa^*[j_{k+1}])$, the inducing process will produce ${\sf suf}(\overline{sa^*}[i_k])<{\sf suf}(\overline{sa^*}[j_k])$ , which implies ${\sf suf}(sa^*[i_k])<{\sf suf}(sa^*[j_k])$ because condition (2), leading to a contradiction.
-\end{IEEEproof}
+	\label{alg:3}
 
-The first condition of Theorem~\ref{theorem:2} is naturally satisfied when computing $sa^*$ from $sa_1$.
-A naive solution for checking the second condition is to keep a copy of $\overline{sa^*}$ when computing $sa$ from $sa^*$ and compare it with $sa^*$ afterward.
-This takes linear time, space and I/O overhead.
-We prefer to check equality of them by comparing their fingerprints, which can be calculated during the scan of these arrays in S\ref{induction_phase:1}'' and S\ref{induction_phase:3}''.
+	%\SetAlgoNoLine
+	\Fn{{\sf CheckByMethodB}}{
 
-\begin{corollary} \label{corollary:2}
-	$sa[0, n)$ is the SA for the input string $x[0, n)$ with a high probability if the following conditions are satisfied for all $i \in [0, n_1)$: \\
-	(1) $sa^*[i]$ is a permutation of all the S*-type suffixes and  $sa^*[0]=n-1$. \\
-	(2) The fingerprints of $sa^*$ and $\overline{sa^*}$ are equal. \\
-\end{corollary}
+        Compute the fingerprints of $sa^*$, ${\sf sa_{I1}}(c)$ and ${\sf sa_{S1}}(c)$ when inducing the order and the LCP-values of L-type suffixes. \\
 
-As can be seen in Fig.~\ref{fig:example1}, both $sa^*$ and $\overline{sa^*}$ contain items $\{14, 8, 5, 2\}$ arranged in the same order.
-By using the fingerprinting function introduced in the next subsection, the fingerprints can be calculated in linear time using constant space.
-This implies another probabilistic algorithm designed by Method B.
+        Check if ${\sf sa_{I1}}(c) = {\sf sa_{S1}}(c)$ and ${\sf lcp_{I1}}(c) = {\sf lcp_{S1}}(c)$ by comparing their fingerprints. \\
+
+        Compute the fingerprints of $\overline{sa^*}$, ${\sf sa_{I2}}(c)$ and ${\sf sa_{S2}}(c)$ when inducing the order and the LCP-values of S-type suffixes. \\
+
+        Check if ${\sf sa_{I2}}(c) = {\sf sa_{S2}}(c)$p by comparing their fingerprints. \\
+
+        Check if $sa^* = \overline{sa^*}$ by comparing their fingerprints. \\
+	}
+\end{algorithm*}
 
 \subsection{Fingerprinting Function}
+
+\begin{formula} \label{formula:1}${\sf fp}(A[0, -1]) = 0$.
+\end{formula}
+
+\begin{formula} \label{formula:2}${\sf fp}(A[0, i]) = {\sf fp}(A[0, i - 1]) \cdot {\delta} + A[i]\mod L$ for $i \ge 0$.
+\end{formula}
+
+The proposed two methods check equality of two arrays by comparing their fingerprints, which can be calculated using a hash function.
+Notice that two equal arrays must share an identical hash value, but the inverse is not always true.
+To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting function to compute the required values following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $delta$ is an integer randomly chosen from $[1, L)$.
+By setting $L$ to a large value, the error probability can be reduced to a negligible level.
+In Fig.~\ref{fig:example2}, we depict an example for computing the integer array $A$, where $A$ is identical to $sa^*$ and $\overline{sa^*}$ in Fig.~\ref{fig:example1}.
+Given $L = 23$ and $\delta = 5$, we first compute the fingerprint of $A[0, 0]$ in line 4.
+Because $\sf fp(A[0, -1]) = 0$, $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$.
+By iteratively computing the fingerprints of the prefixes of $A$ according to Formula~\ref{formula:2}, we finally obtain the fingerprints of the whole array in line 7.
 
@@
 
-	\includegraphics[width = 0.9\columnwidth]{example2.pdf}
+	\includegraphics[width = 1\columnwidth]{example2.pdf}
 
@@
 
-Methods A and B check equality of two integer arrays by computing and comparing their fingerprints.
-In our programs, we select the Karp-Rabin fingerprinting function to compute the fingerprints in need.
-As depicted in Fig.~\ref{fig:example2}, the fingerprint of $A$ is calculated according to Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$.
-It should be noticed that two equal arrays must have an identical fingerprint, but the inverse is not always true.
-Fortunately, the error probability can be reduced to a negligible level by setting $L$ to a large value.
+\subsection{Complexity Analysis}
 
-\begin{formula} \label{formula:1}${\sf fp}(A[0, -1]) = 0$. 
-\end{formula}
-
-\begin{formula} \label{formula:2}${\sf fp}(A[0, i]) = {\sf fp}(A[0, i - 1]) \cdot {\delta} + A[i]\mod L$ for $i \ge 0$.
-\end{formula}
+It takes linear time and constant RAM space for computing the required fingerprints in Algorithm~\ref{alg:2} and Algorithm~\ref{alg:3}.
+Further, the bottleneck of Algorithm~\ref{alg:2} lies in checking $sa^*$ following the idea presented in~\cite{wu2017}.
+Specifically, given two suffixes, say $\sf suf(i)$ and $\sf suf(j)$, and their LCP-value, say $d$, we compute and compare the fingerprints of $\sf sub(i, i + d - 1)$ and $\sf sub(j, j + d - 1)$.
+If the fingerprints are equal and the order of $x[i + d]$ and $x[j + d]$ correspond to that of $\sf suf(i)$ and $\sf suf(j)$, it is believed that the suffixes are correctly sorted and their LCP-value is right.
+When implemented in external memory, the checking processes for any number of suffix pairs can be done in batch within a sorting complexity.
+Because at most one out of every two suffixes is S*-type (typically one-third in real-world data sets) and the checking process is only executed at the top recursion level, the checking consumption taken by Algorithm~\ref{alg:2} is much less than the building consumption taken by any disk-based IS suffix sorting algorithm.
 

### 2017-08-05 new-draft/draft.tex
2017/08/05

@@
 We propose two methods that enable any IS suffix sorter to build and check an SA simultaneously.
-The first method can check both suffix and longest common prefix~(LCP) arrays, while the second method is applicable to checking SA only.
-By combining our methods with the Karp-Rabin fingerprinting function, we design two probablistic algorithms that perform verification with a neligible error probability.
-These algorithms are rather lightweight in terms of time and space.
-Particularly, the algorithm designed by the second method only takes linear time and constant space when running in external memory.
-For a further study, we integrate the checking algorithm designed by the second method into the existing suffix sorting algorithm DSA-IS to evaluate the checking overhead, where the time, space and I/O consumptions for verification are considerably smaller than that for construction in the experiments.
-This convinces us that the proposed checking methods, together with the state-of-the-art IS suffix sorters, could constitute efficient solutions for both construction and verification.
+Particularly, the first method checks both suffix and longest common prefix~(LCP) arrays for an input string drawn from a full-ordered alphabet $\Sigma$.
+By combining our methods with the Karp-Rabin fingerprinting function, we design two probabilistic algorithms that perform verification with a negligible error probability.
+Theoretically, the algorithm designed by the first method has a sorting complexity, while the algorithm designed by the second method takes linear time and $\mathcal{O}(\|\Sigma\|)$ RAM space.
+We integrate the algorithm designed by the second method into the existing suffix sorting algorithm DSA-IS for performance evaluation.
+Our experiments indicate that the time, space and I/O consumptions for verifcation are considerably smaller than that for construction.
+We believe that the proposed methods, together with the state-of-the-art IS suffix sorters (and LCP-array builders), could constitute efficient solutions for building and checking suffix (and LCP) array(s) when $\|\Sigma\|$ is much smaller than the size of input string.
 
@@
 
-A suffix array can be built in linear time and space by the internal-memory algorithm SA-IS~\cite{Nong11}.
-According to the IS principle, the relative order of two unsorted suffixes is determined by sequentially comparing the heading characters and the sorted succeeding suffixes starting with their second characters.
-Recently, the IS method has been applied to designing three disk-based suffix sorting algorithms eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}, which have a better time complexity than the other alternatives~(e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}) but suffer from a bottleneck due to the large disk space for obtaining the heading characters of unsorted suffixes and the ranks of their successors in a disk-friendly way.
-More specifically, the average peak disk use to construct an SA encoded by 40-bit integers for pSAscan is $7.5n$, while that for eSAIS, DSA-IS and SAIS-PQ are $24n$, $20n$ and $15n$, respectively.
-The poor space performance is mainly because that the current programs for these algorithms fail to free the disk space for temporary data even when the data is no longer to use.
-A dramatic improvement can be achieved by splitting a file into multiple pieces and deleting each piece immediately when the temporary data in it is not needed any more.
-This technique has been used to implement a new IS suffix sorting algorithm fSAIS~\cite{Karkkainen2017}.
-As reported, the engineering of fSAIS consumes no more than $8n$ disk space for constructing an SA encoded by 40-bit integers, indicating a great potential for optimizing our programs for DSA-IS and SAIS-PQ.
+A suffix array can be built within linear time and space using the internal-memory algorithm SA-IS~\cite{Nong11}.
+According to the IS principle, the order of two suffixes is determined by sequentially comparing their heading characters and sorted successors.
+Recently, the IS method has been used to design three disk-based suffix sorting algorithms eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}.
+These works have better time complexities than the other alternatives~(e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}), but they suffer from a bottleneck due to the large disk space for obtaining the heading characters of unsorted suffixes and the ranks of their sorted successors in a disk-friendly way.
+As reported, the average peak disk use to construct an SA encoded by 40-bit integers for eSAIS, DSA-IS and SAIS-PQ are $24n$, $20n$ and $15n$, respectively, while that for pSAScan is only $7.5n$.
+The poor space performance of the former three algorithms is mainly because that their current programs fail to free the disk space for temporary data even when the data is no longer to use.
+A dramatic improvement can be achieved by splitting a file into multiple pieces and deleting each piece immediately when the temporary data residing on it is not needed any more.
+This technique has been used to implement a novel IS suffix sorting algorithm fSAIS~\cite{Karkkainen2017}.
+The engineering of fSAIS takes no more than $8n$ disk space, indicating a great potential for optimizing the performances of DSA-IS and SAIS-PQ.
 
-A constructed SA should be checked to avoid computation errors caused by implementation bugs and/or hardware malfunctions.
-Currently, the software packages for DC3 and eSAIS provide users a checker designed by the method proposed in~\cite{Dementiev2008a}.
-The overhead for this checker is rather high because it takes two external-memory sorts for arranging $\mathcal{O}(n)$ fixed-size tuples according to their integer keys.
-Against the background, we propose in this paper two methods that enable any IS suffix sorter to check an SA when it is being built.
-The first method can check both suffix and LCP arrays for strings drawn from constant alphabets, while the second method can check SA for strings drawn from any alphabets.
-We augment these methods with the Karp-Rabin fingerprinting function~\cite{Karp1987} to design two probabilistic algorithms, in terms of that their checking results are wrong with a negligible probability.
-These algorithms are rather lightweight in terms of time and space.
-Particularly, the algorithm designed by the second method only takes linear time and constant space to run using external memory.
-For a further study, we integrate this algorithm into DSA-IS to evaluate its checking overhead, where the design of DSA-IS has been adapted by new substring sorting and naming methods for a better performance.
-As demonstrated in our experiments, the time, space and I/O consumptions for verification by this algorithm are negligible to that for construction by the adapted DSA-IS.
-We are convinced that the proposed SA checkers, together with the state-of-the-art IS suffix sorters, could constitute efficient solutions for both construction and verification.
+A constructed SA should be checked to detect potential errors caused by implementation bugs and hardware malfunctions.
+The software packages for DC3 and eSAIS provide users a checker that implements the algorithm presented in~\cite{Dementiev2008a}.
+When running on external-memory model, the checking overhead is rather high because this checker takes two passes of integer sorts for arranging $\mathcal{O}(n)$ fixed-size tuples.
+In this paper, we describe two methods that enable any IS suffix sorter to build and check an SA at the same time.
+In particular, the first method is proposed for checking both suffix and LCP arrays.
+We apply the Karp-Rabin fingerprinting function~\cite{Karp1987} to designing two probabilistic algorithms, in terms of that their checking results are wrong with a negligible probability.
+We combine the algorithm designed by the second method with DSA-IS and implement the combination to make a performance evaluation, where new substring sorting and naming methods is used to lower the peak disk use and I/O volume of the program for DSA-IS in advance.
+Our experimental results indicate that the time, space and I/O consumptions for verification are negligible to that for construction.
+This implies that the IS method can be used to design efficient solutions for verification as well.
 
 The rest of this paper is organized as follows.
-Section~\ref{sec:preliminaries} introduces some basic notations and symbols.
-Section~\ref{sec:builder} gives an overview of IS suffix sorting algorithms and the details of new substring sorting and naming methods.
-Section~\ref{sec:checkers} describes the ideas of checking methods and algorithms.
+Section~\ref{sec:preliminaries} introduces some notations and symbols for presentation convenience.
+Section~\ref{sec:builder} describes the overview of the existing IS suffix sorting algorithms and the details of our new substring sorting and naming methods specific for DSA-IS.
+Section~\ref{sec:checkers} presents the proposed checking methods and the algorithms designed by them.
 Sections~\ref{sec:experiments} and~\ref{sec:conclusion} show the experimental results and the concluding remarks, respectively.
@@
 A substring or suffix is of the same type as its heading character.
-In this paper, we only consider substrings ending with an S*-type character.
 
@@
 
-We partition $x$ into S*-type substrings and replace each substring with its name to produce the reduced string $x_1$, where the name indicates the rank of the corresponding S*-type substring among all.
+Partition $x$ into multiple successive S*-type substrings, such that substring starts with an S*-type character and ends with the leftmost S*-type character on the right side.
+Produce a reduced string $x_1$ by replacing each S*-type substring with its name, where the name indicats its relative order among all.
 
@@
 	}
+	\Else{
+
+		Compute $sa_1$ from $x_1$. 
+	}
 
@@
 
-	\caption{The overview of IS suffix sorting algorithms	.}
+	\caption{The Framework of an IS suffix sorting algorithm.}
 
@@
 
-As shown in Algorithm~\ref{alg:1}, an IS suffix sorting algorithm first performs a reduction phase to produce the reduced string $x_1$.
-If there exist two equal characters in $x_1$, then it recursively calls the reduction phase with $x_1$ as input; otherwise, all the S*-type suffixes in $x$ are already sorted and it performs an induction phase to produce the final $sa$.
-During the execution of two phases, the order of substrings/suffixes is determined by comparing the heading characters and the ranks of their successors according to the IS principle.
-This involves frequent random accesses to $x$ and $sa$ for retrieving the heading characters of unsorted substrings/suffixes and the ranks of their sorted successors.
-It can be done very quickly if both $x$ and $sa$ are wholly loaded into RAM; otherwise, each disk access takes an individual I/O operation, leading to a severe performance degradation.
-This problem is solved in DSA-IS by conducting two steps:
+Algorithm~\ref{alg:1} shows the framework of an IS suffix sorting algorithm.
+At the very beginning, a reduction phase for sorting and naming S*-type substrings is called to produce the reduced string $x_1$.
+If there exist duplicate characters in $x_1$, the reduction phase is recursively called with $x_1$ as input; otherwise, all the S*-type suffixes are already sorted and $sa_1$ is directly computed from $x_1$.
+Afterward, an induction phase for sorting all the suffixes is called to produce $sa$ for the input string at current level.
+The induction phase is recursively called with $sa$ as input until reaching the top recursion level.
+During the execution of a reduction/induction phase, the order of substrings/suffixes is determined by comparing their heading characters and the ranks of their successors according to the IS principle, which involves a great number of random accesses to $x$ and $sa$.
+This can be done very fast if both $x$ and $sa$ wholly reside on RAM; otherwise, each access takes an individual I/O operation, leading to a severe performance degradation.
+This problem is solved in DSA-IS by performing two steps below during the reduction/induction phase:
 
@@
 As shown in Section~\ref{sec:experiments}, our program for DSA-IS requires less disk space than that for eSAIS, but the former runs slower than the latter caused by the large I/O volume for sorting and naming S*-type substrings during the reduction phase.
-For this, we present new substring sorting and naming methods in the next subsection.
+To improve the performance, we propose new substring sorting and naming methods in the next subsection.
 
@@
 The experimental results in Section~\ref{sec:experiments} indicate that, by using the new substring sorting and naming methods, the adapted DSA-IS, called DSA-IS+, only takes half as much disk space as eSAIS.
-Also, the peak disk use of the program for DSA-IS+ can be further reduced to a level comparable to that of the program for fSAIS.
 
@@
 
-\color{red}{The software packages for DC3 and eSAIS provide users a checker to ensure the correctness of their outputs.
-Below we describe the main idea behind this checker and refer interested readers to ~\cite{Dementiev2008a} for more details.}
+We describe below the main idea of the checking algorithm presented in~\cite{Dementiev2008a}.
 
@@
 Recall that, an IS suffix sorting algorithm recursively perform the reduction phase to compute the reduced string $s_1$ until $s_1$ contains no duplicate characters.
-Afterward, it produces $sa_1$ from $s_1$ and recursively perform the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase executes mainly consists of the following steps:
+Afterward, it produces $sa_1$ from $s_1$ and recursively perform the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase mainly consists of the following steps:
 
@@
 
-A running example of S\ref{induction_phase:2}-Sp\ref{induction_phase:4} is shown in Fig.~\ref{fig:example1}.
+A running example of S\ref{induction_phase:2}-S\ref{induction_phase:4} is shown in Fig.~\ref{fig:example1}.
 Given $sa^*$ is already known, line 6 inserts each S*-type suffix into the right part of the corresponding bucket and arranges them according to their sorted order indicated by $sa^*$.
@@
 	%\SetAlgoNoLine
-	\Fn{{\sf CheckByMethodA}}{
+	\Fn{{\sf CheckByMethodA}($x, sa^*, lcp^*$))}{
 
@@
 
-{\color{red}
 \begin{lemma} \label{lemma:4}
@@
 \end{IEEEproof}
-}
 
@@
 	%\SetAlgoNoLine
-	\Fn{{\sf CheckByMethodB}}{
+	\Fn{{\sf CheckByMethodB}($x, sa^*$)}{
 
@@
 
-It takes linear time and constant RAM space for computing the required fingerprints in Algorithm~\ref{alg:2} and Algorithm~\ref{alg:3}.
-Further, the bottleneck of Algorithm~\ref{alg:2} lies in checking $sa^*$ following the idea presented in~\cite{wu2017}.
-Specifically, given two suffixes, say $\sf suf(i)$ and $\sf suf(j)$, and their LCP-value, say $d$, we compute and compare the fingerprints of $\sf sub(i, i + d - 1)$ and $\sf sub(j, j + d - 1)$.
-If the fingerprints are equal and the order of $x[i + d]$ and $x[j + d]$ correspond to that of $\sf suf(i)$ and $\sf suf(j)$, it is believed that the suffixes are correctly sorted and their LCP-value is right.
-When implemented in external memory, the checking processes for any number of suffix pairs can be done in batch within a sorting complexity.
-Because at most one out of every two suffixes is S*-type (typically one-third in real-world data sets) and the checking process is only executed at the top recursion level, the checking consumption taken by Algorithm~\ref{alg:2} is much less than the building consumption taken by any disk-based IS suffix sorting algorithm.
+It 
+According to Lemma~\ref{lemma:3}, Method A must check both $sa^*$ and $sa$ to accomplish verification.
+Similar to Method B, the overhead for checking $sa$ is mainly caused by fingerprint calculations and thus can be neglected.
+On the other hand, we can apply the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ within sorting complexity, where the time and space in need is proportional to the number of S*-type characters in $x$.
+Because the ratio of S*-type characters to all in $x$ is commonly one-third in real-world datasets, the checking process for $sa^*$ will not become the bottleneck for Method A.
+To achieve a higher speed, we can also parallelize the checking processes for $sa$ and $sa^*$ using more computing resources.
+
+Assuming the induction phase at the top recursion level is correctly implemented, the bottleneck of Algorithm~\ref{alg:2} and~\ref{alg:3} lie in checking the correctness of $sa^*$.
+For Algorithm~\ref{alg:2}, we apply the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ within the sorting complexity.
+Specifically, given two suffixes, i.e.
+$\sf suf(i)$ and $\sf suf(j)$, and their LCP-value $d$, we compute and compare the fingerprints of $\sf sub(i, i + d - 1)$ and $\sf sub(j, j + d - 1)$.
+If the fingerprints are equal and the order of $x[i + d]$ and $x[j + d]$ corresponds to that of $\sf suf(i)$ and $\sf suf(j)$, then the suffixes are correctly sorted and their LCP-value is right with a high probability.
+Because at most one out of every two suffixes is S*-type (commonly one-third in real-world data sets) and the checking process is only executed during the induction phase at the top recursion level, the checking overhead is much less than the building consumption.
+For Algorithm~\ref{alg:3}, the checking overhead is mainly caused by fingerprinting calculations and thus can be neglected compared with the building consumption.
 
@@
 
-According to Corollary~\ref{corollary:2}, Method A must check both $sa^*$ and $sa$ to accomplish verification.
-Similar to Method B, the overhead for checking $sa$ is mainly caused by fingerprint calculations and thus can be neglected.
-On the other hand, we can apply the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ within sorting complexity, where the time and space in need is proportional to the number of S*-type characters in $x$.
-Because the ratio of S*-type characters to all in $x$ is commonly one-third in real-world datasets, the checking process for $sa^*$ will not become the bottleneck for Method A.
-To achieve a higher speed, we can also parallelize the checking processes for $sa$ and $sa^*$ using more computing resources.
-
 \subsection{Discussion}

### 2017-08-05 new-draft/draft.tex
2017/08/06

@@
 Particularly, the first method checks both suffix and longest common prefix~(LCP) arrays for an input string drawn from a full-ordered alphabet $\Sigma$.
-By combining our methods with the Karp-Rabin fingerprinting function, we design two probabilistic algorithms that perform verification with a negligible error probability.
+By combining our methods with the Karp-Rabin fingerprinting function, we design two algorithms that perform verification with a negligible error probability.
 Theoretically, the algorithm designed by the first method has a sorting complexity, while the algorithm designed by the second method takes linear time and $\mathcal{O}(\|\Sigma\|)$ RAM space.
@@
 \begin{IEEEkeywords}
-Suffix and LCP arrays, construction and verification, external memory.
+Suffix and LCP arrays, construction and verification, internal and external memory.
 \end{IEEEkeywords}}
@@
 We apply the Karp-Rabin fingerprinting function~\cite{Karp1987} to designing two probabilistic algorithms, in terms of that their checking results are wrong with a negligible probability.
-We combine the algorithm designed by the second method with DSA-IS and implement the combination to make a performance evaluation, where new substring sorting and naming methods is used to lower the peak disk use and I/O volume of the program for DSA-IS in advance.
+We combine the algorithm designed by the second method into DSA-IS and implement the combination to make a performance evaluation, where new substring sorting and naming methods are used to redesign DSA-IS for lowering the peak disk use and I/O volume taken by SA construction.
 Our experimental results indicate that the time, space and I/O consumptions for verification are negligible to that for construction.
@@
 
-Partition $x$ into multiple successive S*-type substrings, such that substring starts with an S*-type character and ends with the leftmost S*-type character on the right side.
-Produce a reduced string $x_1$ by replacing each S*-type substring with its name, where the name indicats its relative order among all.
+Partition $x$ into multiple successive S*-type substrings, such that eachsubstring starts with an S*-type character and ends with the leftmost S*-type character on the right side.
+Produce a reduced string $x_1$ by replacing each S*-type substring with its name, where the name is the rank indicating the lexical order among all.
 
@@
 If there exist duplicate characters in $x_1$, the reduction phase is recursively called with $x_1$ as input; otherwise, all the S*-type suffixes are already sorted and $sa_1$ is directly computed from $x_1$.
-Afterward, an induction phase for sorting all the suffixes is called to produce $sa$ for the input string at current level.
+Afterward, an induction phase for sorting all the suffixes is called to produce $sa$ for $x$ at current recursion level.
 The induction phase is recursively called with $sa$ as input until reaching the top recursion level.
@@
 After the first two steps, the short S*-type substrings of each block and the long S*-type substrings of $x$ are sorted and separately organized as a sequence in external memory.
-Assume $x$ is split into $k$ blocks, the multi-way sorter in S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap to cache the current smallest of each sequence and continually retrieve the top item from the heap to determine the lexical order of substrings from different sequences.
+Assume $x$ is split into $k$ blocks, the multi-way sorter in S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap to cache the current smallest of each sequence and continually retrieves the top item from the heap to determine the lexical order of substrings from different sequences.
 The heap contains at most $k + 1$ substrings at any point of time and it compares any two substrings in $\mathcal{O}(D)$ time by a literal string comparison\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}.
@@
 The key point here is to check equality of two substrings successively popped from the heap in S\ref{dsaism_sorting_method:3}' as fast as possible.
-If either of the two substrings is short, our solution is to literally compare them in $\mathcal{O}(D)$ time.
-Otherwise, if both of them are long, then we assign names to all the long S*-type substrings using the technique proposed for SAIS-PQ in S\ref{dsaism_sorting_method:2}' and compare two successively popped long S*-type substrings by their names.
+If either of the two substrings is short, we literally compare them in $\mathcal{O}(D)$ time.
+Otherwise, if both of them are long, then we assign names to all the long S*-type substrings using the technique proposed for SAIS-PQ in S\ref{dsaism_sorting_method:2}' and directly compare two successively popped long S*-type substrings by their names.
 This technique is based on the idea that two substrings are of the same name if their heading characters and the names of their successors are both equal\footnote{We refer interested readers to~\cite{Liu15} for more information.}.
@@
 
-The disk-based implementation of this checker conducts two passes of integer sorts and each sort arranges the order of $n$ fixed-size tuples using external memory.
+The disk-based implementation of this checker conducts two passes of integer sorts and each sort arranges the order of $\mathcal{O}(n)$ fixed-size tuples using external memory.
 As can be observed from Section~\ref{sec:experiments}, the peak disk use and the I/O volume for an SA encoded by 40-bit integers are around $26n$ and $53n$, respectively.
@@
 
-Recall that, an IS suffix sorting algorithm recursively perform the reduction phase to compute the reduced string $s_1$ until $s_1$ contains no duplicate characters.
-Afterward, it produces $sa_1$ from $s_1$ and recursively perform the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase mainly consists of the following steps:
+Remember that an IS suffix sorting algorithm recursively performs the reduction phase to compute the reduced string $s_1$ until $s_1$ contains no duplicate characters.
+Afterward, it produces $sa_1$ from $s_1$ and recursively perform the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase consists of the following steps:
 
@@
 Given $sa^*$ is already known, line 6 inserts each S*-type suffix into the right part of the corresponding bucket and arranges them according to their sorted order indicated by $sa^*$.
-For example, $\sf suf(8)$, $\sf suf(5)$, $\sf suf(2)$ are placed into the right part of $\sf sa\_bkt(i)$ and arranged in their sorted order indicated by $sa^*$.
+For example, $\sf suf(2)$, $\sf suf(5)$, $\sf suf(8)$ are placed at the rightmost empty position of $\sf sa\_bkt(i)$ in sequence.
 Then, we find the leftmost position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes.
@@
     (1) $sa^*$ is correctly computed. \\
-    (2) S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented. \\
+    (2) S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented at the top recursion level. \\
 \end{lemma}
 
-\begin{IEEEproof} Given two suffixes placed at $sa[i]$ and $sa[j]$ and their successors placed at $sa[p]$ and $sa[q]$, we prove the statement as follows.
+\begin{IEEEproof} Given two suffixes placed at $sa[i]$ and $sa[j]$ and their successors placed at $sa[p]$ and $sa[q]$ at the top recursion level, we prove the statement as follows.
 
-Because S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented, $i < j \iff (x[sa[i]], p) < (x[sa[j]], q)$. This satisfies the second condition of Lemma~\ref{lemma:1}.
+Because S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented, we have $i < j \iff (x[sa[i]], p) < (x[sa[j]], q)$ following the IS principle.
+This satisfies the second condition of Lemma~\ref{lemma:1}.
 
 Further, suppose $sa[i] = sa[j]$ and $i \ne j$, then $sa[p] = sa[q]$ and $p \ne q$.
-Repeatedly replacing $(i, j)$ with $(p, q)$ until $\sf suf(sa[i])$ is S*-type, then $sa^*$ contain duplicate elements.
+Repeating this reasoning process by replacing $(i, j)$ with $(p, q)$ until $\sf suf(sa[i])$ is S*-type, then $sa^*$ contains duplicate elements.
 Because $sa^*$ is correctly computed, each element in $sa^*$ must be unique.
@@
 This is considerable smaller than the whole program for a disk-based suffix sorting algorithm.
-Following the idea, we assume S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented and propose two methods for checking computation errors caused by implementation bugs and I/O errors caused by hardware malfunctions.
+Following the idea, we assume S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented and propose two methods for checking computation errors caused by implementation bugs and hardware malfunctions.
 The algorithms designed by these methods can be seamlessly integrated into any IS suffix sorting algorithms for building and checking an SA at the same time.
@@
 
-Our first method is directly extended from Lemma~\ref{lemma:2}.
+Method A is based on Lemma~\ref{lemma:3}, which is extended from Lemma~\ref{lemma:2}.
 
 \begin{lemma} \label{lemma:3}
-	Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied: \\
+	Assume the induction phase is correctly implemented at the top recursion level, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied: \\
 	(1) $sa^*$ is correct. \\
-	(2) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}''. \\
+	(2) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' for $i \in [0, n)$. \\
 
@@
 The first condition of Lemma~\ref{lemma:3} can be checked by a sparse SA checker like~\cite{wu2017}.
-Notice that each suffix induced into $sa$ will be latter scanned for inducing the order of its predecessor during S\ref{induction_phase:2}'' and S\ref{induction_phase:4}''.
-The second condition intends for detecting I/O errors and can be checked by determining whether the induced and scanned values for each suffix are equal.
+Notice that each suffix induced into $sa$ will be latter scanned for inducing the order of its predecessor during S\ref{induction_phase:3}''-S\ref{induction_phase:4}''.
+The second condition intends for detecting errors caused by hardware malfunctions and can be checked by determining whether the induced and scanned values for each suffix are equal.
 The problem to be solved here is that when a suffix is induced into a bucket, its corresponding value in $sa$ may not be scanned at once.
 Our solution is to integrate the checking process into the building process, by ensuring the sequence of values placed at a bucket is identical to that scanned later.
-For the purpose, we use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end of the induction phase.
+For the purpose, we use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end.
 If the two fingerprints for each bucket are equal, then the second condition of Lemma~\ref{lemma:3} will be seen with a high probability.
 As a result, $sa$ can be built and probabilistically checked at the same time.
-It should be noticed that this method can be also applied to checking the longest common prefix array when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}.
+It should be noticed that this method can be also applied to checking the LCP array when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}.
+
 As a result, we design Algorithm~\ref{alg:2} to probabilistically check $sa$ and $lcp$ during the induction phase at the top recursion level.
+We point out that ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$.
 
@@
 	%\SetAlgoNoLine
-	\Fn{{\sf CheckByMethodA}($x, sa^*, lcp^*$))}{
+	\Fn{{\sf CheckByMethodA}($x, sa^*, lcp^*$)}{
 
-        Check the correctness of $sa^*$ and $lcp^*$ by the method proposed in~\cite{wu2017}. \\
+        Verify $sa^*$ and $lcp^*$ by the checker presented in~\cite{wu2017}. \\
 
@@
 
-Method B uses a different way to check the correctness of $sa^*$ in Lemma~\ref{lemma:4}.
+As can be seen in Lemma~\ref{lemma:4}, Method B uses a different way to check the correctness of $sa^*$.
 Before our presentation, we first introduce a notation called $\overline{sa^*}$.
 The same as $sa^*$, $\overline{sa^*}$ also represents the suffix array for all the S*-type suffixes.
-Both of them are computed during the induction phase at the top recursion level of an IS suffix sorting algorithm.
-To be specific, $sa^*$ is computed from $sa_1$ and $\overline{sa^*}$ is retrieved from the constructed $sa$.
+Both of them are computed during the induction phase at the top recursion level.
+To be specific, $sa^*$ is computed from $sa_1$ and $\overline{sa^*}$ is retrieved from the constructed final $sa$.
 
 \begin{lemma} \label{lemma:4}
-    Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied for all $i \in [0, n_1)$:
+    Assume the induction phase is correctly implemented at the top recursion level, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied:
 
 	(1) $sa^*[i]$ is a permutation of all the S*-type suffixes. \\
-	(2) $sa^*[i] = \overline{sa^*}[i]$. \\
-	(3) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}''. \\
+	(2) $sa^*[i] = \overline{sa^*}[i]$ for $i \in [0, n_1)$. \\
+	(3) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' for $i \in [0, n)$. \\
 
@@
 Notice that two equal arrays must share an identical hash value, but the inverse is not always true.
-To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting function to compute the required values following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $delta$ is an integer randomly chosen from $[1, L)$.
+To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting function to compute the required values following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$.
 By setting $L$ to a large value, the error probability can be reduced to a negligible level.
@@
 Given $L = 23$ and $\delta = 5$, we first compute the fingerprint of $A[0, 0]$ in line 4.
-Because $\sf fp(A[0, -1]) = 0$, $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$.
+Because $\sf fp(A[0, -1]) = 0$, then $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$.
 By iteratively computing the fingerprints of the prefixes of $A$ according to Formula~\ref{formula:2}, we finally obtain the fingerprints of the whole array in line 7.
@@
 
-It 
-According to Lemma~\ref{lemma:3}, Method A must check both $sa^*$ and $sa$ to accomplish verification.
-Similar to Method B, the overhead for checking $sa$ is mainly caused by fingerprint calculations and thus can be neglected.
-On the other hand, we can apply the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ within sorting complexity, where the time and space in need is proportional to the number of S*-type characters in $x$.
-Because the ratio of S*-type characters to all in $x$ is commonly one-third in real-world datasets, the checking process for $sa^*$ will not become the bottleneck for Method A.
-To achieve a higher speed, we can also parallelize the checking processes for $sa$ and $sa^*$ using more computing resources.
-
-Assuming the induction phase at the top recursion level is correctly implemented, the bottleneck of Algorithm~\ref{alg:2} and~\ref{alg:3} lie in checking the correctness of $sa^*$.
+The bottleneck of Algorithm~\ref{alg:2} and~\ref{alg:3} lie in checking the correctness of $sa^*$.
 For Algorithm~\ref{alg:2}, we apply the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ within the sorting complexity.
@@
 If the fingerprints are equal and the order of $x[i + d]$ and $x[j + d]$ corresponds to that of $\sf suf(i)$ and $\sf suf(j)$, then the suffixes are correctly sorted and their LCP-value is right with a high probability.
-Because at most one out of every two suffixes is S*-type (commonly one-third in real-world data sets) and the checking process is only executed during the induction phase at the top recursion level, the checking overhead is much less than the building consumption.
+Because at most one out of every two suffixes is S*-type (commonly one-third suffixes are S*-type in real-world data sets) and the checking process is only executed during the induction phase at the top recursion level, the checking overhead is much less than the building consumption.
 For Algorithm~\ref{alg:3}, the checking overhead is mainly caused by fingerprinting calculations and thus can be neglected compared with the building consumption.
@@
 In this way, our program can be further improved to achieve a space performance comparable to fSAIS.
+In our next paper, we will propose another disk-based sorting algorithm that has a space performance even better than fSAIS.
+It takes only $1n$ work space to run excluding the input and output.
 
@@
 
-In this paper, we proposed two methods that enable any IS suffix sorting algorithm to build and check SA simultaneously.
-In particular, the probabilistic algorithm designed by the second method is rather lightweight in terms of that it takes negligible time and space compared with the existing IS suffix sorting and checking algorithms.
-We also made an attempt to improve the space performance of DSA-IS using new substring sorting and naming methods.
-The experimental results shows that our program for the adapted algorithm DSA-IS+ runs as fast as that for eSAIS and requires only half as much disk space as the latter on various real-world datasets.
-We also showed that the space performance of this program can be further optimized using some optimization techniques.
-In our next paper, we will propose another disk-based suffix sorting algorithm that takes only $1n$ work space to run.
+We proposed two methods that enable any IS suffix sorting algorithm to build and check SA simultaneously.
+The program for our algorithm designed by Method B consumes negligible time and RAM space in comparison with that for the existing suffix sorting and checking algorithms on various real-world data sets.
+We also made our first attempt to improve the performance of DSA-IS using new substring and naming methods.
+The implementation for the adapted algorithm DSA-IS+ runs as fast as that for eSAIS and requires only half as much disk space as the latter.
+These results indicate the great potential of the IS method to design efficient solutions for SA constructing and verification.
 

### 2017-08-08 new-draft/draft.tex
2017/08/08

@@
 
-\title{Building and Checking Suffix Array Using Induced Sorting Method}
+\title{Building and Checking Suffix and LCP Arrays Using Induced Sorting Method}
 
@@
 
-Suffix array~(SA) can be built by the induced sorting~(IS) method on both internal and external memory models.
-We propose two methods that enable any IS suffix sorter to build and check an SA simultaneously.
-Particularly, the first method checks both suffix and longest common prefix~(LCP) arrays for an input string drawn from a full-ordered alphabet $\Sigma$.
-By combining our methods with the Karp-Rabin fingerprinting function, we design two algorithms that perform verification with a negligible error probability.
-Theoretically, the algorithm designed by the first method has a sorting complexity, while the algorithm designed by the second method takes linear time and $\mathcal{O}(\|\Sigma\|)$ RAM space.
-We integrate the algorithm designed by the second method into the existing suffix sorting algorithm DSA-IS for performance evaluation.
-Our experiments indicate that the time, space and I/O consumptions for verifcation are considerably smaller than that for construction.
-We believe that the proposed methods, together with the state-of-the-art IS suffix sorters (and LCP-array builders), could constitute efficient solutions for building and checking suffix (and LCP) array(s) when $\|\Sigma\|$ is much smaller than the size of input string.
+Suffix and longest common prefix~(LCP) arrays can be built by the induced sorting~(IS) method on both internal and external memory models.
+We propose two methods that enable any IS builder to build and check suffix and LCP arrays simultaneously.
+The first method is for checking both suffix and LCP arrays, while the second method is for checking SA only.
+By combining the Karp-Rabin fingerprinting techinque into our methods, we design two algorithms that perform checking correctly with a negligible error probability.
+Theoretically, the algorithm designed by the first method has a sorting complexity, while the algorithm designed by the second method only takes linear time to run.
+We integrate the algorithm designed by the second method into the existing disk-based suffix sorting algorithm DSA-IS and implement their combination for performance evaluation.
+From our experiments, the checking overhead is considerably smaller than the building consumption.
 
@@
 
-A suffix array can be built within linear time and space using the internal-memory algorithm SA-IS~\cite{Nong11}.
-According to the IS principle, the order of two suffixes is determined by sequentially comparing their heading characters and sorted successors.
-Recently, the IS method has been used to design three disk-based suffix sorting algorithms eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}.
-These works have better time complexities than the other alternatives~(e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}), but they suffer from a bottleneck due to the large disk space for obtaining the heading characters of unsorted suffixes and the ranks of their sorted successors in a disk-friendly way.
-As reported, the average peak disk use to construct an SA encoded by 40-bit integers for eSAIS, DSA-IS and SAIS-PQ are $24n$, $20n$ and $15n$, respectively, while that for pSAScan is only $7.5n$.
-The poor space performance of the former three algorithms is mainly because that their current programs fail to free the disk space for temporary data even when the data is no longer to use.
-A dramatic improvement can be achieved by splitting a file into multiple pieces and deleting each piece immediately when the temporary data residing on it is not needed any more.
-This technique has been used to implement a novel IS suffix sorting algorithm fSAIS~\cite{Karkkainen2017}.
-The engineering of fSAIS takes no more than $8n$ disk space, indicating a great potential for optimizing the performances of DSA-IS and SAIS-PQ.
+A suffix array~(SA) can be built within linear time and space by the internal-memory algorithm SA-IS~\cite{Nong11}.
+According to the IS principle, the order of two suffixes is determined by comparing their heading characters and successors in sequence, where the order of their successors is assumed to be determined in advance.
+Recently, the IS method has been also applied to designing three disk-based suffix sorting algorithms eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}, where eSAIS can build both suffix and LCP arrays at the same time.
+These works have better time complexities than the other alternatives ~(e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}), but they all suffer from a bottleneck due to the large disk space for obtaining the heading characters of unsorted suffixes and the ranks of their sorted successors in a disk-friendly way.
+It is reported that the average peak disk use to construct an SA encoded by 40-bit integers for pSAscan is only $7.5n$, while that for eSAIS, DSA-IS and SAIS-PQ are $24n$, $18n$ and $15n$, respectively.
+The poor space performance of the IS algorithms is mainly because that their current programs fail to free the disk space for temporary data even when the data is no longer to use.
+A dramatic improvement can be achieved by storing temporary data in multiple files and deleting each file when the data residing on it is not needed any more.
+This technique has been used to implement a novel IS suffix sorting algorithm fSAIS~\cite{Karkkainen2017}, where the engineering of fSAIS takes no more than $8n$ disk space.
+This indicates a great potential for optimizing the performances of DSA-IS and SAIS-PQ.
 
-A constructed SA should be checked to detect potential errors caused by implementation bugs and hardware malfunctions.
-The software packages for DC3 and eSAIS provide users a checker that implements the algorithm presented in~\cite{Dementiev2008a}.
-When running on external-memory model, the checking overhead is rather high because this checker takes two passes of integer sorts for arranging $\mathcal{O}(n)$ fixed-size tuples.
-In this paper, we describe two methods that enable any IS suffix sorter to build and check an SA at the same time.
-In particular, the first method is proposed for checking both suffix and LCP arrays.
-We apply the Karp-Rabin fingerprinting function~\cite{Karp1987} to designing two probabilistic algorithms, in terms of that their checking results are wrong with a negligible probability.
-We combine the algorithm designed by the second method into DSA-IS and implement the combination to make a performance evaluation, where new substring sorting and naming methods are used to redesign DSA-IS for lowering the peak disk use and I/O volume taken by SA construction.
-Our experimental results indicate that the time, space and I/O consumptions for verification are negligible to that for construction.
-This implies that the IS method can be used to design efficient solutions for verification as well.
+A constructed SA should be checked to detect potential errors caused by implementation bugs and other malfunctions.
+The software packages for DC3 and eSAIS provide users a checker based on the idea presented in~\cite{Dementiev2008a}.
+When running on external-memory model, the cost taken by this checker is rather high, because it performs two passes of integer sorts for arranging $\mathcal{O}(n)$ fixed-size tuples.
+In this paper, we describe two methods that enable any IS builder to build and check suffix and LCP arrays simultaneously.
+The first method is for checking both suffix and LCP arrays, while the second method is for checking SA only.
+We employ the Karp-Rabin fingerprinting technique~\cite{Karp1987} to design two probabilistic algorithms, in terms of that their checking results are wrong with a negligible probability.
+For analysis, we first use new substring sorting and naming methods to improve the design of DSA-IS and then combine the second checking method into the adapted DSA-IS for evaluating the checking overhead.
+Our experimental results indicate that the time, space and I/O volume for checking SA is considerable smaller than that for building SA.
+This implies that the IS method can be used to design efficient solutions for SA verification as well.
 
@@
 Section~\ref{sec:preliminaries} introduces some notations and symbols for presentation convenience.
-Section~\ref{sec:builder} describes the overview of the existing IS suffix sorting algorithms and the details of our new substring sorting and naming methods specific for DSA-IS.
-Section~\ref{sec:checkers} presents the proposed checking methods and the algorithms designed by them.
+Section~\ref{sec:builder} gives an overview of the existing IS suffix sorting algorithms and show the details of our new substring sorting and naming methods specific for DSA-IS.
+Section~\ref{sec:checkers} presents the proposed checking methods and the probabilistic algorithms designed by them.
 Sections~\ref{sec:experiments} and~\ref{sec:conclusion} show the experimental results and the concluding remarks, respectively.
@@
 We use an array $t$ to record the type of all the characters in $x$, where $t[i] = 1$ or $0$ if $x[i]$ is S-type or L-type, respectively.
-A substring or suffix is of the same type as its heading character.
+The type of a substring or suffix is determined by that of its heading character.
 
 Given two characters $x[i]$ and $x[i + 1]$, we say $x[i]$ is the predecessor of $x[i + 1]$ and $x[i + 1]$ is the successor of $x[i]$.
-Accordingly, we define the predecessor-successor relationship between ${\sf suf}(i)/{\sf sub}(i, j)$ and ${\sf suf}(i + 1)/{\sf sub}(i + 1, j)$.
+We define the predecessor-successor relationship between ${\sf suf}(i)/{\sf sub}(i, j)$ and ${\sf suf}(i + 1)/{\sf sub}(i + 1, j)$.
 
-Partition $x$ into multiple successive S*-type substrings, such that eachsubstring starts with an S*-type character and ends with the leftmost S*-type character on the right side.
-Produce a reduced string $x_1$ by replacing each S*-type substring with its name, where the name is the rank indicating the lexical order among all.
+Partition $x$ into multiple S*-type substrings.
+Each substring only contains two S*-type characters and any two neighboring substrings overlap an S*-type character.
+Produce a reduced string $x_1$ by replacing each S*-type substring with its name, which represents the rank of the corresponding S*-type substring among all.
+In the following paragraphs, we use "rank" and "name" interchangeably to indicate the lexical order of the corresponding substring.
 
-The suffix array $sa$ indicates the lexical order of all the suffixes in $x$, where $sa[i]$ records the starting position of the $(i + 1)$-th smallest suffix.
-We also define the suffix array $sa_1$ for $x_1$ following the same way.
+The suffix array $sa$ arranges all the suffixes of $x$ in their lexical order, where $sa[i]$ records the starting position of the $(i + 1)$-th smallest suffix.
+We also define the suffix array $sa_1$ for the reduced string $x_1$ in the same way.
 
-The LCP array $lcp$ indicates the LCP-values of all the neighboring suffixes in $sa$, where $lcp[0] = 0$ and $lcp[i] = \ell$ for $i \in [1, n)$, where $\ell$ is the length of longest common prefix of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[i - 1])$.
+The LCP array $lcp$ records the LCP-value of each pair of neighboring suffixes in $sa$.
+We assume $lcp[0] = 0$ and let $lcp[i] = \ell$ for $i \in [1, n)$, where $\ell$ is the LCP-value of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[i - 1])$.
 
-All the suffixes in $sa$ are naturally partitioned into multiple buckets.
-Each bucket occupies a contiguous interval in $sa$ and contains all the suffixes starting with a certain character.
+All the suffixes in $sa$ are naturally grouped into multiple buckets.
+Each bucket occupies a contiguous interval in $sa$ and contains all the suffixes with a same heading character.
 Without loss of generality, we denote by ${\sf sa\_bkt}(c_0)$ the bucket for suffixes starting with $c_0$.
-Notice that ${\sf sa\_bkt}(c_0)$ can be further divided into two parts, where the left part ${\sf sa\_bkt_L}(c_0)$ and the right part ${\sf sa\_bkt_S}(c_0)$ contain the L-type and S-type suffixes, respectively.
-We also partition $lcp$ into buckets such that ${\sf lcp\_bkt}(c_0)$, ${\sf lcp\_bkt_L}(c_0)$ and ${\sf lcp\_bkt_S}(c_0)$ record the LCP-values of suffixes in ${\sf sa\_bkt}(c_0)$, ${\sf sa\_bkt_L}(c_0)$ and ${\sf sa\_bkt_S}(c_0)$, respectively.
+It should be noticed that ${\sf sa\_bkt}(c_0)$ can be divided into two parts where the left part ${\sf sa\_bkt_L}(c_0)$ and the right part ${\sf sa\_bkt_S}(c_0)$ contain the L-type and S-type suffixes, respectively.
+We also define ${\sf lcp\_bkt}(c_0)$, ${\sf lcp\_bkt_L}(c_0)$ and ${\sf lcp\_bkt_S}(c_0)$ on $lcp$ for each character in $\Sigma$.
 
@@
 	/* Reduction Phase */ \\
-	Sort S*-type substrings by the IS method.
+	Sort S*-type substrings by the IS method. \label{alg:1:ref:1}
 
-	Name the sorted S*-type substrings to produce $x_1$.
+	Name the sorted S*-type substrings to produce $x_1$. \label{alg:1:ref:2}
 
@@
 
-		Recursively call the reduction phase on $x_1$.
+		Recursively call the reduction phase on $x_1$.  \label{alg:1:ref:3}
 	}
@@
 
-		Compute $sa_1$ from $x_1$.
+		Compute $sa_1$ from $x_1$.  \label{alg:1:ref:4}
 	}
@@
 	/* Induction Phase */ \\
-	Sort suffixes by the IS method.
+	Sort suffixes by the IS method.  \label{alg:1:ref:5}
 
@@
 
-Algorithm~\ref{alg:1} shows the framework of an IS suffix sorting algorithm.
-At the very beginning, a reduction phase for sorting and naming S*-type substrings is called to produce the reduced string $x_1$.
-If there exist duplicate characters in $x_1$, the reduction phase is recursively called with $x_1$ as input; otherwise, all the S*-type suffixes are already sorted and $sa_1$ is directly computed from $x_1$.
-Afterward, an induction phase for sorting all the suffixes is called to produce $sa$ for $x$ at current recursion level.
-The induction phase is recursively called with $sa$ as input until reaching the top recursion level.
-During the execution of a reduction/induction phase, the order of substrings/suffixes is determined by comparing their heading characters and the ranks of their successors according to the IS principle, which involves a great number of random accesses to $x$ and $sa$.
-This can be done very fast if both $x$ and $sa$ wholly reside on RAM; otherwise, each access takes an individual I/O operation, leading to a severe performance degradation.
-This problem is solved in DSA-IS by performing two steps below during the reduction/induction phase:
+Algorithm~\ref{alg:1} shows the framework of any IS suffix sorting algorithm.
+In lines~\ref{alg:1:ref:1}-\ref{alg:1:ref:2}, a reduction phase for sorting and naming S*-type substrings is called to produce the reduced string $x_1$.
+If there exist duplicate characters in $x_1$, the reduction phase is called with $x_1$ as input at the higher recursion level in line~\ref{alg:1:ref:3}; otherwise, all the S*-type suffixes in $x$ are already sorted and $sa_1$ is directly computed from $x_1$ in line~\ref{alg:1:ref:4}.
+Afterward, an induction phase for sorting suffixes is called to produce $sa$ for $x$ at the current recursion level in line~\ref{alg:1:ref:5}.
+The induction phase is recursively called with $sa$ as input at the lower recursion level (if any).
+
+During the execution of a reduction/induction phase, all the substrings/suffixes are sorted by comparing their heading characters and the ranks of their successors according to the IS principle.
+This involves a great number of random accesses to $x$ and $sa$, which can be done very fast if both $x$ and $sa$ can wholly reside on RAM.
+However, if the size of input and output exceeds the capacity of internal memory, each access will take an individual I/O operation, leading to a performance degradation.
+The DSA-IS algorithm solves the problem by performing the following two steps during a reduction/induction phase:
 
 \begin{enumerate}[S1]
-	\item Split $x$ into blocks and sort substrings/suffixes of each block by calling SA-IS, the heading characters in use are copied to external memory in their access order.
+
+	\item Split $x$ into blocks and sort substrings/suffixes of each block by calling SA-IS.
+The heading characters in need are copied to external memory in their access order.
 \label{dsais_sorting_method:1}
-	\item Sort the substrings/suffixes of $x$ by their heading characters and the ranks of their successors.
-Organize the sorted substrings/suffixes in an external-memory heap and induce their predecessors into the heap according to the IS principle, where the heading characters of the induced substrings/suffixes are retrieved from external memory by sequential I/O operations.\label{dsais_sorting_method:2}
+
+	\item Sort the substrings/suffixes by their heading characters and the ranks of their successors in an external-memory heap.
+Scan the sorted substrings/suffixes in the heap and induce their predecessors into the heap, where the heading characters of the induced substrings/suffixes have been retrieved from external memory in advance by sequential I/O operations.
+\label{dsais_sorting_method:2}
 \end{enumerate}
 
-As shown in Section~\ref{sec:experiments}, our program for DSA-IS requires less disk space than that for eSAIS, but the former runs slower than the latter caused by the large I/O volume for sorting and naming S*-type substrings during the reduction phase.
-To improve the performance, we propose new substring sorting and naming methods in the next subsection.
+As shown in Section~\ref{sec:experiments}, our program for DSA-IS requires less disk space than that for eSAIS, but the former runs slower than the latter due to the large I/O volume for sorting and naming S*-type substrings during the reduction phase.
+To improve the performance, we propose new substring sorting and naming methods in the following.
 
@@
 \begin{enumerate}[S1']
-	\item Sort the long of each block by S\ref{dsais_sorting_method:1}. During the process, copy the short to external memory in their sorted order.~\label{dsaism_sorting_method:1}
+	\item Sort the long in each block by S\ref{dsais_sorting_method:1}. During the process, copy the short to external memory in their sorted order.~\label{dsaism_sorting_method:1}
 
-	\item Sort the long of $x$ by S\ref{dsais_sorting_method:2}. During the process, copy the leftmost $D$ characters of the long to external memory in their sorted order.~\label{dsaism_sorting_method:2}
+	\item Sort the long in $x$ by S\ref{dsais_sorting_method:2}. During the process, copy the leftmost $D$ characters of the long to external memory in their sorted order.~\label{dsaism_sorting_method:2}
 
@@
 
-After the first two steps, the short S*-type substrings of each block and the long S*-type substrings of $x$ are sorted and separately organized as a sequence in external memory.
-Assume $x$ is split into $k$ blocks, the multi-way sorter in S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap to cache the current smallest of each sequence and continually retrieves the top item from the heap to determine the lexical order of substrings from different sequences.
-The heap contains at most $k + 1$ substrings at any point of time and it compares any two substrings in $\mathcal{O}(D)$ time by a literal string comparison\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}.
-Theoretically, the sorting method has a good performance if the majority of S*-type substrings in $x$ are short with a small $D$.
-This is commonly satisfied in practice, because the average length of S*-type substrings is typically small in real-world datasets.
+After S\ref{dsaism_sorting_method:1}'-S\ref{dsaism_sorting_method:2}', the short S*-type substrings in each block and the long S*-type substrings in $x$ are sorted and separately organized as a sequence in external memory.
+Assume $x$ is split into $k$ blocks, the multi-way sorter in S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap to cache the current smallest of each sequence and continually retrieve the top item from the heap to determine the lexical order of substrings from different sequences.
+The heap contains at most $k + 1$ substrings at any point of time and it compares any two substrings in $\mathcal{O}(D)$ time by conducting a literal string comparison\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}.
+The above sorting method can achieve a good performance if the majority of S*-type substrings in $x$ are short, provided $D$ is small.
+This is commonly satisfied in real-world datasets.
 
-In what follows, we describe a method for naming the S*-type substrings during the above sorting process.
-The key point here is to check equality of two substrings successively popped from the heap in S\ref{dsaism_sorting_method:3}' as fast as possible.
-If either of the two substrings is short, we literally compare them in $\mathcal{O}(D)$ time.
-Otherwise, if both of them are long, then we assign names to all the long S*-type substrings using the technique proposed for SAIS-PQ in S\ref{dsaism_sorting_method:2}' and directly compare two successively popped long S*-type substrings by their names.
-This technique is based on the idea that two substrings are of the same name if their heading characters and the names of their successors are both equal\footnote{We refer interested readers to~\cite{Liu15} for more information.}.
-
+Next, we describe a method for naming the S*-type substrings when they are being sorted.
+The key point here is to check equality of two substrings successively popped from the heap in S\ref{dsaism_sorting_method:3}' immediately.
+If either of the two substrings is short, then we literally compare them in $\mathcal{O}(D)$ time to check their equality.
+Otherwise, the two substrings are both long, we first determine their names when inducing them in S\ref{dsaism_sorting_method:2}' and compare these names in $\mathcal{O}(1)$ time to check their equality.
+This naming technique was originally proposed for SAIS-PQ for merging the sorting and naming processes into a whole.
 The experimental results in Section~\ref{sec:experiments} indicate that, by using the new substring sorting and naming methods, the adapted DSA-IS, called DSA-IS+, only takes half as much disk space as eSAIS.
@@
 
-We describe below the main idea of the checking algorithm presented in~\cite{Dementiev2008a}.
+We describe below the main idea of the existing checker presented in~\cite{Dementiev2008a}.
 
@@
 The second condition indicates that the order of suffixes in $sa$ corresponds to that of their heading characters and successors.
-Because a comparison between any two suffixes can be done by comparing their heading characters and successors, the above conditions are sufficient and necessary for SA verification.
+Because any two suffixes can be sorted by comparing their heading characters and successors, the above conditions are sufficient and necessary for SA verification.
 
@@
 
-Remember that an IS suffix sorting algorithm recursively performs the reduction phase to compute the reduced string $s_1$ until $s_1$ contains no duplicate characters.
-Afterward, it produces $sa_1$ from $s_1$ and recursively perform the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase consists of the following steps:
+Recall that, an IS suffix sorting algorithm recursively perform the reduction phase to compute the reduced string $x_1$ until $x_1$ contains no duplicate characters.
+Afterward, it produces $sa_1$ from $x_1$ and recursively perform the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase consists of the following steps:
 
@@
 
-A running example of S\ref{induction_phase:2}-S\ref{induction_phase:4} is shown in Fig.~\ref{fig:example1}.
+A running example of S\ref{induction_phase:2}''-S\ref{induction_phase:4}'' is shown in Fig.~\ref{fig:example1}.
 Given $sa^*$ is already known, line 6 inserts each S*-type suffix into the right part of the corresponding bucket and arranges them according to their sorted order indicated by $sa^*$.
-For example, $\sf suf(2)$, $\sf suf(5)$, $\sf suf(8)$ are placed at the rightmost empty position of $\sf sa\_bkt(i)$ in sequence.
-Then, we find the leftmost position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes.
-To do this, we first check $sa[0] = 14$ (marked by the symbol @) and induce the predecessor of $\sf suf(14)$ in lines 8-9.
-Because $x[13] = i$ is L-type, we put $\sf suf(13)$ into the current leftmost empty position in $\sf sa\_bkt_L(i)$.
+For example, ${\sf suf}(8)$, ${\sf suf}(5)$, ${\sf suf}(2)$ are placed into the right part of ${\sf sa\_bkt}(i)$ and arranged in their sorted order indicated by $sa^*$.
+Next, we find the leftmost position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes.
+For this, we first check $sa[0] = 14$ (marked by the symbol @) and induce the predecessor of ${\sf suf}(14)$ in lines 8-9.
+Because $x[13] = i$ is L-type, we put ${\sf suf}(13)$ into the current leftmost empty position in ${\sf sa\_bkt_L}(i)$.
 To step through $sa$ in this way, we get all the L-type suffixes sorted in line 17.
 Afterward, we find the rightmost position of each bucket and scan $sa$ leftward for inducing the order of S-type suffixes.
-When scanning $sa[14] = 4$ in lines 19-20, we see $x[3] = i$ is S-type and thus put $\sf suf(3)$ into the current rightmost empty position in $\sf sa\_bkt_S(i)$.
+When scanning $sa[14] = 4$ in lines 19-20, we see $x[3] = i$ is S-type and thus put ${\sf suf}(3)$ into the current rightmost empty position in ${\sf sa\_bkt_S}(i)$.
 Following the same idea, we get all the S-type suffixes sorted in line 28.
@@
     (1) $sa^*$ is correctly computed. \\
-    (2) S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented at the top recursion level. \\
+    (2) S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented. \\
 \end{lemma}
@@
 
-Because S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented, we have $i < j \iff (x[sa[i]], p) < (x[sa[j]], q)$ following the IS principle.
-This satisfies the second condition of Lemma~\ref{lemma:1}.
+Because S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented, $i < j \iff (x[sa[i]], p) < (x[sa[j]], q)$. This satisfies the second condition of Lemma~\ref{lemma:1}.
 
 Further, suppose $sa[i] = sa[j]$ and $i \ne j$, then $sa[p] = sa[q]$ and $p \ne q$.
-Repeating this reasoning process by replacing $(i, j)$ with $(p, q)$ until $\sf suf(sa[i])$ is S*-type, then $sa^*$ contains duplicate elements.
-Because $sa^*$ is correctly computed, each element in $sa^*$ must be unique.
+Repeating this reasoning process by replacing $(i, j)$ with $(p, q)$ until ${\sf suf}(sa[i])$ is S*-type, then $sa^*$ must contain duplicate elements.
+However, each element in $sa^*$ is unique as $sa^*$ is corrected computed.
 This leads to a contradiction.
+
 \end{IEEEproof}
 
-The above lemma indicates that, if the first condition of Lemma~\ref{lemma:2} is true, we can check the correctness of $sa^*$ instead of $sa$ at the top recursion level to perform SA verification.
-It should be noticed that the code snippet for the induction phase at the top recursion level only consists of tens of C++ code lines.
+It should be noticed that, we can check the correctness of $sa^*$ instead of $sa$ at the top recursion level to perform SA verification if the first condition of Lemma~\ref{lemma:2} is always true.
+In fact, the code snippet for the induction phase at the top recursion level only consists of tens of C++ code lines.
 This is considerable smaller than the whole program for a disk-based suffix sorting algorithm.
-Following the idea, we assume S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented and propose two methods for checking computation errors caused by implementation bugs and hardware malfunctions.
-The algorithms designed by these methods can be seamlessly integrated into any IS suffix sorting algorithms for building and checking an SA at the same time.
+Following the idea, we assume S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented and propose two methods for checking computation errors caused by implementation bugs and other malfunctions.
+The algorithms designed by these methods can be seamlessly integrated into any IS suffix sorting algorithm for building and checking an SA at the same time.
 
@@
 The first condition of Lemma~\ref{lemma:3} can be checked by a sparse SA checker like~\cite{wu2017}.
-Notice that each suffix induced into $sa$ will be latter scanned for inducing the order of its predecessor during S\ref{induction_phase:3}''-S\ref{induction_phase:4}''.
-The second condition intends for detecting errors caused by hardware malfunctions and can be checked by determining whether the induced and scanned values for each suffix are equal.
+The second condition intends for detecting malfunctions other than implementation bugs (e.g., I/O errors).
+This is checked by determining whether the induced and scanned values for each suffix are equal\footnote{Notice that each suffix induced into $sa$ will be latter scanned for inducing the order of its predecessor during S\ref{induction_phase:3}'' and S\ref{induction_phase:4}''.
+}.
 The problem to be solved here is that when a suffix is induced into a bucket, its corresponding value in $sa$ may not be scanned at once.
 Our solution is to integrate the checking process into the building process, by ensuring the sequence of values placed at a bucket is identical to that scanned later.
-For the purpose, we use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end.
+For the purpose, we use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end of the induction phase.
 If the two fingerprints for each bucket are equal, then the second condition of Lemma~\ref{lemma:3} will be seen with a high probability.
@@
 It should be noticed that this method can be also applied to checking the LCP array when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}.
-
-As a result, we design Algorithm~\ref{alg:2} to probabilistically check $sa$ and $lcp$ during the induction phase at the top recursion level.
-We point out that ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$.
+As a result, we design Algorithm~\ref{alg:2} to probabilistically check $sa$ and $lcp$ at the top recursion level.
+In the algorithm, ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$.
 
@@
 
-As can be seen in Lemma~\ref{lemma:4}, Method B uses a different way to check the correctness of $sa^*$.
+According to Lemma~\ref{lemma:4}, Method B uses a different way to check the correctness of $sa^*$.
 Before our presentation, we first introduce a notation called $\overline{sa^*}$.
 The same as $sa^*$, $\overline{sa^*}$ also represents the suffix array for all the S*-type suffixes.
-Both of them are computed during the induction phase at the top recursion level.
-To be specific, $sa^*$ is computed from $sa_1$ and $\overline{sa^*}$ is retrieved from the constructed final $sa$.
+Both $sa^*$ and $\overline{sa^*}$ are computed during the induction phase at the top recursion level, where the former is calculated from $sa_1$ in S\ref{induction_phase:1}'' and the latter is calculated when inducing the S*-type suffixes into $sa$ in S\ref{induction_phase:4}''.
 
@@
 
-    Case 2: Suppose any two S*-type suffixes are not corrected sorted, i.e., $\sf suf(sa^*[i_0]) > \sf suf(sa^*[j_0])$ and $i_0 < j_0$.
-By condition (2), we have $\sf suf(\overline{sa^*}[i_0] > \sf suf(\overline{sa^*}[j_0])$.
-Given that the order of $\sf suf(\overline{sa^*}[i_0]$ and $\sf suf(\overline{sa^*}[j_0])$ are induced from ${\sf suf}(sa^*[i_1])$ and ${\sf suf}(sa^*[j_1])$, then ${\sf sub}(\overline{sa^*}[i_0], sa^*[i_1])$ and ${\sf sub}(\overline{sa^*}[j_0], sa^*[j_1])$ are two LMS-substrings and there must be ${\sf suf}(sa^*[i_1])>{\sf suf}(sa^*[j_1])$ and $i_1 < j_1$.
+    Case 2: Suppose any two S*-type suffixes are not corrected sorted, i.e., ${\sf suf}(sa^*[i_0]) > {\sf suf}(sa^*[j_0])$ and $i_0 < j_0$.
+By condition (2), we have ${\sf suf}(\overline{sa^*}[i_0] > {\sf suf}(\overline{sa^*}[j_0])$.
+Given that the order of ${\sf suf}(\overline{sa^*}[i_0])$ and ${\sf suf}(\overline{sa^*}[j_0])$ are induced from ${\sf suf}(sa^*[i_1])$ and ${\sf suf}(sa^*[j_1])$, then ${\sf sub}(\overline{sa^*}[i_0], sa^*[i_1])$ and ${\sf sub}(\overline{sa^*}[j_0], sa^*[j_1])$ are two S*-type substrings and there must be ${\sf suf}(sa^*[i_1])>{\sf suf}(sa^*[j_1])$ and $i_1 < j_1$.
 Repeating this reasoning process, because condition (1), we must see ${\sf suf}(sa^*[i_k])>{\sf suf}(sa^*[j_k])$ and ${\sf suf}(sa^*[i_{k+1}])<{\sf suf}(sa^*[j_{k+1}])$, where $i_k < j_k$ and $i_{k+1} < j_{k+1}$.
@@
 
-The first condition of Lemma~\ref{lemma:4} is naturally satisfied when computing $sa^*$ from $sa_1$.
-We reuse the fingerprinting function to probabilistically check the third condition.
-This method is also applied to checking the second condition by computing and comparing the fingerprints of $sa^*$ and $\overline{sa^*}$.
-Specifically, the fingerprint of $sa^*$ is calculated when it is being scanned in S$\ref{induction_phase:1}''$.
-Because $\overline{sa^*}$ is a subset of $sa$, the fingerprint of $\overline{sa^*}$ is calculated when the S-type suffixes are being induced in S$\ref{induction_phase:4}''$.
+The first condition of Lemma~\ref{lemma:4} is naturally satisfied when computing $sa^*$ from $sa_1$ in S\ref{induction_phase:1}.
+The second condition is probabilistically checked by computing and comparing the fingerprints of $sa^*$ and $\overline{sa^*}$.
+The fingerprint of $sa^*$ is computed when placing the elements of $sa^*$ into $sa$ in S\ref{induction_phase:2}, while the fingerprint of $\overline{sa^*}$ is computed when inducing the S*-type suffixes into $sa$.
 
@@
 
-        Compute the fingerprints of $sa^*$, ${\sf sa_{I1}}(c)$ and ${\sf sa_{S1}}(c)$ when inducing the order and the LCP-values of L-type suffixes. \\
+        Compute the fingerprints of $sa^*$, ${\sf sa_{I1}}(c)$ and ${\sf sa_{S1}}(c)$ when inducing the order of L-type suffixes. \\
 
-        Check if ${\sf sa_{I1}}(c) = {\sf sa_{S1}}(c)$ and ${\sf lcp_{I1}}(c) = {\sf lcp_{S1}}(c)$ by comparing their fingerprints. \\
+        Check if ${\sf sa_{I1}}(c) = {\sf sa_{S1}}(c)$ by comparing their fingerprints. \\
 
-        Compute the fingerprints of $\overline{sa^*}$, ${\sf sa_{I2}}(c)$ and ${\sf sa_{S2}}(c)$ when inducing the order and the LCP-values of S-type suffixes. \\
+        Compute the fingerprints of $\overline{sa^*}$, ${\sf sa_{I2}}(c)$ and ${\sf sa_{S2}}(c)$ when inducing the order of S-type suffixes. \\
 
-        Check if ${\sf sa_{I2}}(c) = {\sf sa_{S2}}(c)$p by comparing their fingerprints. \\
+        Check if ${\sf sa_{I2}}(c) = {\sf sa_{S2}}(c)$ by comparing their fingerprints. \\
 
@@
 
-The proposed two methods check equality of two arrays by comparing their fingerprints, which can be calculated using a hash function.
+The proposed two methods check equality of two arrays (or sub-arrays) by comparing their fingerprints, which can be calculated using a hash function.
 Notice that two equal arrays must share an identical hash value, but the inverse is not always true.
@@
 Given $L = 23$ and $\delta = 5$, we first compute the fingerprint of $A[0, 0]$ in line 4.
-Because $\sf fp(A[0, -1]) = 0$, then $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$.
-By iteratively computing the fingerprints of the prefixes of $A$ according to Formula~\ref{formula:2}, we finally obtain the fingerprints of the whole array in line 7.
+Because $\sf fp(A[0, -1]) = 0$, $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$.
+By iteratively computing the fingerprints of the prefixes of $A$, we finally obtain the fingerprint of $A$ in line 7.
 
@@
 
-The bottleneck of Algorithm~\ref{alg:2} and~\ref{alg:3} lie in checking the correctness of $sa^*$.
+Assuming the induction phase at the top recursion level is correctly implemented, the bottleneck of Algorithms~\ref{alg:2} and~\ref{alg:3} occur when checking $sa^*$.
 For Algorithm~\ref{alg:2}, we apply the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ within the sorting complexity.
 Specifically, given two suffixes, i.e.
-$\sf suf(i)$ and $\sf suf(j)$, and their LCP-value $d$, we compute and compare the fingerprints of $\sf sub(i, i + d - 1)$ and $\sf sub(j, j + d - 1)$.
-If the fingerprints are equal and the order of $x[i + d]$ and $x[j + d]$ corresponds to that of $\sf suf(i)$ and $\sf suf(j)$, then the suffixes are correctly sorted and their LCP-value is right with a high probability.
-Because at most one out of every two suffixes is S*-type (commonly one-third suffixes are S*-type in real-world data sets) and the checking process is only executed during the induction phase at the top recursion level, the checking overhead is much less than the building consumption.
-For Algorithm~\ref{alg:3}, the checking overhead is mainly caused by fingerprinting calculations and thus can be neglected compared with the building consumption.
+${\sf suf}(i)$ and ${\sf suf}(j)$, and their LCP-value $d$, we compute and compare the fingerprints of ${\sf sub}(i, i + d - 1)$ and ${\sf sub}(j, j + d - 1)$.
+If the fingerprints are equal and the order of $x[i + d]$ and $x[j + d]$ corresponds to that of ${\sf suf}(i)$ and ${\sf suf}(j)$, then the suffixes are correctly sorted and their LCP-value is right with a high probability.
+Because at most one out of every two suffixes is S*-type (commonly one-third in real-world data sets) and the checking process is only executed during the induction phase at the top recursion level, the checking overhead is much less than the building consumption.
+This conclusion is also true for Algorithm~\ref{alg:3}, where the checking overhead is mainly taken by computing fingerprints in need.
 
@@
 
-For implementation simplicity, we engineer DSA-IS and DSA-IS+ by the STXXL's containers~(vector, sorter, priority queue and stream).
+For performance comparison, we engineer DSA-IS and DSA-IS+ by the STXXL's containers~(vector, sorter, priority queue and stream).
 The experimental platform is a desktop computer equipped with an Intel Xeon E3-1220 V2 CPU, 4GiB RAM and 500GiB HD.
@@
 
-Because fSAIS is not available online, we use eSAIS as a baseline for analyzing the performance of DSA-IS and DSA-IS+.
+Because fSAIS is not available online, we use eSAIS as a baseline for analyzing the performance of DSA-IS and DSA-IS+, where the program for eSAIS is also implemented by the STXXL library.
 Fig.~\ref{fig:construction_performance1} shows a comparison between the programs for these three algorithms in terms of the investigated metrics.
@@
 In this way, our program can be further improved to achieve a space performance comparable to fSAIS.
-In our next paper, we will propose another disk-based sorting algorithm that has a space performance even better than fSAIS.
-It takes only $1n$ work space to run excluding the input and output.
+Our next paper will describe a novel disk-based IS suffix sorter that only takes 1n work space excluding the disk space for storing input and output.
 
@@
 
-We proposed two methods that enable any IS suffix sorting algorithm to build and check SA simultaneously.
-The program for our algorithm designed by Method B consumes negligible time and RAM space in comparison with that for the existing suffix sorting and checking algorithms on various real-world data sets.
-We also made our first attempt to improve the performance of DSA-IS using new substring and naming methods.
-The implementation for the adapted algorithm DSA-IS+ runs as fast as that for eSAIS and requires only half as much disk space as the latter.
-These results indicate the great potential of the IS method to design efficient solutions for SA constructing and verification.
+By assuming the induction phase at the top recursion level is correctly implemented, we proposed two methods that enable any IS suffix sorter to build and check an SA simultaneously.
+The probabilistic algorithm designed by the second method is rather lightweight, it takes negligible time and space compared with the existing IS suffix sorting and checking algorithms.
+We also made our first attempt to improve the space performance of DSA-IS using new substring and naming methods.
+Our experimental results show that the program for the adapted algorithm DSA-IS+ runs as fast as that for eSAIS and consumes only half as much disk space as the latter on various real-world datasets.
+We are now designing and implementing a novel IS suffix sorter that takes no more than 1n work space on external memory model.
+Theoretically, this suffix sorter has a better space performance than fSAIS under the same circumstances.
 

### 2017-08-09 new-draft/draft.tex
2018/08/09

@@
 \begin{lemma} \label{lemma:2}
-    $sa[0, n)$ is the SA for $x[0, n)$ if and only if the following conditions are satisfied: \\
+    For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for $x[0, n)$ if and only if the following conditions are satisfied at the top recursion level: \\
     (1) $sa^*$ is correctly computed. \\
@@
 
-    Case 2: Suppose any two S*-type suffixes are not corrected sorted, i.e., ${\sf suf}(sa^*[i_0]) > {\sf suf}(sa^*[j_0])$ and $i_0 < j_0$.
+    Case 2: Suppose any two S*-type suffixes are not correctly sorted, i.e., ${\sf suf}(sa^*[i_0]) > {\sf suf}(sa^*[j_0])$ and $i_0 < j_0$.
 By condition (2), we have ${\sf suf}(\overline{sa^*}[i_0] > {\sf suf}(\overline{sa^*}[j_0])$.
@@
 
-The first condition of Lemma~\ref{lemma:4} is naturally satisfied when computing $sa^*$ from $sa_1$ in S\ref{induction_phase:1}.
+The first condition of Lemma~\ref{lemma:4} is naturally satisfied when computing $sa^*$ from $sa_1$ in S\ref{induction_phase:1}''.
 The second condition is probabilistically checked by computing and comparing the fingerprints of $sa^*$ and $\overline{sa^*}$.
-The fingerprint of $sa^*$ is computed when placing the elements of $sa^*$ into $sa$ in S\ref{induction_phase:2}, while the fingerprint of $\overline{sa^*}$ is computed when inducing the S*-type suffixes into $sa$.
+The fingerprint of $sa^*$ is computed when placing the elements of $sa^*$ into $sa$ in S\ref{induction_phase:2}'', while the fingerprint of $\overline{sa^*}$ is computed when inducing the S*-type suffixes into $sa$.
 
@@
 Specifically, given two suffixes, i.e.
-${\sf suf}(i)$ and ${\sf suf}(j)$, and their LCP-value $d$, we compute and compare the fingerprints of ${\sf sub}(i, i + d - 1)$ and ${\sf sub}(j, j + d - 1)$.
-If the fingerprints are equal and the order of $x[i + d]$ and $x[j + d]$ corresponds to that of ${\sf suf}(i)$ and ${\sf suf}(j)$, then the suffixes are correctly sorted and their LCP-value is right with a high probability.
+${\sf suf}(i)$ and ${\sf suf}(j)$, and their LCP-value $\ell$, we compute and compare the fingerprints of ${\sf sub}(i, i + \ell - 1)$ and ${\sf sub}(j, j + \ell - 1)$.
+If the fingerprints are equal and the order of $x[i + \ell]$ and $x[j + \ell]$ corresponds to that of ${\sf suf}(i)$ and ${\sf suf}(j)$, then the suffixes are correctly sorted and their LCP-value is right with a high probability.
 Because at most one out of every two suffixes is S*-type (commonly one-third in real-world data sets) and the checking process is only executed during the induction phase at the top recursion level, the checking overhead is much less than the building consumption.

### 2017-08-10 new-draft/draft.tex
2017/08/10

@@
 
-\title{Building and Checking Suffix and LCP Arrays Using Induced Sorting Method}
+\title{Building and Checking a Suffix Array by Induced Sorting}
 
@@
 
-Suffix and longest common prefix~(LCP) arrays can be built by the induced sorting~(IS) method on both internal and external memory models.
-We propose two methods that enable any IS builder to build and check suffix and LCP arrays simultaneously.
-The first method is for checking both suffix and LCP arrays, while the second method is for checking SA only.
-By combining the Karp-Rabin fingerprinting techinque into our methods, we design two algorithms that perform checking correctly with a negligible error probability.
-Theoretically, the algorithm designed by the first method has a sorting complexity, while the algorithm designed by the second method only takes linear time to run.
-We integrate the algorithm designed by the second method into the existing disk-based suffix sorting algorithm DSA-IS and implement their combination for performance evaluation.
+A suffix array can be built by the induced sorting~(IS) method on both internal and external memory models.
+We propose two methods for checking a suffix array when it is being built by any IS suffix sorter, i.e.
+build and check a suffix array by the IS method simultaneously.
+In particular, the first method can be extended to check both suffix and LCP arrays.
+By combining the Karp-Rabin fingerprinting technique with our methods, we design two algorithms that perform checking correctly with a negligible error probability.
+The algorithm designed by the first method has an integer sorting complexity, while the algorithm designed by the second method runs in linear time.
+We integrate the algorithm designed by the second method into the existing disk-based suffix sorting algorithm DSA-IS and implement it for performance analysis.
+From our experiments, the checking overhead is considerably smaller than the building consumption.
+
+%A suffix array can be built by the induced sorting~(IS) method on both internal and external memory models.
+We propose two methods to enable the IS to check a suffix array when it is built, i.e.
+build and check a suffix array simultaneously.
+In particular, the first method can be exteneded to check both suffix and LCP arrays.
+By combining the Karp-Rabin fingerprinting techinque with our methods, we design two algorithms that can check correctly with a negligible error probability.
+The algorithm designed by the first method has an integer sorting complexity, while the algorithm designed by the second method runs in linear time.
+We integrate the algorithm designed by the second method into the existing disk-based suffix sorting algorithm DSA-IS and implement it for performance evaluation.
 From our experiments, the checking overhead is considerably smaller than the building consumption.
@@
 
+{\color{red}
+
+In the accepted paper, the LCP-value $\ell$ of ${\sf suf}(i)$ and ${\sf suf}(j)$ is assumed to be right if ${\sf sub}(i, i + \ell - 1) = {\sf sub}(j, j + \ell - 1)$ and $x[i + \ell] \ne x[j + \ell]$.
+Given that $\ell$ is correct, the lexical order of ${\sf suf}(i)$ and ${\sf suf}(j)$ is checked by comparing $x[i + \ell]$ and $x[j + \ell]$.
+
+In this paper, both the lexical order and the LCP-value of two neighboring suffixes in the given suffix array is checked at the top recursion level following the IS principle.
+
+}
+
 \section{Introduction}\label{sec:introduction}
 
-A suffix array~(SA) can be built within linear time and space by the internal-memory algorithm SA-IS~\cite{Nong11}.
-According to the IS principle, the order of two suffixes is determined by comparing their heading characters and successors in sequence, where the order of their successors is assumed to be determined in advance.
+A suffix array~(SA) can be built within linear time and space by the internal-memory algorithm SA-IS~\cite{Nong11} using the induced sorting method.
+According to the IS principle, the order of two suffixes is determined by comparing their heading characters and successors in sequence, where the order of their successors is determined recursively.
 Recently, the IS method has been also applied to designing three disk-based suffix sorting algorithms eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}, where eSAIS can build both suffix and LCP arrays at the same time.
-These works have better time complexities than the other alternatives ~(e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}), but they all suffer from a bottleneck due to the large disk space for obtaining the heading characters of unsorted suffixes and the ranks of their sorted successors in a disk-friendly way.
-It is reported that the average peak disk use to construct an SA encoded by 40-bit integers for pSAscan is only $7.5n$, while that for eSAIS, DSA-IS and SAIS-PQ are $24n$, $18n$ and $15n$, respectively.
-The poor space performance of the IS algorithms is mainly because that their current programs fail to free the disk space for temporary data even when the data is no longer to use.
-A dramatic improvement can be achieved by storing temporary data in multiple files and deleting each file when the data residing on it is not needed any more.
-This technique has been used to implement a novel IS suffix sorting algorithm fSAIS~\cite{Karkkainen2017}, where the engineering of fSAIS takes no more than $8n$ disk space.
-This indicates a great potential for optimizing the performances of DSA-IS and SAIS-PQ.
+Compared to the others designed by different methods (e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}), these algorithms runs fast with linear I/O complexity.
+However, these algorithms currently suffer from a bottleneck due to the large disk space for obtaining the heading characters of unsorted suffixes and the ranks of their sorted successors in a disk-friendly way.
+It is reported that the average disk use for constructing an SA encoded by 40-bit integers for pSAscan is only $7.5n$ bytes, while that for eSAIS, DSA-IS and SAIS-PQ are $24n$, $18n$ and $15n$ bytes, respectively, where $n$ is the size of input string.
+The poor space performance observed from these IS algorithms is mainly because the current programs for these algorithms fail to immediately free the disk space for temporary data even when the data is no longer needed.
+A dramatic improvement can be achieved by storing temporary data in multiple files and deleting a file when the data in the file is not needed any more.
+This trick has been used by fSAIS~\cite{Karkkainen2017} to engineer the external-memory IS method for better performance, where the peak disk use is $8n$ bytes for constructing an SA of 40-bit integers in the given experiments.
 
+While the research on developing various SA builders using the IS method is involving, the developed software are becoming complex and appear as open-source without guarantee.
 A constructed SA should be checked to detect potential errors caused by implementation bugs and other malfunctions.
-The software packages for DC3 and eSAIS provide users a checker based on the idea presented in~\cite{Dementiev2008a}.
-When running on external-memory model, the cost taken by this checker is rather high, because it performs two passes of integer sorts for arranging $\mathcal{O}(n)$ fixed-size tuples.
-In this paper, we describe two methods that enable any IS builder to build and check suffix and LCP arrays simultaneously.
-The first method is for checking both suffix and LCP arrays, while the second method is for checking SA only.
+The software packages for DC3 and eSAIS provide a checker based on the idea presented in~\cite{Dementiev2008a}.
+When running on external-memory model, the cost taken  by this checker is rather high as it performs two passes of integer sorts for arranging $\mathcal{O}(n)$ fixed-size tuples.
+In this paper, we propose two methods to enable the IS method to build and check an SA simultaneously.
+In particular, the first method can be extended to check both suffix and LCP arrays.
 We employ the Karp-Rabin fingerprinting technique~\cite{Karp1987} to design two probabilistic algorithms, in terms of that their checking results are wrong with a negligible probability.
 For analysis, we first use new substring sorting and naming methods to improve the design of DSA-IS and then combine the second checking method into the adapted DSA-IS for evaluating the checking overhead.
-Our experimental results indicate that the time, space and I/O volume for checking SA is considerable smaller than that for building SA.
-This implies that the IS method can be used to design efficient solutions for SA verification as well.
+Our experimental results indicate that the time, space and I/O volume for checking SA is considerably smaller than that for building SA.
 
@@
 
-Partition $x$ into multiple S*-type substrings.
-Each substring only contains two S*-type characters and any two neighboring substrings overlap an S*-type character.
-Produce a reduced string $x_1$ by replacing each S*-type substring with its name, which represents the rank of the corresponding S*-type substring among all.
-In the following paragraphs, we use "rank" and "name" interchangeably to indicate the lexical order of the corresponding substring.
+Partition $x$ into multiple S*-type substrings, each substring contains two S*-type characters and any two neighboring substrings overlap an S*-type character.
+We produce a reduced string $x_1$ by replacing each substring with its name, where the name indicates the rank of the	 substring among all.
+In this paper, we use "rank" and "name" interchangeably to represent the lexical order of substrings.
 
@@
 Each bucket occupies a contiguous interval in $sa$ and contains all the suffixes with a same heading character.
-Without loss of generality, we denote by ${\sf sa\_bkt}(c_0)$ the bucket for suffixes starting with $c_0$.
-It should be noticed that ${\sf sa\_bkt}(c_0)$ can be divided into two parts where the left part ${\sf sa\_bkt_L}(c_0)$ and the right part ${\sf sa\_bkt_S}(c_0)$ contain the L-type and S-type suffixes, respectively.
-We also define ${\sf lcp\_bkt}(c_0)$, ${\sf lcp\_bkt_L}(c_0)$ and ${\sf lcp\_bkt_S}(c_0)$ on $lcp$ for each character in $\Sigma$.
+Without loss of generality, we denote by ${\sf sa\_bkt}(c_0)$ the bucket for suffixes starting with $c_0$, where ${\sf sa\_bkt}(c_0)$ can be naturally divided into two parts ${\sf sa\_bkt_L}(c_0)$ and ${\sf sa\_bkt_S}(c_0)$ containing L-type and S-type suffixes, respectively.
+We also define ${\sf lcp\_bkt}(c_0)$, ${\sf lcp\_bkt_L}(c_0)$ and ${\sf lcp\_bkt_S}(c_0)$ on $lcp$ for $c_0 \in \Sigma$.
 
@@
 
-Algorithm~\ref{alg:1} shows the framework of any IS suffix sorting algorithm.
-In lines~\ref{alg:1:ref:1}-\ref{alg:1:ref:2}, a reduction phase for sorting and naming S*-type substrings is called to produce the reduced string $x_1$.
-If there exist duplicate characters in $x_1$, the reduction phase is called with $x_1$ as input at the higher recursion level in line~\ref{alg:1:ref:3}; otherwise, all the S*-type suffixes in $x$ are already sorted and $sa_1$ is directly computed from $x_1$ in line~\ref{alg:1:ref:4}.
-Afterward, an induction phase for sorting suffixes is called to produce $sa$ for $x$ at the current recursion level in line~\ref{alg:1:ref:5}.
-The induction phase is recursively called with $sa$ as input at the lower recursion level (if any).
+Algorithm~\ref{alg:1} shows the framework of an IS suffix sorting algorithm.
+In lines~\ref{alg:1:ref:1}-\ref{alg:1:ref:2}, a reduction phase for sorting and naming S*-type substrings is executed to produce the reduced string $x_1$.
+If there exist duplicate characters in $x_1$, then the reduction phase is recursively called with $x_1$ as input in line~\ref{alg:1:ref:3} and executed at the higher recursion level; otherwise, the S*-type suffixes in $x$ are already sorted and $sa_1$ is directly computed from $x_1$ in line~\ref{alg:1:ref:4}.
+Afterward, an induction phase for sorting suffixes is executed to produce $sa$ at the current recursion level in line~\ref{alg:1:ref:5}.
+The induction phase is called recursively with $sa$ as input and executed at the lower recursion level until the final $sa$ is generated.
 
@@
 This involves a great number of random accesses to $x$ and $sa$, which can be done very fast if both $x$ and $sa$ can wholly reside on RAM.
-However, if the size of input and output exceeds the capacity of internal memory, each access will take an individual I/O operation, leading to a performance degradation.
-The DSA-IS algorithm solves the problem by performing the following two steps during a reduction/induction phase:
+However, if the size of input and output exceeds the capacity of internal memory, each access will take an individual I/O operation to disk, leading to a performance degradation.
+The DSA-IS algorithm solves this problem by performing the following two steps during a reduction/induction phase:
 
@@
 
-	\item Sort the substrings/suffixes by their heading characters and the ranks of their successors in an external-memory heap.
-Scan the sorted substrings/suffixes in the heap and induce their predecessors into the heap, where the heading characters of the induced substrings/suffixes have been retrieved from external memory in advance by sequential I/O operations.
-\label{dsais_sorting_method:2}
+	\item Sort the substrings/suffixes by their heading characters and the ranks of their successors in an external-memory heap. \label{dsais_sorting_method:2}
 \end{enumerate}
 
+Because all the substrings/suffixes in the heap are sorted already, S\ref{dsais_sorting_method:2} determines the rank of a substring/suffix when popping it from the heap.
+When inducing the predecessor of current top substring/suffix, the heading character in need is retrieved from external memory by a sequential I/O operation.
 As shown in Section~\ref{sec:experiments}, our program for DSA-IS requires less disk space than that for eSAIS, but the former runs slower than the latter due to the large I/O volume for sorting and naming S*-type substrings during the reduction phase.
@@
 \begin{enumerate}[S1']
-	\item Sort the long in each block by S\ref{dsais_sorting_method:1}. During the process, copy the short to external memory in their sorted order.~\label{dsaism_sorting_method:1}
+	\item Sort the long in each block, copy the heading characters in need to external memory by S\ref{dsais_sorting_method:1}.
+Meanwhile, sort the short in each block, copy the short to external memory in their sorted order.~\label{dsaism_sorting_method:1}
 
-	\item Sort the long in $x$ by S\ref{dsais_sorting_method:2}. During the process, copy the leftmost $D$ characters of the long to external memory in their sorted order.~\label{dsaism_sorting_method:2}
+	\item Sort the long in $x$ by S\ref{dsais_sorting_method:2}, copy the leftmost $D$ characters of the long to external memory in their sorted order.~\label{dsaism_sorting_method:2}
 
-	\item Merge the short and long by a multi-way sorter.~\label{dsaism_sorting_method:3}
+	\item Merge the short and the long by a multi-way sorter.~\label{dsaism_sorting_method:3}
 \end{enumerate}
 
-After S\ref{dsaism_sorting_method:1}'-S\ref{dsaism_sorting_method:2}', the short S*-type substrings in each block and the long S*-type substrings in $x$ are sorted and separately organized as a sequence in external memory.
-Assume $x$ is split into $k$ blocks, the multi-way sorter in S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap to cache the current smallest of each sequence and continually retrieve the top item from the heap to determine the lexical order of substrings from different sequences.
-The heap contains at most $k + 1$ substrings at any point of time and it compares any two substrings in $\mathcal{O}(D)$ time by conducting a literal string comparison\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}.
-The above sorting method can achieve a good performance if the majority of S*-type substrings in $x$ are short, provided $D$ is small.
+The sorted short S*-type substrings in each block are organized as a sequence in external memory after S\ref{dsaism_sorting_method:1}', while the sorted long S*-type substrings in $x$ are also organized as a sequence in external memory after S\ref{dsaism_sorting_method:2}'.
+Assume $x$ is split into $k$ blocks, the multi-way sorter in S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap to determine the lexical order of substrings belonging to different sequences.
+The heap contains at most $k + 1$ substrings at any point of time.
+It compares any two substring in $\mathcal{O}(D)$ time by conducting a literal string comparison\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}.
+
+The above sorting method has a good performance if $D$ is small and the majority of S*-type substrings in $x$ are short.
 This is commonly satisfied in real-world datasets.
-
-Next, we describe a method for naming the S*-type substrings when they are being sorted.
-The key point here is to check equality of two substrings successively popped from the heap in S\ref{dsaism_sorting_method:3}' immediately.
-If either of the two substrings is short, then we literally compare them in $\mathcal{O}(D)$ time to check their equality.
-Otherwise, the two substrings are both long, we first determine their names when inducing them in S\ref{dsaism_sorting_method:2}' and compare these names in $\mathcal{O}(1)$ time to check their equality.
-This naming technique was originally proposed for SAIS-PQ for merging the sorting and naming processes into a whole.
-The experimental results in Section~\ref{sec:experiments} indicate that, by using the new substring sorting and naming methods, the adapted DSA-IS, called DSA-IS+, only takes half as much disk space as eSAIS.
+We next describe a method for naming the S*-type substrings when they are being sorted according to S\ref{dsaism_sorting_method:1}'-S\ref{dsaism_sorting_method:3}'.
+The key point is to check equality of two suffixes successively popped from the heap.
+For this, we literally compare the two substrings in $\mathcal{O}(D)$ time if either of them is short; otherwise, we compare their names in $\mathcal{O}(1)$ time instead.
+Recall that the rank of each long S*-type substring is determined when inducing its predecessor, this value is reused to represent its name herein\footnote{A similar technique has been used to merge the substring sorting and naming process in SAIS-PQ.}.
+The experimental results in Section~\ref{sec:experiments} demonstrate that, by using the new substring sorting and naming methods, the adapted DSA-IS, called DSA-IS+, only takes half as much disk space as eSAIS.
 
@@
 
-\begin{IEEEproof}The first condition indicates that $sa$ is a permutation of all the suffixes in $x$.
-The second condition indicates that the order of suffixes in $sa$ corresponds to that of their heading characters and successors.
-Because any two suffixes can be sorted by comparing their heading characters and successors, the above conditions are sufficient and necessary for SA verification.
+\begin{IEEEproof} If condition (1) is true, then $sa$ is a permutation of all the suffixes in $x$.
+If condition (2) is true, then the lexical order of any two neighboring suffixes in $sa$ is determined by comparing their heading characters and the ranks of their successors.
+As a result, all the suffixes in $x$ are sorted according to the lexical order of their heading characters and successors.
 
@@
 
-The disk-based implementation of this checker conducts two passes of integer sorts and each sort arranges the order of $\mathcal{O}(n)$ fixed-size tuples using external memory.
+The disk-based implementation of this checker conducts two passes of integer sorts and each sort arranges the order of $\mathcal{O}(n)$ fixed-size tuples in external memory.
 As can be observed from Section~\ref{sec:experiments}, the peak disk use and the I/O volume for an SA encoded by 40-bit integers are around $26n$ and $53n$, respectively.
@@
 
-Recall that, an IS suffix sorting algorithm recursively perform the reduction phase to compute the reduced string $x_1$ until $x_1$ contains no duplicate characters.
-Afterward, it produces $sa_1$ from $x_1$ and recursively perform the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase consists of the following steps:
+Recall that, an IS suffix sorting algorithm recursively executes the reduction phase to compute the reduced string $x_1$ until $x_1$ contains no duplicate characters.
+Afterward, it produces $sa_1$ from $x_1$ and recursively executes the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase consists of the following steps:
 
@@
 A running example of S\ref{induction_phase:2}''-S\ref{induction_phase:4}'' is shown in Fig.~\ref{fig:example1}.
-Given $sa^*$ is already known, line 6 inserts each S*-type suffix into the right part of the corresponding bucket and arranges them according to their sorted order indicated by $sa^*$.
-For example, ${\sf suf}(8)$, ${\sf suf}(5)$, ${\sf suf}(2)$ are placed into the right part of ${\sf sa\_bkt}(i)$ and arranged in their sorted order indicated by $sa^*$.
-Next, we find the leftmost position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes.
+Given $sa^*$ is already known, line 6 inserts each S*-type suffix into the corresponding bucket.
+For example, ${\sf suf}(2)$, ${\sf suf}(5)$, ${\sf suf}(8)$ are sequentially placed into the current rightmost empty position of ${\sf sa\_bkt}(i)$, where the insertion order corresponds to their sorted order indicated by $sa^*$.
+The next step is to find the leftmost position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes.
 For this, we first check $sa[0] = 14$ (marked by the symbol @) and induce the predecessor of ${\sf suf}(14)$ in lines 8-9.
@@
 Following the same idea, we get all the S-type suffixes sorted in line 28.
-Based on the discussion, we show in Lemma~\ref{lemma:2} a set of sufficient and necessary conditions for SA verification.
+Following the discussion, we show in Lemma~\ref{lemma:2} a set of sufficient and necessary conditions for checking SA.
 
@@
 
-It should be noticed that, we can check the correctness of $sa^*$ instead of $sa$ at the top recursion level to perform SA verification if the first condition of Lemma~\ref{lemma:2} is always true.
+Lemma~\ref{lemma:2} shows that we can check the correctness of $sa^*$ instead of $sa$ at the top recursion level to perform SA verification if the first condition of Lemma~\ref{lemma:2} is always true.
 In fact, the code snippet for the induction phase at the top recursion level only consists of tens of C++ code lines.
-This is considerable smaller than the whole program for a disk-based suffix sorting algorithm.
+This is considerably smaller than the whole program for a disk-based suffix sorting algorithm.
 Following the idea, we assume S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented and propose two methods for checking computation errors caused by implementation bugs and other malfunctions.
@@
 
-Method A is based on Lemma~\ref{lemma:3}, which is extended from Lemma~\ref{lemma:2}.
+Method A is based on Lemma~\ref{lemma:3}, which is an extension of Lemma~\ref{lemma:2}.
 
 \begin{lemma} \label{lemma:3}
-	Assume the induction phase is correctly implemented at the top recursion level, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied: \\
+	Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ of an IS suffix sorter is the SA for the input string $x[0,  n)$ if the following conditions are satisfied: \\
 	(1) $sa^*$ is correct. \\
@@
 The problem to be solved here is that when a suffix is induced into a bucket, its corresponding value in $sa$ may not be scanned at once.
-Our solution is to integrate the checking process into the building process, by ensuring the sequence of values placed at a bucket is identical to that scanned later.
+Our solution is to ensure the sequence of values placed at a bucket is identical to that scanned latter.
 For the purpose, we use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end of the induction phase.
@@
 As a result, $sa$ can be built and probabilistically checked at the same time.
-It should be noticed that this method can be also applied to checking the LCP array when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}.
-As a result, we design Algorithm~\ref{alg:2} to probabilistically check $sa$ and $lcp$ at the top recursion level.
-In the algorithm, ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$.
+
+It was reported in~\cite{Fischer11, Bingmann12} that the IS method can be used to construct LCP array as well.
+For any two neighboring suffixes in $sa$, their LCP-value is zero if they start with different characters; otherwise, the LCP-value of them is one greater than that of their successors.
+This implies the possibility for extending Method A to check both suffix and LCP arrays.
+We design Algorithm~\ref{alg:2} to check $sa$ and $lcp$ when they are induced from $sa^*$ and $lcp^*$ at the top recursion level of an IS builder.
+In this algorithm, ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$.
+These values are used to check the second condition of Lemma~\ref{lemma:3} later.
+We also compute ${\sf lcp_{I1}}(c)/{\sf lcp_{I2}}(c)$ and ${\sf lcp_{S1}}(c)/{\sf lcp_{S2}}(c)$ and check their equality in the meantime.
 
@@
 \begin{lemma} \label{lemma:4}
-    Assume the induction phase is correctly implemented at the top recursion level, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied:
+    Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ of an IS suffix sorter is the SA for the input string $x[0, n)$ if the following conditions are satisfied:
 
@@
 
-    We only prove the sufficiency as the necessity is clear.
+    We prove the sufficiency according to the following two cases.
 
-    Case 1: Suppose all the S*-type suffixes are correctly sorted in $sa^*$, then $sa$ is right according to Lemma~\ref{lemma:3}.
+    Case 1: Suppose all the S*-type suffixes are correctly sorted in $sa^*$, then $sa$ is right based on Lemma~\ref{lemma:3}.
 
@@
 The second condition is probabilistically checked by computing and comparing the fingerprints of $sa^*$ and $\overline{sa^*}$.
-The fingerprint of $sa^*$ is computed when placing the elements of $sa^*$ into $sa$ in S\ref{induction_phase:2}'', while the fingerprint of $\overline{sa^*}$ is computed when inducing the S*-type suffixes into $sa$.
+The fingerprint of $sa^*$ is computed when placing the elements of $sa^*$ into $sa$ in S\ref{induction_phase:2}'', while the fingerprint of $\overline{sa^*}$ is computed when inducing the S*-type suffixes into $sa$ in S\ref{induction_phase:4}''.
+We design Algorithm~\ref{alg:3} to check if the final SA observes Lemma~\ref{lemma:4}.
 
@@
 
+The proposed two methods check equality of two arrays (or sub-arrays) by comparing their fingerprints, which can be calculated using a hash function.
+Notice that two equal arrays must be mapped to an identical hash value, but the inverse is not always true.
+To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting function to compute these hash values, called fingerprints,	 following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$.
+By setting $L$ to a large value, the error probability can be reduced to a negligible level.
+In Fig.~\ref{fig:example2}, we depict an example for computing the integer array $A$ identical to $sa^*$ and $\overline{sa^*}$ in Fig.~\ref{fig:example1}.
+Given $L = 23$ and $\delta = 5$, we first compute the fingerprint of $A[0, 0]$ in line 4.
+Because $\sf fp(A[0, -1]) = 0$, $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$.
+By iteratively computing the fingerprints of the prefixes of $A$, we finally obtain the fingerprint of $A$ in line 7.
+
 \begin{formula} \label{formula:1}${\sf fp}(A[0, -1]) = 0$.
@@
 \end{formula}
-
-The proposed two methods check equality of two arrays (or sub-arrays) by comparing their fingerprints, which can be calculated using a hash function.
-Notice that two equal arrays must share an identical hash value, but the inverse is not always true.
-To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting function to compute the required values following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$.
-By setting $L$ to a large value, the error probability can be reduced to a negligible level.
-In Fig.~\ref{fig:example2}, we depict an example for computing the integer array $A$, where $A$ is identical to $sa^*$ and $\overline{sa^*}$ in Fig.~\ref{fig:example1}.
-Given $L = 23$ and $\delta = 5$, we first compute the fingerprint of $A[0, 0]$ in line 4.
-Because $\sf fp(A[0, -1]) = 0$, $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$.
-By iteratively computing the fingerprints of the prefixes of $A$, we finally obtain the fingerprint of $A$ in line 7.
 
@@
 
-Assuming the induction phase at the top recursion level is correctly implemented, the bottleneck of Algorithms~\ref{alg:2} and~\ref{alg:3} occur when checking $sa^*$.
-For Algorithm~\ref{alg:2}, we apply the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ within the sorting complexity.
-Specifically, given two suffixes, i.e.
-${\sf suf}(i)$ and ${\sf suf}(j)$, and their LCP-value $\ell$, we compute and compare the fingerprints of ${\sf sub}(i, i + \ell - 1)$ and ${\sf sub}(j, j + \ell - 1)$.
-If the fingerprints are equal and the order of $x[i + \ell]$ and $x[j + \ell]$ corresponds to that of ${\sf suf}(i)$ and ${\sf suf}(j)$, then the suffixes are correctly sorted and their LCP-value is right with a high probability.
-Because at most one out of every two suffixes is S*-type (commonly one-third in real-world data sets) and the checking process is only executed during the induction phase at the top recursion level, the checking overhead is much less than the building consumption.
-This conclusion is also true for Algorithm~\ref{alg:3}, where the checking overhead is mainly taken by computing fingerprints in need.
+Assume the induction phase at the top recursion level is correctly implemented, Algorithm~\ref{alg:2} employs the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ and $lcp^*$.
+To be specific, for two neighboring suffixes ${\sf suf}(i)$ and ${\sf suf}(j)$ in $sa^*$, their LCP-value $\ell$ in $lcp^*$ is right if ${\sf sub}(i, i + \ell - 1) = {\sf sub}(j, j + \ell - 1)$ and $x[i + \ell] \ne x[j + \ell]$, where the equality of ${\sf sub}(i, i + \ell - 1)$ and ${\sf sub}(j, j + \ell - 1)$ is checked by comparing the fingerprints of these substrings.
+Given that $\ell$ is correct, the lexical order of ${\sf suf}(i)$ and ${\sf suf}(j)$ is checked by comparing $x[i + \ell]$ with $x[j + \ell]$.
+This method computes the fingerprints of all the target substrings within a sorting complexity proportional to $n_1$.
+Because at most one out of every two suffixes is S*-type, the checking overhead is much less than building consumption.
+By contrast, Algorithm~\ref{alg:3} takes overall $\mathcal{O}(n)$ to compute all the required fingerprints.
 

### 2017-08-11 new-draft/draft.tex
2017/08/11

@@
 
-In the accepted paper, the LCP-value $\ell$ of ${\sf suf}(i)$ and ${\sf suf}(j)$ is assumed to be right if ${\sf sub}(i, i + \ell - 1) = {\sf sub}(j, j + \ell - 1)$ and $x[i + \ell] \ne x[j + \ell]$.
-Given that $\ell$ is correct, the lexical order of ${\sf suf}(i)$ and ${\sf suf}(j)$ is checked by comparing $x[i + \ell]$ and $x[j + \ell]$.
+The LCP-value $\ell$ of ${\sf suf}(i)$ and ${\sf suf}(j)$ is right if ${\sf sub}(i, i + \ell - 1) = {\sf sub}(j, j + \ell - 1)$ and $x[i + \ell] \ne x[j + \ell]$.
+In the accepted paper, we check equality of ${\sf sub}(i, i + \ell - 1)$ and ${\sf sub}(j, j + \ell - 1)$ by comparing their fingerprints.
+Given that $\ell$ is correct, the lexical order of ${\sf suf}(i)$ and ${\sf suf}(j)$ is determined by comparing $x[i + \ell]$ and $x[j + \ell]$.
 
@@
 
-While the research on developing various SA builders using the IS method is involving, the developed software are becoming complex and appear as open-source without guarantee.
+While the research on developing various SA builders using the IS method is evolving, the developed software are becoming complex and appear as open-source without guarantee.
 A constructed SA should be checked to detect potential errors caused by implementation bugs and other malfunctions.
@@
 Partition $x$ into multiple S*-type substrings, each substring contains two S*-type characters and any two neighboring substrings overlap an S*-type character.
-We produce a reduced string $x_1$ by replacing each substring with its name, where the name indicates the rank of the	 substring among all.
-In this paper, we use "rank" and "name" interchangeably to represent the lexical order of substrings.
+We produce a reduced string $x_1$ by replacing each substring with its name, where the name represents the rank of this	 substring among all.
+In this paper, we use "rank" and "name" interchangeably to indicate the lexical order of substrings.
 
@@
 
-	\item Split $x$ into blocks and sort substrings/suffixes of each block by calling SA-IS.
+	\item Split $x$ into blocks and sort substrings/suffixes in each block by calling SA-IS.
 The heading characters in need are copied to external memory in their access order.
@@
 
-	\item Sort the substrings/suffixes by their heading characters and the ranks of their successors in an external-memory heap. \label{dsais_sorting_method:2}
+	\item Sort the substrings/suffixes in $x$ by their heading characters and the ranks of their successors in an external-memory heap. \label{dsais_sorting_method:2}
 \end{enumerate}
@@
 \begin{enumerate}[S1']
-	\item Sort the long in each block, copy the heading characters in need to external memory by S\ref{dsais_sorting_method:1}.
-Meanwhile, sort the short in each block, copy the short to external memory in their sorted order.~\label{dsaism_sorting_method:1}
+	\item Reuse S\ref{dsais_sorting_method:1} to sort the short and the long in each block.
+Copy the heading characters for sorting the long to external memory in their access order and copy the short to external memory in their sorted order.~\label{dsaism_sorting_method:1}
 
-	\item Sort the long in $x$ by S\ref{dsais_sorting_method:2}, copy the leftmost $D$ characters of the long to external memory in their sorted order.~\label{dsaism_sorting_method:2}
+	\item Reuse S\ref{dsais_sorting_method:2} to sort the long in $x$. Copy the leftmost $D$ characters of the long to external memory in their sorted order.~\label{dsaism_sorting_method:2}
 
@@
 
-The sorted short S*-type substrings in each block are organized as a sequence in external memory after S\ref{dsaism_sorting_method:1}', while the sorted long S*-type substrings in $x$ are also organized as a sequence in external memory after S\ref{dsaism_sorting_method:2}'.
+The short S*-type substrings in each block are already sorted and organized as a sequence in external memory after S\ref{dsaism_sorting_method:1}', while the long S*-type substrings in $x$ are already sorted and organized as a sequence in external memory after S\ref{dsaism_sorting_method:2}'.
 Assume $x$ is split into $k$ blocks, the multi-way sorter in S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap to determine the lexical order of substrings belonging to different sequences.
 The heap contains at most $k + 1$ substrings at any point of time.
-It compares any two substring in $\mathcal{O}(D)$ time by conducting a literal string comparison\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}.
+It performs a literal string comparison to compare any two substrings in $\mathcal{O}(D)$\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}.
 
@@
 We next describe a method for naming the S*-type substrings when they are being sorted according to S\ref{dsaism_sorting_method:1}'-S\ref{dsaism_sorting_method:3}'.
-The key point is to check equality of two suffixes successively popped from the heap.
+The key point is to check equality of two substrings successively popped from the heap.
 For this, we literally compare the two substrings in $\mathcal{O}(D)$ time if either of them is short; otherwise, we compare their names in $\mathcal{O}(1)$ time instead.
-Recall that the rank of each long S*-type substring is determined when inducing its predecessor, this value is reused to represent its name herein\footnote{A similar technique has been used to merge the substring sorting and naming process in SAIS-PQ.}.
-The experimental results in Section~\ref{sec:experiments} demonstrate that, by using the new substring sorting and naming methods, the adapted DSA-IS, called DSA-IS+, only takes half as much disk space as eSAIS.
+Recall that the rank of each long S*-type substring is determined when inducing its predecessor, this value is reused to represent its name herein\footnote{A similar technique has been used to merge the substring sorting and naming processes in SAIS-PQ.}.
+Our experiments in Section~\ref{sec:experiments} show that, by using the new substring sorting and naming methods, the adapted DSA-IS, called DSA-IS+, only takes half as much disk space as eSAIS.
 
@@
 Following the same idea, we get all the S-type suffixes sorted in line 28.
-Following the discussion, we show in Lemma~\ref{lemma:2} a set of sufficient and necessary conditions for checking SA.
+Following the discussion, we show in Lemma~\ref{lemma:2} a set of sufficient conditions for checking SA.
 
@@
 \begin{lemma} \label{lemma:2}
-    For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for $x[0, n)$ if and only if the following conditions are satisfied at the top recursion level: \\
+    For any IS suffix sorter, its output $sa[0, n)$ is the SA for $x[0, n)$ if the following conditions are satisfied at the top recursion level: \\
     (1) $sa^*$ is correctly computed. \\
@@
 
-Further, suppose $sa[i] = sa[j]$ and $i \ne j$, then $sa[p] = sa[q]$ and $p \ne q$.
+Suppose $sa[i] = sa[j]$ and $i \ne j$, because S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented, $sa[p] = sa[q]$ and $p \ne q$.
 Repeating this reasoning process by replacing $(i, j)$ with $(p, q)$ until ${\sf suf}(sa[i])$ is S*-type, then $sa^*$ must contain duplicate elements.
-However, each element in $sa^*$ is unique as $sa^*$ is corrected computed.
-This leads to a contradiction.
+However, each element in $sa^*$ is unique as $sa^*$ is correctly computed, leading to a contradiction.
+This satisfies the first condition of Lemma~\ref{lemma:1}.
 
@@
 
-It was reported in~\cite{Fischer11, Bingmann12} that the IS method can be used to construct LCP array as well.
+It was reported in~\cite{Fischer11} and~\cite{Bingmann12} that the IS method can be used to construct LCP array as well.
 For any two neighboring suffixes in $sa$, their LCP-value is zero if they start with different characters; otherwise, the LCP-value of them is one greater than that of their successors.
@@
 In this algorithm, ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$.
-These values are used to check the second condition of Lemma~\ref{lemma:3} later.
-We also compute ${\sf lcp_{I1}}(c)/{\sf lcp_{I2}}(c)$ and ${\sf lcp_{S1}}(c)/{\sf lcp_{S2}}(c)$ and check their equality in the meantime.
+These values are compared for checking the second condition of Lemma~\ref{lemma:3}.
+The algorithm also computes and compares ${\sf lcp_{I1}}(c)/{\sf lcp_{I2}}(c)$ and ${\sf lcp_{S1}}(c)/{\sf lcp_{S2}}(c)$ in the meantime.
 
@@
 
-        Check if ${\sf sa_{I2}}(c) = {\sf sa_{S2}}(c)$, ${\sf lcp_{I2}}(c) = {\sf lcp_{S2}}(c)$. \\
+        Check if ${\sf sa_{I2}}(c) = {\sf sa_{S2}}(c)$ and ${\sf lcp_{I2}}(c) = {\sf lcp_{S2}}(c)$. \\
 	}
@@
 
-Assume the induction phase at the top recursion level is correctly implemented, Algorithm~\ref{alg:2} employs the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ and $lcp^*$.
-To be specific, for two neighboring suffixes ${\sf suf}(i)$ and ${\sf suf}(j)$ in $sa^*$, their LCP-value $\ell$ in $lcp^*$ is right if ${\sf sub}(i, i + \ell - 1) = {\sf sub}(j, j + \ell - 1)$ and $x[i + \ell] \ne x[j + \ell]$, where the equality of ${\sf sub}(i, i + \ell - 1)$ and ${\sf sub}(j, j + \ell - 1)$ is checked by comparing the fingerprints of these substrings.
-Given that $\ell$ is correct, the lexical order of ${\sf suf}(i)$ and ${\sf suf}(j)$ is checked by comparing $x[i + \ell]$ with $x[j + \ell]$.
-This method computes the fingerprints of all the target substrings within a sorting complexity proportional to $n_1$.
-Because at most one out of every two suffixes is S*-type, the checking overhead is much less than building consumption.
-By contrast, Algorithm~\ref{alg:3} takes overall $\mathcal{O}(n)$ to compute all the required fingerprints.
+A key operation of Algorithms~\ref{alg:2} and~\ref{alg:3} is to check equality of two integer sequences by comparing their fingerprints.
+This is done in linear time and constant RAM space by using the Karp-Rabin fingerprinting technique.
+Algorithm~\ref{alg:2} also employs the method presented in~\cite{wu2017} to verify $sa^*$ and $lcp^*$.
+For any two neighboring suffixes in $sa^*$, this method checks their lexical order indicated by $sa^*$ and the LCP-value $\ell$ indicated by $lcp^*$ as follows: (1) check if the fingerprints of their leftmost $\ell$ characters are equal; (2) check if their $(\ell + 1)$-th characters are different; (3) check if their lexical order corresponds to that of their $(\ell + 1)$-th characters.
+It should be noticed that this method checks all the neighboring suffix pairs in batch within a sorting complexity.
+Because at most one out of every two suffixes is S*-type and the checking process is only executed at the top recursion level, the consumption for checking $sa^*$ and $lcp^*$ is much less than that for building the two arrays by the IS method.
 
@@
 The probabilistic algorithm designed by the second method is rather lightweight, it takes negligible time and space compared with the existing IS suffix sorting and checking algorithms.
-We also made our first attempt to improve the space performance of DSA-IS using new substring and naming methods.
-Our experimental results show that the program for the adapted algorithm DSA-IS+ runs as fast as that for eSAIS and consumes only half as much disk space as the latter on various real-world datasets.
+In this paper, we also made the first attempt to improve the performance of DSA-IS using new substring and naming methods.
+Our program for the adapted algorithm DSA-IS+ runs as fast as that for eSAIS and consumes only half as much disk space as the latter on various real-world datasets.
 We are now designing and implementing a novel IS suffix sorter that takes no more than 1n work space on external memory model.

### 2017-08-26 new-draft/draft.tex
2017/08/27

@@
 
-A suffix array can be built by the induced sorting~(IS) method on both internal and external memory models.
-We propose two methods for checking a suffix array when it is being built by any IS suffix sorter, i.e.
-build and check a suffix array by the IS method simultaneously.
-In particular, the first method can be extended to check both suffix and LCP arrays.
-By combining the Karp-Rabin fingerprinting technique with our methods, we design two algorithms that perform checking correctly with a negligible error probability.
-The algorithm designed by the first method has an integer sorting complexity, while the algorithm designed by the second method runs in linear time.
-We integrate the algorithm designed by the second method into the existing disk-based suffix sorting algorithm DSA-IS and implement it for performance analysis.
-From our experiments, the checking overhead is considerably smaller than the building consumption.
-
-%A suffix array can be built by the induced sorting~(IS) method on both internal and external memory models.
-We propose two methods to enable the IS to check a suffix array when it is built, i.e.
-build and check a suffix array simultaneously.
-In particular, the first method can be exteneded to check both suffix and LCP arrays.
-By combining the Karp-Rabin fingerprinting techinque with our methods, we design two algorithms that can check correctly with a negligible error probability.
-The algorithm designed by the first method has an integer sorting complexity, while the algorithm designed by the second method runs in linear time.
-We integrate the algorithm designed by the second method into the existing disk-based suffix sorting algorithm DSA-IS and implement it for performance evaluation.
-From our experiments, the checking overhead is considerably smaller than the building consumption.
+A suffix array~(SA) can be built on both internal and external memory models following the induced sorting~(IS) principle.
+We propose two methods for checking an SA when it is being built by any IS suffix sorter.
+The first method can be used to probabilistically check both suffix and longest common prefix~(LCP) arrays, while the second method is for checking SA deterministically.
+To achieve a good performance, we combine the Karp-Rarbin fingerprinting techinque into our methods to design two algorithms that perform checking correctly with a negligible error probability.
+The algorithm designed by the first method has an integer sorting complexity, while the algorithm designed by the second method runs within linear time and constant RAM space.
+We integrate the algorithm designed by the second method into the existing disk-based suffix sorting algorithm DSA-IS for evaluation, where the experimental results indicate that the checking overhead is considerably smaller than the building consumption.
 
@@
 
-{\color{red}
-
-The LCP-value $\ell$ of ${\sf suf}(i)$ and ${\sf suf}(j)$ is right if ${\sf sub}(i, i + \ell - 1) = {\sf sub}(j, j + \ell - 1)$ and $x[i + \ell] \ne x[j + \ell]$.
-In the accepted paper, we check equality of ${\sf sub}(i, i + \ell - 1)$ and ${\sf sub}(j, j + \ell - 1)$ by comparing their fingerprints.
-Given that $\ell$ is correct, the lexical order of ${\sf suf}(i)$ and ${\sf suf}(j)$ is determined by comparing $x[i + \ell]$ and $x[j + \ell]$.
-
-In this paper, both the lexical order and the LCP-value of two neighboring suffixes in the given suffix array is checked at the top recursion level following the IS principle.
-
-}
-
 \section{Introduction}\label{sec:introduction}
 
-A suffix array~(SA) can be built within linear time and space by the internal-memory algorithm SA-IS~\cite{Nong11} using the induced sorting method.
-According to the IS principle, the order of two suffixes is determined by comparing their heading characters and successors in sequence, where the order of their successors is determined recursively.
-Recently, the IS method has been also applied to designing three disk-based suffix sorting algorithms eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}, where eSAIS can build both suffix and LCP arrays at the same time.
-Compared to the others designed by different methods (e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}), these algorithms runs fast with linear I/O complexity.
-However, these algorithms currently suffer from a bottleneck due to the large disk space for obtaining the heading characters of unsorted suffixes and the ranks of their sorted successors in a disk-friendly way.
-It is reported that the average disk use for constructing an SA encoded by 40-bit integers for pSAscan is only $7.5n$ bytes, while that for eSAIS, DSA-IS and SAIS-PQ are $24n$, $18n$ and $15n$ bytes, respectively, where $n$ is the size of input string.
-The poor space performance observed from these IS algorithms is mainly because the current programs for these algorithms fail to immediately free the disk space for temporary data even when the data is no longer needed.
-A dramatic improvement can be achieved by storing temporary data in multiple files and deleting a file when the data in the file is not needed any more.
-This trick has been used by fSAIS~\cite{Karkkainen2017} to engineer the external-memory IS method for better performance, where the peak disk use is $8n$ bytes for constructing an SA of 40-bit integers in the given experiments.
+The SA-IS algorithm takes linear time and RAM space to build an SA on internal memory model~\cite{Nong11}.
+The IS method determines the lexical order of two suffixes by comparing their heading characters and successors in sequence, where the lexical order of their successors is determined recursively following the same way.
+Recently, the IS method has been applied to designing three disk-based suffix sorting algorithms eSAIS~\cite{Bingmann12}, DSA-IS~\cite{Nong15} and SAIS-PQ~\cite{Liu15}.
+In particular, eSAIS can build both suffix and LCP arrays at the same time.
+These IS algorithms have linear I/O complexity better than the others designed by different methods (e.g., DC3~\cite{Dementiev2008a}, bwt-disk~\cite{Ferragina2012}, SAscan~\cite{Karkkainen2014} and pSAscan~\cite{Karkkainen2015}), but they suffer from a bottleneck due to the large disk space for obtaining the heading characters of unsorted suffixes and the ranks of their sorted successors in a disk-friendly way.
+As reported, the average disk use for pSAscan to build a size-$n$ SA encoded by 40-bit integers is around $7.5n$ bytes, while that for eSAIS, DSA-IS and SAIS-PQ are $24n$, $18n$ and $15n$ bytes, respectively.
+The poor space performance for these IS algorithms is mainly because their current programs fail to free the disk space for temporary data even when the data is no longer needed. 
+A dramatic improvement can be achieved by storing temporary data into multiple files and deleting a file immediately when the data in it is not needed anymore.
+This trick has been employed by fSAIS~\cite{Karkkainen2017} to engineer the external-memory IS method, where the peak disk use is $8n$ bytes for building an SA of 40-bit integers in the given experiments.
 
@@
 When running on external-memory model, the cost taken by this checker is rather high as it performs two passes of integer sorts for arranging $\mathcal{O}(n)$ fixed-size tuples. 
-In this paper, we propose two methods to enable the IS method to build and check an SA simultaneously.
-In particular, the first method can be extended to check both suffix and LCP arrays.
-We employ the Karp-Rabin fingerprinting technique~\cite{Karp1987} to design two probabilistic algorithms, in terms of that their checking results are wrong with a negligible probability.
-For analysis, we first use new substring sorting and naming methods to improve the design of DSA-IS and then combine the second checking method into the adapted DSA-IS for evaluating the checking overhead.
-Our experimental results indicate that the time, space and I/O volume for checking SA is considerably smaller than that for building SA.
+We propose in this paper two methods that enable the IS method to build and check an SA at the same time. 
+Our first method is probabilistic and can be extended to check both suffix and LCP arrays, while the second method is deterministic and designed for checking SA only. 
+By combining the Karp-Rabin fingerprinting technique~\cite{Karp1987}, we design two probabilistic algorithms to perform checking correctly with a negligible error probability.
+The key operation of both algorithms is to check equality of two integer sequences by computing and comparing their fingerprints.
+As will be shown, the algorithm designed by the first method checks an SA within a integer sorting complexity and the algorithm designed by the second method checks an SA in linear time.
+For a further study, we first use new substring sorting and naming methods to improve the design of DSA-IS and then integrate the algorithm designed by the second method into the adapted DSA-IS for evaluation.
+Our experimental results indicate that the time, space and I/O volume for checking SA is negligible in comparison with that for building SA.
 
@@
 The type of a substring or suffix is determined by that of its heading character.
+In the rest of this paper, we only consider substrings ending with an S*-type character.
 
@@
 We produce a reduced string $x_1$ by replacing each substring with its name, where the name represents the rank of this	 substring among all.
-In this paper, we use "rank" and "name" interchangeably to indicate the lexical order of substrings.
+We interchangeably use "rank" and "name"  to indicate the lexical order of a substring.
 
@@
 Each bucket occupies a contiguous interval in $sa$ and contains all the suffixes with a same heading character.
-Without loss of generality, we denote by ${\sf sa\_bkt}(c_0)$ the bucket for suffixes starting with $c_0$, where ${\sf sa\_bkt}(c_0)$ can be naturally divided into two parts ${\sf sa\_bkt_L}(c_0)$ and ${\sf sa\_bkt_S}(c_0)$ containing L-type and S-type suffixes, respectively.
-We also define ${\sf lcp\_bkt}(c_0)$, ${\sf lcp\_bkt_L}(c_0)$ and ${\sf lcp\_bkt_S}(c_0)$ on $lcp$ for $c_0 \in \Sigma$.
+Without loss of generality, we denote by ${\sf sa\_bkt}(c_0)$ the bucket for suffixes starting with $c_0$, where ${\sf sa\_bkt}(c_0)$ can be naturally divided into two parts ${\sf sa\_bkt_L}(c_0)$ and ${\sf sa\_bkt_S}(c_0)$ that contain L-type and S-type suffixes, respectively.
+Similarly, we define ${\sf lcp\_bkt}(c_0)$, ${\sf lcp\_bkt_L}(c_0)$ and ${\sf lcp\_bkt_S}(c_0)$ on $lcp$ for $c_0 \in \Sigma$.
 
@@
 	/* Check Recursion Condition */ \\	
-	\If{exist two equal characters in $x_1$}{
+	\If{exist duplicate characters in $x_1$}{
 
@@
 This involves a great number of random accesses to $x$ and $sa$, which can be done very fast if both $x$ and $sa$ can wholly reside on RAM.
-However, if the size of input and output exceeds the capacity of internal memory, each access will take an individual I/O operation to disk, leading to a performance degradation.
-The DSA-IS algorithm solves this problem by performing the following two steps during a reduction/induction phase:
+However, if the size of input and output exceeds the capacity of internal memory, each access may take an individual I/O operation to disk, leading to a performance degradation.
+The DSA-IS algorithm solves this problem by executing the following two steps during a reduction/induction phase:
 
@@
 
-	\item Sort the substrings/suffixes in $x$ by their heading characters and the ranks of their successors in an external-memory heap. \label{dsais_sorting_method:2}
+	\item Sort the substrings/suffixes in $x$ according to the lexical order of their heading characters and the ranks of their successors by using an external-memory heap. \label{dsais_sorting_method:2}
 \end{enumerate}
 
-Because all the substrings/suffixes in the heap are sorted already, S\ref{dsais_sorting_method:2} determines the rank of a substring/suffix when popping it from the heap.
-When inducing the predecessor of current top substring/suffix, the heading character in need is retrieved from external memory by a sequential I/O operation.
+S\ref{dsais_sorting_method:2} induces the lexical order of substrings/suffixes by using an external-memory heap.
+It continually pops the top substring/suffix from the heap and induces its predecessors into the heap according to the rank of the top substring/suffix and the heading character of the predecessors.
+For all the induced substrings/suffixes, their heading characters are already arranged in access order on disk by S\ref{dsais_sorting_method:1} and thus can be retrieved from external memory by sequential I/O operations before use.
 As shown in Section~\ref{sec:experiments}, our program for DSA-IS requires less disk space than that for eSAIS, but the former runs slower than the latter due to the large I/O volume for sorting and naming S*-type substrings during the reduction phase.
@@
 
-All the S*-type substrings are classified into long and short categories with respect to whether or not containing more than $D$ characters.
+All the substrings are classified into long and short categories with respect to whether or not containing more than $D$ characters.
 The new substring sorting method mainly consists of the three steps below:
@@
 \begin{enumerate}[S1']
-	\item Reuse S\ref{dsais_sorting_method:1} to sort the short and the long in each block.
-Copy the heading characters for sorting the long to external memory in their access order and copy the short to external memory in their sorted order.~\label{dsaism_sorting_method:1}
+	\item Reuse S\ref{dsais_sorting_method:1} to sort all the substrings in each block.
+Copy the heading characters of the induced long substrings to external memory in access order and copy the induced short S*-type substrings to external memory in sorted order.~\label{dsaism_sorting_method:1}
 
-	\item Reuse S\ref{dsais_sorting_method:2} to sort the long in $x$. Copy the leftmost $D$ characters of the long to external memory in their sorted order.~\label{dsaism_sorting_method:2}
+	\item Reuse S\ref{dsais_sorting_method:2} to sort the long substrings in $x$.
+Copy the leftmost $D$ characters of the long S*-type substrings to external memory in sorted order.~\label{dsaism_sorting_method:2}
 
-	\item Merge the short and the long by a multi-way sorter.~\label{dsaism_sorting_method:3}
+	\item Merge all the short and long by a multi-way sorter.~\label{dsaism_sorting_method:3}
 \end{enumerate}
@@
 The heap contains at most $k + 1$ substrings at any point of time.
-It performs a literal string comparison to compare any two substrings in $\mathcal{O}(D)$\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}.
+It performs a literal string comparison to compare any two S*-type substrings in $\mathcal{O}(D)$\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}.
 
@@
 This is commonly satisfied in real-world datasets.
-We next describe a method for naming the S*-type substrings when they are being sorted according to S\ref{dsaism_sorting_method:1}'-S\ref{dsaism_sorting_method:3}'.
-The key point is to check equality of two substrings successively popped from the heap.
-For this, we literally compare the two substrings in $\mathcal{O}(D)$ time if either of them is short; otherwise, we compare their names in $\mathcal{O}(1)$ time instead.
-Recall that the rank of each long S*-type substring is determined when inducing its predecessor, this value is reused to represent its name herein\footnote{A similar technique has been used to merge the substring sorting and naming processes in SAIS-PQ.}.
+We next describe a method for naming the S*-type substrings when they are being sorted by S\ref{dsaism_sorting_method:1}'-S\ref{dsaism_sorting_method:3}'.
+The key operation is to check equality of two substrings successively popped from the heap in S\ref{dsaism_sorting_method:3}'.
+If either of the substrings is short, then we literally compare them in $\mathcal{O}(D)$ time.
+Otherwise, we check whether their ranks are equal by comparing their heading characters and the ranks of their successors in S\ref{dsaism_sorting_method:2}'.
+For the same purpose, SAIS-PQ also applies a similar technique to merge the substring sorting and naming processes.
 Our experiments in Section~\ref{sec:experiments} show that, by using the new substring sorting and naming methods, the adapted DSA-IS, called DSA-IS+, only takes half as much disk space as eSAIS.
@@
 If condition (2) is true, then the lexical order of any two neighboring suffixes in $sa$ is determined by comparing their heading characters and the ranks of their successors.
-As a result, all the suffixes in $x$ are sorted according to the lexical order of their heading characters and successors.
+As a result, all the suffixes in $x$ are sorted following the induced sorting principle.
 
@@
 Following the idea, we assume S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented and propose two methods for checking computation errors caused by implementation bugs and other malfunctions.
-The algorithms designed by these methods can be seamlessly integrated into any IS suffix sorting algorithm for building and checking an SA at the same time.
+These methods adopt different ways to ensure the correctness of $sa^*$.
+Particularly, our first method can be used to check both suffix and LCP arrays.
 
@@
 
-Method A is based on Lemma~\ref{lemma:3}, which is an extension of Lemma~\ref{lemma:2}.
+The induced sorting method can be applied to building LCP array as well~\cite{Fischer11}.
+That is, the LCP-value of two substrings can be induced from that of their successors as following: the LCP-value is equal to zero if the two substrings have different heading characters; otherwise, the LCP-value is one greater than that of their successors.
+Assume the induction phase for inducing $sa$ and $lcp$ from $sa^*$ and $lcp^*$ is correctly implemented at the top recursion level, we show in Lemma~\ref{lemma:3} a set of conditions for checking both $sa$ and $lcp$ when they are built by an IS builder like~\cite{Fischer11, Bingmann12}.
 
 \begin{lemma} \label{lemma:3}
-	Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ of an IS suffix sorter is the SA for the input string $x[0,  n)$ if the following conditions are satisfied: \\
-	(1) $sa^*$ is correct. \\
-	(2) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' for $i \in [0, n)$. \\
-
+	Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ and $lcp[0, n)$ of an IS builder are the suffix and LCP arrays for the input string $x[0, n)$ with a high probability if the following conditions are satisfied: \\
+	(1) $sa^*$ and $lcp^*$ are correct according to the checking result of the sparse suffix and LCP arrays probabilistic checker~\cite{wu2017}. \\
+	(2) $sa[i]$ and $lcp[i]$ are equal to the values induced into the corresponding positions of $sa$ and $lcp$ during the induction phase, for $i \in [0, n)$. \\
 \end{lemma}
 
-The first condition of Lemma~\ref{lemma:3} can be checked by a sparse SA checker like~\cite{wu2017}.
-The second condition intends for detecting malfunctions other than implementation bugs (e.g., I/O errors).
-This is checked by determining whether the induced and scanned values for each suffix are equal\footnote{Notice that each suffix induced into $sa$ will be latter scanned for inducing the order of its predecessor during S\ref{induction_phase:3}'' and S\ref{induction_phase:4}''.
-}.
-The problem to be solved here is that when a suffix is induced into a bucket, its corresponding value in $sa$ may not be scanned at once.
+The checker in~\cite{wu2017} verifies $sa^*$ and $lcp^*$ within an integer sorting complexity.
+For each pair of neighboring suffixes in the given $sa$, it first employs the Karp-Rabin fingerprinting technique to compute the fingerprints of their leftmost $\ell$ characters, where $\ell$ is their LCP-value indicated by the given $lcp$.
+If the two fingerprints are equal, then it continues to check if the lexical order of  their $(\ell + 1)$-th characters corresponds to that of themselves indicated by the given $sa$.
+The second condition in Lemma~\ref{lemma:3} is intended for detecting malfunctions other than implementation bugs.
+Because each suffix induced into $sa/lcp$ will be latter scanned for inducing the order of its predecessor, this condition can be checked by determining whether the induced and scanned values for each suffix are equal.
+The problem here is that when a suffix is induced into a bucket, its corresponding value in $sa$ may not be scanned at once.
 Our solution is to ensure the sequence of values placed at a bucket is identical to that scanned latter.
-For the purpose, we use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end of the induction phase.
+For the purpose, we reuse the Karp-Rabin fingerprinting technique to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end of induction phase.
 If the two fingerprints for each bucket are equal, then the second condition of Lemma~\ref{lemma:3} will be seen with a high probability.
-As a result, $sa$ can be built and probabilistically checked at the same time.
+As a result, $sa$ and $lcp$ can be built and probabilistically checked at the same time.
 
-It was reported in~\cite{Fischer11} and~\cite{Bingmann12} that the IS method can be used to construct LCP array as well.
-For any two neighboring suffixes in $sa$, their LCP-value is zero if they start with different characters; otherwise, the LCP-value of them is one greater than that of their successors.
-This implies the possibility for extending Method A to check both suffix and LCP arrays.
 We design Algorithm~\ref{alg:2} to check $sa$ and $lcp$ when they are induced from $sa^*$ and $lcp^*$ at the top recursion level of an IS builder.
 In this algorithm, ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$.
-These values are compared for checking the second condition of Lemma~\ref{lemma:3}.
-The algorithm also computes and compares ${\sf lcp_{I1}}(c)/{\sf lcp_{I2}}(c)$ and ${\sf lcp_{S1}}(c)/{\sf lcp_{S2}}(c)$ in the meantime.
+The algorithm computes and compares the fingerprints of these sequences and those of ${\sf lcp_{I1}}(c)/{\sf lcp_{I2}}(c)$ and ${\sf lcp_{S1}}(c)/{\sf lcp_{S2}}(c)$ to check the second condition of Lemma~\ref{lemma:3}.
 
@@
 
-According to Lemma~\ref{lemma:4}, Method B uses a different way to check the correctness of $sa^*$.
-Before our presentation, we first introduce a notation called $\overline{sa^*}$.
+Method B is deterministic and specific for checking SA only.
+It uses a different way to ensure the correctness of $sa^*$.
+Before our presentation, we first introduces a notation called $\overline{sa^*}$.
 The same as $sa^*$, $\overline{sa^*}$ also represents the suffix array for all the S*-type suffixes.
-Both $sa^*$ and $\overline{sa^*}$ are computed during the induction phase at the top recursion level, where the former is calculated from $sa_1$ in S\ref{induction_phase:1}'' and the latter is calculated when inducing the S*-type suffixes into $sa$ in S\ref{induction_phase:4}''.
+Both $sa^*$ and $\overline{sa^*}$ are computed during the induction phase at the top recursion level, where the former is calculated from $sa_1$ in S\ref{induction_phase:1}'' and the latter is computed when inducing the S*-type suffixes into $sa$ in S\ref{induction_phase:4}''.
 
@@
 	(2) $sa^*[i] = \overline{sa^*}[i]$ for $i \in [0, n_1)$. \\
-	(3) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' for $i \in [0, n)$. \\
-
 \end{lemma}
@@
 The first condition of Lemma~\ref{lemma:4} is naturally satisfied when computing $sa^*$ from $sa_1$ in S\ref{induction_phase:1}''.
-The second condition is probabilistically checked by computing and comparing the fingerprints of $sa^*$ and $\overline{sa^*}$.
-The fingerprint of $sa^*$ is computed when placing the elements of $sa^*$ into $sa$ in S\ref{induction_phase:2}'', while the fingerprint of $\overline{sa^*}$ is computed when inducing the S*-type suffixes into $sa$ in S\ref{induction_phase:4}''.
-We design Algorithm~\ref{alg:3} to check if the final SA observes Lemma~\ref{lemma:4}.
+One way for deterministically checking the second condition is to copy both $sa^*$ and $\overline{sa^*}$ to external memory and compare their elements one by one after the induction phase.
+Algorithm~\ref{alg:3} applies a more space efficient way to checking equality of the two integer sequences with a negligible error probability.
+It reuses the Karp-Rabin fingerprinting technique to compute and compare the fingerprints of $sa^*$ and $\overline{sa^*}$, where the fingerprint of $sa^*$ is computed when placing the elements of $sa^*$ into $sa$ in S\ref{induction_phase:2}'' and the fingerprint of $\overline{sa^*}$ is computed when inducing the S*-type suffixes into $sa$ in S\ref{induction_phase:4}''.
 
@@
 
-        Compute the fingerprints of $sa^*$, ${\sf sa_{I1}}(c)$ and ${\sf sa_{S1}}(c)$ when inducing the order of L-type suffixes. \\
+        Compute the fingerprints of $sa^*$ when inducing the order of L-type suffixes. \\
 
-        Check if ${\sf sa_{I1}}(c) = {\sf sa_{S1}}(c)$ by comparing their fingerprints. \\
-
-        Compute the fingerprints of $\overline{sa^*}$, ${\sf sa_{I2}}(c)$ and ${\sf sa_{S2}}(c)$ when inducing the order of S-type suffixes. \\
-
-        Check if ${\sf sa_{I2}}(c) = {\sf sa_{S2}}(c)$ by comparing their fingerprints. \\
+        Compute the fingerprints of $\overline{sa^*}$ when inducing the order of S-type suffixes. \\
 
@@
 
-The proposed two methods check equality of two arrays (or sub-arrays) by comparing their fingerprints, which can be calculated using a hash function.
-Notice that two equal arrays must be mapped to an identical hash value, but the inverse is not always true.
-To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting function to compute these hash values, called fingerprints,	 following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$.
+Both Algorithms~\ref{alg:2} and ~\ref{alg:3} check equality of sequences by comparing their fingerprints.
+We can use any rolling hash function to compute the fingerprints in need.
+Notice that two equal sequences must be mapped to an identical hash value, but the inverse is not always true.
+To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting techniques to compute these values following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$.
 By setting $L$ to a large value, the error probability can be reduced to a negligible level.
@@
 Because $\sf fp(A[0, -1]) = 0$, $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$.
-By iteratively computing the fingerprints of the prefixes of $A$, we finally obtain the fingerprint of $A$ in line 7.
+By iteratively computing the fingerprints in this way, we finally obtain the fingerprint of $A$ in line 7.
 
@@
 
-A key operation of Algorithms~\ref{alg:2} and~\ref{alg:3} is to check equality of two integer sequences by comparing their fingerprints.
-This is done in linear time and constant RAM space by using the Karp-Rabin fingerprinting technique.
-Algorithm~\ref{alg:2} also employs the method presented in~\cite{wu2017} to verify $sa^*$ and $lcp^*$.
-For any two neighboring suffixes in $sa^*$, this method checks their lexical order indicated by $sa^*$ and the LCP-value $\ell$ indicated by $lcp^*$ as follows: (1) check if the fingerprints of their leftmost $\ell$ characters are equal; (2) check if their $(\ell + 1)$-th characters are different; (3) check if their lexical order corresponds to that of their $(\ell + 1)$-th characters.
-It should be noticed that this method checks all the neighboring suffix pairs in batch within a sorting complexity.
-Because at most one out of every two suffixes is S*-type and the checking process is only executed at the top recursion level, the consumption for checking $sa^*$ and $lcp^*$ is much less than that for building the two arrays by the IS method.
+Algorithms~\ref{alg:2} and ~\ref{alg:3} take linear time and constant RAM space for computing the target sequences by the Karp-Rabin fingerprinting technique.
+Algorithm~\ref{alg:2} also employs the checker presented in~\cite{wu2017} to verify $sa^*$ and $lcp^*$ within an integer sorting complexity.
+Because at most one out of every two suffixes is S*-type and the checking process is only executed at the top recursion level, the overhead for checking $sa^*$ and $lcp^*$ is considerably smaller than that for constructing the two arrays by an IS builder.
 
@@
 
-By assuming the induction phase at the top recursion level is correctly implemented, we proposed two methods that enable any IS suffix sorter to build and check an SA simultaneously.
-The probabilistic algorithm designed by the second method is rather lightweight, it takes negligible time and space compared with the existing IS suffix sorting and checking algorithms.
-In this paper, we also made the first attempt to improve the performance of DSA-IS using new substring and naming methods.
+By assuming the induction phase at the top recursion level is correctly implemented, we propose two methods that enable an IS builder to build and check an SA simultaneously.
+The probabilistic algorithm designed by Method B is rather lightweight, it takes negligible time and space to run compared to the existing IS suffix sorting and checking algorithms.
+We also made our first attempt to improve the performance of DSA-IS using new substring sorting and naming methods.
 Our program for the adapted algorithm DSA-IS+ runs as fast as that for eSAIS and consumes only half as much disk space as the latter on various real-world datasets.

### 2017-08-27 new-draft/draft.tex
2017/08/27

@@
 
-\title{Building and Checking a Suffix Array by Induced Sorting}
+\title{Building and Checking Suffix Array by Induced Sorting Method}
 
@@
 
-A suffix array~(SA) can be built on both internal and external memory models following the induced sorting~(IS) principle.
-We propose two methods for checking an SA when it is being built by any IS suffix sorter.
-The first method can be used to probabilistically check both suffix and longest common prefix~(LCP) arrays, while the second method is for checking SA deterministically.
+A suffix array~(SA) can be built on both internal and external memory models using the induced sorting~(IS) method.
+We propose two methods for checking an SA when it is being built by any IS builder.
+The first method is probabilistic and can be used to check both suffix and longest common prefix~(LCP) arrays, while the second method is deterministic and designed for checking SA only.
 To achieve a good performance, we combine the Karp-Rarbin fingerprinting techinque into our methods to design two algorithms that perform checking correctly with a negligible error probability.
@@
 When running on external-memory model, the cost taken by this checker is rather high as it performs two passes of integer sorts for arranging $\mathcal{O}(n)$ fixed-size tuples. 
-We propose in this paper two methods that enable the IS method to build and check an SA at the same time. 
-Our first method is probabilistic and can be extended to check both suffix and LCP arrays, while the second method is deterministic and designed for checking SA only. 
-By combining the Karp-Rabin fingerprinting technique~\cite{Karp1987}, we design two probabilistic algorithms to perform checking correctly with a negligible error probability.
-The key operation of both algorithms is to check equality of two integer sequences by computing and comparing their fingerprints.
-As will be shown, the algorithm designed by the first method checks an SA within a integer sorting complexity and the algorithm designed by the second method checks an SA in linear time.
-For a further study, we first use new substring sorting and naming methods to improve the design of DSA-IS and then integrate the algorithm designed by the second method into the adapted DSA-IS for evaluation.
+We propose in this paper two methods that enable the IS method to build and check an SA simultaneously. 
+Our first method is probabilistic and can be used to check both suffix and LCP arrays, while the second method is deterministic and designed for checking SA only. 
+We design two probabilistic algorithms by the proposed methods to perform checking correctly with a negligible error probability.
+A key operation of the algorithms is to check equality of two integer sequences by comparing their fingerprints, where the fingerprints are computed using the Karp-Rabin fingerprinting technique~\cite{Karp1987}.
+As will be shown, the algorithm designed by the first method checks an SA within an integer sorting complexity, while the algorithm designed by the second method checks an SA using linear time and constant RAM space.
+For a further study, we employ new substring sorting and naming methods to improve the design of DSA-IS and integrate the algorithm designed by the second method into the adapted DSA-IS for evaluation.
 Our experimental results indicate that the time, space and I/O volume for checking SA is negligible in comparison with that for building SA.
@@
 Section~\ref{sec:preliminaries} introduces some notations and symbols for presentation convenience.
-Section~\ref{sec:builder} gives an overview of the existing IS suffix sorting algorithms and show the details of our new substring sorting and naming methods specific for DSA-IS.
-Section~\ref{sec:checkers} presents the proposed checking methods and the probabilistic algorithms designed by them.
+Section~\ref{sec:builder} gives an overview of the existing IS suffix sorting algorithms and describes our new substring sorting and naming methods for DSA-IS.
+Section~\ref{sec:checkers} presents the proposed checking methods and the probabilistic algorithms.
 Sections~\ref{sec:experiments} and~\ref{sec:conclusion} show the experimental results and the concluding remarks, respectively.
@@
 Given two characters $x[i]$ and $x[i + 1]$, we say $x[i]$ is the predecessor of $x[i + 1]$ and $x[i + 1]$ is the successor of $x[i]$.
-We define the predecessor-successor relationship between ${\sf suf}(i)/{\sf sub}(i, j)$ and ${\sf suf}(i + 1)/{\sf sub}(i + 1, j)$.
+We define the predecessor-successor relationship between ${\sf suf}(i)/{\sf sub}(i, j)$ and ${\sf suf}(i + 1)/{\sf sub}(i + 1, j)$ as well.
 
@@
 Each bucket occupies a contiguous interval in $sa$ and contains all the suffixes with a same heading character.
-Without loss of generality, we denote by ${\sf sa\_bkt}(c_0)$ the bucket for suffixes starting with $c_0$, where ${\sf sa\_bkt}(c_0)$ can be naturally divided into two parts ${\sf sa\_bkt_L}(c_0)$ and ${\sf sa\_bkt_S}(c_0)$ that contain L-type and S-type suffixes, respectively.
+Without loss of generality, we denote by ${\sf sa\_bkt}(c_0)$ the bucket for suffixes starting with $c_0$, where ${\sf sa\_bkt}(c_0)$ can be naturally divided into left and right parts ${\sf sa\_bkt_L}(c_0)$ and ${\sf sa\_bkt_S}(c_0)$ that contain L-type and S-type suffixes, respectively.
 Similarly, we define ${\sf lcp\_bkt}(c_0)$, ${\sf lcp\_bkt_L}(c_0)$ and ${\sf lcp\_bkt_S}(c_0)$ on $lcp$ for $c_0 \in \Sigma$.
@@
 
-	\item Split $x$ into blocks and sort substrings/suffixes in each block by calling SA-IS.
-The heading characters in need are copied to external memory in their access order.
+	\item Split $x$ into blocks and sort substrings/suffixes in each block by the internal-memory IS method.
+Copy the heading characters of the induced substrings/suffixes to external memory in access order.
 \label{dsais_sorting_method:1}
 
-	\item Sort the substrings/suffixes in $x$ according to the lexical order of their heading characters and the ranks of their successors by using an external-memory heap. \label{dsais_sorting_method:2}
+	\item Sort the substrings/suffixes in $x$ according to the lexical order of their heading characters and the ranks of their successors by a disk-based heap. \label{dsais_sorting_method:2}
 \end{enumerate}
 
-S\ref{dsais_sorting_method:2} induces the lexical order of substrings/suffixes by using an external-memory heap.
-It continually pops the top substring/suffix from the heap and induces its predecessors into the heap according to the rank of the top substring/suffix and the heading character of the predecessors.
-For all the induced substrings/suffixes, their heading characters are already arranged in access order on disk by S\ref{dsais_sorting_method:1} and thus can be retrieved from external memory by sequential I/O operations before use.
-As shown in Section~\ref{sec:experiments}, our program for DSA-IS requires less disk space than that for eSAIS, but the former runs slower than the latter due to the large I/O volume for sorting and naming S*-type substrings during the reduction phase.
+S\ref{dsais_sorting_method:2} adapts the internal-memory IS method for running on external memory model by using a disk-based heap.
+It continually pops the top substring/suffix from the heap and induces its predecessor into the heap.
+For all the induced substrings/suffixes, their heading characters are already arranged in access order on disk at the end of S\ref{dsais_sorting_method:1} and thus can be retrieved from external memory by sequential I/O operations before use.
+As shown in Section~\ref{sec:experiments}, our program for DSA-IS requires less disk space than that for eSAIS, but the former runs slower than the latter due to the large I/O volume for sorting and naming S*-type substrings in the reduction phase.
 To improve the performance, we propose new substring sorting and naming methods in the following.
@@
 
-All the substrings are classified into long and short categories with respect to whether or not containing more than $D$ characters.
+All the S*-type substrings are classified into long and short categories with respect to whether or not containing more than $D$ characters.
 The new substring sorting method mainly consists of the three steps below:
@@
 	\item Reuse S\ref{dsais_sorting_method:1} to sort all the substrings in each block.
-Copy the heading characters of the induced long substrings to external memory in access order and copy the induced short S*-type substrings to external memory in sorted order.~\label{dsaism_sorting_method:1}
+Copy the heading characters of the induced substrings contained in the long S*-type substrings to external memory in access order and copy the induced short S*-type substrings to external memory in sorted order.~\label{dsaism_sorting_method:1}
 
-	\item Reuse S\ref{dsais_sorting_method:2} to sort the long substrings in $x$.
+	\item Reuse S\ref{dsais_sorting_method:2} to sort the substrings contained in the long S*-type substrings in $x$.
 Copy the leftmost $D$ characters of the long S*-type substrings to external memory in sorted order.~\label{dsaism_sorting_method:2}
 
-	\item Merge all the short and long by a multi-way sorter.~\label{dsaism_sorting_method:3}
+	\item Merge all the short and long S*-type substrings using a multi-way sorter.~\label{dsaism_sorting_method:3}
 \end{enumerate}
 
-The short S*-type substrings in each block are already sorted and organized as a sequence in external memory after S\ref{dsaism_sorting_method:1}', while the long S*-type substrings in $x$ are already sorted and organized as a sequence in external memory after S\ref{dsaism_sorting_method:2}'.
-Assume $x$ is split into $k$ blocks, the multi-way sorter in S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap to determine the lexical order of substrings belonging to different sequences.
-The heap contains at most $k + 1$ substrings at any point of time.
-It performs a literal string comparison to compare any two S*-type substrings in $\mathcal{O}(D)$\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}.
+S\ref{dsaism_sorting_method:1}' separately sorts the short S*-type substrings in each block and organizes them as a sequence in external memory, while S\ref{dsaism_sorting_method:2}' sorts the long S*-type substrings in $x$ and organizes them as a sequence in external memory.
+Assume $x$ is partitioned into $k$ blocks, S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap for the multi-way sorter to determine the lexical order of substrings from different sequences.
+The heap contains at most $k + 1$ substrings at any point of time, it performs a literal comparison to compare two S*-type substrings in $\mathcal{O}(D)$ time\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the latter is lexicographically greater than the former.}.
 
-The above sorting method has a good performance if $D$ is small and the majority of S*-type substrings in $x$ are short.
-This is commonly satisfied in real-world datasets.
-We next describe a method for naming the S*-type substrings when they are being sorted by S\ref{dsaism_sorting_method:1}'-S\ref{dsaism_sorting_method:3}'.
-The key operation is to check equality of two substrings successively popped from the heap in S\ref{dsaism_sorting_method:3}'.
-If either of the substrings is short, then we literally compare them in $\mathcal{O}(D)$ time.
-Otherwise, we check whether their ranks are equal by comparing their heading characters and the ranks of their successors in S\ref{dsaism_sorting_method:2}'.
-For the same purpose, SAIS-PQ also applies a similar technique to merge the substring sorting and naming processes.
-Our experiments in Section~\ref{sec:experiments} show that, by using the new substring sorting and naming methods, the adapted DSA-IS, called DSA-IS+, only takes half as much disk space as eSAIS.
+The above sorting method can achieve an ideal performance if $D$ is small and the majority of S*-type substrings in $x$ are short.
+This feature is commonly satisfied in real-world datasets.
+Next, we describe a method for naming all the S*-type substrings when they are being sorted by S\ref{dsaism_sorting_method:1}'-S\ref{dsaism_sorting_method:3}'.
+Specifically, for two substrings successively popped from the heap in S\ref{dsaism_sorting_method:3}', this method checks their equality as following: if either of the substrings is short, then we literally compare them in $\mathcal{O}(D)$ time; otherwise, we check whether their ranks are equal by comparing their heading characters and the ranks of their successors in S\ref{dsaism_sorting_method:2}'.
+It should be noticed that SAIS-PQ also uses a similar technique to merge the substring sorting and naming processes.
+Our experiments in Section~\ref{sec:experiments} demonstrate that, by using the new substring sorting and naming methods, the adapted DSA-IS, called DSA-IS+, only takes half as much disk space as eSAIS.
 
@@
 If condition (2) is true, then the lexical order of any two neighboring suffixes in $sa$ is determined by comparing their heading characters and the ranks of their successors.
-As a result, all the suffixes in $x$ are sorted following the induced sorting principle.
+As a result, all the suffixes in $x$ are sorted following the IS principle.
 
@@
 \begin{enumerate}[S1'']
-	\item sort the starting positions of all the S*-type suffixes with their ranks indicated by $sa_1$ to produce $sa^*$ \label{induction_phase:1}.
+	\item Sort the starting positions of all the S*-type suffixes with their ranks indicated by $sa_1$ to produce $sa^*$ \label{induction_phase:1}.
 
@@
 Following the same idea, we get all the S-type suffixes sorted in line 28.
-Following the discussion, we show in Lemma~\ref{lemma:2} a set of sufficient conditions for checking SA.
+Based on the previous discussion, we show in Lemma~\ref{lemma:2} a set of sufficient conditions for checking SA.
 
@@
 
-\begin{IEEEproof} Given two suffixes placed at $sa[i]$ and $sa[j]$ and their successors placed at $sa[p]$ and $sa[q]$ at the top recursion level, we prove the statement as follows.
+\begin{IEEEproof} Given ${\sf suf}(sa[i])$ and ${\sf suf}(sa[j])$ induced from ${\sf suf}(sa[p])$ and ${\sf suf}(sa[q])$ at the top recursion level, we prove the statement as follows.
 
@@
 
-Lemma~\ref{lemma:2} shows that we can check the correctness of $sa^*$ instead of $sa$ at the top recursion level to perform SA verification if the first condition of Lemma~\ref{lemma:2} is always true.
+Lemma~\ref{lemma:2} indicates that we can check $sa^*$ instead to perform SA verification at the top recursion level if the second condition of Lemma~\ref{lemma:2} is always true.
 In fact, the code snippet for the induction phase at the top recursion level only consists of tens of C++ code lines.
-This is considerably smaller than the whole program for a disk-based suffix sorting algorithm.
-Following the idea, we assume S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented and propose two methods for checking computation errors caused by implementation bugs and other malfunctions.
-These methods adopt different ways to ensure the correctness of $sa^*$.
-Particularly, our first method can be used to check both suffix and LCP arrays.
+This is considerably smaller than the whole program for a disk-based IS suffix sorting algorithm.
+Following the idea, we assume S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented and propose in this section two methods for checking SA, where the methods adopt different approaches to ensure the correctness of $sa^*$.
+In particular, our first method can be used to check both suffix and LCP arrays.
 
@@
 
-The induced sorting method can be applied to building LCP array as well~\cite{Fischer11}.
-That is, the LCP-value of two substrings can be induced from that of their successors as following: the LCP-value is equal to zero if the two substrings have different heading characters; otherwise, the LCP-value is one greater than that of their successors.
+The IS method can be applied to constructing LCP array as well~\cite{Fischer11}.
+The LCP-value of two substrings is induced from that of their successors as following: the LCP-value is equal to zero if the two substrings have different heading characters; otherwise, the LCP-value is one greater than that of their successors.
 Assume the induction phase for inducing $sa$ and $lcp$ from $sa^*$ and $lcp^*$ is correctly implemented at the top recursion level, we show in Lemma~\ref{lemma:3} a set of conditions for checking both $sa$ and $lcp$ when they are built by an IS builder like~\cite{Fischer11, Bingmann12}.
@@
 	Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ and $lcp[0, n)$ of an IS builder are the suffix and LCP arrays for the input string $x[0, n)$ with a high probability if the following conditions are satisfied: \\
-	(1) $sa^*$ and $lcp^*$ are correct according to the checking result of the sparse suffix and LCP arrays probabilistic checker~\cite{wu2017}. \\
+	(1) $sa^*$ and $lcp^*$ are checked to be correct using the sparse suffix and LCP arrays probabilistic checker~\cite{wu2017}. \\
 	(2) $sa[i]$ and $lcp[i]$ are equal to the values induced into the corresponding positions of $sa$ and $lcp$ during the induction phase, for $i \in [0, n)$. \\
@@
 The second condition in Lemma~\ref{lemma:3} is intended for detecting malfunctions other than implementation bugs.
-Because each suffix induced into $sa/lcp$ will be latter scanned for inducing the order of its predecessor, this condition can be checked by determining whether the induced and scanned values for each suffix are equal.
-The problem here is that when a suffix is induced into a bucket, its corresponding value in $sa$ may not be scanned at once.
+Because each suffix induced into $sa/lcp$ will be latter scanned for inducing its predecessor, this condition can be checked by determining whether the induced and scanned values for each item in $sa/lcp$ are equal.
+The problem here is that when an item is induced into a bucket of $sa/lcp$, its corresponding value in $sa/lcp$ may not be scanned at once.
 Our solution is to ensure the sequence of values placed at a bucket is identical to that scanned latter.
@@
 
-We design Algorithm~\ref{alg:2} to check $sa$ and $lcp$ when they are induced from $sa^*$ and $lcp^*$ at the top recursion level of an IS builder.
-In this algorithm, ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$.
-The algorithm computes and compares the fingerprints of these sequences and those of ${\sf lcp_{I1}}(c)/{\sf lcp_{I2}}(c)$ and ${\sf lcp_{S1}}(c)/{\sf lcp_{S2}}(c)$ to check the second condition of Lemma~\ref{lemma:3}.
+We design Algorithm~\ref{alg:2} to check $sa$ and $lcp$ when inducing them from $sa^*$ and $lcp^*$ using the IS method.
+For any $c \in \Sigma$, ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two integer sequences respectively induced into ${\sf sa\_bkt_L}(c)$ and ${\sf sa\_bkt_S}(c)$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two integer sequences respectively scanned from the left and right parts of the same bucket in $sa$.
+Algorithm~\ref{alg:2} computes and compares the fingerprints of these sequences to check the second condition of Lemma~\ref{lemma:3}.
 
@@
 	%\SetAlgoNoLine
-	\Fn{{\sf CheckByMethodA}($x, sa^*, lcp^*$)}{
+	\Fn{{\sf CheckerA}($x, sa^*, lcp^*$)}{
 
@@
 
-Method B is deterministic and specific for checking SA only.
-It uses a different way to ensure the correctness of $sa^*$.
+Method B is deterministic and specific for checking SA.
 Before our presentation, we first introduces a notation called $\overline{sa^*}$.
-The same as $sa^*$, $\overline{sa^*}$ also represents the suffix array for all the S*-type suffixes.
-Both $sa^*$ and $\overline{sa^*}$ are computed during the induction phase at the top recursion level, where the former is calculated from $sa_1$ in S\ref{induction_phase:1}'' and the latter is computed when inducing the S*-type suffixes into $sa$ in S\ref{induction_phase:4}''.
+The same as $sa^*$, $\overline{sa^*}$ also represents the SA for all the S*-type suffixes in $x$.
+Both $sa^*$ and $\overline{sa^*}$ are computed during the induction phase at the top recursion level, where the former is calculated from $sa_1$ in S\ref{induction_phase:1}'' and the latter is figured out when inducing S*-type suffixes into $sa$ in S\ref{induction_phase:4}''.
 
@@
 One way for deterministically checking the second condition is to copy both $sa^*$ and $\overline{sa^*}$ to external memory and compare their elements one by one after the induction phase.
-Algorithm~\ref{alg:3} applies a more space efficient way to checking equality of the two integer sequences with a negligible error probability.
-It reuses the Karp-Rabin fingerprinting technique to compute and compare the fingerprints of $sa^*$ and $\overline{sa^*}$, where the fingerprint of $sa^*$ is computed when placing the elements of $sa^*$ into $sa$ in S\ref{induction_phase:2}'' and the fingerprint of $\overline{sa^*}$ is computed when inducing the S*-type suffixes into $sa$ in S\ref{induction_phase:4}''.
+Algorithm~\ref{alg:3} applies a more space efficient way to checking equality of $sa^*$ and $\overline{sa^*}$ with a negligible error probability.
+It employs the Karp-Rabin fingerprinting technique to compute and compare the fingerprints of the two integer sequences, where the fingerprint of $sa^*$ is computed when placing its elements into $sa$ in S\ref{induction_phase:2}'' and the fingerprint of $\overline{sa^*}$ is computed when inducing the S*-type suffixes into $sa$ in S\ref{induction_phase:4}''.
 
@@
 	%\SetAlgoNoLine
-	\Fn{{\sf CheckByMethodB}($x, sa^*$)}{
+	\Fn{{\sf CheckerB}($x, sa^*$)}{
 
@@
 Notice that two equal sequences must be mapped to an identical hash value, but the inverse is not always true.
-To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting techniques to compute these values following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$.
+To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting technique to compute these values following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$.
 By setting $L$ to a large value, the error probability can be reduced to a negligible level.
-In Fig.~\ref{fig:example2}, we depict an example for computing the integer array $A$ identical to $sa^*$ and $\overline{sa^*}$ in Fig.~\ref{fig:example1}.
+We depict an example in Fig.~\ref{fig:example2} for computing an integer array $A$, where the array is equal to $sa^*$ and $\overline{sa^*}$ shown in Fig.~\ref{fig:example1}.
 Given $L = 23$ and $\delta = 5$, we first compute the fingerprint of $A[0, 0]$ in line 4.
-Because $\sf fp(A[0, -1]) = 0$, $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$.
-By iteratively computing the fingerprints in this way, we finally obtain the fingerprint of $A$ in line 7.
+Because $\sf fp(A[0, -1]) = 0$, we have $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$.
+By iteratively computing the fingerprints of prefixes of $A$ in this way, we finally obtain ${\sf fp}(A[0, 3])$ in line 7.
 
@@
 
-	\caption{An example for calculating fingerprints by Karp-Rabin fingerprinting function.}
+	\caption{An example for calculating fingerprints using the Karp-Rabin fingerprinting technique.}
 
@@
 
-Algorithms~\ref{alg:2} and ~\ref{alg:3} take linear time and constant RAM space for computing the target sequences by the Karp-Rabin fingerprinting technique.
-Algorithm~\ref{alg:2} also employs the checker presented in~\cite{wu2017} to verify $sa^*$ and $lcp^*$ within an integer sorting complexity.
-Because at most one out of every two suffixes is S*-type and the checking process is only executed at the top recursion level, the overhead for checking $sa^*$ and $lcp^*$ is considerably smaller than that for constructing the two arrays by an IS builder.
+Algorithm~\ref{alg:3} takes linear time and constant RAM space to compute the required fingerprints.
+For Algorithm~\ref{alg:2}, it requires $\mathcal{O}(\|\Sigma\|)$ RAM space to cache the fingerprints of induced and scanned values in $sa$ and $lcp$ buckets.
+Besides, Algorithm~\ref{alg:2} also employs the checker presented in~\cite{wu2017} to verify $sa^*$ and $lcp^*$ within an integer sorting complexity.
+Because at most one out of every two suffixes is S*-type and the checking process is only executed at the top recursion level, the overhead for checking $sa^*$ and $lcp^*$ is much less than that for building the two arrays by an IS builder.
 
@@
 
-By assuming the induction phase at the top recursion level is correctly implemented, we propose two methods that enable an IS builder to build and check an SA simultaneously.
+By assuming the induction phase at the top recursion level is correctly implemented, we propose two methods that enable any IS builder to build and check an SA simultaneously.
 The probabilistic algorithm designed by Method B is rather lightweight, it takes negligible time and space to run compared to the existing IS suffix sorting and checking algorithms.
